### 2016-09-06 ms_caf/ms_caf.tex
initial commit

\documentclass[10pt,twocolumn,preprint]{emulateapj}
\usepackage[english]{babel}
\usepackage{amsmath}

\newcommand{\bs}{\boldsymbol}
\newcommand{\diff}{{\mathrm d}}
\newcommand{\Cov}{\mathsf{C}}
\newcommand{\Fish}{\mathsf{F}}
\newcommand{\Cl}{\mathcal {C}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\T}{\mathcal{T}}
\newcommand{\Msun}{M_\odot}

%\shortauthors{Author et al.} \shorttitle{New CubeP3M}
\begin{document}
\title{New $N$-body Algorithm}
\author{Authors}

\affil{Affiliations}

\email{email@email.email}

\begin{abstract}
We present a memory-lite $N$-body algorithm.
\end{abstract}

\keywords{}

\maketitle

\section{Introduction}
$N$-body simulations are important. Citation \cite{2000ApJ...534L..19P}.  Bold type use $\bs{v}=\vec{v}$.

\begin{figure} \centering
  \includegraphics[width=1.0\linewidth]{fig1.pdf}
  \caption{Caption goes here.}
  \label{fig.1}
\end{figure}

\section{Conclusion}
We present a memory-lite $N$-body algorithm.

\acknowledgements
Thanks.

\bibliographystyle{hapj}
\bibliography{../haoran_ref}

\end{document}

### 2016-09-06 ms_disp1/ms_disp1.tex
initial commit

\documentclass[aps,prd,twocolumn,superscriptaddress,amsfont,amssymb,amsmath,nofootinbib,showpacs,balancelastpage]{revtex4-1}
%\documentclass[prd,superscriptaddress,twocolumn,floatfix]{revtex4}

\usepackage{graphicx,longtable,natbib,mathrsfs,color}
\usepackage{txfonts}

\newcommand{\bs}{\boldsymbol}
\newcommand{\diff}{{\mathrm d}}
\newcommand{\Cov}{\mathsf{C}}
\newcommand{\Fish}{\mathsf{F}}
\newcommand{\Cl}{\mathcal {C}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\T}{\mathcal{T}}
\newcommand{\Msun}{M_\odot}

%
\def\apjl{Astrophys. J. Lett.}
\def\apjs{Astrophys. J. Suppl. Ser.}
\def\aj{Astron. J.}
\def\mnras{Mon. Not. R. Astron. Soc.}
\def\aap{Astron. Astrophys.}                % Astronomy and Astrophysics
\def\jcap{J. Cosmology Astropart. Phys.}
\def\aapr{Astron. Astrophys. Rev.}

\begin{document}

\addtolength{\hoffset}{-0.525cm}
\addtolength{\textwidth}{1.05cm}
\title{Displacement field paper}

%\author{Hao-Ran~Yu} \affiliation{}\email{haoran@cita.utoronto.ca}
%\author{Qiaoyin~Pan}
%\author{Ue-Li~Pen} \affiliation{}
%\author{Tong-Jie~Zhang} \affiliation{}
%---

%\date{Received \today; published -- 00, 0000}

\begin{abstract}
Abstract goes here.
\end{abstract}

%\pacs{98.80.Es 95.36.+x}

\maketitle

\section{Introduction}\label{sec.intro}
Introduction goes here. Citation \cite{2000ApJ...534L..19P}. Bold type use $\bs{v}=\vec{v}$.

\section{Method}\label{sec.method}
Method goes here.

\begin{figure} \centering
  \includegraphics[width=1.0\linewidth]{fig1.pdf}
  \caption{Caption goes here.}
  \label{fig.1}
\end{figure}

\section{Discussion and conclusion}\label{sec.discussion}
Discussion goes here.

\section*{Acknowledgements}
Acknowledgements goes here.

%\bibliographystyle{apsrev4-1}
\bibliographystyle{h-physrev3}
\bibliography{../haoran_ref}

\end{document}

### 2016-09-11 ms_caf/ms_caf.tex
caf_paper update

@@
 
-%\shortauthors{Author et al.} \shorttitle{New CubeP3M}
+\shortauthors{Author et al.} \shorttitle{New CubeP3M}
 \begin{document}
-\title{New $N$-body Algorithm}
+\title{Compact Particle-Mesh $N$-body Algorithm}
 \author{Authors}
@@
 
-\email{email@email.email}
+\email{haoran@cita.utoronto.ca}
 
 \begin{abstract}
-We present a memory-lite $N$-body algorithm.
+$N$-body algorithms are computation-light but memory-heavy imbalanced problems in supercomputing.
+Traditional $N$-body simulation algorithms use at least 24-byte per particle in memory, where six 4-byte single precision real numbers keep track of each particle's phase space coordinate.
+This is usually the bottle neck of scaling-up of a $N$-body problem on modern supercomputers.
+Here we present a parallel, memory-lite Particle-Mesh based $N$-body algorithm, in which each particle can occupy as low as 9-byte memory.
+This is accomplished by storing only particles' relative location respect to a meshed grid, and their velocity relative to a predictable distribution function.
+The remaining information is given by the ordering of particles in memory space, and a complimentary density field, which is negligible in memory.
+Our numerical experiments show that this algorithm provides acceptable accuracy compared to traditional algorithm in cosmological simulations.
+
 \end{abstract}
@@
 \section{Introduction}
-$N$-body simulations are important. Citation \cite{2000ApJ...534L..19P}.  Bold type use $\bs{v}=\vec{v}$.
+$N$-body simulation is a powerful tool to solve the highly nonlinear dynamic problems \citep{1988csup.book.....H}.
+It is widely used in cosmology to model the formation and evolution of the large scale structure (LSS).
+With the fast development of parallel supercomputers, we are able to simulate a system of more than a trillion ($10^{12}$) $N$-body particles.
+To date the largest $N$-body simulation in application is the ``TianNu'' simulation run on the TianHe-2 supercomputer by cosmological simulation code {\tt CUBEP3M} \citep{2013MNRAS.436..540H}.
+It uses nearly $3\times 10^{12}$ particles to simulate the cold dark matter (CDM) and cosmic neutrino evolution through the cosmic age.
+However {\tt CUBEP3M} and TianNu did not fully use TianHe-2's computing power -- 87\% of the computing power on TianHe-2 are given by the Intel Xeon-Phi coprocessors while they occupy only 27\% of the total memory.
+Although we accomplished preliminary works on developing native and offload modes of {\tt CUBEP3M} on Xeon-Phi many integrated core (MIC) architecture, the memory problem is still the bottle neck of scaling up the $N$-body problem on the heterogeneous architectures.
+Lower the memory usage per particle will solve this problem, and will also help the $N$-body problem development on GPUs, which are even more massive parallelized and memory limited.
+In this paper we present a $N$-body algorithm which uses as low as 9 byte per particle (here after we use ``bpp'' referring to ``byte per particle'') and show that it can give accurate cosmological LSS simulations.
+The algorithm is presented in \S 2, and a comparison between this method and traditional method is shown in \S 3.
+Discussions and conclusions are in \S 4.
 
-\begin{figure} \centering
-  \includegraphics[width=1.0\linewidth]{fig1.pdf}
-  \caption{Caption goes here.}
-  \label{fig.1}
+\section{Algorithm}
+The most memory consuming part of a $N$-body simulation is usually the phase space coordinates of each $N$-body particles -- 24 bpp (4 single precision real numbers) must be used to store each particles' position and velocity.
+A memory-lite parallel $N$-body code {\tt CUBEP3M} can use as low as 40 bpp in sacrificing the speed.
+This includes the phase coordinates (24 bpp) for particles in physical domain and buffered region, a linked list (4 bpp), and a global coarse mesh and local fine mesh.
+We attempt to replace the coordinates and linked list 24+4=28 bpp memory usage with a integer based storage, described as follows.
+
+\subsection{Particle location storage}
+Instead of storing the particle's global coordinates, we store its relative position respective to the coarse cell who contains the particle.
+We divide the coarse cell, in each dimension, evenly into $2^8=256$ bins, and we use a 1-byte integer (with range 0 to 255 or -128 to 127) to show which bin it locates in this dimension.
+Next, global locations of particles are given by cell-ordered format in memory space, and a complimentary number count of particle number in this coarse mesh (coarse density field) will give complete information of particle distribution in cells.
+Since the this field is usually coarser than particle distribution, it takes negligible memory.
+This coarse density field can be further compressed into 1-byte integer format, such that a 1-byte unsigned integer show the particle number in this coarse cell in range 0 to 255.
+In the densest cells (rarely happened) where there are $\ge 255$ particles, we can just write 255, and write the actual number as a 4-byte integer in another file.
+In a simulation with volume $L^3$ and $N_c^3$ coarse cells, particle positions are stored with a precision of $L/(256N_c)$, which is usually 10 times finer than the force resolution.
+
+\subsection{Particle velocity storage}\label{s.velocity}
+Velocities, in each dimension, are compressed to 2-byte integers representing their index within a histogram containing $2^{16}$ equally spaced bins ranging from $-v_{\rm max}$ to $v_{\rm max}$, where $v_{\rm max}$ is the maximum absolute particle velocity.
+$v_{\rm max}$ can be derived from either computation or prediction.
+We prefer the latter case.
+In cosmological simulations, velocity distribution of matter fluids can be predicted by linear theory, given a model and redshift $z$ (time of the simulation).
+Then by any time step, we have a cumulative distribution function (CDF) $F(|\bs{v}|)$ of particles' velocity (see appendix A).
+We, instead, equally divide the $y$-axis of this CDF into $2^{16}$ bins, and the 2-byte integer shows which velocity bin the particle locates, in each dimension.
+This step does finer sampling on low-velocity end in velocity space and increases the simulation accuracy.
+This CDF may introduce velocity cutoffs, that particle cannot exceed some certain $\bs{v}_{\rm max}(z)$ given by the model.
+This process prevents the nonphysical scattering behaviours of particles, which do not exist in fluids.
+Practically, simulation results are very insensitive to the choice of velocity distribution.
+Even a top-hat distribution in range $-v_{\rm max}$ to $v_{\rm max}$ give accurate results.
+One can just assume a Gaussian distribution and whose standard deviation is given by the model and redshift $\sigma(\bs{\theta},z)$.
+Proper choice of model only optimizes the velocity space sampling, but does not affect the physics.
+
+\subsection{Spacial decomposition}
+{\tt CUBE} uses cubic decomposition structures.
+The global simulation volume is decomposed into ${\tt nn}^3$ cubic sub-volumes, and each of these are assigned to a coarray {\it image}\footnote{Images are the concept of computing nodes or MPI tasks in Coarray Fortran.
+We use this terminology in this paper.}.
+Inside of an image, the sub-volume is further decomposed into ${\tt nnt}^3$ cubic {\it tiles}.
+Each tile is surrounded by a {\it buffer} region which is usually ${\tt ncb}$ coarse cells thick.
+The buffer is designed for two reasons: (1) computing the fine mesh force, whose cut-off ${\tt nforce\_cutoff}\leq{\tt ncb}$, and (2) collecting all possible particles travelling from a tiles buffer region to its center, {\it physical} region.
+Thus, a integer-based coarse mesh density is defined as
+\begin{equation*}
+{\tt integer(1)\ rho\_c(nex,nex,nex,nnt,nnt,nnt)[nn,nn,*]}\\
+\end{equation*}
+where ${\tt nex}={\tt nt}+2{\tt ncb}$ covers the buffer region on both sides, {\tt nnt} is the tile dimemsions, and {\tt nn} is the image {\it codimensions}\footnote{Coarray Fortran concept.
+Codimensions can do communications between images.}.
+Particles' phase coordinates {\tt x} and {\tt v} are copied between tiles and images, when necessary.
+\begin{figure}[h]
+\begin{verbatim}
+    program CUBE
+      call initialize
+      call particle_initialize
+      sync all
+      call buffer_density
+      call buffer_x
+      call buffer_v
+      do
+        call timestep
+        call update_x
+        call buffer_density
+        call buffer_x
+        call PM ! update_v
+        call buffer_v
+        if(checkpoint_step) then
+           call checkpoint
+           if (final_step) exit
+        endif
+      enddo
+      call finalize
+    endprogram
+\end{verbatim}
+\caption{Overall structure of {\tt CUBE}.}
+\label{fig.code}
 \end{figure}
 
-\section{Conclusion}
-We present a memory-lite $N$-body algorithm.
+\subsection{Code overview}
+By using the previous compact format storage, we modified the {\tt CUBEP3M} and utilized it in the outputs TianNu simulation.
+On average, 9.125 bpp is used for all outputs and saves 62 \% hard drive space on Tianhe-2.
+The compact version of {\tt CUBEP3M}, named {\tt CUBE}, accomplished the compact format in memory, and the code keeps only integer format for storing particles' phase space coordinates.
+{\tt CUBE} is written in Coarray Fortran, where the coarray features are used to do MPI communications.
+Here we review the methodology of solving this integer based $N$-body problem.
+Like its predecessor {\tt CUBEP3M}, {\tt CUBE} uses two-level mesh grids for solving the Poisson equations.
+The advantages of this are described in \cite{2013MNRAS.436..540H}.
+
+Figure \ref{fig.code} shows the overall structure of the code.
+Subroutine {\tt initialize} creates necessary FFT plans and read in configuration files telling the program at which redshifts we need to do checkpoints, halofinds, or stop the simulation.
+Force kernels are also created or read in here.
+In {\tt particle\_initialize}, for each image, we read in initial positions and velocities in the compressed format.
+When this step is done by all images (synchronized by ``{\tt sync all}'', which is equivalent to a {\tt mpi\_barrier}), the density fields {\tt rho\_c} are buffered between tiles (communications between images are needed) by subroutine {\tt buffer\_density}.
+Then, particles' phase space coordinates {\tt x} and {\tt v} are also buffered, by {\tt buffer\_x} and {\tt buffer\_v} respectively.
+In these steps, {\tt x} and {\tt v}'s ordering is changed according to the updated {\tt rho\_c}.
+
+In each iteration of the main loop, we firstly call {\tt timestep}, where a increment of time {\tt dt} is controlled by particles' maximum velocities, accelerations, and cosmic expansions.
+According to {\tt dt}, subroutine {\tt update\_x} updates the positions of particles in a ``{\it gather}'' algorithm.
+${\tt x}={\tt x}+{\tt v}*{\tt dt}$ is executed twice, first time to determine a updated density field on the tile, {\tt rho\_c\_new}, and second time to generate a new particle list {\tt x\_new} and {\tt v\_new} on the tile.
+The reason for this repeated execution is the dependence of {\tt x\_new} and {\tt v\_new}'s value and ordering on both the old {\tt x}, {\tt v} and the new {\tt rho\_c\_new}.
+However, ${\tt x}={\tt x}+{\tt v}*{\tt dt}$ scales as $o(N)$ and is computational inexpensive.
+Although {\tt x\_new} and {\tt v\_new} arrays take extra memory, they do not appear simultaneously for multiple tiles, and compared to {\tt x} and {\tt v}, the extra memory overhead is by factor of $1/{\tt nnt}^3$.
+This requires the ``gather'' algorithm to move particles\footnote{In contrast, {\tt CUBEP3M} uses a ``{\it scatter}'' algorithm.} -- as we synchronize the correct \{{\tt rho\_c}, {\tt x}, {\tt v}\} in the buffer region of each tile, and the particles in the extended (physical+buffer) region are a superset of particles locating in the physical region in the next time step, given that ${\tt v}*{\tt dt}$ is not greater than the buffer depth.
+In {\tt update\_x}, \{{\tt rho\_c\_new}, {\tt x\_new}, {\tt v\_new}\} can be set to collect only the physical region of the tile, with the rest of this subset discarded.
+At the end, \{{\tt rho\_c}, {\tt x}, {\tt v}\} is replaced by \{{\tt rho\_c\_new}, {\tt x\_new}, {\tt v\_new}\}, and the memory of latter is freed up.
+
+Next, {\tt buffer\_density} and {\tt buffer\_x} are called again to synchronize the updated particle positions in order that tile-based local fine forces are computed correctly.
+
+A standard particle-mesh scheme is then performed in the {\tt PM} subroutine.
+First, by looping over tiles, fine force is calculated and velocities are updated on the physical region of each tile.
+The memory overhead of fine-mesh arrays are well controlled by $1/{\tt nnt}^3$.
+Next, global coarse force is solved by using a coarser (usually by factor of 4) mesh by dimensional splitting -- a distributed-memory cubic decomposed 3D coarse density field is interpolated by particles, and we Fourier transform data in consecutive three dimensions with global transposition in between (known as the pencil decomposition).
+After the multiplication of force kernels, the inverse transform takes place to get the cubic distributed coarse force field, upon which velocities are updated again.
+The memory usage of coarse arrays are negligible.
+An optional particle-particle (PP) force can be called to increase the force resolution and the velocities are updated last time according to this.
+As discussed in \S\ref{s.velocity}, updating velocities use model predicted velocity distribution, thus ${\tt v}={\tt v}+{\tt dv}$ is executed only once, scaling directly to the velocity distribution of the next time step ${\tt d}+{\tt dt}$.
+During each of these force calculations, maximum accelerations are collected serving for the {\tt timestep} in the next iteration, controlling next {\tt dt}.
+
+By far, particle velocities in the physical regions of each tile are updated.
+Particle locations in the buffer regions has updated before {\tt PM} and remained unchanged.
+So we simply call {\tt buffer\_v} again, such that the {\tt update\_x} in the next iteration will be done correctly.
+
+If a desired redshift is reached, we call {\tt checkpoint} and/or exit the main loop.
+The corresponding logical variables are controlled in {\tt timestep}.
+Finally, in {\tt finalize} subroutine we destroy all the FFT plans and finish up any timing or statistics taken in the simulation.
+
+\subsection{Memory layout}
+
+\section{Results}
+
+\section{Discussion and Conclusion}
+
+\appendix
+\section{Velocity distribution function}
+
+%\begin{figure} \centering
+%  \includegraphics[width=1.0\linewidth]{delta_mesh.pdf}
+%  \caption{Caption goes here.}
+%  \label{fig.1}
+%\end{figure}
 

### 2016-09-12 ms_caf/ms_caf.tex
haoran

@@
 
+\begin{figure}[t]
+\includegraphics[width=1.0\linewidth]{fig_tile.pdf}
+\caption{Spacial decomposition.}
+\label{fig.tile}
+\end{figure}
+
 \subsection{Spacial decomposition}
@@
 Particles' phase coordinates {\tt x} and {\tt v} are copied between tiles and images, when necessary.
-\begin{figure}[h]
+Figure \ref{fig.tile} shows the spacial decomposition in a 2-dimensional analogy, with ${\tt nn=2}$ and ${\tt nnt=2}$.
+
+\begin{figure}[t]
 \begin{verbatim}
+
     program CUBE
@@
 In {\tt particle\_initialize}, for each image, we read in initial positions and velocities in the compressed format.
+A initial condition generator is also written in the compressed format (see Appendix B).
 When this step is done by all images (synchronized by ``{\tt sync all}'', which is equivalent to a {\tt mpi\_barrier}), the density fields {\tt rho\_c} are buffered between tiles (communications between images are needed) by subroutine {\tt buffer\_density}.
@@
 \subsection{Memory layout}
+Here we summarize the memory usage in unit of bpp (byte per particle).
 
 \section{Results}
+We use the same initial conditions, and use {\tt CUBEP3M} and {\tt CUBE} to simulate the large scale structure formation separately and compare their results.
+
+Results of {\tt izipx=1}
+
+Results of {\tt izipx=2}
+
+\begin{figure}[h]
+\begin{verbatim}
+
+\end{verbatim}
+
+\caption{Power spectrum differential plot}
+\label{fig.power}
+\end{figure}
 
 \section{Discussion and Conclusion}
+Cosmological simulation, PM method.
+
+PID.
+
+Summarize memory.
+
+Phi, GPU.
+
+\clearpage
 
 \appendix
-\section{Velocity distribution function}
+\section{A. Velocity distribution function}
 
-%\begin{figure} \centering
-%  \includegraphics[width=1.0\linewidth]{delta_mesh.pdf}
-%  \caption{Caption goes here.}
-%  \label{fig.1}
-%\end{figure}
+\section{B. Initial condition generator}
 

### 2016-09-12 ms_caf/ms_caf.tex
update velocity dispersion formula by Xin

@@
 \section{A. Velocity distribution function}
+To optimize the storage of the velocity field, we need to estimate the velocity dispersion at each snapshot.
+At non-linear scale, the velocity distribution function is surely non-Gaussian.
+However, to the first order approximation, we simply assume it as Gaussian and characterized by the variance
+\begin{eqnarray}
+\sigma_v = \frac{1}{3}  (a H f D)^2 \int d^3q \frac{P_{L}(q)}{q^2} 
+\end{eqnarray}
+where $a(z)$ is the scale factor, $H(z)$ the Hubble parameter, $D$ is the linear growth factor, $f=d \ln D/d\ln a$, 
+and $P_L(k)$ is the linear power spectrum of density contrast at redshift zero.
 

### 2016-09-12 ms_disp1/ms_disp1.tex
displacement field Wiener filters

@@
 \begin{abstract}
-Abstract goes here.
+In this paper, we discover the ability of using displacement field to reconstruct initial linear perturbations in the early universe.
 \end{abstract}
@@
 \section{Introduction}\label{sec.intro}
-Introduction goes here. Citation \cite{2000ApJ...534L..19P}. Bold type use $\bs{v}=\vec{v}$.
+Introduction goes here. Citation example \cite{2013MNRAS.436..540H}.
+Cite Hong-Ming's 1D paper.
+
+The rest of this paper is structured as follows.
+In section \ref{sec.method} we describe the simulation and reconstruction algorithm.
+In section \ref{sec.results} we show the results of the reconstruction.
+Discussion and conclusion are in \ref{sec.discussion}.
 
 \section{Method}\label{sec.method}
-Method goes here.
+Method goes there.
 
-\begin{figure} \centering
-  \includegraphics[width=1.0\linewidth]{fig1.pdf}
-  \caption{Caption goes here.}
-  \label{fig.1}
-\end{figure}
+\subsection{Simulation}
+Cosmological parameters, simulation parameters.
+
+Initial conditions, get linear density field $\delta_L$.
+Run simulation by code {\tt CUBE} and get the nonlinear density field $\delta$ at $z=0$.
+{\tt CUBE} uses particle-ID (PID) to record the initial location of particles, and the information is tracked until the $z=0$ and we can get the displacement vector ${\bs s}\equiv{\bs x}-{\bs q}$ for every particle.
+Then these vectors are interpolated onto the initial Lagrangian coordinates ${\bs q}$ of particles and we get the displacement field ${\bs s}({\bs q})$.
+
+\subsection{Reconstruction}
+Get the raw reconstructed density field $\delta_R=-\nabla\cdot{\bs s}({\bs q})$.
+
+%\begin{figure}[t] \centering
+%  \includegraphics[width=1.0\linewidth]{fig1.pdf}
+%  \caption{Caption goes here.}
+%  \label{fig.1}
+%\end{figure}
+
+\section{Results}\label{sec.results}
+
+Wiener filter.
+
+We try to decompose $\delta_L$ into two components,
+\begin{equation}\label{eq.decompose}
+    \delta_L=r\delta_R+\delta_N,
+\end{equation}
+where $r\delta_R$ is completely correlated with $\delta_L$, and the uncorrelated part $\delta_N$ is generated by the nonlinear evolution.
+To do this, we first correlate equation (\ref{eq.decompose}) with $\delta_R$,
+\begin{equation}\label{eq.correlate}
+    \left\langle\delta_L\delta_R\right\rangle
+    =r\left\langle\delta_R\delta_R\right\rangle
+    +\left\langle\delta_N\delta_R\right\rangle,
+\end{equation}
+and vanish the second term in the right-hand-side.
+Then the $r$ is given by $\left\langle\delta_L\delta_R\right\rangle / \left\langle\delta_R\delta_R\right\rangle$ and mode-coupling term $\delta_N=\delta_L-r\delta_R$.
+We next square equation (\ref{eq.decompose}) and gives
+\begin{equation}\label{eq.square}
+    \left\langle\delta_L\delta_L\right\rangle
+    =r^2\left\langle\delta_R\delta_R\right\rangle
+    +\left\langle\delta_N\delta_N\right\rangle
+    +2r\left\langle\delta_R\delta_N\right\rangle,
+\end{equation}
+and the third term of right-hand-side also vanishes. By defining
+\begin{equation}
+    \left\langle\delta_L\delta_L\right\rangle
+    =b^2\left\langle\delta_R\delta_R\right\rangle,
+\end{equation}
+we write
+\begin{equation}
+    \left\langle\delta_N\delta_N\right\rangle=(b^2-r^2)\left\langle\delta_R\delta_R\right\rangle
+\end{equation}
+and a Wiener filter can be constructed as
+\begin{equation}
+    W=\frac{r^2\left\langle\delta_R\delta_R\right\rangle}
+    {r^2\left\langle\delta_R\delta_R\right\rangle+\left\langle\delta_N\delta_N\right\rangle}
+    =\frac{r^2}{b^2}.
+\end{equation}
+
+From equation (\ref{eq.decompose}), multiplying $W$, we obtain the optimal reconstructed linear density field by
+\begin{equation}
+    \tilde\delta_L=rW\delta_R=\frac{r^3}{b^2}\delta_R.
+\end{equation}
+From equation (\ref{eq.square}), multiplying $W^2$, the reconstructed linear power spectrum is giving by
+\begin{equation}
+\begin{aligned}
+    \tilde{P}&=W^2\left\langle\delta_L\delta_L\right\rangle  \\
+    &=r^2W^2\left\langle\delta_R\delta_R\right\rangle +W^2\left\langle\delta_N\delta_N\right\rangle  \\
+    &=b^2W^2\left\langle\delta_R\delta_R\right\rangle  \\
+    &=r^4b^{-2}\left\langle\delta_R\delta_R\right\rangle.
+\end{aligned}
+\end{equation}
 

### 2016-09-13 ms_caf/ms_caf.tex
haoran

@@
 \begin{abstract}
-$N$-body algorithms are computation-light but memory-heavy imbalanced problems in supercomputing.
+Cosmological large scale structure $N$-body simulations are computation-light but memory-heavy imbalanced problems in supercomputing.
 Traditional $N$-body simulation algorithms use at least 24-byte per particle in memory, where six 4-byte single precision real numbers keep track of each particle's phase space coordinate.
-This is usually the bottle neck of scaling-up of a $N$-body problem on modern supercomputers.
-Here we present a parallel, memory-lite Particle-Mesh based $N$-body algorithm, in which each particle can occupy as low as 9-byte memory.
+This is usually the bottleneck of scaling-up of a $N$-body problem on modern supercomputers whose core/memory ratio is high.
+Here we present a parallel, memory-lite Particle-Mesh based $N$-body code, in which each particle can occupy as low as 6-byte memory.
 This is accomplished by storing only particles' relative location respect to a meshed grid, and their velocity relative to a predictable distribution function.
@@
 Lower the memory usage per particle will solve this problem, and will also help the $N$-body problem development on GPUs, which are even more massive parallelized and memory limited.
-In this paper we present a $N$-body algorithm which uses as low as 9 byte per particle (here after we use ``bpp'' referring to ``byte per particle'') and show that it can give accurate cosmological LSS simulations.
+In this paper we present a $N$-body algorithm which uses as low as 6 byte per particle (here after we use ``bpp'' referring to ``byte per particle'') and show that it can give accurate cosmological LSS simulations.
 The algorithm is presented in \S 2, and a comparison between this method and traditional method is shown in \S 3.

### 2016-09-13 ms_disp1/ms_disp1.tex
haoran

@@
 Wiener filter.
+Define
+\begin{equation}
+    r\equiv \frac{\left\langle\delta_L\delta_R\right\rangle}
+    {\sqrt{\left\langle\delta_L\delta_L\right\rangle
+    \left\langle\delta_R\delta_R\right\rangle}}
+\end{equation}
+and
+\begin{equation}
+    b^2\equiv \frac{\left\langle\delta_R\delta_R\right\rangle}
+    {\left\langle\delta_L\delta_L\right\rangle}.
+\end{equation}
 
@@
 \begin{equation}\label{eq.decompose}
-    \delta_L=r\delta_R+\delta_N,
+    \delta_L=r'\delta_R+\delta_N,
 \end{equation}
-where $r\delta_R$ is completely correlated with $\delta_L$, and the uncorrelated part $\delta_N$ is generated by the nonlinear evolution.
+where $r'\delta_R$ is completely correlated with $\delta_L$, and the uncorrelated part $\delta_N$ is generated by the nonlinear evolution.
 To do this, we first correlate equation (\ref{eq.decompose}) with $\delta_R$,
@@
     \left\langle\delta_L\delta_R\right\rangle
-    =r\left\langle\delta_R\delta_R\right\rangle
+    =r'\left\langle\delta_R\delta_R\right\rangle
     +\left\langle\delta_N\delta_R\right\rangle,
 \end{equation}
-and vanish the second term in the right-hand-side.
-Then the $r$ is given by $\left\langle\delta_L\delta_R\right\rangle / \left\langle\delta_R\delta_R\right\rangle$ and mode-coupling term $\delta_N=\delta_L-r\delta_R$.
+and vanish the second term in the right-hand-side. Then
+\begin{equation}
+    r'=\frac{\left\langle\delta_L\delta_R\right\rangle}
+    {\left\langle\delta_R\delta_R\right\rangle}=\frac{r}{b}.
+\end{equation}
+Then the mode-coupling term 
+\begin{equation}
+    \delta_N=\delta_L-\frac{r}{b}\delta_R.
+\end{equation}
 We next square equation (\ref{eq.decompose}) and gives
@@
     \left\langle\delta_L\delta_L\right\rangle
-    =r^2\left\langle\delta_R\delta_R\right\rangle
+    =\frac{r^2}{b^2}\left\langle\delta_R\delta_R\right\rangle
     +\left\langle\delta_N\delta_N\right\rangle
-    +2r\left\langle\delta_R\delta_N\right\rangle,
+    +2\frac{r}{b}\left\langle\delta_R\delta_N\right\rangle,
 \end{equation}
-and the third term of right-hand-side also vanishes. By defining
+and the third term of right-hand-side also vanishes. As a result,
 \begin{equation}
-    \left\langle\delta_L\delta_L\right\rangle
-    =b^2\left\langle\delta_R\delta_R\right\rangle,
-\end{equation}
-we write
-\begin{equation}
-    \left\langle\delta_N\delta_N\right\rangle=(b^2-r^2)\left\langle\delta_R\delta_R\right\rangle
+    \left\langle\delta_N\delta_N\right\rangle
+    =\frac{1-r^2}{b^2}
+    \left\langle\delta_R\delta_R\right\rangle
 \end{equation}
@@
 \begin{equation}
-    W=\frac{r^2\left\langle\delta_R\delta_R\right\rangle}
-    {r^2\left\langle\delta_R\delta_R\right\rangle+\left\langle\delta_N\delta_N\right\rangle}
-    =\frac{r^2}{b^2}.
+    W=\frac{r^2}{r^2+1-r^2}=r^2.
 \end{equation}
@@
 \begin{equation}
-    \tilde\delta_L=rW\delta_R=\frac{r^3}{b^2}\delta_R.
+    \tilde\delta_L=W\frac{r}{b}\delta_R=\frac{r^3}{b}\delta_R.
 \end{equation}
@@
     \tilde{P}&=W^2\left\langle\delta_L\delta_L\right\rangle  \\
-    &=r^2W^2\left\langle\delta_R\delta_R\right\rangle +W^2\left\langle\delta_N\delta_N\right\rangle  \\
-    &=b^2W^2\left\langle\delta_R\delta_R\right\rangle  \\
-    &=r^4b^{-2}\left\langle\delta_R\delta_R\right\rangle.
+    &=r^2\frac{r^2}{b^2}\left\langle\delta_R\delta_R\right\rangle +r^2\left\langle\delta_N\delta_N\right\rangle  \\
+    &=\frac{r^2}{b^2}\left\langle\delta_R\delta_R\right\rangle.
 \end{aligned}

### 2016-09-20 ms_disp1/ms_disp1.tex
Sept 19 by haoran: disp1 paper; caf & init compile

@@
 \addtolength{\textwidth}{1.05cm}
-\title{Displacement field paper}
+\title{Reconstruction of Large Scale Structure by Nonlinear Displacement Fields}
 
@@
 \begin{abstract}
-In this paper, we discover the ability of using displacement field to reconstruct initial linear perturbations in the early universe.
+We investigate the ability of using nonlinear displacement field to reconstruct the primordial linear perturbations of the large scale structure.
+According to the linear Lagrangian perturbation theory, the first order primordial linear density field is given by the negative divergence of the displacement field.
+By running $N$-body simulations, we show that this reconstruction algorithm recovers great amount of nonlinear information in the early universe.
+In reality, the displacement fields can be solved from only the final stage of the density fields.
+This has potential to reconstruct baryonic acoustic oscillation (BAO) from current and future large scale structure surveys.
 \end{abstract}
@@
 \section{Introduction}\label{sec.intro}
-Introduction goes here. Citation example \cite{2013MNRAS.436..540H}.
-Cite Hong-Ming's 1D paper.
+Our universe starts from primordial Gaussian perturbations at a very early stage, and from those fluctuations, the gravitational instability drives the formation of the large scale structure (LSS) distribution of matter.
+These structures grow linearly until the perturbations are large enough such that first order perturbation theories are unable to analytically describe the LSS distributions.
+Inversely, the final stage nonlinear LSS distribution contains higher order statistics, and thus makes it more challenging to interpret the basic cosmological parameters.
+This leads to various attempts to recover earlier stages of LSS, in which statistics are closer to Gaussian.
+Because Gaussian fields can be adequately described by two-point statistics, ideally after some recovery algorithms, more information can be extracted, more straightforwardly, by power spectra or two-point correlation functions.
 
+These algorithms include nonlinear Wiener filters in wavelet space, distribution function transforms etc.
+Standard BAO reconstruction algorithms smooths the nonlinear density field on linear scale ($\sim$10 Mpc$/h$) and reverse the large scale bulk flows by a negative Zel'dovich linear displacement.
+Here we propose a new reconstruction method that uses the nonlinear displacement field to recover the primordial density field.
+In the linear Lagrangian perturbation theory, the negative divergence of the the displacement field respect to Lagrangian coordinates gives the linear density field.
+Of course, the full displacement field is non-observable, as it requires the initial distributions of matter, however there are many techniques to estimate the nonlinear displacement field from a final distribution of matter.
+For example, when a homogeneous initial matter distribution is assumed, there is a unique solution of curl-less displacement field to relate the initial and final distributions without shell-crossing.
+This solution can be solved by a metric transformation equation \citep{1995ApJS..100..269P,1998ApJS..115...19P}.
+In 1-dimensional case, the solution simplifies to a reordering of matter elements by Eulerian positions, and [Hong-ming et.
+al.] studies this case.
+The homogeneous and curl-less assumptions are generally valid in cosmology [cite or proof].
+Shell-crossing does exist, and this effect can be quantified by $N$-body simulations.
+
+In this paper we track the initial and final distributions of particles in LSS $N$-body simulations, and use the displacement field to reconstruct the initial linear density field.
 The rest of this paper is structured as follows.
@@
 \section{Method}\label{sec.method}
-Method goes there.
 
 \subsection{Simulation}
-Cosmological parameters, simulation parameters.
+We use the public cosmological simulation code {\tt CUBE}.
+Cosmological parameters are in accordance with Planck.
+Initial conditions are generated at redshift $z=100$ using Zel'dovich approximation.
+$N_p=512^3$ $N$-body particles are evolved via their mutual gravitational interactions to $z=0$, in a periodic box with $L=300$ Mpc$/h$ per side.
+The code is set to use standard a particle-mesh (PM) algorithm \cite{1988csup.book.....H} on a two-level mesh grids (details see \cite{2013MNRAS.436..540H}) and cloud-in-cell (CIC) is used in particle interpolations in force calculation and obtaining the density field $\rho$ at late stages.
 

### 2016-09-20 cafcube_code/compile_caf.sh

@@ -2,6 +2,7 @@ module purge
 module load gcc/5.4.0
 module load openmpi/2.0.0-gcc-5.4.0
 module load fftw/3.3.5-gcc-5.4.0-openmpi-2.0.0
+module list
 
 make clean
 make

### 2016-09-20 cafcube_code/init/compile_init.sh

@@ -1,7 +1,14 @@
 #/opt/fftw/3.3.5-gcc-5.4.0-openmpi-2.0.0/lib/
 #/opt/fftw/3.3.5-gcc-5.4.0-openmpi-2.0.0/include/
+module purge
+module load gcc/5.4.0
+module load openmpi/2.0.0-gcc-5.4.0
+module load fftw/3.3.5-gcc-5.4.0-openmpi-2.0.0
 
-rm *.o *.out
+module list
+
+rm -f *.o *.out *.mod *~
+
 mpif90 -O3 -cpp -fcoarray=single -mcmodel=medium -c ../parameters.f90 -o parameters.o
 
 mpif90 -O3 -cpp -fcoarray=single -mcmodel=medium -Dpenfft_4x -c penfft_config.f90

### 2016-09-20 ms_disp1/ms_disp1.tex
haoran: bug fix in initial_conditions.f90; disp1 paper update

@@
 \newcommand{\Msun}{M_\odot}
+\newcommand{\lb}{\left\langle}
+\newcommand{\rb}{\right\rangle}
 
@@
 Cosmological parameters are in accordance with Planck.
-Initial conditions are generated at redshift $z=100$ using Zel'dovich approximation.
+Initial conditions are generated at redshift $z=50$ using Zel'dovich approximation.
 $N_p=512^3$ $N$-body particles are evolved via their mutual gravitational interactions to $z=0$, in a periodic box with $L=300$ Mpc$/h$ per side.
-The code is set to use standard a particle-mesh (PM) algorithm \cite{1988csup.book.....H} on a two-level mesh grids (details see \cite{2013MNRAS.436..540H}) and cloud-in-cell (CIC) is used in particle interpolations in force calculation and obtaining the density field $\rho$ at late stages.
-
-Initial conditions, get linear density field $\delta_L$.
-Run simulation by code {\tt CUBE} and get the nonlinear density field $\delta$ at $z=0$.
-{\tt CUBE} uses particle-ID (PID) to record the initial location of particles, and the information is tracked until the $z=0$ and we can get the displacement vector ${\bs s}\equiv{\bs x}-{\bs q}$ for every particle.
-Then these vectors are interpolated onto the initial Lagrangian coordinates ${\bs q}$ of particles and we get the displacement field ${\bs s}({\bs q})$.
+The code is set to use standard a particle-mesh (PM) algorithm \cite{1988csup.book.....H} on a two-level mesh grids (details see \cite{2013MNRAS.436..540H}) and cloud-in-cell (CIC) is used in particle interpolations in force calculation and obtaining the density field $\rho({\bs x})$ in Eulerian coordinates ${\bs x}$ at late stages.
+We use density contrast $\delta\equiv\rho/\lb\rho\rb-1$ to describe the density fluctuations.
+The primordial linear density field $\delta_L$ is given by the initial stage and scaled to $z=0$ by the linear growth factor.
 
 \subsection{Reconstruction}
-Get the raw reconstructed density field $\delta_R=-\nabla\cdot{\bs s}({\bs q})$.
+In the simulation, we uses particle-ID (PID) to record the initial location ${\bs q}$ of particles, and the information is tracked until the $z=0$ and we can get the Lagrangian displacement vector ${\bs \Psi}\equiv{\bs x}-{\bs q}$ for every particle.
+Then these vectors are interpolated onto the initial Lagrangian coordinates ${\bs q}$ of particles and we get the displacement field ${\bs \Psi}({\bs q})$.
+The raw reconstructed density field is given by the differential motion of matter elements,
+\begin{equation}
+    \delta_R=-\nabla\cdot{\bs \Psi}({\bs q}).
+\end{equation}
+
+To quantify the linear information in the reconstructed density field, we decompose $\delta_R$ in Fourier space as
+\begin{equation}\label{eq.decompose}
+    \delta_R(k)=r'\delta_L+\delta_N,
+\end{equation}
+where $r'\delta_R$ is completely correlated with linear density $\delta_L$. Correlating equation (\ref{eq.decompose}) with $\delta_L$ gives
+\begin{equation}
+    P_{LR}=r'P_{LL}+P_{LN}
+\end{equation}
+where $P_{ij}\equiv\lb\delta_i\delta_j\rb$ denotes the cross power spectrum.
+Since $\delta_N$ is uncorrelated with $\delta_L$, $P_{LN}=0$.
+With the definition of cross correlation coefficient $r\equiv P_{LR}/\sqrt{P_{LL}P_{RR}}$ and bias $b^2=P_{RR}/P_{LL}$, we solve $r'=P_{LR}/P_{LL}=rb$.
+According to equation (\ref{eq.decompose}), the auto power spectrum is decomposed as
+\begin{equation}
+    P_{RR}=r^2b^2P_{LL}+P_{NN},
+\end{equation}
+and $P_{NN}=(1-r^2)P_{RR}$. Then we construct a Wiener filter to filter out the uncorrelated part in $\delta_R$:
+\begin{equation}
+    W(k)=\frac{r^2b^2P_{LL}}{r^2b^2P_{LL}+P_{NN}}=r^2.
+\end{equation}
+The estimated linear density, or the optimal reconstructed density is given by
+\begin{equation}
+    \tilde\delta_R=Wb^{-1}\delta_R,
+\end{equation}
+and the optimal reconstructed power spectrum is given by
+\begin{equation}
+    \tilde P=W^2b^{-2}P_{RR}=W^2P_{LL}+W^2b^{-2}P_{NN}.
+\end{equation}
 
@@
 \section{Results}\label{sec.results}
-
-Wiener filter.
-Define
-\begin{equation}
-    r\equiv \frac{\left\langle\delta_L\delta_R\right\rangle}
-    {\sqrt{\left\langle\delta_L\delta_L\right\rangle
-    \left\langle\delta_R\delta_R\right\rangle}}
-\end{equation}
-and
-\begin{equation}
-    b^2\equiv \frac{\left\langle\delta_R\delta_R\right\rangle}
-    {\left\langle\delta_L\delta_L\right\rangle}.
-\end{equation}
-
-We try to decompose $\delta_L$ into two components,
-\begin{equation}\label{eq.decompose}
-    \delta_L=r'\delta_R+\delta_N,
-\end{equation}
-where $r'\delta_R$ is completely correlated with $\delta_L$, and the uncorrelated part $\delta_N$ is generated by the nonlinear evolution.
-To do this, we first correlate equation (\ref{eq.decompose}) with $\delta_R$,
-\begin{equation}\label{eq.correlate}
-    \left\langle\delta_L\delta_R\right\rangle
-    =r'\left\langle\delta_R\delta_R\right\rangle
-    +\left\langle\delta_N\delta_R\right\rangle,
-\end{equation}
-and vanish the second term in the right-hand-side. Then
-\begin{equation}
-    r'=\frac{\left\langle\delta_L\delta_R\right\rangle}
-    {\left\langle\delta_R\delta_R\right\rangle}=\frac{r}{b}.
-\end{equation}
-Then the mode-coupling term 
-\begin{equation}
-    \delta_N=\delta_L-\frac{r}{b}\delta_R.
-\end{equation}
-We next square equation (\ref{eq.decompose}) and gives
-\begin{equation}\label{eq.square}
-    \left\langle\delta_L\delta_L\right\rangle
-    =\frac{r^2}{b^2}\left\langle\delta_R\delta_R\right\rangle
-    +\left\langle\delta_N\delta_N\right\rangle
-    +2\frac{r}{b}\left\langle\delta_R\delta_N\right\rangle,
-\end{equation}
-and the third term of right-hand-side also vanishes. As a result,
-\begin{equation}
-    \left\langle\delta_N\delta_N\right\rangle
-    =\frac{1-r^2}{b^2}
-    \left\langle\delta_R\delta_R\right\rangle
-\end{equation}
-and a Wiener filter can be constructed as
-\begin{equation}
-    W=\frac{r^2}{r^2+1-r^2}=r^2.
-\end{equation}
-
-From equation (\ref{eq.decompose}), multiplying $W$, we obtain the optimal reconstructed linear density field by
-\begin{equation}
-    \tilde\delta_L=W\frac{r}{b}\delta_R=\frac{r^3}{b}\delta_R.
-\end{equation}
-From equation (\ref{eq.square}), multiplying $W^2$, the reconstructed linear power spectrum is giving by
-\begin{equation}
-\begin{aligned}
-    \tilde{P}&=W^2\left\langle\delta_L\delta_L\right\rangle  \\
-    &=r^2\frac{r^2}{b^2}\left\langle\delta_R\delta_R\right\rangle +r^2\left\langle\delta_N\delta_N\right\rangle  \\
-    &=\frac{r^2}{b^2}\left\langle\delta_R\delta_R\right\rangle.
-\end{aligned}
-\end{equation}
 

### 2016-09-20 cafcube_code/init/initial_conditions.f90

@@ -207,10 +207,10 @@ enddo
   !!!!! test only
 #endif
 
-call cross_power(xi,cube,cube)
-open(11,file='initpower.dat',access='stream')
-write(11) xi
-close(11)
+!call cross_power(xi,cube,cube)
+!open(11,file='initpower.dat',access='stream')
+!write(11) xi
+!close(11)
 
 call fft_cube2pencil_fine
 call trans_zxy2xyz_fine
@@ -286,9 +286,9 @@ if (correct_kernel) then
   call trans_xyz2zxy_fine
   call ifft_pencil2cube_fine
   
-  open(11,file='laplace.dat',access='stream')
-  write(11) cube
-  close(11)
+  !open(11,file='laplace.dat',access='stream')
+  !write(11) cube
+  !close(11)
 
   sync all
 
@@ -339,15 +339,15 @@ cx=real(cx)*cx_temp
 cx_temp=cx  ! backup phi(k)
 
 !!!!!! output phi in real space
-call trans_xyz2zxy_fine
-call ifft_pencil2cube_fine
-phi=0
-phi(1:nf,1:nf,1:nf)=cube ! phi1
-open(11,file='phi1.dat',access='stream')
-write(11) cube
-close(11)
-call fft_cube2pencil_fine
-call trans_zxy2xyz_fine
+!call trans_xyz2zxy_fine
+!call ifft_pencil2cube_fine
+!phi=0
+!phi(1:nf,1:nf,1:nf)=cube ! phi1
+!open(11,file='phi1.dat',access='stream')
+!write(11) cube
+!close(11)
+!call fft_cube2pencil_fine
+!call trans_zxy2xyz_fine
 !!!!!!
 
 #ifndef ZA
@@ -501,9 +501,9 @@ call trans_zxy2xyz_fine
 
   !cube=cube-phixy**2-phiyz**2-phizx**2 ! source term of phi2
 
-  open(11,file='phi2_source.dat',access='stream')
-  write(11) cube
-  close(11)
+  !open(11,file='phi2_source.dat',access='stream')
+  !write(11) cube
+  !close(11)
 
   sync all
 
@@ -535,9 +535,9 @@ call trans_zxy2xyz_fine
   call trans_xyz2zxy_fine
   call ifft_pencil2cube_fine
 
-  open(11,file='phi2.dat',access='stream')
-  write(11) cube
-  close(11)
[…]

### 2016-09-21 ms_disp1/ms_disp1.tex
disp paper figures preliminary

@@
 In section \ref{sec.results} we show the results of the reconstruction.
-Discussion and conclusion are in \ref{sec.discussion}.
+Discussion and conclusion are in section \ref{sec.discussion}.
+
+\begin{figure}[b] \centering
+  \includegraphics[width=1.0\linewidth]{fig1.pdf}
+  \caption{Dimensionless auto- and cross- power spectra between nonlinear density field $\delta$, linear density field $\delta_L$ and reconstructed density field $\delta_R$.}
+  \label{fig.1}
+\end{figure}
+\begin{figure}[t] \centering
+  \includegraphics[width=1.0\linewidth]{fig2.pdf}
+  \caption{Caption goes here.}
+  \label{fig.2}
+\end{figure}
 
@@
 
+Two-point statistics of the density fields are quantified by the cross power spectrum $P_{ij}(k)\equiv(2\pi)^{-3}\langle|\delta_i(k)\delta_j(k)|\rangle$, where subscripts $i,j$ may refer to linear, nonlinear, or reconstructed density fields.
+When $i=j$ it reduces to the auto power spectrum $P_{ii}(k)$ or $P(k)$.
+We usually plot the dimensionless power spectrum $\Delta^2(k)\equiv k^3P(k)/2\pi^2$.
+The blue solid and black dashed curves in Fig.\ref{fig.1} show the power spectra of $\delta$ and $\delta_L$.
+Their difference shows the nonlinear evolution of LSS on small scales.
+Their cross power drops to a very low value, indicating a loss of linear information in the nonlinear power spectrum.
+
 \subsection{Reconstruction}
@@
 \end{equation}
+In Fig.\ref{fig.1} we show the power spectrum of $\delta_R$ and its cross power with $\delta_L$.
+Despite of a lowered power of $\delta_R$ compared to $\delta$, it has a much higher cross power with $\delta_L$, up to a relatively smaller scale (higher $k$).
 
@@
 \begin{equation}
-    P_{LR}=r'P_{LL}+P_{LN}
+    P_{LR}=r'P_{LL}+P_{LN},
 \end{equation}
-where $P_{ij}\equiv\lb\delta_i\delta_j\rb$ denotes the cross power spectrum.
+where $P_{ij}\equiv\langle\delta_i\delta_j\rangle$ denotes the cross power spectrum.
 Since $\delta_N$ is uncorrelated with $\delta_L$, $P_{LN}=0$.
 With the definition of cross correlation coefficient $r\equiv P_{LR}/\sqrt{P_{LL}P_{RR}}$ and bias $b^2=P_{RR}/P_{LL}$, we solve $r'=P_{LR}/P_{LL}=rb$.
+We plot the cross correlation coefficient $r_{LN}$ and $r_{LR}$ in Fig.\ref{fig.2}.
+Clearly, $\delta_R$ contains much more linear information on smaller scales.
+
 According to equation (\ref{eq.decompose}), the auto power spectrum is decomposed as
@@
 \end{equation}
-The estimated linear density, or the optimal reconstructed density is given by
+The Wiener filters are also plot in Fig.\ref{fig.2}. The estimated linear density, or the optimal reconstructed density is given by
 \begin{equation}
-    \tilde\delta_R=Wb^{-1}\delta_R,
+    \tilde\delta_R=Wb^{-1}\delta_R.
 \end{equation}
-and the optimal reconstructed power spectrum is given by
+The full filters $Wb^{-1}$ are also plotted in Fig.\ref{fig.2}. The optimal reconstructed power spectrum is given by
 \begin{equation}
@@
 \end{equation}
+The $W^2$ describes the damping of the linear power spectrum.
 
-%\begin{figure}[t] \centering
-%  \includegraphics[width=1.0\linewidth]{fig1.pdf}
-%  \caption{Caption goes here.}
-%  \label{fig.1}
-%\end{figure}
+\begin{figure}[t] \centering
+  \includegraphics[width=1.0\linewidth]{fig3.pdf}
+  \caption{Caption goes here.}
+  \label{fig.3}
+\end{figure}
+
+\begin{figure}[t] \centering
+  \includegraphics[width=1.0\linewidth]{fig4.pdf}
+  \caption{Caption goes here.}
+  \label{fig.4}
+\end{figure}
 
 \section{Results}\label{sec.results}
+Fig.\ref{fig.2} shows the damping factors $W^2(k)$ for the optimal filtered nonlinear and reconstructed density fields.
+We fit the Gaussian BAO damping model ${\mathcal D}(k)=\exp(-k^2\Sigma^2/2)$ and give $\Sigma=1.8$ Mpc$/h$ and $\Sigma=10$ Mpc$/h$ for nonlinear and reconstructed fields.
 

### 2016-09-24 ms_disp1/ms_disp1.tex
haoran

@@
 \begin{abstract}
-We investigate the ability of using nonlinear displacement field to reconstruct the primordial linear perturbations of the large scale structure.
-According to the linear Lagrangian perturbation theory, the first order primordial linear density field is given by the negative divergence of the displacement field.
-By running $N$-body simulations, we show that this reconstruction algorithm recovers great amount of nonlinear information in the early universe.
-In reality, the displacement fields can be solved from only the final stage of the density fields.
-This has potential to reconstruct baryonic acoustic oscillation (BAO) from current and future large scale structure surveys.
+We investigate the ability of using nonlinear displacement field to reconstruct 
+the primordial linear perturbations of the large scale structure. According to 
+the linear Lagrangian perturbation theory, the first order primordial linear 
+density field is given by the negative divergence of the displacement field. By 
+running $N$-body simulations, we show that this reconstruction algorithm recovers 
+great amount of nonlinear information in the early universe. In reality, the 
+displacement fields can be solved from only the final stage of the density 
+fields. This has potential to reconstruct baryonic acoustic oscillation (BAO) 
+from current and future large scale structure surveys.
+
 \end{abstract}
@@
 \section{Introduction}\label{sec.intro}
-Our universe starts from primordial Gaussian perturbations at a very early stage, and from those fluctuations, the gravitational instability drives the formation of the large scale structure (LSS) distribution of matter.
-These structures grow linearly until the perturbations are large enough such that first order perturbation theories are unable to analytically describe the LSS distributions.
-Inversely, the final stage nonlinear LSS distribution contains higher order statistics, and thus makes it more challenging to interpret the basic cosmological parameters.
-This leads to various attempts to recover earlier stages of LSS, in which statistics are closer to Gaussian.
-Because Gaussian fields can be adequately described by two-point statistics, ideally after some recovery algorithms, more information can be extracted, more straightforwardly, by power spectra or two-point correlation functions.
+Our universe starts from primordial Gaussian perturbations at a very early stage, 
+and from those fluctuations, the gravitational instability drives the formation of 
+the large scale structure (LSS) distribution of matter. These structures grow 
+linearly until the perturbations are large enough such that first order 
+perturbation theories are unable to analytically describe the LSS distributions. 
+Inversely, the final stage nonlinear LSS distribution contains higher order 
+statistics, and thus makes it more challenging to interpret the basic cosmological 
+parameters. This leads to various attempts to recover earlier stages of LSS, in 
+which statistics are closer to Gaussian. Because Gaussian fields can be adequately 
+described by two-point statistics, ideally after some recovery algorithms, more 
+information can be extracted, more straightforwardly, by power spectra or two-point 
+correlation functions.
 
-These algorithms include nonlinear Wiener filters in wavelet space, distribution function transforms etc.
-Standard BAO reconstruction algorithms smooths the nonlinear density field on linear scale ($\sim$10 Mpc$/h$) and reverse the large scale bulk flows by a negative Zel'dovich linear displacement.
-Here we propose a new reconstruction method that uses the nonlinear displacement field to recover the primordial density field.
-In the linear Lagrangian perturbation theory, the negative divergence of the the displacement field respect to Lagrangian coordinates gives the linear density field.
-Of course, the full displacement field is non-observable, as it requires the initial distributions of matter, however there are many techniques to estimate the nonlinear displacement field from a final distribution of matter.
-For example, when a homogeneous initial matter distribution is assumed, there is a unique solution of curl-less displacement field to relate the initial and final distributions without shell-crossing.
-This solution can be solved by a metric transformation equation \citep{1995ApJS..100..269P,1998ApJS..115...19P}.
-In 1-dimensional case, the solution simplifies to a reordering of matter elements by Eulerian positions, and [Hong-ming et.
-al.] studies this case.
-The homogeneous and curl-less assumptions are generally valid in cosmology [cite or proof].
-Shell-crossing does exist, and this effect can be quantified by $N$-body simulations.
+These algorithms include nonlinear Wiener filters in wavelet space, distribution 
+function transforms etc. Standard BAO reconstruction algorithms smooths the 
+nonlinear density field on linear scale ($\sim$10 Mpc$/h$) and reverse the large 
+scale bulk flows by a negative Zel'dovich linear displacement. Here we propose a 
+new reconstruction method that uses the nonlinear displacement field to recover the 
+primordial density field. In the linear Lagrangian perturbation theory, the 
+negative divergence of the the displacement field respect to Lagrangian coordinates 
+gives the linear density field. Of course, the full displacement field is non-
+observable, as it requires the initial distributions of matter, however there are 
+many techniques to estimate the nonlinear displacement field from a final 
+distribution of matter. For example, when a homogeneous initial matter distribution 
+is assumed, there is a unique solution of curl-less displacement field to relate 
+the initial and final distributions without shell-crossing. This solution can be 
+solved by a metric transformation equation \citep{1995ApJS..100..269P,1998ApJS..
+115...19P}. In 1-dimensional case, the solution simplifies to a reordering of 
+matter elements by Eulerian positions, and \cite{2016arXiv160907041Z} studies this 
+case. The homogeneous and curl-less assumptions are generally valid in cosmology 
+[cite or proof]. Shell-crossing does exist, and this effect can be quantified by 
+$N$-body simulations.
 
-In this paper we track the initial and final distributions of particles in LSS $N$-body simulations, and use the displacement field to reconstruct the initial linear density field.
-The rest of this paper is structured as follows.
-In section \ref{sec.method} we describe the simulation and reconstruction algorithm.
-In section \ref{sec.results} we show the results of the reconstruction.
-Discussion and conclusion are in section \ref{sec.discussion}.
+In this paper we track the initial and final distributions of particles in LSS $N$-
+body simulations, and use the displacement field to reconstruct the initial linear 
+density field. The rest of this paper is structured as follows. In section 
+\ref{sec.method} we describe the simulation and reconstruction algorithm. In 
+section \ref{sec.results} we show the results of the reconstruction. Discussion and 
+conclusion are in section \ref{sec.discussion}.
 
@@
   \includegraphics[width=1.0\linewidth]{fig2.pdf}
-  \caption{Caption goes here.}
+  \caption{Correlation functions $r(\delta_L,\delta)$ and $r(\delta_L,\delta_R)$.}
   \label{fig.2}
@@
 \subsection{Simulation}
-We use the public cosmological simulation code {\tt CUBE}.
-Cosmological parameters are in accordance with Planck.
-Initial conditions are generated at redshift $z=50$ using Zel'dovich approximation.
-$N_p=512^3$ $N$-body particles are evolved via their mutual gravitational interactions to $z=0$, in a periodic box with $L=300$ Mpc$/h$ per side.
-The code is set to use standard a particle-mesh (PM) algorithm \cite{1988csup.book.....H} on a two-level mesh grids (details see \cite{2013MNRAS.436..540H}) and cloud-in-cell (CIC) is used in particle interpolations in force calculation and obtaining the density field $\rho({\bs x})$ in Eulerian coordinates ${\bs x}$ at late stages.
-We use density contrast $\delta\equiv\rho/\lb\rho\rb-1$ to describe the density fluctuations.
-The primordial linear density field $\delta_L$ is given by the initial stage and scaled to $z=0$ by the linear growth factor.
+We use the public cosmological simulation code {\tt CUBE}. Cosmological parameters 
+are in accordance with Planck. Initial conditions are generated at redshift $z=50$ 
+using Zel'dovich approximation. $N_p=512^3$ $N$-body particles are evolved via 
+their mutual gravitational interactions to $z=0$, in a periodic box with $L=300$ 
+Mpc$/h$ per side. The code is set to use standard a particle-mesh (PM) algorithm 
+\cite{1988csup.book.....H} on a two-level mesh grids (details see \cite{2013MNRAS.
+436..540H}) and cloud-in-cell (CIC) is used in particle interpolations in force 
+calculation and obtaining the density field $\rho({\bs x})$ in Eulerian coordinates 
+${\bs x}$ at late stages. We use density contrast $\delta\equiv\rho/\lb\rho\rb-1$ 
+to describe the density fluctuations. The primordial linear density field $
+\delta_L$ is given by the initial stage and scaled to $z=0$ by the linear growth 
+factor.
 
-Two-point statistics of the density fields are quantified by the cross power spectrum $P_{ij}(k)\equiv(2\pi)^{-3}\langle|\delta_i(k)\delta_j(k)|\rangle$, where subscripts $i,j$ may refer to linear, nonlinear, or reconstructed density fields.
-When $i=j$ it reduces to the auto power spectrum $P_{ii}(k)$ or $P(k)$.
-We usually plot the dimensionless power spectrum $\Delta^2(k)\equiv k^3P(k)/2\pi^2$.
-The blue solid and black dashed curves in Fig.\ref{fig.1} show the power spectra of $\delta$ and $\delta_L$.
-Their difference shows the nonlinear evolution of LSS on small scales.
-Their cross power drops to a very low value, indicating a loss of linear information in the nonlinear power spectrum.
+Two-point statistics of the density fields are quantified by the cross power 
+spectrum $P_{ij}(k)\equiv(2\pi)^{-3}\langle|\delta_i(k)||\delta_j(k)|\rangle$, 
+where subscripts $i,j$ may refer to linear, nonlinear, or reconstructed density 
+fields. When $i=j$ it reduces to the auto power spectrum $P_{ii}(k)$ or $P(k)$. We 
+usually plot the dimensionless power spectrum $\Delta^2(k)\equiv k^3P(k)/2\pi^2$. 
+The blue solid and black dashed curves in Fig.\ref{fig.1} show the power spectra of 
+$\delta$ and $\delta_L$. Their difference shows the nonlinear evolution of LSS on 
+small scales. Their cross power drops to a very low value, indicating a loss of 
+linear information in the nonlinear power spectrum.
 
 \subsection{Reconstruction}
-In the simulation, we uses particle-ID (PID) to record the initial location ${\bs q}$ of particles, and the information is tracked until the $z=0$ and we can get the Lagrangian displacement vector ${\bs \Psi}\equiv{\bs x}-{\bs q}$ for every particle.
-Then these vectors are interpolated onto the initial Lagrangian coordinates ${\bs q}$ of particles and we get the displacement field ${\bs \Psi}({\bs q})$.
-The raw reconstructed density field is given by the differential motion of matter elements,
+In the simulation, we uses particle-ID (PID) to record the initial location ${\bs 
+q}$ of particles, and the information is tracked until the $z=0$ and we can get the 
+Lagrangian displacement vector ${\bs \Psi}\equiv{\bs x}-{\bs q}$ for every 
+particle. Then these vectors are interpolated onto the initial Lagrangian 
+coordinates ${\bs q}$ of particles and we get the displacement field ${\bs \Psi}
+({\bs q})$.
+The raw reconstructed density field is given by the differential motion of matter 
+elements,
 \begin{equation}
@@
 \end{equation}
-where $P_{ij}\equiv\langle\delta_i\delta_j\rangle$ denotes the cross power spectrum.
-Since $\delta_N$ is uncorrelated with $\delta_L$, $P_{LN}=0$.
-With the definition of cross correlation coefficient $r\equiv P_{LR}/\sqrt{P_{LL}P_{RR}}$ and bias $b^2=P_{RR}/P_{LL}$, we solve $r'=P_{LR}/P_{LL}=rb$.
-We plot the cross correlation coefficient $r_{LN}$ and $r_{LR}$ in Fig.\ref{fig.2}.
-Clearly, $\delta_R$ contains much more linear information on smaller scales.
+where $P_{ij}\equiv\langle\delta_i\delta_j\rangle$ denotes the cross power 
+spectrum. Since $\delta_N$ is uncorrelated with $\delta_L$, $P_{LN}=0$. With the 
+definition of cross correlation coefficient $r(\delta_L,\delta_R)\equiv P_{LR}/\sqrt{P_{LL}P_{RR}}$ 
+and bias $b^2=P_{RR}/P_{LL}$, we solve $r'=P_{LR}/P_{LL}=rb$. We plot the cross 
+correlation coefficient $r_{LN}$ and $r_{LR}$ in Fig.\ref{fig.2}. Clearly, $
+\delta_R$ contains much more linear information on smaller scales.
 
@@
   \includegraphics[width=1.0\linewidth]{fig3.pdf}
-  \caption{Caption goes here.}
+  \caption{Power spectra of the nonlinear density field and its
+  filtered density field.}
   \label{fig.3}
@@
   \includegraphics[width=1.0\linewidth]{fig4.pdf}
-  \caption{Caption goes here.}
+  \caption{Power spectra of the reconstructed density field and its
+  filtered density field.}
   \label{fig.4}
@@
 \section{Results}\label{sec.results}
-Fig.\ref{fig.2} shows the damping factors $W^2(k)$ for the optimal filtered nonlinear and reconstructed density fields.
-We fit the Gaussian BAO damping model ${\mathcal D}(k)=\exp(-k^2\Sigma^2/2)$ and give $\Sigma=1.8$ Mpc$/h$ and $\Sigma=10$ Mpc$/h$ for nonlinear and reconstructed fields.
+Fig.\ref{fig.2} shows the cross correlation functions,  damping factors $W^2(k)$ for the optimal filtered 
+nonlinear and reconstructed density fields. We fit the Gaussian BAO damping model $
+{\mathcal D}(k)=\exp(-k^2\Sigma^2/2)$ and give $\Sigma=1.8$ Mpc$/h$ and $\Sigma=10$ 
+Mpc$/h$ for nonlinear and reconstructed fields.
 

### 2016-09-24 ms_disp1/ms_disp1.tex
haoran

@@
 displacement fields can be solved from only the final stage of the density 
-fields. This has potential to reconstruct baryonic acoustic oscillation (BAO) 
+fields. This demonstrates the potential to reconstruct baryonic acoustic oscillation (BAO) 
 from current and future large scale structure surveys.
@@
 and from those fluctuations, the gravitational instability drives the formation of 
-the large scale structure (LSS) distribution of matter. These structures grow 
-linearly until the perturbations are large enough such that first order 
-perturbation theories are unable to analytically describe the LSS distributions. 
-Inversely, the final stage nonlinear LSS distribution contains higher order 
-statistics, and thus makes it more challenging to interpret the basic cosmological 
-parameters. This leads to various attempts to recover earlier stages of LSS, in 
-which statistics are closer to Gaussian. Because Gaussian fields can be adequately 
-described by two-point statistics, ideally after some recovery algorithms, more 
-information can be extracted, more straightforwardly, by power spectra or two-point 
-correlation functions.
+the large scale structure (LSS) distribution of matter [cite LSS]. These structures grow 
+linearly until the perturbations are large enough so that the first order 
+perturbation theories are unable to analytically describe the LSS distributions [cite PT]. 
+As a result, the final nonlinear LSS distribution contains higher order 
+statistics, and thus makes it more challenging to be interpreted into basic
+cosmological parameters [cite constraint]. For example, the baryonic acoustic oscillation (BAO)
+scale can be used as a ``standard ruler'' to constrain the cosmic expansion history
+and thus probes the dark energy properties [cite BAO], and nonlinear evolution smears the BAO
+features [cite smear BAO]. These leads to various attempts to recover earlier stages of LSS, in 
+which statistics are closer to Gaussian [cite Gaussianize, Reco examples].
+Because Gaussian fields can be adequately described by two-point statistics,
+ideally after some recovery algorithms, more information can be extracted,
+more straightforwardly, by power spectra or two-point correlation functions [can cite friends].
 
-These algorithms include nonlinear Wiener filters in wavelet space, distribution 
-function transforms etc. Standard BAO reconstruction algorithms smooths the 
+These algorithms include nonlinear Wiener filters in wavelet space [cite myself], distribution 
+function transforms etc [cite Pengjie, etc.]. Standard BAO reconstruction algorithms smooth the 
 nonlinear density field on linear scale ($\sim$10 Mpc$/h$) and reverse the large 
-scale bulk flows by a negative Zel'dovich linear displacement. Here we propose a 
+scale bulk flows by a negative Zel'dovich linear displacement [cite Eisenstein]. Here we propose a 
 new reconstruction method that uses the nonlinear displacement field to recover the 
@@
 negative divergence of the the displacement field respect to Lagrangian coordinates 
-gives the linear density field. Of course, the full displacement field is non-
+gives the linear density field [cite Jeong]. Of course, the full displacement field is non-
 observable, as it requires the initial distributions of matter, however there are 
@@
 the initial and final distributions without shell-crossing. This solution can be 
-solved by a metric transformation equation \citep{1995ApJS..100..269P,1998ApJS..
-115...19P}. In 1-dimensional case, the solution simplifies to a reordering of 
+solved by a metric transformation equation \citep{1995ApJS..100..269P,1998ApJS..115...19P}. In 1-dimensional case, the solution simplifies to a reordering of 
 matter elements by Eulerian positions, and \cite{2016arXiv160907041Z} studies this 
@@
 body simulations, and use the displacement field to reconstruct the initial linear 
-density field. The rest of this paper is structured as follows. In section 
+density field. In section 
 \ref{sec.method} we describe the simulation and reconstruction algorithm. In 
@@
 \section{Method}\label{sec.method}
+We show the LSS simulation and displacement field setups in section \ref{ss.sim},
+and the reconstruction processes are presented in section \ref{ss.reco}.
 
-\subsection{Simulation}
-We use the public cosmological simulation code {\tt CUBE}. Cosmological parameters 
-are in accordance with Planck. Initial conditions are generated at redshift $z=50$ 
+\subsection{Simulation}\label{ss.sim}
+We use the open source cosmological simulation code {\tt CUBE}\footnote{github link of cafcube}
+[cite Yu et al. in prep].
+Cosmological parameters 
+are [TBD before submission]. Initial conditions are generated at redshift $z=50$ 
 using Zel'dovich approximation. $N_p=512^3$ $N$-body particles are evolved via 
@@
 Mpc$/h$ per side. The code is set to use standard a particle-mesh (PM) algorithm 
-\cite{1988csup.book.....H} on a two-level mesh grids (details see \cite{2013MNRAS.
-436..540H}) and cloud-in-cell (CIC) is used in particle interpolations in force 
+\cite{1988csup.book.....H} on a two-level mesh grids
+(details see \cite{2013MNRAS.436..540H}) and cloud-in-cell
+(CIC) is used in particle interpolations in force 
 calculation and obtaining the density field $\rho({\bs x})$ in Eulerian coordinates 
@@
 spectrum $P_{ij}(k)\equiv(2\pi)^{-3}\langle|\delta_i(k)||\delta_j(k)|\rangle$, 
-where subscripts $i,j$ may refer to linear, nonlinear, or reconstructed density 
+where subscripts $i,j$ may refer to linear ($L$), nonlinear ($N$), or reconstructed ($R$) density 
 fields. When $i=j$ it reduces to the auto power spectrum $P_{ii}(k)$ or $P(k)$. We 
@@
 
-\subsection{Reconstruction}
-In the simulation, we uses particle-ID (PID) to record the initial location ${\bs 
+\subsection{Reconstruction}\label{ss.reco}
+In the simulation, we use particle-ID (PID) to record the initial (Lagrangian) location ${\bs 
 q}$ of particles, and the information is tracked until the $z=0$ and we can get the 
@@
 \end{equation}
+Because the reconstruction processes are implemented on Lagrangian coordinates,
+$\delta_R$ takes the coordinates of $\bs q$ instead of $\bs x$.
+We just write $\bs q$'s Fourier wave number
+$k_q$ as $k$ to simplify the expression.
 In Fig.\ref{fig.1} we show the power spectrum of $\delta_R$ and its cross power with $\delta_L$.

### 2016-09-25 ms_disp1/ms_disp1.tex
haoran

@@
 $N$-body simulations.
+\begin{figure}[b] \centering
+  \includegraphics[width=1.0\linewidth]{fig1.pdf}
+  \caption{Dimensionless auto- and cross- power spectra between nonlinear density field $\delta$, linear density field $\delta_L$ and reconstructed density field $\delta_R$.}
+  \label{fig.1}
+\end{figure}
+\begin{figure}[t] \centering
+  \includegraphics[width=1.0\linewidth]{fig2.pdf}
+  \caption{Correlation functions $r(\delta_L,\delta)$ and $r(\delta_L,\delta_R)$
+  (solid lines). Their
+  damping of the power spectra $W^2$ (dashed lines),
+  and their best fit BAO damping models (dotted lines).}
+  \label{fig.2}
+\end{figure}
 
@@
 conclusion are in section \ref{sec.discussion}.
-
-\begin{figure}[b] \centering
-  \includegraphics[width=1.0\linewidth]{fig1.pdf}
-  \caption{Dimensionless auto- and cross- power spectra between nonlinear density field $\delta$, linear density field $\delta_L$ and reconstructed density field $\delta_R$.}
-  \label{fig.1}
-\end{figure}
-\begin{figure}[t] \centering
-  \includegraphics[width=1.0\linewidth]{fig2.pdf}
-  \caption{Correlation functions $r(\delta_L,\delta)$ and $r(\delta_L,\delta_R)$.}
-  \label{fig.2}
-\end{figure}
 
@@
 $k_q$ as $k$ to simplify the expression.
-In Fig.\ref{fig.1} we show the power spectrum of $\delta_R$ and its cross power with $\delta_L$.
-Despite of a lowered power of $\delta_R$ compared to $\delta$, it has a much higher cross power with $\delta_L$, up to a relatively smaller scale (higher $k$).
+In Fig.\ref{fig.1} we show the power spectrum of $\delta_R$ and its cross power
+with $\delta_L$. Despite of a lowered power of $\delta_R$ compared to $\delta$, it has a much higher 
+cross power with $\delta_L$, up to a relatively smaller scale (higher $k$).
+\begin{figure}[t] \centering
+  \includegraphics[width=1.0\linewidth]{fig3.pdf}
+  \caption{Power spectra of the nonlinear density field and its
+  filtered density field.}
+  \label{fig.3}
+\end{figure}
+
+\begin{figure}[t] \centering
+  \includegraphics[width=1.0\linewidth]{fig4.pdf}
+  \caption{Power spectra of the reconstructed density field and its
+  filtered density field.}
+  \label{fig.4}
+\end{figure}
 
@@
 \end{equation}
-The full filters $Wb^{-1}$ are also plotted in Fig.\ref{fig.2}. The optimal reconstructed power spectrum is given by
+The full filters $Wb^{-1}$ are also plotted in Fig.\ref{fig.2}.
+The optimal reconstructed power spectrum is given by
 \begin{equation}
@@
 The $W^2$ describes the damping of the linear power spectrum.
-
-\begin{figure}[t] \centering
-  \includegraphics[width=1.0\linewidth]{fig3.pdf}
-  \caption{Power spectra of the nonlinear density field and its
-  filtered density field.}
-  \label{fig.3}
-\end{figure}
-
-\begin{figure}[t] \centering
-  \includegraphics[width=1.0\linewidth]{fig4.pdf}
-  \caption{Power spectra of the reconstructed density field and its
-  filtered density field.}
-  \label{fig.4}
-\end{figure}
 
@@
 
+Fig.\ref{fig.3} and Fig.\ref{fig.4} show their filtered power spectra.
+
 \section{Discussion and conclusion}\label{sec.discussion}
-Discussion goes here.
+We extract the actual displacement field of matter elements in cosmological $N$-body
+simulations, and use this displacement field in Lagrangian coordinates to reconstruct
+the primordial linear perturbations. The result shows a prominent improvement from
+$r_{LN}$ to $r_{LR}$ in Fig.\ref{fig.2} -- recovering the lost linear information on
+nearly an order of magnitude smaller scales. more tomorrow...
 

### 2016-09-27 ms_disp1/ms_disp1.tex
introduction modified

@@
 \newcommand{\rb}{\right\rangle}
+\newcommand{\tcb}{\textcolor{blue}}
+\newcommand{\tcr}{\textcolor{red}}
 
@@
 \def\mnras{Mon. Not. R. Astron. Soc.}
-\def\aap{Astron. Astrophys.}                % Astronomy and Astrophysics
+\def\aap{Astron. Astrophys.}
 \def\jcap{J. Cosmology Astropart. Phys.}
@@
 \addtolength{\textwidth}{1.05cm}
-\title{Reconstruction of Large Scale Structure by Nonlinear Displacement Fields}
+\title{Reconstruction of Large Scale Structure by Nonlinear \tcr{Lagrangian} Displacement Fields}
 
-%\author{Hao-Ran~Yu} \affiliation{}\email{haoran@cita.utoronto.ca}
-%\author{Qiaoyin~Pan}
-%\author{Ue-Li~Pen} \affiliation{}
-%\author{Tong-Jie~Zhang} \affiliation{}
-%---
+\author{Hao-Ran~Yu}%\email{haoran@cita.utoronto.ca}
+\affiliation{Kavli Institute for Astronomy and Astrophysics, Peking University, Beijing 100871, China}
+\affiliation{Canadian Institute for Theoretical Astrophysics, University of Toronto, 60 St. George Street, Toronto, Ontario M5S 3H8, Canada}
 
+\author{Ue-Li~Pen}%\email{pen@cita.utoronto.ca}
+\affiliation{Canadian Institute for Theoretical Astrophysics, University of Toronto, 60 St. George Street, Toronto, Ontario M5S 3H8, Canada}
+\affiliation{Dunlap Institute for Astronomy and Astrophysics, University of Toronto, 50 St. George Street, Toronto, Ontario M5S 3H4, Canada}
+\affiliation{Canadian Institute for Advanced Research, CIFAR Program in Gravitation and Cosmology, Toronto, Ontario M5G 1Z8, Canada}
+\affiliation{Perimeter Institute for Theoretical Physics, 31 Caroline Street North, Waterloo, Ontario, N2L 2Y5, Canada}
+
+\date{\today}
 %\date{Received \today; published -- 00, 0000}
@@
 density field is given by the negative divergence of the displacement field. By 
-running $N$-body simulations, we show that this reconstruction algorithm recovers 
+running $N$-body simulations, we show that this \tcr{new} reconstruction algorithm recovers 
 great amount of nonlinear information in the early universe. In reality, the 
@@
 and from those fluctuations, the gravitational instability drives the formation of 
-the large scale structure (LSS) distribution of matter [cite LSS]. These structures grow 
+the large scale structure (LSS) distribution of matter
+\citep{1970A&A.....5...84Z,1985ApJ...292..371D}.
+These structures grow 
 linearly until the perturbations are large enough so that the first order 
-perturbation theories are unable to analytically describe the LSS distributions [cite PT]. 
+perturbation theories are unable to analytically describe the LSS distributions 
+\citep{2016JCAP...01..043M}. 
 As a result, the final nonlinear LSS distribution contains higher order 
 statistics, and thus makes it more challenging to be interpreted into basic
-cosmological parameters [cite constraint]. For example, the baryonic acoustic oscillation (BAO)
+cosmological parameters. One such example is that, the baryonic acoustic oscillation (BAO)
 scale can be used as a ``standard ruler'' to constrain the cosmic expansion history
-and thus probes the dark energy properties [cite BAO], and nonlinear evolution smears the BAO
-features [cite smear BAO]. These leads to various attempts to recover earlier stages of LSS, in 
-which statistics are closer to Gaussian [cite Gaussianize, Reco examples].
+and thus probes the dark energy properties \citep{2005NewAR..49..360E}, but
+nonlinear evolution smears the BAO features and lowers the measurement
+accuracy \citep{2005ApJ...633..560E,2012MNRAS.419.2949N}.
+There are various attempts to recover earlier stages
+of LSS, in which statistics are closer to Gaussian 
+\citep{1992MNRAS.254..315W,2013MNRAS.436..759H}.
 Because Gaussian fields can be adequately described by two-point statistics,
 ideally after some recovery algorithms, more information can be extracted,
-more straightforwardly, by power spectra or two-point correlation functions [can cite friends].
+more straightforwardly, by power spectra or two-point correlation functions
+\citep{2005MNRAS.360L..82R,2012MNRAS.421..832Y}.
 
-These algorithms include nonlinear Wiener filters in wavelet space [cite myself], distribution 
-function transforms etc [cite Pengjie, etc.]. Standard BAO reconstruction algorithms smooth the 
+Standard BAO reconstruction algorithms smooth the 
 nonlinear density field on linear scale ($\sim$10 Mpc$/h$) and reverse the large 
-scale bulk flows by a negative Zel'dovich linear displacement [cite Eisenstein]. Here we propose a 
-new reconstruction method that uses the nonlinear displacement field to recover the 
+scale bulk flows by a negative Zel'dovich linear displacement
+\citep{2007ApJ...664..675E,2009PhRvD..80l3501N,2009PhRvD..79f3523P}. Here we propose a 
+\tcr{new reconstruction method} that uses the nonlinear displacement field to recover the 
 primordial density field. In the linear Lagrangian perturbation theory, the 
-negative divergence of the the displacement field respect to Lagrangian coordinates 
-gives the linear density field [cite Jeong]. Of course, the full displacement field is non-
-observable, as it requires the initial distributions of matter, however there are 
+negative divergence of the the displacement field
+$\bs\Psi(\bs q)$ respect to Lagrangian coordinates $\bs q$
+gives the linear density field \tcr{\citep{2010PhDT.........4J}}.
+Of course, the full displacement field $\bs\Psi(\bs q)$
+is non-observable, as it requires the initial distributions of matter, however there are 
 many techniques to estimate the nonlinear displacement field from a final 
@@
 the initial and final distributions without shell-crossing. This solution can be 
-solved by a metric transformation equation \citep{1995ApJS..100..269P,1998ApJS..115...19P}. In 1-dimensional case, the solution simplifies to a reordering of 
-matter elements by Eulerian positions, and \cite{2016arXiv160907041Z} studies this 
-case. The homogeneous and curl-less assumptions are generally valid in cosmology 
-[cite or proof]. Shell-crossing does exist, and this effect can be quantified by 
-$N$-body simulations.
+solved by a metric transformation equation \citep{1995ApJS..100..269P,1998ApJS..115...19P}.
+\tcr{[Peacock, Goldberg?]}
+In 1-dimensional (1D) case, the exact solution of \cite{1995ApJS..100..269P}
+simplifies to an ordering of matter elements by Eulerian coordinates.
+Zhu {\it et al.} \cite{2016arXiv160907041Z} apply this algorithm to the result
+of an 1D simulation \citep{2016JCAP...01..043M} and obtain an estimated
+displacement field $s(q)$, and find that \tcr{this new method} well recovers the linear
+information and reconstructs BAO peak in correlation function. 
+In 3D case, there are various techniques to obtain $\bs s(\bs q)$.
+However one needs to carefully consider effects of curl, shell-crossing in 3D case.
+
+Before these steps, we need to quantify the amount of linear information
+that can be recovered from the full nonlinear displacement field $\bs\Psi(\bs q)$,
+and further estimations $\bs s(\bs q)$ can be compared with this result.
+In this paper we run a LSS $N$-body simulation and track the motion of
+particles to obtain $\bs\Psi(\bs q)$. According to this we reconstruct the linear density
+field and compare to the primordial density field of the initial
+conditions. We describe the simulation and reconstruction algorithm in section
+\ref{sec.method}, and we show the results in section \ref{sec.results}.
+Discussion and conclusion are in section \ref{sec.discussion}.
+
 \begin{figure}[b] \centering
@@
 
-In this paper we track the initial and final distributions of particles in LSS $N$-
-body simulations, and use the displacement field to reconstruct the initial linear 
-density field. In section 
-\ref{sec.method} we describe the simulation and reconstruction algorithm. In 
-section \ref{sec.results} we show the results of the reconstruction. Discussion and 
-conclusion are in section \ref{sec.discussion}.
-
 \section{Method}\label{sec.method}
@@
 \subsection{Simulation}\label{ss.sim}
-We use the open source cosmological simulation code {\tt CUBE}\footnote{github link of cafcube}
-[cite Yu et al. in prep].
+We use the open source cosmological simulation code {\tt CUBE}%\footnote{github link of cafcube}
+\citep{cafcube}.
 Cosmological parameters 

### 2016-09-28 ms_disp1/ms_disp1.tex
revised conclusion

@@
 \citep{cafcube}.
-Cosmological parameters 
-are [TBD before submission]. Initial conditions are generated at redshift $z=50$ 
-using Zel'dovich approximation. $N_p=512^3$ $N$-body particles are evolved via 
+Cosmological parameters are set as
+$\Omega_m=0.27$, $\Omega_\Lambda=0.73$, $h_0=0.68$, $n_s=0.96$ and $\sigma_8=0.83$.
+Initial conditions are generated at redshift $z=50$ 
+using Zel'dovich approximation, and using a CLASS transfer function.
+$N_p=512^3$ $N$-body particles are evolved via 
 their mutual gravitational interactions to $z=0$, in a periodic box with $L=300$ 
@@
 In Fig.\ref{fig.1} we show the power spectrum of $\delta_R$ and its cross power
-with $\delta_L$. Despite of a lowered power of $\delta_R$ compared to $\delta$, it has a much higher 
-cross power with $\delta_L$, up to a relatively smaller scale (higher $k$).
+with $\delta_L$. Despite of a lowered power of $\delta_R$ compared to $\delta$,
+it has a much higher cross power with $\delta_L$ compared to $\delta_N$,
+up to a relatively smaller scale (higher $k$).
 \begin{figure}[t] \centering
@@
 According to equation (\ref{eq.decompose}), the auto power spectrum is decomposed as
-\begin{equation}
+\begin{equation}\label{eq.power}
     P_{RR}=r^2b^2P_{LL}+P_{NN},
@@
 The optimal reconstructed power spectrum is given by
-\begin{equation}
+\begin{equation}\label{eq.opt}
     \tilde P=W^2b^{-2}P_{RR}=W^2P_{LL}+W^2b^{-2}P_{NN}.
@@
 \section{Results}\label{sec.results}
-Fig.\ref{fig.2} shows the cross correlation functions,  damping factors $W^2(k)$ for the optimal filtered 
-nonlinear and reconstructed density fields. We fit the Gaussian BAO damping model $
+Fig.\ref{fig.2} shows the cross correlation functions,
+damping factors $W^2(k)$ for the optimal filtered 
+nonlinear and reconstructed density fields.
+We can see that the cross correlation $r_{LR}$ is recovered to
+nearly an order of magnitude smaller scales compared to $r_{LN}$.
+We also fit the Gaussian BAO damping model $
 {\mathcal D}(k)=\exp(-k^2\Sigma^2/2)$ and give $\Sigma=1.8$ Mpc$/h$ and $\Sigma=10$ 
 Mpc$/h$ for nonlinear and reconstructed fields.
+We repeat the analysis with various box sizes,
+and the results are consistent.
 
-Fig.\ref{fig.3} and Fig.\ref{fig.4} show their filtered power spectra.
+In Fig.\ref{fig.3} and Fig.\ref{fig.4} we show the power spectrum
+decomposition by equation (\ref{eq.power}) for both $\delta_N$ 
+(with subscripts ``$_R$'' replaced by ``$_N$'') and $\delta_R$.
+In Fig.\ref{fig.3} we can see that, on scales $k>0.2\ h/$Mpc,
+$P_{NN}$ term dominates the power spectrum due to a low correlation
+between $\delta_L$ and $delta_N$. The optimal filtered power spectrum
+damps quickly on small scales. In comparison, in Fig.\ref{fig.4},
+$P_{NN}$ starts to dominate on scales $k>1\ h/$Mpc. As a result, the
+optimal filtered reconstructed power spectrum
+(by equation (\ref{eq.opt})) is much higher.
 
@@
 $r_{LN}$ to $r_{LR}$ in Fig.\ref{fig.2} -- recovering the lost linear information on
-nearly an order of magnitude smaller scales. more tomorrow...
+nearly an order of magnitude smaller scales.
+This is achieved by implementing differential movement information
+of matter elements on Lagrangian coordinates, rather than on
+Eulerian coordinates. This result illustrates the feasibility
+of using estimated displacement field $\bs s(\bs q)$ to reconstruct primordial linear
+density field. A straightforward example of a estimation of $\bs s(\bs q)$
+is given by \cite{1995ApJS..100..269P,1998ApJS..115...19P}.
+More complicated algorithms can be found in \tcr{[cite]}, where
+shell-crossing effects are also considered. In reality, one needs
+to consider all aspects including vorticity, shell-crossing, bias, noise
+and data complexities. The impact of these factors can be quantitatively
+compared with the impact of different estimation methods, and with
+the exact solution by $N$-body simulations.
+
+\tcr{Is this reconstruction method optimal? Is there any other better methods?}
+
+\tcr{A proper name of this method?}
 
 \section*{Acknowledgements}
-Acknowledgements goes here.
+HRY acknowledges General Financial Grant from the China Postdoctoal Science Foundation No.2015M570884.
 

### 2016-10-04 ms_disp1/ms_disp1.tex
new figures

@@
 \usepackage{graphicx,longtable,natbib,mathrsfs,color}
-\usepackage{txfonts}
+\usepackage{txfonts,aas_macros}
 
@@
 
-%
-\def\apjl{Astrophys. J. Lett.}
-\def\apjs{Astrophys. J. Suppl. Ser.}
-\def\aj{Astron. J.}
-\def\mnras{Mon. Not. R. Astron. Soc.}
-\def\aap{Astron. Astrophys.}
-\def\jcap{J. Cosmology Astropart. Phys.}
-\def\aapr{Astron. Astrophys. Rev.}
-
 \begin{document}
@@
 \addtolength{\textwidth}{1.05cm}
-\title{Reconstruction of Large Scale Structure by Nonlinear \tcr{Lagrangian} Displacement Fields}
+\title{Reconstruction of Large Scale Structure by Nonlinear \tcr{Lagrangian} Displacement Fields - I}
 
@@
 \maketitle
-
+\begin{figure}
+ \centering
+  \includegraphics[width=0.99\linewidth]{denx3.pdf}
+  \caption{}
+  \label{fig.1}
+\end{figure}
 \section{Introduction}\label{sec.intro}
@@
 solved by a metric transformation equation \citep{1995ApJS..100..269P,1998ApJS..115...19P}.
-\tcr{[Peacock, Goldberg?]}
+\cite{1989ApJ...344L..53P,2000ApJ...544...21G}
+\tcr{[methods of Peebles, david Goldberg? least action ]}
 In 1-dimensional (1D) case, the exact solution of \cite{1995ApJS..100..269P}
@@
 
-\begin{figure}[b] \centering
-  \includegraphics[width=1.0\linewidth]{fig1.pdf}
-  \caption{Dimensionless auto- and cross- power spectra between nonlinear density field $\delta$, linear density field $\delta_L$ and reconstructed density field $\delta_R$.}
-  \label{fig.1}
-\end{figure}
 \begin{figure}[t] \centering
@@
 \begin{figure}[t] \centering
-  \includegraphics[width=1.0\linewidth]{fig3.pdf}
+  \includegraphics[width=1.0\linewidth]{recopower.pdf}
   \caption{Power spectra of the nonlinear density field and its
@@
 \begin{figure}[t] \centering
-  \includegraphics[width=1.0\linewidth]{fig4.pdf}
+  \includegraphics[width=1.0\linewidth]{pdfs.pdf}
   \caption{Power spectra of the reconstructed density field and its

### 2016-10-04 ms_disp1/ms_disp1.tex
oct 4

@@
 \addtolength{\textwidth}{1.05cm}
-\title{Reconstruction of Large Scale Structure by Nonlinear \tcr{Lagrangian} Displacement Fields - I}
+\title{Reconstruction of Large Scale Structure by Nonlinear \tcr{Lagrangian} Displacement Fields}
 
@@
 solved by a metric transformation equation \citep{1995ApJS..100..269P,1998ApJS..115...19P}.
-\cite{1989ApJ...344L..53P,2000ApJ...544...21G}
-\tcr{[methods of Peebles, david Goldberg? least action ]}
 In 1-dimensional (1D) case, the exact solution of \cite{1995ApJS..100..269P}
@@
 of an 1D simulation \citep{2016JCAP...01..043M} and obtain an estimated
-displacement field $s(q)$, and find that \tcr{this new method} well recovers the linear
+displacement field $\tilde\Psi(q)$, and find that \tcr{this new method} well recovers the linear
 information and reconstructs BAO peak in correlation function. 
-In 3D case, there are various techniques to obtain $\bs s(\bs q)$.
+In 3D case, there are various techniques to obtain $\tilde{\bs \Psi}(\bs q)$.
 However one needs to carefully consider effects of curl, shell-crossing in 3D case.
@@
 that can be recovered from the full nonlinear displacement field $\bs\Psi(\bs q)$,
-and further estimations $\bs s(\bs q)$ can be compared with this result.
+and further estimations $\tilde{\bs \Psi}(\bs q)$ can be compared with this result.
 In this paper we run a LSS $N$-body simulation and track the motion of
@@
 \subsection{Simulation}\label{ss.sim}
-We use the open source cosmological simulation code {\tt CUBE}%\footnote{github link of cafcube}
+We use the open source cosmological simulation code {\tt CUBE} %\footnote{github link of cafcube}
 \citep{cafcube}.
@@
 {\mathcal D}(k)=\exp(-k^2\Sigma^2/2)$ and give $\Sigma=1.8$ Mpc$/h$ and $\Sigma=10$ 
-Mpc$/h$ for nonlinear and reconstructed fields.
+Mpc$/h$ for reconstructed and nonlinear fields.
 We repeat the analysis with various box sizes,
@@
 $P_{NN}$ term dominates the power spectrum due to a low correlation
-between $\delta_L$ and $delta_N$. The optimal filtered power spectrum
+between $\delta_L$ and $\delta_N$. The optimal filtered power spectrum
 damps quickly on small scales. In comparison, in Fig.\ref{fig.4},
@@
 Eulerian coordinates. This result illustrates the feasibility
-of using estimated displacement field $\bs s(\bs q)$ to reconstruct primordial linear
-density field. A straightforward example of a estimation of $\bs s(\bs q)$
+of using estimated displacement field $\tilde{\bs \Psi}(\bs q)$ to reconstruct primordial linear
+density field. A straightforward example of a estimation of $\tilde{\bs \Psi}(\bs q)$
 is given by \cite{1995ApJS..100..269P,1998ApJS..115...19P}.
@@
 
+\cite{1989ApJ...344L..53P,2000ApJ...544...21G}
+\tcr{[methods of Peebles, david Goldberg? least action ]}
+
 \tcr{Is this reconstruction method optimal? Is there any other better methods?}
@@
 \section*{Acknowledgements}
-HRY acknowledges General Financial Grant from the China Postdoctoal Science Foundation No.2015M570884.
+We thank Hong-Ming Zhu for helpful discussions.
+HRY acknowledges General Financial Grant No.2015M570884 and Special Financial Grant No.2016T90009 from the China Postdoctoal Science Foundation.
 

### 2016-10-04 ms_disp1/ms_disp1.tex
results, discussion and conclusion

@@
  \centering
-  \includegraphics[width=0.99\linewidth]{denx3.pdf}
-  \caption{}
-  \label{fig.1}
+  \includegraphics[width=0.95\linewidth]{denx3.pdf}
+  \caption{Visualization of the nonlinear density field $\delta_N$ (top),
+  linear density field $\delta_L$ (middle) and the raw reconstructed density
+  field $\delta_R$ (bottom). These projections have 9.375 Mpc$/h$ thickness
+  and 300 Mpc$/h$ per side. The top panel shows the nonlinear displacement
+  field $\bs\Psi$ by the deformed mesh, which traces the LSS of $\delta_N$.}
+  \label{fig.projection}
 \end{figure}
@@
   \includegraphics[width=1.0\linewidth]{fig2.pdf}
-  \caption{Correlation functions $r(\delta_L,\delta)$ and $r(\delta_L,\delta_R)$
+  \caption{Correlation functions $r(\delta_L,\delta_N)$ and $r(\delta_L,\delta_R)$
   (solid lines). Their
@@
   and their best fit BAO damping models (dotted lines).}
-  \label{fig.2}
+  \label{fig.corr}
 \end{figure}
@@
 \delta_L$ is given by the initial stage and scaled to $z=0$ by the linear growth 
-factor.
+factor. In the top and middle panels of Fig.1 we show projections of the nonlinear density field
+$\delta_N$ given by the simulation and the linear density field $\delta_L$.
+$\delta_N$ is obtained by the particle distribution at redshift $z=0$, and
+the particles are interpolated by using the cloud-in-cell (CIC) algorithm.
+Because $\delta_N$ is highly nonlinear and follows an approximate
+log-normal distribution, we plot $\log_{10}(\delta_N+1)$ instead.
+The nonlinear evolution of $\delta_N$ makes it very different from $\delta_L$
+in apprearence.
 
-Two-point statistics of the density fields are quantified by the cross power 
+The two-point statistics of these density fields are quantified by the cross power 
 spectrum $P_{ij}(k)\equiv(2\pi)^{-3}\langle|\delta_i(k)||\delta_j(k)|\rangle$, 
@@
 usually plot the dimensionless power spectrum $\Delta^2(k)\equiv k^3P(k)/2\pi^2$.
-The blue solid and black dashed curves in Fig.\ref{fig.1} show the power spectra of 
-$\delta$ and $\delta_L$. Their difference shows the nonlinear evolution of LSS on 
-small scales. Their cross power drops to a very low value, indicating a loss of 
-linear information in the nonlinear power spectrum.
 
@@
 ({\bs q})$.
+To visualize the $\bs\Psi$ field, we draw a 3D uniform mesh over the volume,
+and use the given $\bs\Psi$ field to deform the mesh according to the direction
+and physical amplitude of $\bs\Psi$. In the top panel of Fig.\ref{fig.projection},
+The resulting mesh illustrates a pseudo
+``curvilinear coordinate'' similar to \cite{1995ApJS..100..269P},
+however the mesh can be overlapped due to shell-crossing. The densest mesh grids
+trace the densest structures of $\delta_N$, whereas the undeformed grid positions
+are the Lagrangian coordinates in which we do the reconstruction.
 The raw reconstructed density field is given by the differential motion of matter 
@@
 $k_q$ as $k$ to simplify the expression.
-In Fig.\ref{fig.1} we show the power spectrum of $\delta_R$ and its cross power
-with $\delta_L$. Despite of a lowered power of $\delta_R$ compared to $\delta$,
-it has a much higher cross power with $\delta_L$ compared to $\delta_N$,
-up to a relatively smaller scale (higher $k$).
+
 \begin{figure}[t] \centering
@@
   filtered density field.}
-  \label{fig.3}
+  \label{fig.recopower}
 \end{figure}
@@
   filtered density field.}
-  \label{fig.4}
+  \label{fig.pdfs}
 \end{figure}
@@
 and bias $b^2=P_{RR}/P_{LL}$, we solve $r'=P_{LR}/P_{LL}=rb$. We plot the cross 
-correlation coefficient $r_{LN}$ and $r_{LR}$ in Fig.\ref{fig.2}. Clearly, $
+correlation coefficient $r_{LN}$ and $r_{LR}$ in Fig.\ref{fig.corr}. Clearly, $
 \delta_R$ contains much more linear information on smaller scales.
@@
 \end{equation}
-The Wiener filters are also plot in Fig.\ref{fig.2}. The estimated linear density, or the optimal reconstructed density is given by
+The Wiener filters are also plot in Fig.\ref{fig.corr}. The estimated linear density, or the optimal reconstructed density is given by
 \begin{equation}
@@
 \end{equation}
-The full filters $Wb^{-1}$ are also plotted in Fig.\ref{fig.2}.
+%The full filters $Wb^{-1}$ are also plotted in Fig.\ref{fig.corr}.
 The optimal reconstructed power spectrum is given by
@@
 \section{Results}\label{sec.results}
-Fig.\ref{fig.2} shows the cross correlation functions,
+
+From the above algorithm, the projection of $\delta_R$ is plotted in
+the bottom panel of Fig.\ref{fig.projection}, which looks closer to 
+$\delta_N$. In the top panel of Fig.\ref{fig.recopower}, the blue solid and
+black dashed curves show the auto power spectra of 
+nonlinear density field $\delta_N$ and $\delta_L$. Their difference shows the nonlinear evolution of LSS on 
+small scales. Their cross power drops to a very low value, indicating a loss of 
+linear information in the nonlinear power spectrum.
+In the bottom panel of Fig.\ref{fig.recopower} 
+we show the auto power spectrum of $\delta_R$ and its cross power
+with $\delta_L$. Despite of a lowered power of $\delta_R$ compared to $\delta$,
+it has a much higher cross power with $\delta_L$ compared to $\delta_N$,
+up to a relatively smaller scale (higher $k$).
+
+Fig.\ref{fig.corr} shows the cross correlation functions,
 damping factors $W^2(k)$ for the optimal filtered 
@@
 
-In Fig.\ref{fig.3} and Fig.\ref{fig.4} we show the power spectrum
+In the two panels of Fig.\ref{fig.recopower} we show the power spectrum
 decomposition by equation (\ref{eq.power}) for both $\delta_N$ 
 (with subscripts ``$_R$'' replaced by ``$_N$'') and $\delta_R$.
-In Fig.\ref{fig.3} we can see that, on scales $k>0.2\ h/$Mpc,
+In the top panel, on scales $k>0.2\ h/$Mpc,
 $P_{NN}$ term dominates the power spectrum due to a low correlation
 between $\delta_L$ and $\delta_N$. The optimal filtered power spectrum
-damps quickly on small scales. In comparison, in Fig.\ref{fig.4},
+damps quickly on small scales. In comparison, in the bottom panel,
 $P_{NN}$ starts to dominate on scales $k>1\ h/$Mpc. As a result, the
@@
 (by equation (\ref{eq.opt})) is much higher.
+
+In Fig.\ref{fig.pdfs} we use the probability distribution
+as functions of $(\delta_L,\delta_N)$ and $(\delta_L,\delta_R)$
+to show the point-point correlation between these two pairs of density
+fields. In the top panel, $\delta_N$ shows an approximate log-normal
+distribution, and shows no correlation with $\delta_L$. Although
+$\delta_L$ and $\delta_N$ have correlations on linear scales (Fig.\ref{fig.corr}),
+they have no correlation in real space, because initial density fluctuations
+in Lagrangian coordinates are evolved/transformed to Eulerian coordinates.
+As the reconstruction is done in Lagrangian space, it recovers
+certain amount of correlation as shown in the bottom panel
+of Fig.\ref{fig.pdfs}. One can also see that $\delta_R$ follows
+a much closer Gaussian distribution.
+\tcr{We notice that the less-dense regions of $\delta_L$ are more
+correlated with $\delta_R$, whereas denser regions of $\delta_L$ are
+reconstructed worse, and $\delta_R$ shows more dispersion.}
 
@@
 the primordial linear perturbations. The result shows a prominent improvement from
-$r_{LN}$ to $r_{LR}$ in Fig.\ref{fig.2} -- recovering the lost linear information on
+$r_{LN}$ to $r_{LR}$ in Fig.\ref{fig.corr} -- recovering the lost linear information on
 nearly an order of magnitude smaller scales.
@@
 is given by \cite{1995ApJS..100..269P,1998ApJS..115...19P}.
-More complicated algorithms can be found in \tcr{[cite]}, where
+\tcr{More complicated reconstruction algorithms, e.g., least action
+algorithms \cite{1989ApJ...344L..53P,2000ApJ...544...21G} can give
+the exact $\delta_L$}, where
 shell-crossing effects are also considered. In reality, one needs
@@
 
-\cite{1989ApJ...344L..53P,2000ApJ...544...21G}
-\tcr{[methods of Peebles, david Goldberg? least action ]}
+\tcr{A proper name of this method (Lagrangian displacement)?}
 
-\tcr{Is this reconstruction method optimal? Is there any other better methods?}
+\tcr{The advantage of using displacement field in reconstruction is
+its insensitive response from highly nonlinearities --
+densest regions form because matter elements, after shell crossing,
+stop experiencing the linear extrapolated of Zel'dovich approximation.
+i.e. the displacement fields are dominated by early stage linear processes,
+which is the Lagrangian-Eulerian coordinate transform,
+while late stage shell-crossing, nonlinear and baryonic processes
+only fine-tunes the final position $\bs x$. In contrast, traditional
+treatments in reconstruction deals directly on density fields which sensitively
+relies on nonlinear processes --  density values can vary by orders
+of magnitude due to nonlinear/baryonic physics and many sources of errors.}
 
-\tcr{A proper name of this method?}
+\tcr{Advantages of this method (simpler?) over least action?}
 

### 2016-10-13 ms_disp1/ms_disp1.tex
figure captions
update figure
title, abstract, figures, box size

@@
 \addtolength{\textwidth}{1.05cm}
-\title{Reconstruction of Large Scale Structure by Nonlinear \tcr{Lagrangian} Displacement Fields}
+\title{Lagrangian Space Nonlinear $E$-mode clustering}
 
@@
 
+\author{Hong-Ming Zhu}%email{hmzhu@bao.ac.cn}
+\affiliation{Key Laboratory for Computational Astrophysics, National Astronomical Observatories, Chinese Academy of Sciences, 20A Datun Road, Beijing 100012, China}
+\affiliation{University of Chinese Academy of Sciences, Beijing 100049, China}
+
 \date{\today}
@@
 \begin{abstract}
-We investigate the ability of using nonlinear displacement field to reconstruct 
-the primordial linear perturbations of the large scale structure. According to 
-the linear Lagrangian perturbation theory, the first order primordial linear 
-density field is given by the negative divergence of the displacement field. By 
-running $N$-body simulations, we show that this \tcr{new} reconstruction algorithm recovers 
-great amount of nonlinear information in the early universe. In reality, the 
-displacement fields can be solved from only the final stage of the density 
-fields. This demonstrates the potential to reconstruct baryonic acoustic oscillation (BAO) 
+We study the nonlinear $E$-mode clustering in Lagrangian space
+by using large scale structure (LSS) $N$-body simulations
+and use the displacement field information in Lagrangian space
+to estimate the primordial linear density field.
+We find that, compared to Eulerian nonlinear density fields,
+the negative divergence of $E$-mode displacement fields
+in Lagrangian space improves the cross-correlation with
+initial density field by factor of 6 $\sim$ 7,
+containing 2 orders of magnitude more primordial information. 
+In reality, the displacement field is dominated by linear evolution
+and insensitive of nonlinear shell-crossing, vorticity or
+baryonic physics, and thus can be stably solved from only
+the final matter distributions.
+This illustrates ability of potential reconstruction algorithms,
+to reconstruct baryonic acoustic oscillation (BAO) 
 from current and future large scale structure surveys.
@@
  \centering
-  \includegraphics[width=0.95\linewidth]{denx3.pdf}
+  \includegraphics[width=0.95\linewidth]{proj.pdf}
   \caption{Visualization of the nonlinear density field $\delta_N$ (top),
@@
   field $\delta_R$ (bottom). These projections have 9.375 Mpc$/h$ thickness
-  and 300 Mpc$/h$ per side. The top panel shows the nonlinear displacement
+  and 400 Mpc$/h$ per side. The top panel shows the nonlinear displacement
   field $\bs\Psi$ by the deformed mesh, which traces the LSS of $\delta_N$.}
@@
 \citep{2007ApJ...664..675E,2009PhRvD..80l3501N,2009PhRvD..79f3523P}. Here we propose a 
-\tcr{new reconstruction method} that uses the nonlinear displacement field to recover the 
+new reconstruction method that uses the nonlinear displacement field to recover the 
 primordial density field. In the linear Lagrangian perturbation theory, the 
@@
 $\bs\Psi(\bs q)$ respect to Lagrangian coordinates $\bs q$
-gives the linear density field \tcr{\citep{2010PhDT.........4J}}.
+gives the linear density field \citep{2010PhDT.........4J}.
 Of course, the full displacement field $\bs\Psi(\bs q)$
@@
 of an 1D simulation \citep{2016JCAP...01..043M} and obtain an estimated
-displacement field $\tilde\Psi(q)$, and find that \tcr{this new method} well recovers the linear
+displacement field $\tilde\Psi(q)$, and find that this new method well recovers the linear
 information and reconstructs BAO peak in correlation function. 
@@
 \begin{figure}[t] \centering
-  \includegraphics[width=1.0\linewidth]{fig2.pdf}
+  \includegraphics[width=1.0\linewidth]{filter.pdf}
   \caption{Correlation functions $r(\delta_L,\delta_N)$ and $r(\delta_L,\delta_R)$
-  (solid lines). Their
-  damping of the power spectra $W^2$ (dashed lines),
-  and their best fit BAO damping models (dotted lines).}
+  (solid lines) and their scaled BAO damping models (dotted lines).}
   \label{fig.corr}
@@
 $N_p=512^3$ $N$-body particles are evolved via 
-their mutual gravitational interactions to $z=0$, in a periodic box with $L=300$ 
+their mutual gravitational interactions to $z=0$, in a periodic box with $L=400$ 
 Mpc$/h$ per side. The code is set to use standard a particle-mesh (PM) algorithm 
@@
 \begin{figure}[t] \centering
-  \includegraphics[width=1.0\linewidth]{recopower.pdf}
-  \caption{Power spectra of the nonlinear density field and its
+  \includegraphics[width=1.0\linewidth]{power.pdf}
+  \caption{Power spectra decomposition of density fields and their
   filtered density field.}
@@
   \includegraphics[width=1.0\linewidth]{pdfs.pdf}
-  \caption{Power spectra of the reconstructed density field and its
-  filtered density field.}
+  \caption{2D-PDF (preliminary)}
   \label{fig.pdfs}
@@
 We also fit the Gaussian BAO damping model $
-{\mathcal D}(k)=\exp(-k^2\Sigma^2/2)$ and give $\Sigma=1.8$ Mpc$/h$ and $\Sigma=10$ 
+{\mathcal D}(k)=\exp(-k^2\Sigma^2/2)$ and give $\Sigma=1.8$ Mpc$/h$ and $\Sigma=12$ 
 Mpc$/h$ for reconstructed and nonlinear fields.
@@
 a much closer Gaussian distribution.
-\tcr{We notice that the less-dense regions of $\delta_L$ are more
+We notice that the less-dense regions of $\delta_L$ are more
 correlated with $\delta_R$, whereas denser regions of $\delta_L$ are
-reconstructed worse, and $\delta_R$ shows more dispersion.}
+reconstructed worse, and $\delta_R$ shows more dispersion.
+This is caused by nonlinear effects and it damps out in
+higher redshifts.
 
@@
 is given by \cite{1995ApJS..100..269P,1998ApJS..115...19P}.
-\tcr{More complicated reconstruction algorithms, e.g., least action
-algorithms \cite{1989ApJ...344L..53P,2000ApJ...544...21G} can give
-the exact $\delta_L$}, where
-shell-crossing effects are also considered. In reality, one needs
+%\tcr{More complicated reconstruction algorithms, e.g., least action
+%algorithms \cite{1989ApJ...344L..53P,2000ApJ...544...21G} can give
+%the exact $\delta_L$}, where
+%shell-crossing effects are also considered.
+In reality, one needs
 to consider all aspects including vorticity, shell-crossing, bias, noise
@@
 
-\tcr{A proper name of this method (Lagrangian displacement)?}
-
-\tcr{The advantage of using displacement field in reconstruction is
+The advantage of using displacement field in reconstruction is
 its insensitive response from highly nonlinearities --
@@
 relies on nonlinear processes --  density values can vary by orders
-of magnitude due to nonlinear/baryonic physics and many sources of errors.}
-
-\tcr{Advantages of this method (simpler?) over least action?}
+of magnitude due to nonlinear/baryonic physics and many sources of errors.
 
 \section*{Acknowledgements}
-We thank Hong-Ming Zhu for helpful discussions.
+We thank Pengjie Zhang for helpful discussions which triggered
+the inspiration of the work.
 HRY acknowledges General Financial Grant No.2015M570884 and Special Financial Grant No.2016T90009 from the China Postdoctoal Science Foundation.
+[ZHM's acknowledgements.]
+The Dunlap Institute is funded through an endowment established by the David Dunlap family and the University of Toronto.
+Research at the Perimeter Institute is supported by the Government of Canada
+through Industry Canada and by the Province of Ontario through the Ministry of
+Research $\&$ Innovation.
 

### 2016-10-14 ms_disp1/ms_disp1.tex
updated text

@@
 
-Standard BAO reconstruction algorithms smooth the 
-nonlinear density field on linear scale ($\sim$10 Mpc$/h$) and reverse the large 
+Standard BAO reconstruction algorithms apply in Eulerian space.
+They usually smooth the nonlinear density field on
+linear scale ($\sim$10 Mpc$/h$) and reverse the large 
 scale bulk flows by a negative Zel'dovich linear displacement
-\citep{2007ApJ...664..675E,2009PhRvD..80l3501N,2009PhRvD..79f3523P}. Here we propose a 
-new reconstruction method that uses the nonlinear displacement field to recover the 
-primordial density field. In the linear Lagrangian perturbation theory, the 
+\citep{2007ApJ...664..675E,2009PhRvD..80l3501N,2009PhRvD..79f3523P}.
+In the linear Lagrangian perturbation theory, the 
 negative divergence of the the displacement field
@@
 gives the linear density field \citep{2010PhDT.........4J}.
+Here we study the Lagrangian space clustering of dark matter
+and use this information to estimate the initial density field.
 Of course, the full displacement field $\bs\Psi(\bs q)$
 is non-observable, as it requires the initial distributions of matter, however there are 
-many techniques to estimate the nonlinear displacement field from a final 
+many techniques to estimate the displacement field from a final 
 distribution of matter. For example, when a homogeneous initial matter distribution 
@@
 solved by a metric transformation equation \citep{1995ApJS..100..269P,1998ApJS..115...19P}.
-In 1-dimensional (1D) case, the exact solution of \cite{1995ApJS..100..269P}
+In 1-dimensional (1D) case, the exact solution 
 simplifies to an ordering of matter elements by Eulerian coordinates.
@@
 displacement field $\tilde\Psi(q)$, and find that this new method well recovers the linear
-information and reconstructs BAO peak in correlation function. 
-In 3D case, there are various techniques to obtain $\tilde{\bs \Psi}(\bs q)$.
-However one needs to carefully consider effects of curl, shell-crossing in 3D case.
+information and reconstructs 1D BAO peak in correlation function. 
+In a bit more complicated 3D case, there are various
+techniques to estimate $\tilde{\bs \Psi}(\bs q)$,
+and we need to carefully consider effects of curl, shell-crossing in 3D cases.
 
@@
 and further estimations $\tilde{\bs \Psi}(\bs q)$ can be compared with this result.
-In this paper we run a LSS $N$-body simulation and track the motion of
-particles to obtain $\bs\Psi(\bs q)$. According to this we reconstruct the linear density
+In this paper we run LSS $N$-body simulations and track the motion of
+particles to obtain $\bs\Psi(\bs q)$. According to this we estimate the linear density
 field and compare to the primordial density field of the initial
@@
 \section{Method}\label{sec.method}
-We show the LSS simulation and displacement field setups in section \ref{ss.sim},
-and the reconstruction processes are presented in section \ref{ss.reco}.
+We show the LSS simulation and displacement field setups in section \ref{ss.sim}.
+In section \ref{ss.reco}, we estimate the linear density field from the
+displacement field $\bs\Psi(\bs q)$ from simulations.
+Potential reconstruction algorithms are
+based on an estimated displacement field $\tilde{\bs \Psi}(\bs q)$
+instead of $\bs\Psi(\bs q)$. In the following sections we
+simply call the estimation of linear density field
+from $\tilde{\bs \Psi}(\bs q)$ ``reconstruction''.
 
@@
 The nonlinear evolution of $\delta_N$ makes it very different from $\delta_L$
-in apprearence.
+in appearance.
 
@@
 spectrum $P_{ij}(k)\equiv(2\pi)^{-3}\langle|\delta_i(k)||\delta_j(k)|\rangle$, 
-where subscripts $i,j$ may refer to linear ($L$), nonlinear ($N$), or reconstructed ($R$) density 
+where subscripts $i,j$ may refer to linear ($L$), nonlinear ($N$),
+reconstructed ($R$), or noise ($n$) density 
 fields. When $i=j$ it reduces to the auto power spectrum $P_{ii}(k)$ or $P(k)$. We 
@@
   \includegraphics[width=1.0\linewidth]{power.pdf}
-  \caption{Power spectra decomposition of density fields and their
-  filtered density field.}
+  \caption{Power spectra of $\delta_L$, $\delta_N$, and the decomposition
+  of $P_{LL}$ into correlated parts and noise terms according to equation
+  (\ref{eq.power}).}
   \label{fig.recopower}
@@
   \includegraphics[width=1.0\linewidth]{pdfs.pdf}
-  \caption{2D-PDF (preliminary)}
+  \caption{Probability distribution functions (PDFs)
+  $f(\delta_L,\delta_N)$ and $f(\delta_L,\delta_R)$, showing in the upper
+  and lower panels respectively. Both red curves on $x$-axes show $f(\delta_L)$, following
+  a Gaussian distribution. The two curves on $y$-axes in upper and lower panel
+  show respectively $f(\delta_N)$ and $f(\delta_R)$.}
   \label{fig.pdfs}
@@
 
-To quantify the linear information in the reconstructed density field, we decompose $\delta_R$ in Fourier space as
+To quantify the linear information in the reconstructed density field, we decompose $\delta_L$ in Fourier space as
 \begin{equation}\label{eq.decompose}
-    \delta_R(k)=r'\delta_L+\delta_N,
+    \delta_L(k)=r'\delta_R+\delta_n,
 \end{equation}
-where $r'\delta_R$ is completely correlated with linear density $\delta_L$. Correlating equation (\ref{eq.decompose}) with $\delta_L$ gives
+where $r'\delta_R$ is completely correlated with linear density $\delta_L$. Correlating equation (\ref{eq.decompose}) with $\delta_R$ gives
 \begin{equation}
-    P_{LR}=r'P_{LL}+P_{LN},
+    P_{LR}=r'P_{RR}+P_{nR},
 \end{equation}
 where $P_{ij}\equiv\langle\delta_i\delta_j\rangle$ denotes the cross power 
-spectrum. Since $\delta_N$ is uncorrelated with $\delta_L$, $P_{LN}=0$. With the 
+spectrum. Since $\delta_n$ is uncorrelated with $\delta_R$, $P_{nR}=0$. With the 
 definition of cross correlation coefficient $r(\delta_L,\delta_R)\equiv P_{LR}/\sqrt{P_{LL}P_{RR}}$ 
-and bias $b^2=P_{RR}/P_{LL}$, we solve $r'=P_{LR}/P_{LL}=rb$. We plot the cross 
-correlation coefficient $r_{LN}$ and $r_{LR}$ in Fig.\ref{fig.corr}. Clearly, $
+and bias $b^2=P_{RR}/P_{LL}$, we solve $r'=P_{LR}/P_{RR}=rb^{-1}$.
+Note that these computations above and below can also be applied on $\delta_N$ for
+comparison, by replacing ``$_R$'' with ``$_N$'' in the equations,
+while we do not rewrite them explicitly
+in the paper for simplicity. From these, we plot the cross 
+correlation coefficient $r_{LN}=r(\delta_L,\delta_N)$ and
+$r_{LR}=r(\delta_L,\delta_R)$ in Fig.\ref{fig.corr}. Clearly, $
 \delta_R$ contains much more linear information on smaller scales.
@@
 \begin{equation}\label{eq.power}
-    P_{RR}=r^2b^2P_{LL}+P_{NN},
+    P_{LL}=r^2b^{-2}P_{RR}+P_{nn},
 \end{equation}
-and $P_{NN}=(1-r^2)P_{RR}$. Then we construct a Wiener filter to filter out the uncorrelated part in $\delta_R$:
+and $P_{nn}=(1-r^2)P_{LL}$. Then we construct a Wiener filter to filter out the uncorrelated part in $\delta_R$:
 \begin{equation}
-    W(k)=\frac{r^2b^2P_{LL}}{r^2b^2P_{LL}+P_{NN}}=r^2.
+    W(k)=\frac{r^2b^{-2}P_{RR}}{r^2b^{-2}P_{RR}+P_{nn}}=r^2.
 \end{equation}
-The Wiener filters are also plot in Fig.\ref{fig.corr}. The estimated linear density, or the optimal reconstructed density is given by
+The optimal reconstructed density is given by
 \begin{equation}
-    \tilde\delta_R=Wb^{-1}\delta_R.
+    \tilde\delta_R=Wb^{-1}\delta_R,
 \end{equation}
-%The full filters $Wb^{-1}$ are also plotted in Fig.\ref{fig.corr}.
-The optimal reconstructed power spectrum is given by
+and the optimal reconstructed power spectrum is given by
 \begin{equation}\label{eq.opt}
@@
 \end{equation}
-The $W^2$ describes the damping of the linear power spectrum.
+Here $W^2$ describes the damping of the linear power spectrum.
 
@@
 
-From the above algorithm, the projection of $\delta_R$ is plotted in
+To visualize the above algorithms, a projection of $\delta_R$ is plotted in
 the bottom panel of Fig.\ref{fig.projection}, which looks closer to 
-$\delta_N$. In the top panel of Fig.\ref{fig.recopower}, the blue solid and
-black dashed curves show the auto power spectra of 
-nonlinear density field $\delta_N$ and $\delta_L$. Their difference shows the nonlinear evolution of LSS on 
-small scales. Their cross power drops to a very low value, indicating a loss of 
-linear information in the nonlinear power spectrum.
-In the bottom panel of Fig.\ref{fig.recopower} 
-we show the auto power spectrum of $\delta_R$ and its cross power
-with $\delta_L$. Despite of a lowered power of $\delta_R$ compared to $\delta$,
-it has a much higher cross power with $\delta_L$ compared to $\delta_N$,
-up to a relatively smaller scale (higher $k$).
+$\delta_L$ compared to $\delta_N$. However the smallest scales are unable
+to be reconstructed.
 
-Fig.\ref{fig.corr} shows the cross correlation functions,
-damping factors $W^2(k)$ for the optimal filtered 
-nonlinear and reconstructed density fields.
-We can see that the cross correlation $r_{LR}$ is recovered to
-nearly an order of magnitude smaller scales compared to $r_{LN}$.
-We also fit the Gaussian BAO damping model $
-{\mathcal D}(k)=\exp(-k^2\Sigma^2/2)$ and give $\Sigma=1.8$ Mpc$/h$ and $\Sigma=12$ 
+As discussed in section \ref{ss.reco},
+Fig.\ref{fig.corr} shows the cross correlation functions
+$r_{LN}$ and $r_{LR}$. The latter extends the correlation
+with $\delta_L$ to smaller scales by nearly an order of
+magnitude. The extra correlation scales well cover the BAO
+scales of our interest.
+
+In Fig.\ref{fig.recopower}, we show the auto power spectra of
+$\delta_L$ and $\delta_N$ in black dashed and blue solid curves.
+Their difference shows the nonlinear evolution of LSS on 
+small scales. Their cross power (not shown for clarity of the figure)
+drops to a very low value on scales $k\gtrsim 0.1h/$Mpc,
+indicating a loss of linear information in the
+nonlinear power spectrum $P_{NN}$. This scenario directly
+leads to how $\P_{LL}$ is decomposed according to equation
+(\ref{eq.power}). In nonlinear case (with ``$_R$'' replaced
+by ``$_N$'' equation (\ref{eq.power})), on small
+scales $k\gtrsim 0.1h/$Mpc, $P_{LL}$ is dominated by
+uncorrelated, nonlinear noise, shown in the green
+dotted line. In reconstructed case, however, $P_{LL}$
+is decomposed into orange dash-dotted correlated
+part and purple dotted uncorrelated part according to
+equation (\ref{eq.power}). The correlated power spectrum
+is dominated on BAO scales of our interest.
+
+To quantify the improvement of cross-correlation in the power
+spectrum, we compute the 
+damping factors $W^2(k)$ respectively for the optimal filtered 
+nonlinear and reconstructed density fields $\tilde\delta_N$
+and $\tilde\delta_R$.
+We fit Gaussian BAO damping models
+${\mathcal D}(k)=\exp(-k^2\Sigma^2/2)$ to these $W^2(k)$'s
+and give $\Sigma=1.8$ Mpc$/h$ and $\Sigma=12$ 
 Mpc$/h$ for reconstructed and nonlinear fields.
-We repeat the analysis with various box sizes,
-and the results are consistent.
+Since ${\mathcal D}(k)=W^2=r^4$, we plot
+${\mathcal D}_N^{1/4}$ and ${\mathcal D}_R^{1/4}$
+over $r_{LN}$ and $r_{LR}$ in Fig.\ref{fig.corr}.
+The analyses are repeated with various box sizes (100, 300, 800 Mpc$/h$ per side)
+and give consistent results.
 
-In the two panels of Fig.\ref{fig.recopower} we show the power spectrum
-decomposition by equation (\ref{eq.power}) for both $\delta_N$ 
-(with subscripts ``$_R$'' replaced by ``$_N$'') and $\delta_R$.
-In the top panel, on scales $k>0.2\ h/$Mpc,
-$P_{NN}$ term dominates the power spectrum due to a low correlation
-between $\delta_L$ and $\delta_N$. The optimal filtered power spectrum
-damps quickly on small scales. In comparison, in the bottom panel,
-$P_{NN}$ starts to dominate on scales $k>1\ h/$Mpc. As a result, the
-optimal filtered reconstructed power spectrum
-(by equation (\ref{eq.opt})) is much higher.
-
-In Fig.\ref{fig.pdfs} we use the probability distribution
-as functions of $(\delta_L,\delta_N)$ and $(\delta_L,\delta_R)$
+To further illustrate the improvement in real space
+one point function correlations,
+in Fig.\ref{fig.pdfs} we use the probability distribution
+as functions (PDFs) of $(\delta_L,\delta_N)$ and $(\delta_L,\delta_R)$
 to show the point-point correlation between these two pairs of density
-fields. In the top panel, $\delta_N$ shows an approximate log-normal
-distribution, and shows no correlation with $\delta_L$. Although
-$\delta_L$ and $\delta_N$ have correlations on linear scales (Fig.\ref{fig.corr}),
-they have no correlation in real space, because initial density fluctuations
+fields. Since $\delta_n$ in equation (\ref{eq.decompose}) is
+uncorrelated, we use Wiener filtered fields. To keep the
+consistency over $\delta_L$, $\delta_N$ and $\delta_R$, we
+use the $W(k)=r^2_{LR}$ as the Wiener filter.
+The grey-scaled plots in the center of both panels show the 2-variable PDFs,
+whereas their projections onto each variable are just
+1-variable PDFs -- $f(\delta_L)$, $f(\delta_N)$ and $f(\delta_R)$, shown
+as red/blue curves on the axes of Fig.\ref{fig.pdfs}.
+In the top panel, $\delta_N$ shows an approximate log-normal
+distribution (blue curve) and $\delta_L$ follows an expected
+Gaussian distribution. They show tiny positive correlation in the
+2D PDF.
+Because in Fourier space, $\delta_L$ and $\delta_N$ have
+correlations on only very large, linear scales (Fig.\ref{fig.corr}),
+they result in little correlation in real space -- initial density fluctuations
 in Lagrangian coordinates are evolved/transformed to Eulerian coordinates.
 As the reconstruction is done in Lagrangian space, it recovers
-certain amount of correlation as shown in the bottom panel
+certain amount of correlation, as shown in the 2D PDF of the bottom panel
 of Fig.\ref{fig.pdfs}. One can also see that $\delta_R$ follows
-a much closer Gaussian distribution.
-We notice that the less-dense regions of $\delta_L$ are more
+a much closer Gaussian distribution (blue curve of the bottom panel).
+We notice that the less-dense regions of $\delta_L$ are strongly
 correlated with $\delta_R$, whereas denser regions of $\delta_L$ are
-reconstructed worse, and $\delta_R$ shows more dispersion.
-This is caused by nonlinear effects and it damps out in
-higher redshifts.
+reconstructed less better, and $\delta_R$ shows more dispersion.
+These second uncorrelated peaks in the 2D PDF and $f(\delta_R)$
+are caused by nonlinear effects. We examine the reconstruction on
+higher redshifts, and confirm that these second uncorrelated peaks
+damp out as we go to higher redshifts.
 
@@
 We extract the actual displacement field of matter elements in cosmological $N$-body
-simulations, and use this displacement field in Lagrangian coordinates to reconstruct
+simulations, and use this displacement field to study the LSS nonlinear
+clustering in Lagrangian space. The displacement information is used to reconstruct
 the primordial linear perturbations. The result shows a prominent improvement from
@@
 HRY acknowledges General Financial Grant No.2015M570884 and Special Financial Grant No.2016T90009 from the China Postdoctoal Science Foundation.
-[ZHM's acknowledgements.]
+HMZ acknowledges the support of the Chinese MoST 863 program under Grant
+No. 2012AA121701, the CAS Science Strategic Priority Research Program
+XDB09000000 and the NSFC under Grant No. 11373030.
 The Dunlap Institute is funded through an endowment established by the David Dunlap family and the University of Toronto.

### 2016-10-20 ms_disp1/ms_disp1.tex
oct 20

@@
 and use the displacement field information in Lagrangian space
-to estimate the primordial linear density field.
+to recover the primordial linear density field.
 We find that, compared to Eulerian nonlinear density fields,
@@
 the final matter distributions.
-This illustrates ability of potential reconstruction algorithms,
-to reconstruct baryonic acoustic oscillation (BAO) 
+This illustrates ability of potential density reconstruction algorithms,
+to improve the baryonic acoustic oscillation (BAO) mesurements
 from current and future large scale structure surveys.
@@
   \caption{Visualization of the nonlinear density field $\delta_N$ (top),
-  linear density field $\delta_L$ (middle) and the raw reconstructed density
+  linear density field $\delta_L$ (middle) and the raw recovered density
   field $\delta_R$ (bottom). These projections have 9.375 Mpc$/h$ thickness
@@
 In this paper we run LSS $N$-body simulations and track the motion of
-particles to obtain $\bs\Psi(\bs q)$. According to this we estimate the linear density
+particles to obtain $\bs\Psi(\bs q)$. According to this we recover the linear density
 field and compare to the primordial density field of the initial
-conditions. We describe the simulation and reconstruction algorithm in section
+conditions. We describe the simulation and density recovery algorithm in section
 \ref{sec.method}, and we show the results in section \ref{sec.results}.
@@
 We show the LSS simulation and displacement field setups in section \ref{ss.sim}.
-In section \ref{ss.reco}, we estimate the linear density field from the
+In section \ref{ss.reco}, we recover the linear density field from the
 displacement field $\bs\Psi(\bs q)$ from simulations.
-Potential reconstruction algorithms are
+Note that potential reconstruction algorithms are
 based on an estimated displacement field $\tilde{\bs \Psi}(\bs q)$
 instead of $\bs\Psi(\bs q)$. In the following sections we
-simply call the estimation of linear density field
-from $\tilde{\bs \Psi}(\bs q)$ ``reconstruction''.
+use $\delta_R$ to label the recovered linear density field
+from $\tilde{\bs \Psi}(\bs q)$.
 
@@
 where subscripts $i,j$ may refer to linear ($L$), nonlinear ($N$),
-reconstructed ($R$), or noise ($n$) density 
+recovered ($R$), or noise ($n$) density 
 fields. When $i=j$ it reduces to the auto power spectrum $P_{ii}(k)$ or $P(k)$. We 

### 2016-10-22 ms_disp1/ms_disp1.tex
i

@@
 \addtolength{\textwidth}{1.05cm}
-\title{Lagrangian Space Nonlinear $E$-mode clustering}
+\title{\tcr{Recovery of Primordial Density Field by Reversing the Lagrangian Space Nonlinear $E$-mode clustering}}
 
@@
 
-Standard BAO reconstruction algorithms apply in Eulerian space.
+\tcb{Standard BAO reconstruction algorithms apply in Eulerian space.
 They usually smooth the nonlinear density field on
@@
 scale bulk flows by a negative Zel'dovich linear displacement
-\citep{2007ApJ...664..675E,2009PhRvD..80l3501N,2009PhRvD..79f3523P}.
-In the linear Lagrangian perturbation theory, the 
+\citep{2007ApJ...664..675E,2009PhRvD..80l3501N,2009PhRvD..79f3523P}.}
+In the linear Lagrangian perturbation theory (LPT), the 
 negative divergence of the the displacement field
-$\bs\Psi(\bs q)$ respect to Lagrangian coordinates $\bs q$
-gives the linear density field \citep{2010PhDT.........4J}.
-Here we study the Lagrangian space clustering of dark matter
-and use this information to estimate the initial density field.
-Of course, the full displacement field $\bs\Psi(\bs q)$
-is non-observable, as it requires the initial distributions of matter, however there are 
-many techniques to estimate the displacement field from a final 
-distribution of matter. For example, when a homogeneous initial matter distribution 
+$-\nabla_q\cdot\bs\Psi(\bs q)$ respect to Lagrangian coordinates $\bs q$
+gives the linear density field \citep{2010PhDT.........4J},
+\tcr{and in many literatures
+\cite{2013MNRAS.428..141N,2014PhRvD..89h3515C,2016JCAP...03..017B}
+$\nabla_q\cdot\bs\Psi(\bs q)$ is compared with various standard LPT
+and corrected LPT initial conditions. Their motivation was
+trying to correct or modify the LPT displacement fields
+in order that the final positions of particles are brought
+closer to $N$-body results. Their results show that 
+all LPT methods are still unable to mimic the shell-crossing
+scales given by $N$-body simulation. Because of the absence
+of Lagrangian space information in observations, few density
+reconstruction algorithms are developed according to the relation
+between displacement and linear density. However,
+there are techniques to estimate the displacement field from a final 
+distribution of matter.}
+For example, when a homogeneous initial matter distribution 
 is assumed, there is a unique solution of curl-less displacement field to relate 
 the initial and final distributions without shell-crossing. This solution can be 
-solved by a metric transformation equation \citep{1995ApJS..100..269P,1998ApJS..115...19P}.
+solved by a metric transformation equation
+\citep{1995ApJS..100..269P,1998ApJS..115...19P}.
 In 1-dimensional (1D) case, the exact solution 
@@
 of an 1D simulation \citep{2016JCAP...01..043M} and obtain an estimated
-displacement field $\tilde\Psi(q)$, and find that this new method well recovers the linear
+displacement field $\tilde\Psi(q)$, and find
+that this new method well recovers the linear
 information and reconstructs 1D BAO peak in correlation function.
-In a bit more complicated 3D case, there are various
-techniques to estimate $\tilde{\bs \Psi}(\bs q)$,
-and we need to carefully consider effects of curl, shell-crossing in 3D cases.
+\tcr{In 3D and more realistic cases, one needs to
+carefully consider effects of curl, shell-crossing, 
+complicated baryonic physics and biased tracers (galaxies).}
 
-Before these steps, we need to quantify the amount of linear information
-that can be recovered from the full nonlinear displacement field $\bs\Psi(\bs q)$,
-and further estimations $\tilde{\bs \Psi}(\bs q)$ can be compared with this result.
-In this paper we run LSS $N$-body simulations and track the motion of
-particles to obtain $\bs\Psi(\bs q)$. According to this we recover the linear density
-field and compare to the primordial density field of the initial
-conditions. We describe the simulation and density recovery algorithm in section
-\ref{sec.method}, and we show the results in section \ref{sec.results}.
+\tcr{Before these steps, we need to quantify the amount of linear information
+that can be recovered\footnote{To avoid ambiguity, we call
+$\delta_R=-\nabla_q\cdot\bs\Psi(\bs q)$ recovered linear density field, which
+requires $\bs\Psi(\bs q)$ from $N$-body simulations, while
+$-\nabla_q\cdot\tilde{\bs \Psi}(\bs q)$ is called the
+reconstructed density field.}
+ from the full nonlinear displacement field $\bs\Psi(\bs q)$.
+In $\bs\Psi(\bs q)$ from $N$-body simulations,
+(1) shell-crossing effects are fully considered; (2) $E$-mode displacement
+field is used for recovering the linear information (although $B$-mode is
+also present); (3) the setup is clean in absence of baryons and realistic
+observables. So it sets the best quality that potential reconstruction
+algorithms can achieve and further reconstructions by $\tilde{\bs \Psi}(\bs q)$
+can be compared with this result. Actually, in the later discussions and
+our next works we will show that, compared to $\tilde{\bs \Psi}(\bs q)$,
+additional information (shell-crossing) do not enhance the reconstruction
+prominently. Further more, we study the scale dependent cross-correlation
+between $-\nabla_q\cdot\bs\Psi(\bs q)$ and linear density field
+and construct optimal Wiener filters to get the optimal filtered
+recovered linear density field and recovered linear power spectrum.}
+
+In the rest part of the paper, we describe the simulation,
+density recovery algorithm and Wiener filter setups in section
+\ref{sec.method}. We show results in section \ref{sec.results}.
 Discussion and conclusion are in section \ref{sec.discussion}.
@@
 use $\delta_R$ to label the recovered linear density field
-from $\tilde{\bs \Psi}(\bs q)$.
+from $\bs\Psi(\bs q)$.
 
@@
 
-\subsection{Reconstruction}\label{ss.reco}
+\subsection{Density Recovery}\label{ss.reco}
 In the simulation, we use particle-ID (PID) to record the initial (Lagrangian) location ${\bs 
@@
 trace the densest structures of $\delta_N$, whereas the undeformed grid positions
-are the Lagrangian coordinates in which we do the reconstruction.
-The raw reconstructed density field is given by the differential motion of matter 
+are the Lagrangian coordinates in which we do the density recovery.
+The raw recovered density field is given by the differential motion of matter 
 elements,
@@
 \end{equation}
-Because the reconstruction processes are implemented on Lagrangian coordinates,
+Because the recovery processes are implemented on Lagrangian coordinates,
 $\delta_R$ takes the coordinates of $\bs q$ instead of $\bs x$.
@@
   \includegraphics[width=1.0\linewidth]{power.pdf}
-  \caption{Power spectra of $\delta_L$, $\delta_N$, and the decomposition
+  \caption{\tcr{Power spectra of $\delta_L$, $\delta_N$, and the decomposition
   of $P_{LL}$ into correlated parts and noise terms according to equation
-  (\ref{eq.power}).}
+  (\ref{eq.power}).}}
   \label{fig.recopower}
@@
 
-To quantify the linear information in the reconstructed density field, we decompose $\delta_L$ in Fourier space as
+To quantify the linear information in the recovered density field, we decompose $\delta_L$ in Fourier space as
 \begin{equation}\label{eq.decompose}
@@
 \end{equation}
-The optimal reconstructed density is given by
+The optimal recovered density is given by
 \begin{equation}
@@
 \end{equation}
-and the optimal reconstructed power spectrum is given by
+and the optimal recovered power spectrum is given by
 \begin{equation}\label{eq.opt}
@@
 the bottom panel of Fig.\ref{fig.projection}, which looks closer to 
-$\delta_L$ compared to $\delta_N$. However the smallest scales are unable
-to be reconstructed.
+$\delta_L$ compared to $\delta_N$. However the smallest scale structures are unable
+to be recovered.
 
@@
 nonlinear power spectrum $P_{NN}$. This scenario directly
-leads to how $\P_{LL}$ is decomposed according to equation
+leads to how $P_{LL}$ is decomposed according to equation
 (\ref{eq.power}). In nonlinear case (with ``$_R$'' replaced
@@
 uncorrelated, nonlinear noise, shown in the green
-dotted line. In reconstructed case, however, $P_{LL}$
+dotted line. In the case of $\delta_R$, however, $P_{LL}$
 is decomposed into orange dash-dotted correlated
@@
 damping factors $W^2(k)$ respectively for the optimal filtered 
-nonlinear and reconstructed density fields $\tilde\delta_N$
+nonlinear and recovered density fields $\tilde\delta_N$
 and $\tilde\delta_R$.
@@
 and give $\Sigma=1.8$ Mpc$/h$ and $\Sigma=12$ 
-Mpc$/h$ for reconstructed and nonlinear fields.
+Mpc$/h$ for $\delta_R$ and $\delta_N$.
 Since ${\mathcal D}(k)=W^2=r^4$, we plot
@@
 in Lagrangian coordinates are evolved/transformed to Eulerian coordinates.
-As the reconstruction is done in Lagrangian space, it recovers
+As the recovery is done in Lagrangian space, it recovers
 certain amount of correlation, as shown in the 2D PDF of the bottom panel
@@
 a much closer Gaussian distribution (blue curve of the bottom panel).
-We notice that the less-dense regions of $\delta_L$ are strongly
-correlated with $\delta_R$, whereas denser regions of $\delta_L$ are
-reconstructed less better, and $\delta_R$ shows more dispersion.
-These second uncorrelated peaks in the 2D PDF and $f(\delta_R)$
-are caused by nonlinear effects. We examine the reconstruction on
-higher redshifts, and confirm that these second uncorrelated peaks
-damp out as we go to higher redshifts.
+In denser regions of $\delta_L$, $\delta_R$ is saturated at
+$\delta_R=3$, signifying the extreme collapsing of matter \citep{2013MNRAS.428..141N}:
+$\delta_R=-\nabla\cdot\bs\Psi=\nabla\cdot\bs q=3$. Shell crossing
+makes $\delta_R$ oscillate around 3. These second uncorrelated peaks
+damp out as we go to higher redshifts. \tcr{more?}
 
@@
 simulations, and use this displacement field to study the LSS nonlinear
-clustering in Lagrangian space. The displacement information is used to reconstruct
+clustering in Lagrangian space. The displacement information is used to recover
 the primordial linear perturbations. The result shows a prominent improvement from
@@
 
-The advantage of using displacement field in reconstruction is
+\tcr{The advantage of using displacement field in reconstruction is
 its insensitive response from highly nonlinearities --
@@
 while late stage shell-crossing, nonlinear and baryonic processes
-only fine-tunes the final position $\bs x$. In contrast, traditional
+only fine-tunes the final position $\bs x$
+\citep{2014PhRvD..89h3515C}. Compared with estimated displacement fields $\tilde{\bs \Psi}(\bs q)$
+by \cite{1995ApJS..100..269P}, which do not have shell crossing,
+the additional shell-crossing information in $\bs\Psi(\bs q)$
+is uncorrelated with $\delta_L$. This insensitive response from
+nonlinearities enables the stability of reconstruction algorithms,
+which are expected to give similar results of this paper. 
+In contrast, traditional
 treatments in reconstruction deals directly on density fields which sensitively
 relies on nonlinear processes --  density values can vary by orders
-of magnitude due to nonlinear/baryonic physics and many sources of errors.
+of magnitude due to nonlinear/baryonic physics and many sources of errors.}
 

### 2016-10-23 ms_disp1/ms_disp1.tex
arxiv version

@@
 \addtolength{\textwidth}{1.05cm}
-\title{\tcr{Recovery of Primordial Density Field by Reversing the Lagrangian Space Nonlinear $E$-mode clustering}}
+\title{Lagrangian Space Nonlinear $E$-mode clustering}
 
-\author{Hao-Ran~Yu}%\email{haoran@cita.utoronto.ca}
+\author{Hao-Ran~Yu}\email{haoran@cita.utoronto.ca}
 \affiliation{Kavli Institute for Astronomy and Astrophysics, Peking University, Beijing 100871, China}
@@
 
-\author{Ue-Li~Pen}%\email{pen@cita.utoronto.ca}
+\author{Ue-Li~Pen}\email{pen@cita.utoronto.ca}
 \affiliation{Canadian Institute for Theoretical Astrophysics, University of Toronto, 60 St. George Street, Toronto, Ontario M5S 3H8, Canada}
@@
 
-\author{Hong-Ming Zhu}%email{hmzhu@bao.ac.cn}
+\author{Hong-Ming Zhu}\email{hmzhu@bao.ac.cn}
 \affiliation{Key Laboratory for Computational Astrophysics, National Astronomical Observatories, Chinese Academy of Sciences, 20A Datun Road, Beijing 100012, China}
@@
 We find that, compared to Eulerian nonlinear density fields,
-the negative divergence of $E$-mode displacement fields
-in Lagrangian space improves the cross-correlation with
+the $E$-mode displacement fields
+in Lagrangian space improves the cross-correlation scale $k$ with
 initial density field by factor of 6 $\sim$ 7,
 containing 2 orders of magnitude more primordial information.
-In reality, the displacement field is dominated by linear evolution
-and insensitive of nonlinear shell-crossing, vorticity or
-baryonic physics, and thus can be stably solved from only
-the final matter distributions.
 This illustrates ability of potential density reconstruction algorithms,
@@
 \maketitle
-\begin{figure}
- \centering
-  \includegraphics[width=0.95\linewidth]{proj.pdf}
-  \caption{Visualization of the nonlinear density field $\delta_N$ (top),
-  linear density field $\delta_L$ (middle) and the raw recovered density
-  field $\delta_R$ (bottom). These projections have 9.375 Mpc$/h$ thickness
-  and 400 Mpc$/h$ per side. The top panel shows the nonlinear displacement
-  field $\bs\Psi$ by the deformed mesh, which traces the LSS of $\delta_N$.}
-  \label{fig.projection}
-\end{figure}
+
 \section{Introduction}\label{sec.intro}
@@
 
-\tcb{Standard BAO reconstruction algorithms apply in Eulerian space.
+Standard BAO reconstruction algorithms apply in Eulerian space.
 They usually smooth the nonlinear density field on
@@
 scale bulk flows by a negative Zel'dovich linear displacement
-\citep{2007ApJ...664..675E,2009PhRvD..80l3501N,2009PhRvD..79f3523P}.}
+\citep{2007ApJ...664..675E,2009PhRvD..80l3501N,2009PhRvD..79f3523P}.
 In the linear Lagrangian perturbation theory (LPT), the 
@@
 gives the linear density field \citep{2010PhDT.........4J},
-\tcr{and in many literatures
+and in many literatures
 \cite{2013MNRAS.428..141N,2014PhRvD..89h3515C,2016JCAP...03..017B}
@@
 in order that the final positions of particles are brought
-closer to $N$-body results. Their results show that 
-all LPT methods are still unable to mimic the shell-crossing
-scales given by $N$-body simulation. Because of the absence
+closer to $N$-body results. Because of the absence
 of Lagrangian space information in observations, few density
@@
 there are techniques to estimate the displacement field from a final 
-distribution of matter.}
+distribution of matter.
 For example, when a homogeneous initial matter distribution 
@@
 information and reconstructs 1D BAO peak in correlation function.
-\tcr{In 3D and more realistic cases, one needs to
+In 3D and more realistic cases, one needs to
 carefully consider effects of curl, shell-crossing, 
-complicated baryonic physics and biased tracers (galaxies).}
+complicated baryonic physics and biased tracers (galaxies).
 
-\tcr{Before these steps, we need to quantify the amount of linear information
+\begin{figure}
+ \centering
+  \includegraphics[width=0.95\linewidth]{proj.pdf}
+  \caption{Visualization of the nonlinear density field $\delta_N$ (top),
+  linear density field $\delta_L$ (middle) and the raw recovered density
+  field $\delta_R$ (bottom). These projections have 9.375 Mpc$/h$ thickness
+  and 400 Mpc$/h$ per side. The top panel shows the nonlinear displacement
+  field $\bs\Psi$ by the deformed mesh, which traces the LSS of $\delta_N$.}
+  \label{fig.projection}
+\end{figure}
+
+Before these steps, we need to quantify the amount of linear information
 that can be recovered\footnote{To avoid ambiguity, we call
-$\delta_R=-\nabla_q\cdot\bs\Psi(\bs q)$ recovered linear density field, which
+$\delta_R\equiv-\nabla_q\cdot\bs\Psi(\bs q)$ recovered linear density field, which
 requires $\bs\Psi(\bs q)$ from $N$-body simulations, while
 $-\nabla_q\cdot\tilde{\bs \Psi}(\bs q)$ is called the
-reconstructed density field.}
+reconstructed density field, which uses the estimated displacement field.}
  from the full nonlinear displacement field $\bs\Psi(\bs q)$.
@@
 also present); (3) the setup is clean in absence of baryons and realistic
-observables. So it sets the best quality that potential reconstruction
-algorithms can achieve and further reconstructions by $\tilde{\bs \Psi}(\bs q)$
-can be compared with this result. Actually, in the later discussions and
-our next works we will show that, compared to $\tilde{\bs \Psi}(\bs q)$,
-additional information (shell-crossing) do not enhance the reconstruction
-prominently. Further more, we study the scale dependent cross-correlation
+observables. Further reconstructions by $\tilde{\bs \Psi}(\bs q)$
+can be compared with this result.
+Further more, we study the scale dependent cross-correlation
 between $-\nabla_q\cdot\bs\Psi(\bs q)$ and linear density field
 and construct optimal Wiener filters to get the optimal filtered
-recovered linear density field and recovered linear power spectrum.}
+recovered linear density field and recovered linear power spectrum.
 
@@
 Discussion and conclusion are in section \ref{sec.discussion}.
-
-\begin{figure}[t] \centering
-  \includegraphics[width=1.0\linewidth]{filter.pdf}
-  \caption{Correlation functions $r(\delta_L,\delta_N)$ and $r(\delta_L,\delta_R)$
-  (solid lines) and their scaled BAO damping models (dotted lines).}
-  \label{fig.corr}
-\end{figure}
 
@@
 
+\begin{figure}[t] \centering
+  \includegraphics[width=1.0\linewidth]{filter.pdf}
+  \caption{Correlation functions $r(\delta_L,\delta_N)$ and $r(\delta_L,\delta_R)$
+  (solid lines) and their scaled BAO damping models (dotted lines).}
+  \label{fig.corr}
+\end{figure}
+
 The two-point statistics of these density fields are quantified by the cross power 
@@
 
-\begin{figure}[t] \centering
-  \includegraphics[width=1.0\linewidth]{power.pdf}
-  \caption{\tcr{Power spectra of $\delta_L$, $\delta_N$, and the decomposition
-  of $P_{LL}$ into correlated parts and noise terms according to equation
-  (\ref{eq.power}).}}
-  \label{fig.recopower}
-\end{figure}
-
-\begin{figure}[t] \centering
-  \includegraphics[width=1.0\linewidth]{pdfs.pdf}
-  \caption{Probability distribution functions (PDFs)
-  $f(\delta_L,\delta_N)$ and $f(\delta_L,\delta_R)$, showing in the upper
-  and lower panels respectively. Both red curves on $x$-axes show $f(\delta_L)$, following
-  a Gaussian distribution. The two curves on $y$-axes in upper and lower panel
-  show respectively $f(\delta_N)$ and $f(\delta_R)$.}
-  \label{fig.pdfs}
-\end{figure}
-
 To quantify the linear information in the recovered density field, we decompose $\delta_L$ in Fourier space as
@@
 correlation coefficient $r_{LN}=r(\delta_L,\delta_N)$ and
-$r_{LR}=r(\delta_L,\delta_R)$ in Fig.\ref{fig.corr}. Clearly, $
+$r_{LR}=r(\delta_L,\delta_R)$ in Fig.\ref{fig.corr}.
+$r_{LN}$ shows no correlation starting from $k\simeq 0.4 h/$Mpc \citep{2004MNRAS.355..129S}.
+Clearly, $
 \delta_R$ contains much more linear information on smaller scales.
@@
 scales of our interest.
+
+\begin{figure}[t] \centering
+  \includegraphics[width=1.0\linewidth]{power.pdf}
+  \caption{Power spectra of $\delta_L$, $\delta_N$, and the decomposition
+  of $P_{LL}$ into correlated parts and noise terms according to equation
+  (\ref{eq.power}).}
+  \label{fig.recopower}
+\end{figure}
+
+\begin{figure}[t] \centering
+  \includegraphics[width=1.0\linewidth]{pdfs.pdf}
+  \caption{Probability distribution functions (PDFs)
+  $f(\delta_L,\delta_N)$ and $f(\delta_L,\delta_R)$, showing in the upper
+  and lower panels respectively. Both red curves on $x$-axes show $f(\delta_L)$, following
+  a Gaussian distribution. The two curves on $y$-axes in upper and lower panel
+  show respectively $f(\delta_N)$ and $f(\delta_R)$.}
+  \label{fig.pdfs}
+\end{figure}
 
@@
 makes $\delta_R$ oscillate around 3. These second uncorrelated peaks
-damp out as we go to higher redshifts. \tcr{more?}
+damp out as we go to higher redshifts.
 
@@
 
-\tcr{The advantage of using displacement field in reconstruction is
-its insensitive response from highly nonlinearities --
-densest regions form because matter elements, after shell crossing,
-stop experiencing the linear extrapolated of Zel'dovich approximation.
-i.e. the displacement fields are dominated by early stage linear processes,
+The advantage of using displacement field in reconstruction is
+its insensitive response from highly nonlinearities.
+nonlinear densities $\delta_N$ can be arbitrarily large -- 
+one expect virialized regions to be observable, where nonlinear
+density is given by the inverse determinant of the Jacobian
+$\delta_N=|\bs J|^{-1}\gtrsim 200$.
+However, reconstructed densities $\delta_R$ is given by
+the trace $\rm{tr}(\bs J)$ and saturates at 3.
+Actually, the displacement fields are dominated by early stage linear processes,
 which is the Lagrangian-Eulerian coordinate transform,
@@
 relies on nonlinear processes --  density values can vary by orders
-of magnitude due to nonlinear/baryonic physics and many sources of errors.}
+of magnitude due to nonlinear/baryonic physics and many sources of errors.
 
 \section*{Acknowledgements}
-We thank Pengjie Zhang for helpful discussions which triggered
-the inspiration of the work.
+We thank Pengjie Zhang for helpful discussions and comments.
+We acknowledge funding from NSERC.
 HRY acknowledges General Financial Grant No.2015M570884 and Special Financial Grant No.2016T90009 from the China Postdoctoal Science Foundation.
@@
 XDB09000000 and the NSFC under Grant No. 11373030.
-The Dunlap Institute is funded through an endowment established by the David Dunlap family and the University of Toronto.
 Research at the Perimeter Institute is supported by the Government of Canada

### 2016-10-23 ms_disp1/ms_disp1.tex
typo

@@
 This illustrates ability of potential density reconstruction algorithms,
-to improve the baryonic acoustic oscillation (BAO) mesurements
+to improve the baryonic acoustic oscillation (BAO) measurements
 from current and future large scale structure surveys.

### 2016-11-25 ms_disp1/ms_disp1.tex
added MM code

@@
 
-\author{Hong-Ming Zhu}\email{hmzhu@bao.ac.cn}
+\author{Hong-Ming Zhu}\email{hmzhu@nao.cas.cn}
 \affiliation{Key Laboratory for Computational Astrophysics, National Astronomical Observatories, Chinese Academy of Sciences, 20A Datun Road, Beijing 100012, China}
@@
 Initial conditions are generated at redshift $z=50$ 
-using Zel'dovich approximation, and using a CLASS transfer function.
+using Zel'dovich approximation, and using a CLASS transfer function
+\citep{2011JCAP...07..034B}.
 $N_p=512^3$ $N$-body particles are evolved via 

### 2016-11-25 MM_code/column_density.f90

@@ -1,0 +1,59 @@
+integer,parameter :: ng=128 ! number of grid per dim
+integer,parameter :: nplot=4 ! thickness of the column
+
+integer i,j,k
+
+real den(ng,ng,ng)
+real def(ng,ng,ng)
+real dspz(ng,ng,ng)
+real proj(ng,ng)
+
+! read in density (for delta, please +1)
+open(11,file='den.dat',status='old',access='stream')
+read(11) den
+close(11)
+
+! read in deformation potential
+open(11,file='def.dat',status='old',access='stream')
+read(11) def
+close(11)
+
+! get displacement field
+do k=1,ng
+do j=1,ng
+do i=1,ng
+  kp=mod(k,ng)+1
+  jp=mod(j,ng)+1
+  ip=mod(i,ng)+1
+  dspz(i,j,k)=def(i,j,kp)+def(ip,j,kp)+def(i,jp,kp)+def(ip,jp,kp)-def(i,j,k)-def(ip,j,k)-def(i,jp,k)-def(ip,jp,k)
+enddo
+enddo
+enddo
+dspz=dspz/4
+
+! get corrected column density
+k1=ng/2
+k2=k1+nplot-1
+do j=1,ng
+do i=1,ng
+  jm=mod(j-2,ng)+1
+  im=mod(i-2,ng)+1
+  k1m=mod(k1-2,ng)+1
+  kk1=k1-1+(dsp(3,im,jm,k1m)+dsp(3,i,jm,k1m)+dsp(3,im,j,k1m)+dsp(3,i,j,k1m))/4
+  kk2=k2+(dsp(3,im,jm,k2)+dsp(3,i,jm,k2)+dsp(3,im,j,k2)+dsp(3,i,j,k2))/4
+  nk1=floor(kk1)+1
+  nk2=floor(kk2)+1
+  r1=nk1-kk1
+  r2=kk2-nk2+1
+  proj(i,j)=r1*den(i,j,nk1)+sum(den(i,j,nk1+1:nk2-1))+r2*den(i,j,nk2)-sum(den(i,j,nk2:nk1))
+enddo 
+enddo
+proj=proj/nplot
+
+open(11,file='proj.dat',status='replace',access='stream')
+write(11) proj
+close(11)
+
+print*, proj(1,1), proj(ng,ng)
+
+end

### 2016-11-25 MM_code/plotgridden.m

@@ -1,0 +1,89 @@
+SetDefault
+clear; clc; close all
+ng=128;
+nplot=4;
+
+fid=fopen('0.000den128_2.bin');
+den=reshape(fread(fid,'real*4'),ng,ng,ng);
+fclose(fid);
+
+fid=fopen('0.000def_0128_2.bin');
+def=reshape(fread(fid,'real*4'),ng,ng,ng);
+fclose(fid);
+
+% differential
+dsp=zeros(3,ng,ng,ng);
+for k=1:ng
+for j=1:ng
+for i=1:ng
+  kp=mod(k,ng)+1;
+  jp=mod(j,ng)+1;
+  ip=mod(i,ng)+1;
+  dsp(1,i,j,k)=def(ip,j,k)+def(ip,jp,k)+def(ip,j,kp)+def(ip,jp,kp)-def(i,j,k)-def(i,jp,k)-def(i,j,kp)-def(i,jp,kp);
+  dsp(2,i,j,k)=def(i,jp,k)+def(ip,jp,k)+def(i,jp,kp)+def(ip,jp,kp)-def(i,j,k)-def(ip,j,k)-def(i,j,kp)-def(ip,j,kp);
+  dsp(3,i,j,k)=def(i,j,kp)+def(ip,j,kp)+def(i,jp,kp)+def(ip,jp,kp)-def(i,j,k)-def(ip,j,k)-def(i,jp,k)-def(ip,jp,k);
+end
+end
+end
+dsp=dsp/4;
+
+% get corrected den
+k1=ng/2;
+k2=k1+nplot-1;
+den_corr=zeros(ng,ng);
+for j=1:ng
+for i=1:ng
+  jm=mod(j-2,ng)+1;
+  im=mod(i-2,ng)+1;
+  k1m=mod(k1-2,ng)+1;
+  %k2m=mod(k2-2,ng)+1;
+  kk1=k1-1+(dsp(3,im,jm,k1m)+dsp(3,i,jm,k1m)+dsp(3,im,j,k1m)+dsp(3,i,j,k1m))/4;
+  kk2=k2+(dsp(3,im,jm,k2)+dsp(3,i,jm,k2)+dsp(3,im,j,k2)+dsp(3,i,j,k2))/4;
+  nk1=floor(kk1)+1;
+  nk2=floor(kk2)+1;
+  r1=nk1-kk1;
+  r2=kk2-nk2+1;
[…]

### 2016-11-25 MM_code/column_density.f90
debug column_density. now working

@@ -1,7 +1,10 @@
+implicit none
+
 integer,parameter :: ng=128 ! number of grid per dim
 integer,parameter :: nplot=4 ! thickness of the column
 
-integer i,j,k
+integer i,j,k,ip,jp,kp,k1,k2,im,jm,k1m,nk1,nk2
+real kk1,kk2,r1,r2
 
 real den(ng,ng,ng)
 real def(ng,ng,ng)
@@ -22,25 +25,26 @@ close(11)
 do k=1,ng
 do j=1,ng
 do i=1,ng
-  kp=mod(k,ng)+1
-  jp=mod(j,ng)+1
-  ip=mod(i,ng)+1
+  kp=modulo(k,ng)+1
+  jp=modulo(j,ng)+1
+  ip=modulo(i,ng)+1
   dspz(i,j,k)=def(i,j,kp)+def(ip,j,kp)+def(i,jp,kp)+def(ip,jp,kp)-def(i,j,k)-def(ip,j,k)-def(i,jp,k)-def(ip,jp,k)
 enddo
 enddo
 enddo
 dspz=dspz/4
 
+
 ! get corrected column density
 k1=ng/2
 k2=k1+nplot-1
 do j=1,ng
 do i=1,ng
-  jm=mod(j-2,ng)+1
-  im=mod(i-2,ng)+1
-  k1m=mod(k1-2,ng)+1
-  kk1=k1-1+(dsp(3,im,jm,k1m)+dsp(3,i,jm,k1m)+dsp(3,im,j,k1m)+dsp(3,i,j,k1m))/4
-  kk2=k2+(dsp(3,im,jm,k2)+dsp(3,i,jm,k2)+dsp(3,im,j,k2)+dsp(3,i,j,k2))/4
+  jm=modulo(j-2,ng)+1
+  im=modulo(i-2,ng)+1
+  k1m=modulo(k1-2,ng)+1
+  kk1=k1-1+(dspz(im,jm,k1m)+dspz(i,jm,k1m)+dspz(im,j,k1m)+dspz(i,j,k1m))/4
+  kk2=k2+(dspz(im,jm,k2)+dspz(i,jm,k2)+dspz(im,j,k2)+dspz(i,j,k2))/4
   nk1=floor(kk1)+1
   nk2=floor(kk2)+1
   r1=nk1-kk1
@@ -54,6 +58,5 @@ open(11,file='proj.dat',status='replace',access='stream')
 write(11) proj
[…]

### 2016-12-18 ms_disp1/ms_disp1.tex
change text according to the referee

@@
 \addtolength{\textwidth}{1.05cm}
-\title{Lagrangian Space Nonlinear $E$-mode clustering}
+\title{Nonlinear ${\bs E}$-mode clustering in Lagrangian Space}
 
@@
 \citep{1995ApJS..100..269P,1998ApJS..115...19P}.
-In 1-dimensional (1D) case, the exact solution 
-simplifies to an ordering of matter elements by Eulerian coordinates.
+In 1-dimensional (1D) case, this solution 
+simplifies to an ordering of mass elements by their final, Eulerian coordinates.
 Zhu {\it et al.} \cite{2016arXiv160907041Z} apply this algorithm to the result
@@
 from the full nonlinear displacement field $\bs\Psi(\bs q)$.
+This field can be decomposed into a curl-less ``$E$-mode'' component
+and a divergence-less ``$B$-mode'' component, $\bs\Psi=\bs\Psi_E+\bs\Psi_B$,
+where the $B$-mode is raised by late stage non-Gaussianities and is not
+dominant in $\bs\Psi$ \citep{2014PhRvD..89h3515C}.
 In $\bs\Psi(\bs q)$ from $N$-body simulations,
-(1) shell-crossing effects are fully considered; (2) $E$-mode displacement
-field is used for recovering the linear information (although $B$-mode is
-also present); (3) the setup is clean in absence of baryons and realistic
+(1) $E$-mode displacement field is used for recovering the linear information
+(although $B$-mode is also present); (2) shell-crossing effects are fully considered;
+(3) the setup is clean in absence of baryons and realistic
 observables. Further reconstructions by $\tilde{\bs \Psi}(\bs q)$
@@
 Because $\delta_N$ is highly nonlinear and follows an approximate
-log-normal distribution, we plot $\log_{10}(\delta_N+1)$ instead.
+log-normal distribution, we plot $\log_{10}(\delta_N+1)$ instead, and show
+the color scale $\log_{10}(\delta_N+1)\in[-1,1]$ (or $\rho_N/\langle\rho_N\rangle\in[0.1,10]$)
+only, for a better visualization.
 The nonlinear evolution of $\delta_N$ makes it very different from $\delta_L$
@@
 
-To quantify the linear information in the recovered density field, we decompose $\delta_L$ in Fourier space as
+To quantify the linear information in the recovered density field $\delta_R$,
+we decompose $\delta_L$ in Fourier space into two uncorrelated parts,
 \begin{equation}\label{eq.decompose}
@@
 \end{equation}
-where $r'\delta_R$ is completely correlated with linear density $\delta_L$. Correlating equation (\ref{eq.decompose}) with $\delta_R$ gives
+where the first term $r'\delta_R$ is completely correlated with $\delta_R$,
+meaning the linear information we can extract from $\delta_R$.
+The noise term $\delta_n$ is uncorrelated with $\delta_R$, being
+the rest of linear information which is not contained in $\delta_R$.
+Correlating equation (\ref{eq.decompose}) with $\delta_R$ gives
 \begin{equation}
@@
 
-According to equation (\ref{eq.decompose}), the auto power spectrum is decomposed as
+According to equation (\ref{eq.decompose}), the auto power spectrum of $\delta_L$ is decomposed as
 \begin{equation}\label{eq.power}
@@
 \end{equation}
-and $P_{nn}=(1-r^2)P_{LL}$. Then we construct a Wiener filter to filter out the uncorrelated part in $\delta_R$:
+and $P_{nn}=(1-r^2)P_{LL}$. We also explicitly
+compute the cross power spectrum between $\delta_R$ and $\delta_n=\delta_L-rb^{-1}\delta_R$,
+and found that $rb^{-1}P_{nR}$ is about two orders of magnitude lower than $P_{LL}$,
+being consistent with zero. This confirms that
+the signal term $rb^{-1}\delta_R$ and the noise term $\delta_n$
+is indeed uncorrelated and validates equation (\ref{eq.power}).
+According to these two terms, we construct a Wiener filter to filter
+out the uncorrelated part in $\delta_R$:
 \begin{equation}
@@
 its insensitive response from highly nonlinearities.
-nonlinear densities $\delta_N$ can be arbitrarily large -- 
+Nonlinear densities $\delta_N$ can be arbitrarily large -- 
 one expect virialized regions to be observable, where nonlinear
@@
 \section*{Acknowledgements}
-We thank Pengjie Zhang for helpful discussions and comments.
+We thank Pengjie Zhang and Kwan Chuen Chan for helpful discussions and comments.
 We acknowledge funding from NSERC.

### 2017-01-10 ms_disp1/ms_disp1.tex
Jan 10

@@
 reconstructed density field, which uses the estimated displacement field.}
-from the full nonlinear displacement field $\bs\Psi(\bs q)$.
-This field can be decomposed into a curl-less ``$E$-mode'' component
+from the full nonlinear displacement field $\bs\Psi(\bs q)$,
+\tcb{which is the limit of all possible Lagrangian reconstruction
+formulations.}
+The displacement field can be decomposed into a curl-less ``$E$-mode'' component
 and a divergence-less ``$B$-mode'' component, $\bs\Psi=\bs\Psi_E+\bs\Psi_B$,
@@
 Eulerian coordinates. This result illustrates the feasibility
-of using estimated displacement field $\tilde{\bs \Psi}(\bs q)$ to reconstruct primordial linear
-density field. A straightforward example of a estimation of $\tilde{\bs \Psi}(\bs q)$
+of using estimated displacement field $\tilde{\bs \Psi}(\bs q)$
+to reconstruct primordial linear density field,
+\tcb{and this Lagrangian $E$-mode recovery
+ quantifies the upper limit of Lagrangian reconstruction methods.}
+A straightforward example of a estimation of $\tilde{\bs \Psi}(\bs q)$
 is given by \cite{1995ApJS..100..269P,1998ApJS..115...19P}.

### 2017-01-10 CUBE/fftw3_example/compile_intel.sh

@@ -4,4 +4,4 @@ module list
 
 rm -f *.o *.out
 
-ifort -O3 -xHost -fpp -mcmodel=medium test.f90 -lfftw3f -I -L -lm -ldl
+ifort -i8 -O3 -xHost -fpp -mcmodel=medium test.f90 -lfftw3f -I -L -lm -ldl

### 2017-01-10 CUBE/fftw3_example/test.f90

@@ -1,14 +1,30 @@
 implicit none
 include 'fftw3.f'
-integer,parameter :: ng=512 ! number of grid per dimension
+integer,parameter :: ng=2048 ! number of grid per dimension
 integer(8) plan_fft,plan_ifft ! fft plans
 real den(ng+2,ng,ng) ! fft array, additional "+2" for storing Nyquist frequency
+
+integer i
+integer(4) i4
+
+print*, 'ilp64 test: default integer'
+i=2147483647
+i=i+1
+print*,i
+
+
+print*, 'call forward plan'
 call sfftw_plan_dft_r2c_3d(plan_fft,ng,ng,ng,den,den,FFTW_MEASURE)
+print*, 'call backward plan'
 call sfftw_plan_dft_c2r_3d(plan_ifft,ng,ng,ng,den,den,FFTW_MEASURE)
+print*, 'initialize den'
 den(:ng,:,:)=1.23
+print*, 'call forward fft'
 call sfftw_execute(plan_fft) ! forward transform
+print*, 'call backward plan'
 call sfftw_execute(plan_ifft) ! backward transform
 den=den/ng/ng/ng ! normalize
+print*, 'destroy'
 call sfftw_destroy_plan(plan_fft)
 call sfftw_destroy_plan(plan_ifft)
 end

### 2017-01-10 CUBE/penfft_example/test.f90

@@ -1,0 +1,293 @@
+implicit none
+
+integer(8) planx,plany,planz,iplanx,iplany,iplanz
+integer,parameter :: NULL=0
+
+integer,parameter :: nn=1
+integer,parameter :: nc=512 ! nc/image/dim, in physical volume, >=24
+integer,parameter :: npen=nc/nn ! nc /dim in shorter side of the pencil, for pencil decomposition
+
+!! MPI images !!
+integer,parameter :: rank=0                         ! MPI_rank
+integer,parameter :: icz=rank/(nn**2)+1             ! image_z
+integer,parameter :: icy=(rank-nn**2*(icz-1))/nn+1  ! image_y
+integer,parameter :: icx=mod(rank,nn)+1             ! image_x
+
+real r3(nc,nc,nc)[*]
+real rxlong(nc*nn+2,nc,npen)[*]
+complex cx(nc*nn/2+1,nc,npen)[*]
+real crho_c(nc*nn+2,nc,npen)
+complex cy(npen,nn,nn,nc/2+1,npen)[*]
+complex cz(npen,nn,nn,nc/2+1,npen)[*]
+complex cyxz(npen*nn,npen*nn/2+1,npen)
+complex cyyxz(npen,nn,npen*nn/2+1,npen)
+equivalence(cyxz,cyyxz)
+
+print*, 'call fft plan'
+call create_penfft_plan
+print*, 'initialize den'
+r3=1.23
+print*, 'call ftran'
+call fft_cube2pencil
+print*, 'transpose to xyz'
+!call trans_zxy2xyz
+print*, 'transpose to zxy'
+!call trans_xyz2zxy
+print*, 'call btran'
+r3=0
+!call ifft_pencil2cube
+
+print*, r3(1:4,1,1)
+
+call destroy_penfft_plan
+
+
+contains
+
+subroutine fft_cube2pencil
+!use variables
+implicit none
+save
+integer m,m1,m2,m3,i0,i1,i2
+
+m1=icx
+m2=icy
+m3=icz
+
+m=num_images()
+
+do i0=1,nn ! cube->x
[…]

### 2017-01-10 CUBE/penfft_example/variables.f90

@@ -1,0 +1,105 @@
+!#define zipconvert
+module variables
+use parameters
+implicit none
+save
+
+! parameters
+integer,parameter :: npnode=nf**3
+real,parameter :: density_buffer=1.1
+integer,parameter :: npmax=npnode*(nte*1./nt)**3*density_buffer
+integer,parameter ::  nseedmax=200
+real,parameter :: vbuf=0.9
+real,parameter :: dt_max=1
+real,parameter :: dt_scale=1
+integer(8),parameter :: unit8=1
+integer,parameter :: NULL=0
+real,parameter :: pi=3.141592654
+real,parameter :: GG=1.0/6.0/pi
+
+! checkpoint variables
+integer,parameter :: nmax_redshift=100
+integer cur_checkpoint[*], n_checkpoint[*]
+real z_checkpoint(nmax_redshift)[*]
+logical checkpoint_step[*], final_step[*]
+
+! variables
+integer its[*]
+real dt[*],dt_old[*],dt_mid[*]
+real dt_fine(nn**3),dt_pp(nn**3),dt_coarse(nn**3),dt_vmax(nn**3)
+real a[*],da[*],a_mid[*],tau[*],t[*] ! time step
+real f2_max_fine(nnt,nnt,nnt)[*],f2_max_pp(nnt,nnt,nnt)[*],f2_max_coarse[*]
+
+integer iseed(nseedmax), iseedsize
+integer itx,ity,itz,ix,iy,iz,i_dim
+integer i,j,k,l,ip,ipp,pp
+!integer nplocal[*], nptile(nnt,nnt,nnt)
+!integer(8) nptotal, npcheck
+
+!real mass_p
+
+! FFT plans
+integer(8) plan_fft_fine,plan_ifft_fine
+integer(8) planx,plany,planz,iplanx,iplany,iplanz
+
+real v_i2r(3)[*],v_i2r_new(3)[*]
+real vmax(3)[*],vmax_new(3)[*]
+! n^3
+!#ifdef zipconvert
+!  integer(1) xic_new(3,npmax)
+!  integer(2) vic_new(3,npmax)
+!#endif
[…]

### 2017-01-10 CUBE/init/rhof.f90
xp

@@ -1,3 +1,5 @@
+#define write_xp
+
 program rhof
 use parameters
 implicit none
@@ -7,7 +9,7 @@ integer i,j,k,l
 integer nplocal
 integer,parameter :: ng=nf/8
 integer,parameter :: npnode=nf**3
-real,parameter :: density_buffer=1.5
+real,parameter :: density_buffer=1.2
 integer,parameter :: npmax=npnode*density_buffer
 integer ind,dx,dxy,kg,mg,jg,ig,i0,j0,k0,itx,ity,itz,idx,imove,nf_shake,ibin
 integer nshift,ifrom,ileft,iright,nlen,nlast,g(3)
@@ -16,7 +18,7 @@ real kr,kx(3), sincx,sincy,sincz,sinc, rbin
 integer rhoc(nt,nt,nt,nnt,nnt,nnt)
 real rho_f(0:ng+1,0:ng+1,0:ng+1)[*]
 integer idx1(3),idx2(3),ip,np
-real mass_p,dx1(3),dx2(3),tempx(3)
+real mass_p,dx1(3),dx2(3),xp(3)
 
 integer(izipx) x(3,npmax)[*]
 
@@ -52,12 +54,12 @@ sync all
 do cur_checkpoint=n_checkpoint,n_checkpoint
   fn0='.'//opath//'/node'//image2str(this_image()-1)//'/'//z2str(z_checkpoint(cur_checkpoint)) &
   //'zip0_'//image2str(this_image()-1)//'.dat'
-  fn1='.'//opath//'/node'//image2str(this_image()-1)//'/'//z2str(z_checkpoint(cur_checkpoint)) &
-  //'zip1_'//image2str(this_image()-1)//'.dat'
+  !fn1='.'//opath//'/node'//image2str(this_image()-1)//'/'//z2str(z_checkpoint(cur_checkpoint)) &
+  !//'zip1_'//image2str(this_image()-1)//'.dat'
   fn2='.'//opath//'/node'//image2str(this_image()-1)//'/'//z2str(z_checkpoint(cur_checkpoint)) &
   //'zip2_'//image2str(this_image()-1)//'.dat'
-  fn3='.'//opath//'/node'//image2str(this_image()-1)//'/'//z2str(z_checkpoint(cur_checkpoint)) &
-  //'zip3_'//image2str(this_image()-1)//'.dat'
[…]

### 2017-01-10 CUBE/init/voronoi.f90

@@ -1,187 +1,120 @@
-use parameters
-use penfft_config
+! read zip format checkpoint, and write xv files
+! generate Voronoi density fields
+program voronoi
 implicit none
+integer,parameter :: np=223479 ! total number of particles
+integer,parameter :: nf=4608 ! number of fine cells per dim
+integer,parameter :: ng=128 ! number of grids per dim in output resolution
 
-integer,parameter :: nmax_redshift=100
-
-integer nplocal
-integer,parameter :: npnode=nf**3 ! only true for this project
-real,parameter :: density_buffer=1.2
-integer,parameter :: npmax=npnode*density_buffer
-integer i,j,k,l,np,nlast,ind,ip,dx,dxy,kg,mg,jg,ig,i0,j0,k0,pp,itx,ity,itz,idx,imove,nf_shake,ibin
+integer i,j,k,ip,ig,pp
 integer i_s,j_s,k_s,ii_s,jj_s,kk_s,r
-integer wp(npmax),pp_s
+integer wp(np),pp_s
+integer hoc_g(ng,ng,ng),ll_p(np) ! linked list
+integer hoc_p(np),ll_g(ng**3) ! inverse linked list
+real xp(3,np),den(ng,ng,ng),den_ngp(ng,ng,ng)
+real r2,r2min,gpos(3),hpos(3),dpos(3)
 
-integer cur_checkpoint, n_checkpoint
-integer rhoc(nt,nt,nt,nnt,nnt,nnt)
-
-integer(izipx) x(3,npmax)
-integer(2)   pid(4,npmax)
-real xp(3,npmax),pos1(3),gpos(3),dpos(3),r2,r2min
-real den(ng,ng,ng)
-
-integer hoc_g(ng,ng,ng),ll_p(npmax) ! linked list
-integer hoc_p(npmax),ll_g(ng**3) ! inverse linked list
-
-real z_checkpoint(nmax_redshift)
-character (200) :: fn0,fn2
 logical search
 
+! read in particle positions
+open(11,file='xhalo.dat',status='old',access='stream')
+read(11) xp
+close(11)
 
-print*, 'Voronoi density field interpolation'
-print*, 'ng=',ng
-print*, 'npmax=',npmax
-do cur_checkpoint= n_checkpoint,n_checkpoint
[…]

### 2017-02-17 ms_disp1/ms_disp1.tex
published version

--- 2016-12-18 ms_disp1/ms_disp1.tex
@@
 \addtolength{\textwidth}{1.05cm}
-\title{Nonlinear ${\bs E}$-mode clustering in Lagrangian Space}
+\title{Nonlinear ${\bs E}$-mode clustering in Lagrangian space}
 
@@
 We study the nonlinear $E$-mode clustering in Lagrangian space
-by using large scale structure (LSS) $N$-body simulations
+by using large scale structure $N$-body simulations
 and use the displacement field information in Lagrangian space
@@
 in Lagrangian space improves the cross-correlation scale $k$ with
-initial density field by factor of 6 $\sim$ 7,
-containing 2 orders of magnitude more primordial information.
+initial density field by a factor of 6-7,
+containing two orders of magnitude more primordial information.
 This illustrates ability of potential density reconstruction algorithms,
-to improve the baryonic acoustic oscillation (BAO) measurements
+to improve the baryonic acoustic oscillation measurements
 from current and future large scale structure surveys.
@@
 \section{Introduction}\label{sec.intro}
-Our universe starts from primordial Gaussian perturbations at a very early stage, 
+Our Universe starts from primordial Gaussian perturbations at a very early stage, 
 and from those fluctuations, the gravitational instability drives the formation of 
@@
 In the linear Lagrangian perturbation theory (LPT), the 
-negative divergence of the the displacement field
-$-\nabla_q\cdot\bs\Psi(\bs q)$ respect to Lagrangian coordinates $\bs q$
+negative divergence of the displacement field
+$-\nabla_q\cdot\bs\Psi(\bs q)$ with respect to Lagrangian coordinates $\bs q$
 gives the linear density field \citep{2010PhDT.........4J},
@@
 is assumed, there is a unique solution of curl-less displacement field to relate 
-the initial and final distributions without shell-crossing. This solution can be 
+the initial and final distributions without shell crossing. This solution can be 
 solved by a metric transformation equation
 \citep{1995ApJS..100..269P,1998ApJS..115...19P}.
-In 1-dimensional (1D) case, this solution 
+In the one-dimensional (1D) case, this solution 
 simplifies to an ordering of mass elements by their final, Eulerian coordinates.
 Zhu {\it et al.} \cite{2016arXiv160907041Z} apply this algorithm to the result
-of an 1D simulation \citep{2016JCAP...01..043M} and obtain an estimated
+of a 1D simulation \citep{2016JCAP...01..043M} and obtain an estimated
 displacement field $\tilde\Psi(q)$, and find
 that this new method well recovers the linear
-information and reconstructs 1D BAO peak in correlation function.
+information and reconstructs 1D BAO peak in the correlation function.
 In 3D and more realistic cases, one needs to
-carefully consider effects of curl, shell-crossing, 
+carefully consider effects of curl, shell crossing, 
 complicated baryonic physics and biased tracers (galaxies).
@@
 reconstructed density field, which uses the estimated displacement field.}
-from the full nonlinear displacement field $\bs\Psi(\bs q)$.
+from the full nonlinear displacement field $\bs\Psi(\bs q)$,
 This field can be decomposed into a curl-less ``$E$-mode'' component
-and a divergence-less ``$B$-mode'' component, $\bs\Psi=\bs\Psi_E+\bs\Psi_B$,
+and a divergence less ``$B$-mode'' component, $\bs\Psi=\bs\Psi_E+\bs\Psi_B$,
 where the $B$-mode is raised by late stage non-Gaussianities and is not
@@
 (1) $E$-mode displacement field is used for recovering the linear information
-(although $B$-mode is also present); (2) shell-crossing effects are fully considered;
+(although $B$-mode is also present); (2) shell crossing effects are fully considered;
 (3) the setup is clean in absence of baryons and realistic
@@
 can be compared with this result.
-Further more, we study the scale dependent cross-correlation
+Furthermore, we study the scale dependent cross correlation
 between $-\nabla_q\cdot\bs\Psi(\bs q)$ and linear density field
@@
 
-In the rest part of the paper, we describe the simulation,
+In the rest of the paper, we describe the simulation,
 density recovery algorithm and Wiener filter setups in section
@@
 \cite{1988csup.book.....H} on a two-level mesh grids
-(details see \cite{2013MNRAS.436..540H}) and cloud-in-cell
+(for details see \cite{2013MNRAS.436..540H}) and cloud-in-cell
 (CIC) is used in particle interpolations in force 
@@
 
-\subsection{Density Recovery}\label{ss.reco}
+\subsection{Density recovery}\label{ss.reco}
 In the simulation, we use particle-ID (PID) to record the initial (Lagrangian) location ${\bs 
@@
 and physical amplitude of $\bs\Psi$. In the top panel of Fig.\ref{fig.projection},
-The resulting mesh illustrates a pseudo
-``curvilinear coordinate'' similar to \cite{1995ApJS..100..269P},
-however the mesh can be overlapped due to shell-crossing. The densest mesh grids
+The resulting mesh illustrates a ``pseudocurvilinear coordinate'' similar to \cite{1995ApJS..100..269P},
+however the mesh can be overlapped due to shell crossing. The densest mesh grids
 trace the densest structures of $\delta_N$, whereas the undeformed grid positions
@@
   $f(\delta_L,\delta_N)$ and $f(\delta_L,\delta_R)$, showing in the upper
-  and lower panels respectively. Both red curves on $x$-axes show $f(\delta_L)$, following
-  a Gaussian distribution. The two curves on $y$-axes in upper and lower panel
+  and lower panels respectively. Both red curves on the $x$-axes show $f(\delta_L)$, following
+  a Gaussian distribution. The two curves on the $y$-axes in the upper and the lower panel
   show respectively $f(\delta_N)$ and $f(\delta_R)$.}
@@
 dotted line. In the case of $\delta_R$, however, $P_{LL}$
-is decomposed into orange dash-dotted correlated
-part and purple dotted uncorrelated part according to
+is decomposed into the orange dash-dotted correlated
+part and the purple dotted uncorrelated part according to
 equation (\ref{eq.power}). The correlated power spectrum
@@
 use the $W(k)=r^2_{LR}$ as the Wiener filter.
-The grey-scaled plots in the center of both panels show the 2-variable PDFs,
+The grey-scaled plots in the center of both panels show the two-variable PDFs,
 whereas their projections onto each variable are just
-1-variable PDFs -- $f(\delta_L)$, $f(\delta_N)$ and $f(\delta_R)$, shown
+one-variable PDFs -- $f(\delta_L)$, $f(\delta_N)$ and $f(\delta_R)$, shown
 as red/blue curves on the axes of Fig.\ref{fig.pdfs}.
@@
 Eulerian coordinates. This result illustrates the feasibility
-of using estimated displacement field $\tilde{\bs \Psi}(\bs q)$ to reconstruct primordial linear
-density field. A straightforward example of a estimation of $\tilde{\bs \Psi}(\bs q)$
+of using estimated displacement field $\tilde{\bs \Psi}(\bs q)$
+to reconstruct primordial linear density field.
+A straightforward example of a estimation of $\tilde{\bs \Psi}(\bs q)$
 is given by \cite{1995ApJS..100..269P,1998ApJS..115...19P}.
-%\tcr{More complicated reconstruction algorithms, e.g., least action
-%algorithms \cite{1989ApJ...344L..53P,2000ApJ...544...21G} can give
-%the exact $\delta_L$}, where
-%shell-crossing effects are also considered.
 In reality, one needs
-to consider all aspects including vorticity, shell-crossing, bias, noise
+to consider all aspects including vorticity, shell crossing, bias, noise
 and data complexities. The impact of these factors can be quantitatively
@@
 Nonlinear densities $\delta_N$ can be arbitrarily large -- 
-one expect virialized regions to be observable, where nonlinear
+one expects virialized regions to be observable, where nonlinear
 density is given by the inverse determinant of the Jacobian
 $\delta_N=|\bs J|^{-1}\gtrsim 200$.
-However, reconstructed densities $\delta_R$ is given by
+However, reconstructed densities $\delta_R$ are given by
 the trace $\rm{tr}(\bs J)$ and saturates at 3.
@@
 which is the Lagrangian-Eulerian coordinate transform,
-while late stage shell-crossing, nonlinear and baryonic processes
-only fine-tunes the final position $\bs x$
+while late stage shell crossing, nonlinear and baryonic processes
+only fine-tune the final position $\bs x$
 \citep{2014PhRvD..89h3515C}. Compared with estimated displacement fields $\tilde{\bs \Psi}(\bs q)$
 by \cite{1995ApJS..100..269P}, which do not have shell crossing,
-the additional shell-crossing information in $\bs\Psi(\bs q)$
+the additional shell crossing information in $\bs\Psi(\bs q)$
 is uncorrelated with $\delta_L$. This insensitive response from
@@
 
-\section*{Acknowledgements}
+\section*{ACKNOWLEDGMENTS}
 We thank Pengjie Zhang and Kwan Chuen Chan for helpful discussions and comments.
 We acknowledge funding from NSERC.
-HRY acknowledges General Financial Grant No.2015M570884 and Special Financial Grant No.2016T90009 from the China Postdoctoal Science Foundation.
-HMZ acknowledges the support of the Chinese MoST 863 program under Grant
+H.R.Y. acknowledges General Financial Grant No. 2015M570884 and Special Financial Grant No. 2016T90009 from the China PostdoctRoal Science Foundation.
+H.M.Z. acknowledges the support of the Chinese MoST 863 program under Grant
 No. 2012AA121701, the CAS Science Strategic Priority Research Program

### 2017-05-30 ms_caf/ms_caf.tex
haoran
add figure
cicpower (cross with x4v4)

@@
 
-\shortauthors{Author et al.} \shorttitle{New CubeP3M}
+\shortauthors{Yu et al.} \shorttitle{CUBE}
 \begin{document}
-\title{Compact Particle-Mesh $N$-body Algorithm}
-\author{Authors}
+\title{CUBE: Compact Particle-Mesh $N$-body Algorithm}
+\author{
+Hao-Ran Yu$^{1,2}$,
+Ue-Li Pen$^{2,3,4,5}$,
+Xin Wang$^{2}$
+}
 
-\affil{Affiliations}
+\affil{
+$^1$Kavli Institute for Astronomy and Astrophysics, Peking University, Beijing 100871, China\\
+$^2$Canadian Institute for Theoretical Astrophysics, University of Toronto, Toronto, ON M5H 3H8 Canada\\
+$^3$Dunlap Institute for Astronomy and Astrophysics, University of Toronto, Toronto, ON M5S 3H4, Canada\\
+$^4$Canadian Institute for Advanced Research, CIFAR Program in Gravitation and Cosmology, Toronto, ON M5G 1Z8, Canada\\
+$^5$Perimeter Institute for Theoretical Physics, Waterloo, ON N2L 2Y5, Canada
+}
 
@@
 Here we present a parallel, memory-lite Particle-Mesh based $N$-body code, in which each particle can occupy as low as 6-byte memory.
-This is accomplished by storing only particles' relative location respect to a meshed grid, and their velocity relative to a predictable distribution function.
-The remaining information is given by the ordering of particles in memory space, and a complimentary density field, which is negligible in memory.
+This is accomplished by storing only particles' relative location respect to a mesh-grid, and their velocity relative to a predictable distribution function.
+The remaining information is given by the ordering of particles, and a complimentary density field, which is negligible in memory space.
 Our numerical experiments show that this algorithm provides acceptable accuracy compared to traditional algorithm in cosmological simulations.
@@
 
+\begin{figure}[t]
+\centering
+  \includegraphics[width=0.95\linewidth]{ms_caf.pdf}
+\end{figure}
 \section{Introduction}
@@
 \begin{figure}[t]
-\includegraphics[width=1.0\linewidth]{fig_tile.pdf}
+\centering
+  \includegraphics[width=0.95\linewidth]{fig_tile.pdf}
  \caption{Spacial decomposition.}
@@
 
+\begin{figure}[t]
+\centering
+  \includegraphics[width=0.95\linewidth]{ddsp_pcdf_256_box80.pdf}
+ \caption{Displacement.}
+\label{fig.dsp}
+\end{figure}
+
 \subsection{Memory layout}
@@
 \bibliographystyle{hapj}
-\bibliography{../haoran_ref}
+\bibliography{haoran_ref}
 

### 2017-05-30 CUBE/main/parameters.f90

@@ -6,8 +6,8 @@ module parameters
   character(*),parameter :: opath='../output/universe5/'
 
   ! simulation parameters
-  integer(8),parameter :: izipx=2 ! 1 or 2, for particle location
-  integer(8),parameter :: izipv=2 ! 1 or 2, for particle velocity
+  integer(8),parameter :: izipx=1 ! 1 or 2, for particle location
+  integer(8),parameter :: izipv=1 ! 1 or 2, for particle velocity
   integer(8),parameter :: nvbin=int(2,8)**(8*izipv)
   integer(8),parameter :: ishift=-(int(2,8)**(izipx*8-1))
   real(8),parameter :: rshift=0.5-ishift
@@ -59,7 +59,7 @@ module parameters
   real,parameter :: z_i_nu=z_i ! initial redshift for neutrinos
   real,parameter :: a_i=1/(1+z_i) ! initial scale factor
 
-  real,parameter :: box=200.0  ! simulation scale /dim, in unit of Mpc/h
+  real,parameter :: box=250.0  ! simulation scale /dim, in unit of Mpc/h
   real,parameter :: h0=67.74    ! Hubble constant
   real,parameter :: s8=0.8276   ! \sigma_8
   real,parameter :: ratio_nudm_dim=2 ! ratio of number of particles for neutrino/CDM, /dim
@@ -85,7 +85,7 @@ module parameters
   real,parameter :: power_index=2
   real,parameter :: wde=-1
 
-  real,parameter :: f_nl=0.004
+  real,parameter :: f_nl=0
   real,parameter :: g_nl=0
   real,parameter :: n_s=0.96
   real,parameter :: scalar_amp=2.142e-9

### 2017-05-30 CUBE/main/run_mac.sh

@@ -5,6 +5,8 @@ source ../utilities/module_load_mac.sh
 cd ../utilities/
 make clean
 make
+
+# Main N-body code
 cd ../main/
 make clean
 make
@@ -15,5 +17,8 @@ cd ../main/
 ./cafcube.x
 cd ../utilities/
 ./dsp.x
+cd ../utilities/
+./cicpower.x
+
 
 cd ../main/

### 2017-05-30 CUBE/utilities/cicpower.f90

@@ -1,0 +1,147 @@
+program cicpower
+  use parameters
+  use pencil_fft
+  use powerspectrum
+  implicit none
+  save
+  ! nc: coarse grid per node per dim
+  ! nf: fine grid per node per dim
+  integer(8),parameter :: npnode=nf**3 ! only true for this project
+  real,parameter :: density_buffer=1.2
+  integer(8),parameter :: npmax=npnode*density_buffer
+  integer(8) i,j,k,l,i_dim,iq(3),nplocal,itx,ity,itz,nlast,ip,np,idx1(3),idx2(3)
+  integer(4) rhoc(nt,nt,nt,nnt,nnt,nnt)
+  real(4) rho_grid(0:ng+1,0:ng+1,0:ng+1)[*],rho0(ng,ng,ng),rho1(ng,ng,ng)
+  real(4) mass_p,pos1(3),dx1(3),dx2(3)
+  integer(izipx) x(3,npmax)
+
+  real xi(10,nbin)[*]
+
+  call geometry
+
+  if (head) then
+    print*, 'cicpower on resolution:'
+    print*, 'ng=',ng
+  endif
+  sync all
+
+  if (head) then
+    print*, 'checkpoint at:'
+    open(16,file='../main/redshifts.txt',status='old')
+    do i=1,nmax_redshift
+      read(16,end=71,fmt='(f8.4)') z_checkpoint(i)
+      print*, z_checkpoint(i)
+    enddo
+    71 n_checkpoint=i-1
+    close(16)
+    print*,''
+  endif
+
+  sync all
+  n_checkpoint=n_checkpoint[1]
+  z_checkpoint(:)=z_checkpoint(:)[1]
+  sync all
+
+  call create_penfft_plan
+  
+  do cur_checkpoint= 1,n_checkpoint
+    if (head) print*, 'Start analyzing redshift ',z2str(z_checkpoint(cur_checkpoint))
+    open(12,file=output_name('zip2'),status='old',action='read',access='stream')
+    read(12) sim
+    ! check zip format and read rhoc
+    if (sim%izipx/=izipx .or. sim%izipv/=izipv) then
+      print*, 'zip format incompatable'
+      close(12)
+      stop
+    endif
+    read(12) rhoc ! coarse grid density
[…]

### 2017-05-30 CUBE/utilities/module_load_mac.sh

@@ -1,5 +1,6 @@
 export FC='mpif90'
-export XFLAG='-O3 -cpp -fcoarray=single -mcmodel=medium'
+#export XFLAG='-O3 -cpp -fcoarray=single -mcmodel=medium'
+export XFLAG='-O3 -cpp -fcoarray=single'
 #export XFLAG='-O3 -cpp -fcoarray=single -fopenmp'
 export OFLAG=${XFLAG}' -c'
 export FFTFLAG='-lfftw3f -lm -ldl'

### 2017-05-30 CUBE/utilities/powerspectrum.f90

@@ -66,9 +66,9 @@ real,parameter :: nexp=4.0 ! CIC kernel
 #endif
     xi(1,ibin)=xi(1,ibin)+1 ! number count
     xi(2,ibin)=xi(2,ibin)+kr ! k count
-    amp11=real(cx1(i,j,k)*conjg(cx1(i,j,k)))/(ng**3)/(ng**3)/(sinc**0.0)*4*pi*kr**3 ! linear density field
+    amp11=real(cx1(i,j,k)*conjg(cx1(i,j,k)))/(ng**3)/(ng**3)/(sinc**4.0)*4*pi*kr**3 ! linear density field
     amp22=real(cx2(i,j,k)*conjg(cx2(i,j,k)))/(ng**3)/(ng**3)/(sinc**4.0)*4*pi*kr**3 ! CIC'ed density field
-    amp12=real(cx1(i,j,k)*conjg(cx2(i,j,k)))/(ng**3)/(ng**3)/(sinc**2.0)*4*pi*kr**3 ! cross
+    amp12=real(cx1(i,j,k)*conjg(cx2(i,j,k)))/(ng**3)/(ng**3)/(sinc**4.0)*4*pi*kr**3 ! cross
 
     xi(3,ibin)=xi(3,ibin)+amp11 ! auto power 1
     xi(4,ibin)=xi(4,ibin)+amp22 ! auto power 2

### 2017-05-30 CUBE/main/cafcube.f90
vdisp_boost

@@ -1,5 +1,5 @@
 !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
-!       Coarray CubeP3M       !
+!   CUBE™ in Coarray Fortran  !
 !   haoran@cita.utoronto.ca   !
 !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
 
@@ -17,16 +17,16 @@ program cafcube
 
   call initialize
   call particle_initialization
-  print*,v(:,1),v(:,nplocal)
+  if (head) print*,v(:,1),v(:,nplocal)
   call buffer_density
-  print*,v(:,1),v(:,nplocal)
+  if (head) print*,v(:,1),v(:,nplocal)
   call buffer_x
-  print*,v(:,1),v(:,nplocal)
+  if (head) print*,v(:,1),v(:,nplocal)
   call buffer_v
-  print*,v(:,1),v(:,nplocal)
-  print*,'done'
+  if (head) print*,v(:,1),v(:,nplocal)
+  if (head) print*,' buffer before main loop done'
 
-    if (head) open(77,file='vel_info.bin',access='stream',status='replace')
+  if (head) open(77,file=output_dir()//'vinfo'//output_suffix(),access='stream',status='replace')
 
   if (head) print*, '---------- starting main loop ----------'
   DO istep=1,istep_max
@@ -34,10 +34,10 @@ program cafcube
     call update_particle
 
       ! velocity analysis
-      print*,'velocity analysis'
-      print*,'  scale factor',a,a_mid
+      if (head) print*,'velocity analysis'
+      if (head) print*,'  scale factor',a,a_mid
       max_vsim=0; std_vsim=0; kurt_vsim=0
-      print*, 'nplocal',nplocal
+      if (head) print*, 'nplocal',nplocal
       do ip=1,nplocal
         vreal=tan(pi*real(v(:,ip))/real(nvbin-1))/(sqrt(pi/2)/sigma_vi_old)
         abs_vsim=sqrt(sum(vreal**2))
@@ -50,9 +50,10 @@ program cafcube
       kurt_vsim=kurt_vsim/nplocal/std_vsim**4
       print*,'  vmax',max_vsim
       print*,'   std',std_vsim
-      print*,'  kurt',kurt_vsim
+      print*,' std_L',sigma_vi_old*sqrt(3.)
+      !print*,'  kurt',kurt_vsim
 
[…]

### 2017-05-30 CUBE/main/initialize.f90

@@ -83,6 +83,9 @@ subroutine initialize
   open(23,file='../velocity_conversion/vdisp.bin',access='stream')
   read(23) vdisp
   close(23)
+
+  vdisp(:,2)=vdisp(:,2)*vdisp_boost
+  ! convert to sim unit, and in each direction v_i
   vdisp(:,2)=vdisp(:,2)*vdisp(:,1)/1.5/box/h0/sqrt(omega_m)*real(nf_global)/sqrt(3.)
   sigma_vi_old=interp_vdisp(a_i)
   sigma_vi=interp_vdisp(a_i)

### 2017-05-30 CUBE/main/parameters.f90

@@ -59,7 +59,7 @@ module parameters
   real,parameter :: z_i_nu=z_i ! initial redshift for neutrinos
   real,parameter :: a_i=1/(1+z_i) ! initial scale factor
 
-  real,parameter :: box=250.0  ! simulation scale /dim, in unit of Mpc/h
+  real,parameter :: box=100.0  ! simulation scale /dim, in unit of Mpc/h
   real,parameter :: h0=67.74    ! Hubble constant
   real,parameter :: s8=0.8276   ! \sigma_8
   real,parameter :: ratio_nudm_dim=2 ! ratio of number of particles for neutrino/CDM, /dim
@@ -94,6 +94,7 @@ module parameters
   real,parameter :: ra_max=0.2
   real(8),parameter :: v_resolution=2.1/(int(2,8)**(izipv*8))
   real(8),parameter :: x_resolution=1.0/(int(2,8)**(izipx*8))
+  real(8),parameter :: vdisp_boost=1.75
 
   ! transfer function
   integer(8), parameter      :: nk_tf=2000

### 2017-05-30 CUBE/main/pm.f90

@@ -46,11 +46,11 @@ subroutine particle_mesh
   do itz=1,nnt
   do ity=1,nnt
   do itx=1,nnt
-    if (head) print*,'    tile',int(itx,1),int(ity,1),int(itz,1)
+    !if (head) print*,'    tile',int(itx,1),int(ity,1),int(itz,1)
     ! fine_cic_mass ------------------------------------------------------------
     rho_f=0
     crho_f=0
-    if (head) print*,'      fine_cic_mass'
+    !if (head) print*,'      fine_cic_mass'
     do k=2-ncb,nt+ncb-1
     do j=2-ncb,nt+ncb-1
     do i=2-ncb,nt+ncb-1
@@ -78,11 +78,11 @@ subroutine particle_mesh
     enddo
     enddo
     ! fine force ---------------------------------------------------------------
-    if (head) print*,'      fine_fft'
+    !if (head) print*,'      fine_fft'
     call sfftw_execute(plan_fft_fine)
     crho_f(:,:,:)=rho_f(:,:,:) ! back up
     do i_dim=1,3
-      if (head) print*,'      fine_ifft dim',int(i_dim,1)
+      !if (head) print*,'      fine_ifft dim',int(i_dim,1)
       rho_f(::2,:,:)=-crho_f(2::2,:,:)*kern_f(:,:,:,i_dim)
       rho_f(2::2,:,:)=crho_f(::2,:,:)*kern_f(:,:,:,i_dim)
       call sfftw_execute(plan_ifft_fine)
@@ -91,7 +91,7 @@ subroutine particle_mesh
     enddo
     f2_max_fine(itx,ity,itz)=maxval(sum(force_f(:,:,:,:)**2,1))
     ! fine velocity ------------------------------------------------------------
-    if (head) print*,'      fine velocity'
+    !if (head) print*,'      fine velocity'
     do k=1,nt
     do j=1,nt
     do i=1,nt ! loop over coarse cell

### 2017-05-30 CUBE/main/run_many_mac.sh

@@ -1,17 +1,17 @@
 source ../utilities/module_load_mac.sh
 
-cp parameters.f90_44 parameters.f90
-source run_mac.sh > log44
+#cp parameters.f90_1 parameters.f90
+#source run_mac.sh > log1
 
-cp parameters.f90_22 parameters.f90
-source run_mac.sh > log22
+#cp parameters.f90_2 parameters.f90
+#source run_mac.sh > log2
 
-cp parameters.f90_21 parameters.f90
-source run_mac.sh > log21
+#cp parameters.f90_3 parameters.f90
+#source run_mac.sh > log3
 
-cp parameters.f90_12 parameters.f90
-source run_mac.sh > log12
+cp parameters.f90_4 parameters.f90
+source run_mac.sh > log4
 
-cp parameters.f90_11 parameters.f90
-source run_mac.sh > log11
+cp parameters.f90_5 parameters.f90
+source run_mac.sh > log5
 

### 2017-05-30 CUBE/main/update_particle.f90

@@ -22,7 +22,7 @@ subroutine update_particle
   do itz=1,nnt ! loop over tile
   do ity=1,nnt
   do itx=1,nnt
-  if (head) print*,'  tile',int(itx,1),int(ity,1),int(itz,1)
+  !if (head) print*,'  tile',int(itx,1),int(ity,1),int(itz,1)
     rhoce=0
     rholocal=0
     x_new=0
@@ -31,7 +31,7 @@ subroutine update_particle
     pid_new=0
 #   endif
 
-    if (head) print*,'    density loop'
+    !if (head) print*,'    density loop'
     do k=1-ncb,nt+ncb ! loop over coarse grid
     do j=1-ncb,nt+ncb
     do i=1-ncb,nt+ncb
@@ -51,7 +51,7 @@ subroutine update_particle
     enddo
     enddo
 
-    if (head) print*,'    cubesum3'
+    !if (head) print*,'    cubesum3'
     cume=cumsum3(rhoce)
     overhead_tile=max(overhead_tile,cume(nt+2*ncb,nt+2*ncb,nt+2*ncb)/real(np_tile_max))
 
@@ -65,7 +65,7 @@ subroutine update_particle
 
     ! create a new x and v for this local tile
 
-  if (head) print*,'    xvnew loop'
+  !if (head) print*,'    xvnew loop'
     do k=1-ncb,nt+ncb ! update particle
     do j=1-ncb,nt+ncb
     do i=1-ncb,nt+ncb
@@ -111,7 +111,7 @@ subroutine update_particle
 
     sync all
 
-  if (head) print*,'    delete_particle loop'
+  !if (head) print*,'    delete_particle loop'
     do k=1,nt
     do j=1,nt
       ileft=iright+1
@@ -165,7 +165,7 @@ subroutine update_particle
 #ifdef PID
   !pid(:,nplocal+1:)=0
 #endif
-  print*,'  check xv of first particle',x(:,1),v(:,1)
+  !print*,'  check xv of first particle',x(:,1),v(:,1)
 
   sync all
 

### 2017-05-30 CUBE/utilities/initial_conditions.f90

@@ -454,6 +454,7 @@ print*, r3(1:2,1,1)**2
   open(23,file='../velocity_conversion/vdisp.bin',access='stream')
   read(23) vdisp
   close(23)
+  vdisp(:,2)=vdisp(:,2)*vdisp_boost
   vdisp(:,2)=vdisp(:,2)*vdisp(:,1)/1.5/box/h0/sqrt(omega_m)*real(nf_global)/sqrt(3.)
   sigma_vi=interp_vdisp(a)
   if (head) print*,'sigma_vi',sigma_vi

### 2017-06-09 CUBE/main/cafcube.f90
use nonlinear vdisp measured from simulation

@@ -8,8 +8,6 @@ program cafcube
   implicit none
   save
 
-  real(8) abs_vsim, std_vsim, max_vsim, kurt_vsim
-
   if (this_image()==1) print*, 'Coarray CUBE on',nn**3,'  images'
   sync all
   call system('hostname')
@@ -17,14 +15,9 @@ program cafcube
 
   call initialize
   call particle_initialization
-  if (head) print*,v(:,1),v(:,nplocal)
   call buffer_density
-  if (head) print*,v(:,1),v(:,nplocal)
   call buffer_x
-  if (head) print*,v(:,1),v(:,nplocal)
   call buffer_v
-  if (head) print*,v(:,1),v(:,nplocal)
-  if (head) print*,' buffer before main loop done'
 
   if (head) open(77,file=output_dir()//'vinfo'//output_suffix(),access='stream',status='replace')
 
@@ -32,49 +25,6 @@ program cafcube
   DO istep=1,istep_max
     call timestep
     call update_particle
-
-      ! velocity analysis
-      if (head) print*,'velocity analysis'
-      if (head) print*,'  scale factor',a,a_mid
-      max_vsim=0; std_vsim=0; kurt_vsim=0
-      if (head) print*, '  nplocal',nplocal
-      !do ip=1,nplocal
-      ip=0
-      do itz=1,nnt
-      do ity=1,nnt
-      do itx=1,nnt
-      do k=1,nt
-      do j=1,nt
-      do i=1,nt
-      do l=1,rhoc(i,j,k,itx,ity,itz)
-        ip=ip+1
-        vreal=tan(pi*real(v(:,ip))/real(nvbin-1))/(sqrt(pi/2)/(sigma_vi_old*vrel_boost))
-        vreal=vreal+vfield(:,i,j,k,itx,ity,itz)
-        abs_vsim=sqrt(sum(vreal**2))
-        max_vsim=max(max_vsim,abs_vsim)
-        std_vsim=std_vsim+abs_vsim**2!         sqrt(sum(abs_vsim**2/nplocal*1d0))
-        kurt_vsim=kurt_vsim+abs_vsim**4!       sum(abs_vsim**4*1d0/nplocal)/std_vsim**4
-        !print*,abs_vsim,max_vsim,std_vsim,kurt_vsim
-      enddo
-      enddo
-      enddo
-      enddo
-      enddo
-      enddo
[…]

### 2017-06-09 CUBE/main/initialize.f90

@@ -80,15 +80,22 @@ subroutine initialize
 #endif
   sync all
 
-  open(23,file='../velocity_conversion/vdisp.bin',access='stream')
-  read(23) vdisp
-  close(23)
+  !open(23,file='../velocity_conversion/vdisp.bin',access='stream')
+  !read(23) vdisp
+  !close(23)
 
-  vdisp(:,2)=vdisp(:,2)*vdisp_boost
-  ! convert to sim unit, and in each direction v_i
-  vdisp(:,2)=vdisp(:,2)*vdisp(:,1)/1.5/box/h0/sqrt(omega_m)*real(nf_global)/sqrt(3.)
-  sigma_vi_old=interp_vdisp(a_i)
-  sigma_vi=interp_vdisp(a_i)
+  !open(23,file='../velocity_conversion/sigmav_z.bin',access='stream')
+  !read(23) svz
+  !close(23)
+
+  !open(23,file='../velocity_conversion/sigmav_r.bin',access='stream')
+  !read(23) svr
+  !close(23)
+
+  !svz(:,2)=svz(:,2)*svz(:,1)/1.5/box/h0/sqrt(omega_m)*real(nf_global)/sqrt(3.)
+
+  !sigma_vi_old=interp_vdisp(a_i)
+  !sigma_vi=interp_vdisp(a_i)
   !if (head) print*,'sigma_vi',sigma_vi
   sync all
 

### 2017-06-09 CUBE/main/parameters.f90

@@ -3,11 +3,11 @@ module parameters
   save
 
   ! output directory for both IC and snapshots
-  character(*),parameter :: opath='../output/universe5/'
+  character(*),parameter :: opath='../output/universe2/'
 
   ! simulation parameters
-  integer(8),parameter :: izipx=1 ! 1 or 2, for particle location
-  integer(8),parameter :: izipv=1 ! 1 or 2, for particle velocity
+  integer(8),parameter :: izipx=1 
+  integer(8),parameter :: izipv=1
   integer(8),parameter :: nvbin=int(2,8)**(8*izipv)
   integer(8),parameter :: ishift=-(int(2,8)**(izipx*8-1))
   real(8),parameter :: rshift=0.5-ishift
@@ -23,6 +23,7 @@ module parameters
 
   integer(8),parameter :: nf=nc*ncell ! >=96
   integer(8),parameter :: nf_global=nf*nn
+  integer(8),parameter :: nc_global=nc*nn
   integer(8),parameter :: nft=nt*ncell ! >=48
 
   ! ngrid /image/dim for pencil-fft
@@ -94,8 +95,8 @@ module parameters
   real,parameter :: ra_max=0.2
   real(8),parameter :: v_resolution=2.1/(int(2,8)**(izipv*8))
   real(8),parameter :: x_resolution=1.0/(int(2,8)**(izipx*8))
-  real(8),parameter :: vdisp_boost=0.5
-  real(8),parameter :: vrel_boost=2.0
+  !real(8),parameter :: vdisp_boost=1.0
+  real(8),parameter :: vrel_boost=2.7
 
   ! transfer function
   integer(8), parameter      :: nk_tf=2000
@@ -129,7 +130,9 @@ module parameters
     real omega_l
     real s8
     real vsim2phys
-    real z_i ! 40*4 bytes
+    real sigma_vres
+    real sigma_vi
+    real z_i ! 42*4 bytes
   endtype
 
   type(sim_header) sim
@@ -160,6 +163,8 @@ module parameters
       print*,'| omega_l      =',s%omega_l
       print*,'| sigma_8      =',s%s8
       print*,'| vsim2phys    =',s%vsim2phys, '(km/s)/(1.0)'
+      print*,'| sigma_vres   =',s%sigma_vres,'(km/s)'
+      print*,'| sigma_vi     =',s%sigma_vi,'(simulation unit)'
       print*,'| z_i          =',s%z_i
       print*,'------------------------------------------------------------------------------'
       endif

### 2017-06-09 CUBE/main/particle_initialization.f90

@@ -73,10 +73,26 @@ subroutine particle_initialization
 
   print*,'  from image',this_image(),'read',nplocal,' particles'
   sync all
+
+  !sigma_vfi=interp_sigmav(a_i,box/nf_global) ! sigma(v) on scale of fine grid
+  !sigma_vci=interp_sigmav(a_i,box/nc_global) ! sigma(v) on scale of coarse grid
+
+  !print*, sigma_vfi, sim%vsim2phys; stop
+
+  sigma_vi=sim%sigma_vi
+  !sigma_vres=sqrt(sigma_vfi**2-sigma_vci**2) ! sigma(v) residual
+  !sigma_vci_old=sigma_vci
+  !sigma_vfi_old=sigma_vfi
+  !sigma_vres_old=sigma_vres
+
   if (head) then
-    print*, '  npglobal =', npglobal
-    print*, '  mass_p=', mass_p
-    print*, '  vsim2phy =',sim%vsim2phys, ' (km/s)/(1.0)'
+    print*,'  npglobal =', npglobal
+    print*,'  mass_p=', mass_p
+    print*,'  vsim2phy =',sim%vsim2phys, ' (km/s)/(1.0)'
+    !print*,'  std_vf(a=',a_i,', r=',box/nf_global,'Mpc/h)',sqrt(3.)*sigma_vfi*sim%vsim2phys,'km/s'
+    !print*,'  std_vc(a=',a_i,', r=',box/nc_global,'Mpc/h)',sqrt(3.)*sigma_vci*sim%vsim2phys,'km/s'
+    !print*,'  std_vres',sqrt(3.)*sigma_vres*sim%vsim2phys,'km/s'
+    print*,'  sigma_vi =',sigma_vi,'(simulation unit)'
   endif
   sync all
 endsubroutine

### 2017-06-09 CUBE/main/pm.f90

@@ -34,13 +34,6 @@ subroutine particle_mesh
   f2_max_pp=0
   f2_max_coarse=0
 
-  sigma_vi=interp_vdisp(a+da)
-  if (head) then
-    print*, '  a, sigma_vi_old =',a,sigma_vi_old
-    print*, '  a+da,  sigma_vi =',a+da, sigma_vi
-  endif
-  sync all
-
   if (head) print*, '  pm fine over',nnt**3,'tiles'
 
   do itz=1,nnt
@@ -106,12 +99,8 @@ subroutine particle_mesh
         dx2 = 1 - dx1
         idx1=idx1+nfb
         idx2=idx2+nfb
-        vreal=tan(pi*real(v(:,ip))/real(nvbin-1))/(sqrt(pi/2)/(sigma_vi_old*vrel_boost))
-        !print*,v(:,ip)
-        !print*,vreal
-        !print*,vfield(:,i,j,k,itx,ity,itz)
-        !print*,vreal+vfield(:,i,j,k,itx,ity,itz)
-        !print*,'force'
+        vreal=tan(pi*real(v(:,ip))/real(nvbin-1))/(sqrt(pi/2)/(sigma_vi*vrel_boost))
+
         vreal=vreal+force_f(:,idx1(1),idx1(2),idx1(3))*a_mid*dt/6/pi*dx1(1)*dx1(2)*dx1(3)
         vreal=vreal+force_f(:,idx2(1),idx1(2),idx1(3))*a_mid*dt/6/pi*dx2(1)*dx1(2)*dx1(3)
         vreal=vreal+force_f(:,idx1(1),idx2(2),idx1(3))*a_mid*dt/6/pi*dx1(1)*dx2(2)*dx1(3)
@@ -120,18 +109,8 @@ subroutine particle_mesh
         vreal=vreal+force_f(:,idx2(1),idx1(2),idx2(3))*a_mid*dt/6/pi*dx2(1)*dx1(2)*dx2(3)
         vreal=vreal+force_f(:,idx2(1),idx2(2),idx1(3))*a_mid*dt/6/pi*dx2(1)*dx2(2)*dx1(3)
         vreal=vreal+force_f(:,idx2(1),idx2(2),idx2(3))*a_mid*dt/6/pi*dx2(1)*dx2(2)*dx2(3)
-        !print*,vreal
-        !print*,vfield(:,i,j,k,itx,ity,itz)
-        !print*,vreal+vfield(:,i,j,k,itx,ity,itz)
[…]

### 2017-06-09 CUBE/main/update_particle.f90

@@ -11,14 +11,9 @@ subroutine update_particle
   real(4) :: weight_v=0.1 ! how previous-step vfield is mostly weighted
   integer(8) cume(1-2*ncb:nt+2*ncb,1-2*ncb:nt+2*ncb,1-2*ncb:nt+2*ncb)
   integer(4) rholocal(1-2*ncb:nt+2*ncb,1-2*ncb:nt+2*ncb,1-2*ncb:nt+2*ncb) ! count writing
-  integer(8) ii,jj,kk
-  integer(8) iii(3)
-  integer(8) pr1,pr2,pr3,pr4
 
   if (head) print*,'update_particle'
   dt_mid=(dt_old+dt)/2
-
-
   overhead_tile=0
   iright=0
   do itz=1,nnt ! loop over tile
@@ -34,7 +29,6 @@ subroutine update_particle
 #   ifdef PID
     pid_new=0
 #   endif
-
     !if (head) print*,'    density loop'
     do k=1-ncb,nt+ncb ! loop over coarse grid
     do j=1-ncb,nt+ncb
@@ -44,29 +38,25 @@ subroutine update_particle
       do l=1,np
         ip=nlast-np+l
         xq=((/i,j,k/)-1d0) + (int(x(:,ip)+ishift,izipx)+rshift)*x_resolution
-        vreal=tan((pi*real(v(:,ip)))/real(nvbin-1)) / (sqrt(pi/2)/(sigma_vi_old*vrel_boost))
+        vreal=tan((pi*real(v(:,ip)))/real(nvbin-1)) / (sqrt(pi/2)/(sigma_vi*vrel_boost))
         vreal=vreal+vfield(:,i,j,k,itx,ity,itz)
         deltax=(dt_mid*vreal)/ncell
         g=ceiling(xq+deltax)
         rhoce(g(1),g(2),g(3))=rhoce(g(1),g(2),g(3))+1 ! update mesh
         vfield_new(:,g(1),g(2),g(3))=vfield_new(:,g(1),g(2),g(3))+vreal
-        !print*,x(:,ip)
-        !print*,xq,deltax,g;stop
       enddo
     enddo
     enddo
     enddo
     ! vfield_new is kept the same as previous-step if the grid is empty.
     ! vfield_new is at most weight_v weighted by previous-step for non-empty grids.
-    ! this also gets rid of "if"s.
[…]

### 2017-06-09 CUBE/main/variables.f90

@@ -32,8 +32,11 @@ module variables
 
   ! FFT plans
   integer(8) plan_fft_fine,plan_ifft_fine
-  real vmax,overhead_tile[*],overhead_image[*]
-  real vdisp(506,2),sigma_vi_old,sigma_vi
+  real vmax,overhead_tile[*],overhead_image[*],sigma_vi,sigma_vi_new
+  !real vdisp(506,2),sigma_vi_old,sigma_vi
+  real(4) svz(500,2),svr(100,2)
+  real(8) sigma_vci,sigma_vfi,sigma_vres,sigma_vci_old,sigma_vfi_old,sigma_vres_old
+  real(8) std_vsim_c,std_vsim_res,std_vsim
   ! n^3
   integer(izipx) x(3,np_image_max)[*], x_new(3,np_tile_max)
   integer(izipv) v(3,np_image_max)[*], v_new(3,np_tile_max)
@@ -101,22 +104,34 @@ contains
     enddo
   endfunction
 
-  real function interp_vdisp(aa)
+  real function interp_sigmav(aa,rr)
     implicit none
     integer(8) ii,i1,i2
-    real aa
+    real aa,rr,term_z,term_r
     i1=1
-    i2=506
+    i2=500
     do while (i2-i1>1)
       ii=(i1+i2)/2
-      if (aa>vdisp(ii,1)) then
+      if (aa>svz(ii,1)) then
         i1=ii
       else
         i2=ii
       endif
     enddo
-    interp_vdisp=vdisp(i1,2)+(vdisp(i2,2)-vdisp(i1,2))*(aa-vdisp(i1,1))/(vdisp(i2,1)-vdisp(i1,1))
+    term_z=svz(i1,2)+(svz(i2,2)-svz(i1,2))*(aa-svz(i1,1))/(svz(i2,1)-svz(i1,1))
+    i1=1
+    i2=100
+    do while (i2-i1>1)
+      ii=(i1+i2)/2
+      if (rr>svz(ii,1)) then
+        i1=ii
+      else
+        i2=ii
+      endif
+    enddo
+    term_r=svr(i1,2)+(svr(i2,2)-svr(i1,2))*(rr-svr(i1,1))/(svr(i2,1)-svr(i1,1))
+    interp_sigmav=term_z*term_r
+    print*,term_z,term_r
   endfunction
-
 
 endmodule

### 2017-06-09 CUBE/utilities/initial_conditions.f90

@@ -26,7 +26,7 @@ program initial_conditions
   real(8) v8, norm, xq(3),gradphi(3),vreal(3), dvar[*], dvarg
   integer(int64) :: time64
 
-  integer(8) nplocal[*],npglobal
+  integer(8) nplocal[*],npglobal,ip,l
   ! power spectrum arrays
   real, dimension(7,nk) :: tf    !CAMB
   real, dimension(2,nc) :: pkm,pkn
@@ -62,7 +62,10 @@ program initial_conditions
     integer(8) iq(3)
 #endif
   real grad_max(3)[*],vmax(3),vf
-  real vdisp(506,2),sigma_vi
+  !real vdisp(506,2),sigma_vi
+  real(4) svz(500,2),svr(100,2)
+  real(8) sigma_vc,sigma_vf,sigma_vres,sigma_vi
+  real(8) std_vsim_c,std_vsim_res,std_vsim
 
   character (10) :: img_s, z_s
 
@@ -452,13 +455,28 @@ print*, r3(1:2,1,1)**2
 
   sync all
 
-  open(23,file='../velocity_conversion/vdisp.bin',access='stream')
-  read(23) vdisp
+  open(23,file='../velocity_conversion/sigmav_z.bin',access='stream')
+  read(23) svz
   close(23)
-  vdisp(:,2)=vdisp(:,2)*vdisp_boost
-  vdisp(:,2)=vdisp(:,2)*vdisp(:,1)/1.5/box/h0/sqrt(omega_m)*real(nf_global)/sqrt(3.)
-  sigma_vi=interp_vdisp(a)
-  if (head) print*,'sigma_vi',sigma_vi
+  open(24,file='../velocity_conversion/sigmav_r.bin',access='stream')
+  read(24) svr
+  close(43)
+
+  !svz(:,2)=svz(:,2)/sim%vsim2phys/sqrt(3.)
+  sigma_vf=interp_sigmav(a,box/nf_global) ! sigma(v) on scale of fine grid, in km/s
+  sigma_vc=interp_sigmav(a,box/nc_global) ! sigma(v) on scale of coarse grid, in km/s
+  sigma_vres=sqrt(sigma_vf**2-sigma_vc**2) ! sigma(v) residual, in km/s
[…]

### 2017-06-27 CUBE/main/buffer_density.f90
using nonlinear v-field in simulation

@@ -87,31 +87,17 @@ subroutine buffer_density
     stop
   endif
 
-  !print*, 'x before shift =', sum(x*unit8)
-  !print*, 'v before shift =', sum(v*unit8)
 
-  ! move xv to the end, leave enough space for buffer
-  !print*,sum(rhoc),np_image_max
   nshift=np_image_max-nplocal
   x(:,nshift+1:np_image_max)=x(:,1:nplocal)
   v(:,nshift+1:np_image_max)=v(:,1:nplocal)
-  !x(:,:nshift)=0 ! redundant
-  !v(:,:nshift)=0 ! redundant
 
-#ifdef PID
-  pid(nshift+1:np_image_max)=pid(1:nplocal)
-#endif
-
-  !print*, 'x after shift=', sum(x*unit8)
+# ifdef PID
+    pid(nshift+1:np_image_max)=pid(1:nplocal)
+# endif
 
   cum=cumsum6(rhoc)
-  !print*, 'cumsum of all particles =', cum(nt+ncb,nt+ncb,nt+ncb,nnt,nnt,nnt)
-  ! create extended zip
-
-  ! redistribute local particles -----------------
-
-  ! nshift is the offset between local (backed up at the end) and extented
-  ifrom=nshift ! ifrom is the index of shifted particles, contiuous
+  ifrom=nshift
 
   do itz=1,nnt
   do ity=1,nnt
@@ -124,9 +110,9 @@ subroutine buffer_density
       nlen=nlast-cum(0,iy,iz,itx,ity,itz)
       x(:,nlast-nlen+1:nlast)=x(:,ifrom+1:ifrom+nlen)
       v(:,nlast-nlen+1:nlast)=v(:,ifrom+1:ifrom+nlen)
-#ifdef PID
+#     ifdef PID
         pid(nlast-nlen+1:nlast)=pid(ifrom+1:ifrom+nlen)
-#endif
+#     endif
       ifrom=ifrom+nlen
     enddo
     enddo

### 2017-06-27 CUBE/main/initialize.f90

@@ -1,7 +1,3 @@
-!#define readkernel
-!#define penffttest
-#define mkdir
-
 subroutine initialize
   use variables
   use cubefft
@@ -22,28 +18,9 @@ subroutine initialize
   sync all
   ! omp_init
 
-#ifdef penffttest
-    if (head) print*, 'pencil_fft test...'
-    call random_number(r3)
-    call pencil_fft_forward
-    call pencil_fft_backward
-    if (head) print*, 'penfft test done.'
-    sync all
-#endif
 
-
-#ifndef readkernel
-    call kernel_f
-    call kernel_c
-#else
-  ! now works only for single image
-    open(14,file='cubep3m_kern_f.dat',status='old',access='stream')
-    read(14) kern_f
-    close(14)
-    open(15,file='cubep3m_kern_c.dat',status='old',access='stream')
-    read(15) kern_c
-    close(15)
-#endif
+  call kernel_f
+  call kernel_c
 
   istep=0
   t=0
@@ -73,30 +50,7 @@ subroutine initialize
   sync all
   n_checkpoint=n_checkpoint[1]
   z_checkpoint(:)=z_checkpoint(:)[1]
-  sync all
-
-#ifdef mkdir
-    call system('mkdir -p '//opath//'/image'//image2str(image))
-#endif
-  sync all
-
-  !open(23,file='../velocity_conversion/vdisp.bin',access='stream')
-  !read(23) vdisp
-  !close(23)
-
-  !open(23,file='../velocity_conversion/sigmav_z.bin',access='stream')
-  !read(23) svz
-  !close(23)
-
-  !open(23,file='../velocity_conversion/sigmav_r.bin',access='stream')
-  !read(23) svr
-  !close(23)
-
-  !svz(:,2)=svz(:,2)*svz(:,1)/1.5/box/h0/sqrt(omega_m)*real(nf_global)/sqrt(3.)
-
-  !sigma_vi_old=interp_vdisp(a_i)
-  !sigma_vi=interp_vdisp(a_i)
-  !if (head) print*,'sigma_vi',sigma_vi
+  call system('mkdir -p '//opath//'/image'//image2str(image))
   sync all
 
 endsubroutine

### 2017-06-27 CUBE/main/parameters.f90

@@ -3,10 +3,10 @@ module parameters
   save
 
   ! output directory for both IC and snapshots
-  character(*),parameter :: opath='../output/universe2/'
+  character(*),parameter :: opath='../output/universe5/'
 
   ! simulation parameters
-  integer(8),parameter :: izipx=1 
+  integer(8),parameter :: izipx=1
   integer(8),parameter :: izipv=1
   integer(8),parameter :: nvbin=int(2,8)**(8*izipv)
   integer(8),parameter :: ishift=-(int(2,8)**(izipx*8-1))
@@ -96,7 +96,7 @@ module parameters
   real(8),parameter :: v_resolution=2.1/(int(2,8)**(izipv*8))
   real(8),parameter :: x_resolution=1.0/(int(2,8)**(izipx*8))
   !real(8),parameter :: vdisp_boost=1.0
-  real(8),parameter :: vrel_boost=2.7
+  real(8),parameter :: vrel_boost=2.5
 
   ! transfer function
   integer(8), parameter      :: nk_tf=2000

### 2017-09-13 ms_caf/ms_caf.tex
use nonlinear vdisp measured from simulation
using nonlinear v-field in simulation
power spectrum figures in the paper
new figure for two box sizes and two redshifts.
paper: title, abstract & introduction

@@
 \begin{document}
-\title{CUBE: Compact Particle-Mesh $N$-body Algorithm}
+\title{CUBE: A Memory-lite Cosmological $N$-body Algorithm}
 \author{
@@
 \begin{abstract}
-Cosmological large scale structure $N$-body simulations are computation-light but memory-heavy imbalanced problems in supercomputing.
-Traditional $N$-body simulation algorithms use at least 24-byte per particle in memory, where six 4-byte single precision real numbers keep track of each particle's phase space coordinate.
-This is usually the bottleneck of scaling-up of a $N$-body problem on modern supercomputers whose core/memory ratio is high.
-Here we present a parallel, memory-lite Particle-Mesh based $N$-body code, in which each particle can occupy as low as 6-byte memory.
-This is accomplished by storing only particles' relative location respect to a mesh-grid, and their velocity relative to a predictable distribution function.
-The remaining information is given by the ordering of particles, and a complimentary density field, which is negligible in memory space.
-Our numerical experiments show that this algorithm provides acceptable accuracy compared to traditional algorithm in cosmological simulations.
+Cosmological large scale structure $N$-body simulations are computation-light,
+memory-heavy problems in supercomputing. Traditional $N$-body simulation
+algorithms use at least 24-byte memory per particle, of which six 4-byte single 
+precision floating point numbers keep track phase space coordinates of each particle.
+Here we present the algorithm and accuracy of a new parallel, memory-lite, Particle-
+Mesh based $N$-body code, where each particle can occupy as low as 6-byte memory. This 
+is accomplished by storing relative position and relative velocity of each particle, 
+in the format of 1-byte-integer, respect to their averaged value of a mesh-grid. The 
+remaining information is given by complimentary density and velocity fields, which are 
+negligible in memory space, and proper ordering of particles, which gives no extra 
+memory. Our numerical experiments show that this integer based $N$-body algorithm 
+provides acceptable accuracy compared to traditional algorithm in cosmological 
+simulations. This significant lowering of memory-to-computation ratio breaks the 
+bottleneck of scaling up and speeding up large cosmological $N$-body simulations on 
+multi-core and heterogenous computing systems.
 
@@
 
-\begin{figure}[t]
-\centering
-  \includegraphics[width=0.95\linewidth]{ms_caf.pdf}
-\end{figure}
 \section{Introduction}
-$N$-body simulation is a powerful tool to solve the highly nonlinear dynamic problems \citep{1988csup.book.....H}.
-It is widely used in cosmology to model the formation and evolution of the large scale structure (LSS).
-With the fast development of parallel supercomputers, we are able to simulate a system of more than a trillion ($10^{12}$) $N$-body particles.
-To date the largest $N$-body simulation in application is the ``TianNu'' simulation run on the TianHe-2 supercomputer by cosmological simulation code {\tt CUBEP3M} \citep{2013MNRAS.436..540H}.
-It uses nearly $3\times 10^{12}$ particles to simulate the cold dark matter (CDM) and cosmic neutrino evolution through the cosmic age.
-However {\tt CUBEP3M} and TianNu did not fully use TianHe-2's computing power -- 87\% of the computing power on TianHe-2 are given by the Intel Xeon-Phi coprocessors while they occupy only 27\% of the total memory.
-Although we accomplished preliminary works on developing native and offload modes of {\tt CUBEP3M} on Xeon-Phi many integrated core (MIC) architecture, the memory problem is still the bottle neck of scaling up the $N$-body problem on the heterogeneous architectures.
-Lower the memory usage per particle will solve this problem, and will also help the $N$-body problem development on GPUs, which are even more massive parallelized and memory limited.
-In this paper we present a $N$-body algorithm which uses as low as 6 byte per particle (here after we use ``bpp'' referring to ``byte per particle'') and show that it can give accurate cosmological LSS simulations.
+$N$-body simulation is a powerful tool to solve the highly nonlinear dynamic problems 
+\citep{1988csup.book.....H}. It is widely used in cosmology to model the formation and 
+evolution of the large scale structure (LSS). With the fast development of parallel 
+supercomputers, we are able to simulate a system of more than a trillion ($10^{12}$) 
+$N$-body particles. To date the largest $N$-body simulation in application is the 
+``TianNu'' simulation \citep{2017NatAs...1E.143Y,2017RAA....17...85E} run on the 
+TianHe-2 supercomputer by cosmological simulation code 
+{\tt CUBEP3M} \citep{2013MNRAS.436..540H}. It uses nearly $3\times 10^{12}$ particles 
+to simulate the cold dark matter (CDM) and cosmic neutrino evolution through the 
+cosmic age.
+
+The $N$-body simulations use considerable amount of memory, because the phase space 
+coordinates $(x,y,z,v_x,v_y,v_z)$ of each $N$-body particle must be stored as, at 
+least, six single precision floating point numbers (24 bytes). Contrarily, their 
+computing workload can be alleviated by many algorithms, like Particle-Mesh [cite] and 
+Tree [cite], scale as $o(N\log N)$ or even $o(N)$. On the other hand, modern 
+supercomputer systems use multi cores, many integrated cores (MIC) and even densely 
+parallelized GPUs, bringing orders of magnitude higher computing power, whereas these 
+architectures usually have limited memory allocation. Thus, the computation-light but 
+memory-heavy applications, compared to matrix multiplication and decomposition 
+calculations, are less suitable for fully usage of the computing power of modern 
+supercomputers. For example, although native and offload modes of {\tt CUBEP3M} are 
+able to run on the Intel Xeon-PHI MIC architectures, with the requirement of enough 
+memory, TianNu simulation were done on TianHe-2 with only its CPUs -- 73\% of the 
+total memory but only 13\% of the total computing power.
+
+We present a new $N$-body simulation code {\tt CUBE}, using as low as 6 byte per particle (here after we use ``bpp'' referring to ``byte per particle'').
+We show that it gives accurate results cosmological LSS simulations.
 The algorithm is presented in \S 2, and a comparison between this method and traditional method is shown in \S 3.
@@
 \section{Algorithm}
-The most memory consuming part of a $N$-body simulation is usually the phase space coordinates of each $N$-body particles -- 24 bpp (4 single precision real numbers) must be used to store each particles' position and velocity.
-A memory-lite parallel $N$-body code {\tt CUBEP3M} can use as low as 40 bpp in sacrificing the speed.
-This includes the phase coordinates (24 bpp) for particles in physical domain and buffered region, a linked list (4 bpp), and a global coarse mesh and local fine mesh.
-We attempt to replace the coordinates and linked list 24+4=28 bpp memory usage with a integer based storage, described as follows.
+The most memory consuming part of a $N$-body simulation is usually the phase space 
+coordinates of $N$-body particles -- 24 bpp (4 single precision floating numbers) must 
+be used to store each particles' 3D position and velocity vectors. {\tt CUBEP3M}, an 
+example of a memory-lite parallel $N$-body code, can use as low as 40 bpp in 
+sacrificing computing speed \citep{2013MNRAS.436..540H}. This includes the phase 
+coordinates (24 bpp) for particles in physical domain and buffered region, a linked 
+list (4 bpp), and a global coarse mesh and local fine mesh. We attempt to replace the 
+coordinates and linked list 24+4=28 bpp memory usage with a integer based storage, 
+described as follows.
+\begin{figure}[t]
+\centering
+  \includegraphics[width=0.95\linewidth]{fig_tile.pdf}
+ \caption{Spacial decomposition.}
+\label{fig.tile}
+\end{figure}
 
 \subsection{Particle location storage}
-Instead of storing the particle's global coordinates, we store its relative position respective to the coarse cell who contains the particle.
-We divide the coarse cell, in each dimension, evenly into $2^8=256$ bins, and we use a 1-byte integer (with range 0 to 255 or -128 to 127) to show which bin it locates in this dimension.
+Instead of storing the global coordinates of each particle, we store its offset relative to its parent coarse cell which contains the particle.
+We divide the coarse cell, in each dimension, evenly into $2^8=256$ bins, and use a 1-byte integer (with range 0 to 255 or -128 to 127) to indicate which bin it locates in this dimension.
 Next, global locations of particles are given by cell-ordered format in memory space, and a complimentary number count of particle number in this coarse mesh (coarse density field) will give complete information of particle distribution in cells.
@@
 \centering
-  \includegraphics[width=0.95\linewidth]{fig_tile.pdf}
- \caption{Spacial decomposition.}
+  \includegraphics[width=1.0\linewidth]{vdisp.pdf}
+ \caption{Velocity dispersion.}
 \label{fig.tile}
@@
 
+\begin{figure}[t]
+\centering
+  \includegraphics[width=0.95\linewidth]{overplotparticles.pdf}
+ \caption{particles.}
+\label{fig.particles}
+\end{figure}
+
 By far, particle velocities in the physical regions of each tile are updated.
@@
 
-\begin{figure}[h]
-\begin{verbatim}
+\begin{figure}
+\centering
+  \includegraphics[width=0.95\linewidth]{power_universe1-8_z0and1.pdf}
+ \caption{power spectrum.}
+\label{fig.power}
+\end{figure}
 
-\end{verbatim}
-
-\caption{Power spectrum differential plot}
-\label{fig.power}
+\begin{figure}
+\centering
+  \includegraphics[width=0.95\linewidth]{diff_power.pdf}
+ \caption{power spectrum and cross correlation.}
+\label{fig.ccc}
 \end{figure}

### 2017-09-17 ms_caf/ms_caf.tex
position and velocity sections

@@
 \affil{
-$^1$Kavli Institute for Astronomy and Astrophysics, Peking University, Beijing 100871, China\\
+$^1$Tsung-Dao Lee Institute, Shanghai Jiao Tong University, Shanghai, 200240, China\\
 $^2$Canadian Institute for Theoretical Astrophysics, University of Toronto, Toronto, ON M5H 3H8 Canada\\
@@
 supercomputers. For example, although native and offload modes of {\tt CUBEP3M} are 
-able to run on the Intel Xeon-PHI MIC architectures, with the requirement of enough 
+able to run on the Intel Xeon-PHI MIC architectures, with the require ment of enough 
 memory, TianNu simulation were done on TianHe-2 with only its CPUs -- 73\% of the 
@@
 
-We present a new $N$-body simulation code {\tt CUBE}, using as low as 6 byte per particle (here after we use ``bpp'' referring to ``byte per particle'').
-We show that it gives accurate results cosmological LSS simulations.
-The algorithm is presented in \S 2, and a comparison between this method and traditional method is shown in \S 3.
-Discussions and conclusions are in \S 4.
+We present a new $N$-body simulation code {\tt CUBE}, using as low as 6 byte per particle 
+(here after we use ``bpp'' referring to ``byte per particle''). We show that it gives 
+accurate results in cosmological LSS simulations. The algorithm is presented in \S 2, and 
+a comparison between this method and traditional method is shown in \S 3. Discussions and 
+conclusions are in \S 4.
 
@@
 coordinates (24 bpp) for particles in physical domain and buffered region, a linked 
-list (4 bpp), and a global coarse mesh and local fine mesh. We attempt to replace the 
-coordinates and linked list 24+4=28 bpp memory usage with a integer based storage, 
-described as follows.
-\begin{figure}[t]
-\centering
-  \includegraphics[width=0.95\linewidth]{fig_tile.pdf}
- \caption{Spacial decomposition.}
-\label{fig.tile}
-\end{figure}
+list (4 bpp), and a global coarse mesh and local fine mesh. 4-byte real numbers are 
+not necessarily adequate in representing the {\it global} coordinates in simulations.
+If the box size is many orders of magnitude larger than interactive distance
+between particles, especially in the field of re-simulation of dense subregions,
+double precision (8-byte) coordinates are needed to avoid truncation errors.
+Another solution is to record {\it relative} coordinates for both position and velocity.
+{\tt CUBE} replaces the coordinates and linked list 24+4=28 bpp memory usage
+with an integer based storage, reduces the basic memory usage from 28 bpp down to
+6 bpp, described as following subsections.
 
-\subsection{Particle location storage}
-Instead of storing the global coordinates of each particle, we store its offset relative to its parent coarse cell which contains the particle.
-We divide the coarse cell, in each dimension, evenly into $2^8=256$ bins, and use a 1-byte integer (with range 0 to 255 or -128 to 127) to indicate which bin it locates in this dimension.
-Next, global locations of particles are given by cell-ordered format in memory space, and a complimentary number count of particle number in this coarse mesh (coarse density field) will give complete information of particle distribution in cells.
-Since the this field is usually coarser than particle distribution, it takes negligible memory.
-This coarse density field can be further compressed into 1-byte integer format, such that a 1-byte unsigned integer show the particle number in this coarse cell in range 0 to 255.
-In the densest cells (rarely happened) where there are $\ge 255$ particles, we can just write 255, and write the actual number as a 4-byte integer in another file.
-In a simulation with volume $L^3$ and $N_c^3$ coarse cells, particle positions are stored with a precision of $L/(256N_c)$, which is usually 10 times finer than the force resolution.
+\subsection{Particle position storage}
+We construct a uniform mesh throughout the space and each particle belongs 
+to its parent cell of the mesh. Instead of storing global coordinates of 
+each particle, we store its offset relative to its parent cell which 
+contains the particle. We divide the cell, in each dimension $d$, evenly into 
+$2^8=256$ bins, and use a 1-byte (8 bits) integer $\chi_d \in \{-128,-127,...,127\}$ to 
+indicate which bin it locates in this dimension. The global locations of 
+particles are given by cell-ordered format in memory space, and a 
+complimentary number count of particle number in this mesh (density field) 
+will give complete information of particle distribution in the mesh. Then 
+the global coordinate in $d$th dimension $x_d$ is given by $x_d=(n_c-1)+
+(\chi_d+128+1/2)/256$, where $n_c=1,2,...,N_c$ 
+is the index of the coarse grid. The mesh 
+is chosen to be coarse enough such that the density field takes negligible 
+memory. This coarse density field can be further compressed into 1-byte 
+integer format, such that a 1-byte integer show the particle number in this 
+coarse cell in range 0 to 255. In the densest cells (rarely happened) where 
+there are $\ge 255$ particles, we can just write 255, and write the actual 
+number as a 4-byte integer in another file.
+
+In a simulation with volume $L^3$ and $N_c^3$ coarse cells, particle 
+positions are stored with a precision of $L/(256N_c)$. The force calculation 
+(e.g. softening length) should be configured much finer than this 
+resolution, discussed in later sections. On the other hand, particle 
+position can also be stored as 2-byte (16 bits) integers to increase the resolution. 
+In this case, each coarse cell is divided into $2^{16}=65536$ bins and the 
+position precision is $L/(65536N_c)$, precise enough compared to using
+4-byte global coordinates, see later results. We denote this case ``x2'' and 
+denote using 1-byte integers for positions ``x1''.
+
+We collectively write the general position conversion formulae
+\begin{equation}
+	\chi_d=\left[2^{8n_\chi}(x_d-\left[x_d\right])\right]-2^{8n_\chi-1},
+\end{equation}
+\begin{equation}
+	x_d=(n_c-1)+2^{-8n_\chi}\left(\chi_d+2^{8n_\chi-1}+1/2\right),
+\end{equation}
+where $[\ ]$ is the operator to take the integer part.
+$n_\chi\in\{1,2\}$ is the number of bytes used for each integer, $x_d$ and $\chi_d$ are floating and integer version of the coordinate.
+Similarly, in section \ref{s.velocity}, the velocity counterpart of them are $n_\nu=1,2$, $v_d$ and $\nu_d$.
+Position resolution for $n_\chi$-byte integer, ``x$n_\chi$'', is $2^{-8n_\chi}L/N_c$.
 
 \subsection{Particle velocity storage}\label{s.velocity}
-Velocities, in each dimension, are compressed to 2-byte integers representing their index within a histogram containing $2^{16}$ equally spaced bins ranging from $-v_{\rm max}$ to $v_{\rm max}$, where $v_{\rm max}$ is the maximum absolute particle velocity.
-$v_{\rm max}$ can be derived from either computation or prediction.
-We prefer the latter case.
-In cosmological simulations, velocity distribution of matter fluids can be predicted by linear theory, given a model and redshift $z$ (time of the simulation).
-Then by any time step, we have a cumulative distribution function (CDF) $F(|\bs{v}|)$ of particles' velocity (see appendix A).
-We, instead, equally divide the $y$-axis of this CDF into $2^{16}$ bins, and the 2-byte integer shows which velocity bin the particle locates, in each dimension.
-This step does finer sampling on low-velocity end in velocity space and increases the simulation accuracy.
-This CDF may introduce velocity cutoffs, that particle cannot exceed some certain $\bs{v}_{\rm max}(z)$ given by the model.
-This process prevents the nonphysical scattering behaviours of particles, which do not exist in fluids.
-Practically, simulation results are very insensitive to the choice of velocity distribution.
-Even a top-hat distribution in range $-v_{\rm max}$ to $v_{\rm max}$ give accurate results.
-One can just assume a Gaussian distribution and whose standard deviation is given by the model and redshift $\sigma(\bs{\theta},z)$.
-Proper choice of model only optimizes the velocity space sampling, but does not affect the physics.
+Similarly, actual velocity in $d$th dimension $v_d$ is decomposed into an 
+averaged velocity field on the same coarse grid $v_c$ and a residual $\Delta v$ 
+relative to this field: $v_d=v_c+\Delta v$. $v_c$ is always recorded and 
+kept updated, and occupies negligible memory. We then divide velocity space 
+$\Delta v$ into uneven bins, and use a $n_\nu$-byte integer to indicate
+which $\Delta v$ bin the particle is located.
 
-\begin{figure}[t]
-\centering
-  \includegraphics[width=1.0\linewidth]{vdisp.pdf}
- \caption{Velocity dispersion.}
-\label{fig.tile}
-\end{figure}
+The reason why we use uneven bins is that, slower particles are more 
+abundant compared to faster ones, and one should better resolve slower 
+particles tracing at least linear evolution. On the other hand, there could 
+be extreme scattering particles (in case of particle-particle force), and we 
+can safely ignore or less resolve those nonphysical particles. One of the 
+solution is that, if we know the probability distribution function (PDF) 
+$f(\Delta v)$ we divide its cumulative distribution function (CDF) $F(\Delta 
+v)\in(0,1)$ into $2^{8n_\nu}$ bins to determine the boundary of $\Delta v$ 
+bins, and particles should evenly distribute in the corresponding uneven $
+\Delta v$ bins. Practically we find that either $f(v_d)$ or $f(\Delta v)$ is 
+close to Gaussian, so we can use Gaussian CDF, or any convenient analytic 
+functions which are close to Gaussian, to convert velocity between real 
+numbers and integers.
+
+The essential parameter of the velocity distribution is its variance. On 
+non-linear scale, the velocity distribution function is non-Gaussian. 
+However, to the first order approximation, we simply assume it as Gaussian 
+and characterized by the variance
+\begin{eqnarray}\label{eq.vdisp}
+\sigma_v(a,k) = \frac{1}{3}  (a H f D)^2 \int_0^k d^3q \frac{P_{L}(q)}{q^2},
+\end{eqnarray}
+where $a(z)$ is the scale factor, $H(z)$ the Hubble parameter, $D$ is the 
+linear growth factor, $f=d \ln D/d\ln a$, and $P_L(k)$ is the linear power 
+spectrum of density contrast at redshift zero.
+$\sigma_v(a,k)$ is a function of cosmic evolution $a$ and a smoothing scale 
+$k$, or $r$. $\Delta v$
+is the velocity dispersion relative to the coarse grid, so we approximate 
+its variance as
+\begin{equation}\label{eq.vdelta}
+	\sigma^2_{\Delta}(a)=\sigma^2_v(a,r_c)-\sigma^2_v(a,r_p), 
+\end{equation}
+where $r_c$ is the scale of
+coarse grid, and $r_p$ is the scale of average particle separation. In each
+dimension of 3D velocity field, we use $\sigma^2_{\Delta}(a)/3$ according 
+to the equipartition therorem. On different scales, we measure the 
+statistics of $v_d$, $v_c$ and $\Delta v$ and find good agreement with the 
+above model.
+
+The simulation results are very insensitive if we manually tune the variance of the model $\sigma_\Delta$ within an order of magnitude.
+However, the method of using uneven bins gets much better results than simply using equal bins between minimum and maximum values $[\min(\Delta v),\max(\Delta v)]$.
+So, one can safely use a standard $\Lambda$CDM (Cold Dark Matter with a cosmological constant) for slightly different cosmological models, in equation (\ref{eq.vdisp}).
+In {\tt CUBE}, the velocity conversion takes the formula
+\begin{eqnarray}\label{eq.vconvert}
+	&&\nu_d=\left\lfloor(2^{8n_\nu}-1)\pi^{-1}\tan^{-1}\left(v_d\sqrt{\pi/2\sigma_\Delta^2}\right)\right\rfloor,\\
+	&&v_d=\tan\left(\frac{\pi \nu_d}{2^{8n_\nu}-1}\right)\sqrt{2\sigma_\Delta^2/\pi},
+\end{eqnarray}
+where $\lfloor\ \rfloor$ is the operator to take the nearest integer. 
+Tangent functions are convenient and compute very fast. Compared to error 
+functions used in Gaussian case, they take the same variance at $v_d=0$ but 
+resolve high velocities relatively better.
+Note again that proper choice of conversion formulae and $\sigma_\Delta$ 
+only optimizes the velocity space sampling, but does not affect the physics.
+
+Initially, particles are generated by initial condition generator, at a higher redshift.
+The coarse grid density field $v_c$ is also generated at this step by averaging all particles in the coarse cell.
+A global $\sigma_\Delta$ is calculated by equation (\ref{eq.vdelta}), where linear approximation is hold.
+Then velocities are stored by equation (\ref{eq.vconvert}).
+During the simulation, $v_c$ is updated every time step, and a nonlinear $\sigma_\Delta$ is measured directly from the simulation, and can be simply used in the next time step, after scaled by the ratio of growth factors between two adjacent time steps.
+More details see section \ref{ss.code}.
 
@@
 Codimensions can do communications between images.}.
-Particles' phase coordinates {\tt x} and {\tt v} are copied between tiles and images, when necessary.
+Particles' phase coordinates {\tt xp} and {\tt vp} are copied between tiles and images, when necessary.
 Figure \ref{fig.tile} shows the spacial decomposition in a 2-dimensional analogy, with ${\tt nn=2}$ and ${\tt nnt=2}$.
+\begin{figure}
+\centering
+  \includegraphics[width=0.95\linewidth]{fig_tile.pdf}
+ \caption{Spacial decomposition.}
+\label{fig.tile}
+\end{figure}
+
+\begin{figure}
+\centering
+  \includegraphics[width=1.0\linewidth]{vdisp.pdf}
+ \caption{Velocity dispersion.}
+\label{fig.tile}
+\end{figure}
 
@@
 
-\subsection{Code overview}
+\subsection{Code overview}\label{ss.code}
 By using the previous compact format storage, we modified the {\tt CUBEP3M} and utilized it in the outputs TianNu simulation.
@@
 \section{A. Velocity distribution function}
-To optimize the storage of the velocity field, we need to estimate the velocity dispersion at each snapshot.
-At non-linear scale, the velocity distribution function is surely non-Gaussian.
-However, to the first order approximation, we simply assume it as Gaussian and characterized by the variance
-\begin{eqnarray}
-\sigma_v = \frac{1}{3}  (a H f D)^2 \int d^3q \frac{P_{L}(q)}{q^2}
-\end{eqnarray}
-where $a(z)$ is the scale factor, $H(z)$ the Hubble parameter, $D$ is the linear growth factor, $f=d \ln D/d\ln a$, 
-and $P_L(k)$ is the linear power spectrum of density contrast at redshift zero.
 

### 2017-09-22 ms_caf/ms_caf.tex
code overview section

@@
 \usepackage[english]{babel}
-\usepackage{amsmath}
+\usepackage{amsmath,color}
 
@@
 \newcommand{\Msun}{M_\odot}
+
+\newcommand{\tcb}{\textcolor{blue}}
+\newcommand{\tcr}{\textcolor{red}}
 
@@
 
-\section{Introduction}
+\section{Introduction}\label{s.intro}
 $N$-body simulation is a powerful tool to solve the highly nonlinear dynamic problems 
@@
 
-We present a new $N$-body simulation code {\tt CUBE}, using as low as 6 byte per particle 
-(here after we use ``bpp'' referring to ``byte per particle''). We show that it gives 
-accurate results in cosmological LSS simulations. The algorithm is presented in \S 2, and 
-a comparison between this method and traditional method is shown in \S 3. Discussions and 
-conclusions are in \S 4.
+We present a new $N$-body simulation code {\tt CUBE}, using as low as 6 byte
+per particle (here after we use ``bpp'' referring to ``byte per particle'').
+We show that it gives accurate results in cosmological LSS simulations.
+The method is presented in \S\ref{s.method}, and a comparison between this method
+and traditional method is shown in \S\ref{s.results}. Discussions and 
+conclusions are in \S\ref{s.discussion}.
 
-\section{Algorithm}
+\section{Method}\label{s.method}
 The most memory consuming part of a $N$-body simulation is usually the phase space 
@@
 be used to store each particles' 3D position and velocity vectors. {\tt CUBEP3M}, an 
-example of a memory-lite parallel $N$-body code, can use as low as 40 bpp in 
+example of a memory-efficient parallel $N$-body code, can use as low as 40 bpp in 
 sacrificing computing speed \citep{2013MNRAS.436..540H}. This includes the phase 
@@
 with an integer based storage, reduces the basic memory usage from 28 bpp down to
-6 bpp, described as following subsections.
+6 bpp, described as following \ref{ss.position} and \ref{ss.velocity}. The algorithm is described in \ref{ss.overview}.
 
-\subsection{Particle position storage}
+\subsection{Particle position storage}\label{ss.position}
 We construct a uniform mesh throughout the space and each particle belongs 
@@
 We collectively write the general position conversion formulae
-\begin{equation}
+\begin{equation}\label{eq.chi}
 	\chi_d=\left[2^{8n_\chi}(x_d-\left[x_d\right])\right]-2^{8n_\chi-1},
 \end{equation}
-\begin{equation}
+\begin{equation}\label{eq.x}
 	x_d=(n_c-1)+2^{-8n_\chi}\left(\chi_d+2^{8n_\chi-1}+1/2\right),
@@
 $n_\chi\in\{1,2\}$ is the number of bytes used for each integer, $x_d$ and $\chi_d$ are floating and integer version of the coordinate.
-Similarly, in section \ref{s.velocity}, the velocity counterpart of them are $n_\nu=1,2$, $v_d$ and $\nu_d$.
-Position resolution for $n_\chi$-byte integer, ``x$n_\chi$'', is $2^{-8n_\chi}L/N_c$.
+The velocity counterpart of them are $n_\nu=1,2$, $v_d$ and $\nu_d$.
+$n_c=1,2,...$ is the coarse grid index, given by the ordering of particles and a integer based particle density field (see \ref{sss.spatial_decomposition}).
+The position resolution for $n_\chi$-byte integer, ``x$n_\chi$'', is $2^{-8n_\chi}L/N_c$.
 
-\subsection{Particle velocity storage}\label{s.velocity}
+As a $n_\chi=1$, 1D ($d=1$), 4-coarse-cell ($N_c=4$) example, if $$\chi_1=(-128,127,0,60),$$ particle number density $${\rho_c^{\rm 1D}}=(1,0,2,1),$$ then in unit of coarse cells, $$x_1=(0.001953125, 2.998046875, 2.501953125, 3.736328125).$$
+
+\begin{figure}
+\centering
+  \includegraphics[width=1.0\linewidth]{vdisp.pdf}
+ \caption{\tcb{please check units}}
+\label{fig.vdisp}
+\end{figure}
+
+\subsection{Particle velocity storage}\label{ss.velocity}
 Similarly, actual velocity in $d$th dimension $v_d$ is decomposed into an 
@@
 and characterized by the variance
-\begin{eqnarray}\label{eq.vdisp}
-\sigma_v(a,k) = \frac{1}{3}  (a H f D)^2 \int_0^k d^3q \frac{P_{L}(q)}{q^2},
-\end{eqnarray}
+\begin{equation}\label{eq.vdisp}
+	\sigma_v(a,k) = \frac{1}{3}  (a H f D)^2 \int_0^kd^3q\frac{P_{L}(q)}{q^2},
+\end{equation}
 where $a(z)$ is the scale factor, $H(z)$ the Hubble parameter, $D$ is the 
@@
 $\sigma_v(a,k)$ is a function of cosmic evolution $a$ and a smoothing scale 
-$k$, or $r$. $\Delta v$
+$k$, or $r$ (see Figure \ref{fig.vdisp}). $\Delta v$
 is the velocity dispersion relative to the coarse grid, so we approximate 
@@
 
+\begin{figure}
+\centering
+  \includegraphics[width=0.95\linewidth]{fig_tile.pdf}
+ \caption{Spacial decomposition in {\tt CUBE} in a 2D analogy.
+In this example, there are 2 images per dimension (${\tt nn}=2$), and 2 tiles per image per dimension (${\tt nnt}=2$).
+The orange boxes show the overlapped physical+buffer regions, inside of which one physical regions is indicated with green.}
+\label{fig.tile}
+\end{figure}
+
 The simulation results are very insensitive if we manually tune the variance of the model $\sigma_\Delta$ within an order of magnitude.
@@
 In {\tt CUBE}, the velocity conversion takes the formula
-\begin{eqnarray}\label{eq.vconvert}
-	&&\nu_d=\left\lfloor(2^{8n_\nu}-1)\pi^{-1}\tan^{-1}\left(v_d\sqrt{\pi/2\sigma_\Delta^2}\right)\right\rfloor,\\
-	&&v_d=\tan\left(\frac{\pi \nu_d}{2^{8n_\nu}-1}\right)\sqrt{2\sigma_\Delta^2/\pi},
-\end{eqnarray}
+\begin{equation}\label{eq.nu}
+	\nu_d=\left\lfloor(2^{8n_\nu}-1)\pi^{-1}\tan^{-1}\left(v_d\sqrt{\pi/2\sigma_\Delta^2}\right)\right\rfloor,
+\end{equation}
+\begin{equation}\label{eq.v}
+	v_d=\tan\left(\frac{\pi\nu_d}{2^{8n_\nu}-1}\right)\sqrt{2\sigma_\Delta^2/\pi},
+\end{equation}
 where $\lfloor\ \rfloor$ is the operator to take the nearest integer. 
@@
 A global $\sigma_\Delta$ is calculated by equation (\ref{eq.vdelta}), where linear approximation is hold.
-Then velocities are stored by equation (\ref{eq.vconvert}).
+Then velocities are stored by equation (\ref{eq.nu}).
 During the simulation, $v_c$ is updated every time step, and a nonlinear $\sigma_\Delta$ is measured directly from the simulation, and can be simply used in the next time step, after scaled by the ratio of growth factors between two adjacent time steps.
-More details see section \ref{ss.code}.
+More details see section \ref{sss.algorithm}.
 
-\subsection{Spacial decomposition}
+\subsection{Code overview}\label{ss.overview}
+{\tt CUBE} uses a 2-level PM force calculation, same as {\tt CUBEP3M}, However, in order to apply the integer based format to the $N$-body simulation, substantial structural changes need to be done.
+{\tt CUBE} is written in Coarray Fortran, where Coarray features implement MPI communication between computation nodes/images\footnote{Images are the concept of computing nodes or MPI tasks in Coarray Fortran.
+We use this terminology in this paper.}.
+The algorithm is described in this language.
+
+\subsubsection{Spacial decomposition}\label{sss.spatial_decomposition}
 {\tt CUBE} uses cubic decomposition structures.
-The global simulation volume is decomposed into ${\tt nn}^3$ cubic sub-volumes, and each of these are assigned to a coarray {\it image}\footnote{Images are the concept of computing nodes or MPI tasks in Coarray Fortran.
-We use this terminology in this paper.}.
+The global simulation volume is decomposed into ${\tt nn}^3$ cubic sub-volumes, and each of these are assigned to a coarray {\it image}.
 Inside of an image, the sub-volume is further decomposed into ${\tt nnt}^3$ cubic {\it tiles}.
-Each tile is surrounded by a {\it buffer} region which is usually ${\tt ncb}$ coarse cells thick.
-The buffer is designed for two reasons: (1) computing the fine mesh force, whose cut-off ${\tt nforce\_cutoff}\leq{\tt ncb}$, and (2) collecting all possible particles travelling from a tiles buffer region to its center, {\it physical} region.
-Thus, a integer-based coarse mesh density is defined as
+Each tile is surrounded by a {\it buffer} region which is ${\tt ncb}$ coarse cells thick.
+The buffer is designed for two reasons: (1) computing the fine mesh force, whose cut-off ${\tt nforce\_cutoff}\leq{\tt ncb}$, and (2) collecting all possible particles travelling from a tile's buffer region to its center, {\it physical} region.
+Thus, a integer-based coarse mesh particle number density ($\rho_c$) is defined as
 \begin{equation*}
@@
 \end{equation*}
-where ${\tt nex}={\tt nt}+2{\tt ncb}$ covers the buffer region on both sides, {\tt nnt} is the tile dimemsions, and {\tt nn} is the image {\it codimensions}\footnote{Coarray Fortran concept.
+where ${\tt nex}={\tt nt}+2\times{\tt ncb}$ covers the buffer region on both sides, {\tt nnt} is the tile dimemsions, and {\tt nn} is the image {\it codimensions}\footnote{Coarray Fortran concept.
 Codimensions can do communications between images.}.
-Particles' phase coordinates {\tt xp} and {\tt vp} are copied between tiles and images, when necessary.
+In the buffer region, phase space coordinates of particles {\tt xp} and {\tt vp} are copied from adjacent tiles and images, when necessary.
 Figure \ref{fig.tile} shows the spacial decomposition in a 2-dimensional analogy, with ${\tt nn=2}$ and ${\tt nnt=2}$.
-\begin{figure}
-\centering
-  \includegraphics[width=0.95\linewidth]{fig_tile.pdf}
- \caption{Spacial decomposition.}
-\label{fig.tile}
-\end{figure}
 
-\begin{figure}
-\centering
-  \includegraphics[width=1.0\linewidth]{vdisp.pdf}
- \caption{Velocity dispersion.}
-\label{fig.tile}
-\end{figure}
+The particle position and velocity arrays {\tt xp} and {\tt vp} (equivalent to $\chi_d$ and $\nu_d$ in \ref{ss.position} and \ref{ss.velocity} respectively) are required to be sorted according to the same memory layout of {\tt rho\_c}, such that $n_c$ and thus global positions of particles $x_d$ can be obtained by equation (\ref{eq.x}).
+
+\subsubsection{Algorithm}\label{sss.algorithm}
+Figure \ref{fig.code} shows the overall structure of the code and these subroutines are described in following paragraphs.
+
+{\tt initialize} and {\tt read\_particles} -- The subroutine {\tt initialize} creates necessary FFT plans and read in configuration files telling the program at which redshifts we need to do checkpoints, halofinds, or stop the simulation.
+Force kernels are also created or read in here.
+In {\tt read\_particles}, for each image, we read in all particles in {\it physical} regions of every tile (indicated in Figure \ref{fig.tile}), i.e., their initial positions {\tt xp}, velocities {\tt vp} in the compressed format, and their corresponding, physical part of the number density field {\tt rho\_c}.
+These are obtained by the initial condition generator (see Appendix A).
+Because physical regions of tile are {\it complete} and {\it disjoint} in space, particles at this stage are also complete and disjoint.
+We call it ``disjoint state''.
+At this stage, {\tt rho\_c}'s buffer regions of each tile are 0, and the elements of {\tt xp} and {\tt vp} beyond physical number can be arbitrary and are not used.
+
+{\tt buffer\_density}, {\tt buffer\_x} and {\tt buffer\_v} -- In order to use the integer based format, {\tt xp} and {\tt vp} must always be ordered, and their number density field {\tt rho\_c} must always be present.
+In updating arrays of {\tt xp} and {\tt vp} (not simultaneously) of a local tile, a vicinity (buffer) region of the physical region is also needed.
+First, buffer regions of {\tt rho\_c} is synchronized between tiles and images by subroutine {\tt buffer\_density}.
+Then, by subroutines {\tt buffer\_x} and {\tt buffer\_v} respectively, {\tt xp} and {\tt vp} are updated to contain common, buffered particles, in an order according to the new, buffered {\tt rho\_c}.
+We call this stage ``buffered state''.
+After {\tt particle\_initialize} is done by all images (synchronized by ``{\tt sync all}'', which is equivalent to a {\tt mpi\_barrier}), these three subroutines are called and particles are converted from disjoint state to buffered state.
 
@@
       call initialize
-      call particle_initialize
+      call read_particles
       sync all
       call buffer_density
-      call buffer_x
-      call buffer_v
+      call buffer_xp
+      call buffer_vp
       do
         call timestep
-        call update_x
-        call buffer_density
-        call buffer_x
-        call PM ! update_v
-        call buffer_v
+        call update_xp
+sys
+        call buffer_xp
+        call update_vp
+        call buffer_vp
         if(checkpoint_step) then
+           call update_xp
            call checkpoint
            if (final_step) exit
+           call buffer_density
+           call buffer_xp
+           call buffer_vp
         endif
@@
       call finalize
-    endprogram
+    end
 \end{verbatim}
@@
 
-\subsection{Code overview}\label{ss.code}
-By using the previous compact format storage, we modified the {\tt CUBEP3M} and utilized it in the outputs TianNu simulation.
-On average, 9.125 bpp is used for all outputs and saves 62 \% hard drive space on Tianhe-2.
-The compact version of {\tt CUBEP3M}, named {\tt CUBE}, accomplished the compact format in memory, and the code keeps only integer format for storing particles' phase space coordinates.
-{\tt CUBE} is written in Coarray Fortran, where the coarray features are used to do MPI communications.
-Here we review the methodology of solving this integer based $N$-body problem.
-Like its predecessor {\tt CUBEP3M}, {\tt CUBE} uses two-level mesh grids for solving the Poisson equations.
-The advantages of this are described in \cite{2013MNRAS.436..540H}.
+{\tt timestep} --
+We operate a Runge-Kutta 2 method for time integration.
+i.e., we update position ($D$=drift) and velocity ($K$=kick) every half time step.
+For $n$ time steps, the operation would be $(DKKD)^n$ which is 2nd order accurate.
+The actual simulation applies varied time steps.
+In each iteration of the main loop, we firstly call {\tt timestep}, where a increment of time {\tt dt} is controlled by particles' maximum velocities, accelerations, and cosmic expansions.
 
-Figure \ref{fig.code} shows the overall structure of the code.
-Subroutine {\tt initialize} creates necessary FFT plans and read in configuration files telling the program at which redshifts we need to do checkpoints, halofinds, or stop the simulation.
-Force kernels are also created or read in here.
-In {\tt particle\_initialize}, for each image, we read in initial positions and velocities in the compressed format.
-A initial condition generator is also written in the compressed format (see Appendix B).
-When this step is done by all images (synchronized by ``{\tt sync all}'', which is equivalent to a {\tt mpi\_barrier}), the density fields {\tt rho\_c} are buffered between tiles (communications between images are needed) by subroutine {\tt buffer\_density}.
-Then, particles' phase space coordinates {\tt x} and {\tt v} are also buffered, by {\tt buffer\_x} and {\tt buffer\_v} respectively.
-In these steps, {\tt x} and {\tt v}'s ordering is changed according to the updated {\tt rho\_c}.
+{\tt update\_xp} --
+According to {\tt dt}, subroutine {\tt update\_xp} updates the positions of particles in a ``{\it gather}'' algorithm (in contrast, {\tt CUBEP3M} uses ``scatter'' algorithm) tile by tile.
+Because each tile is in the buffered state with buffer depth {\tt ncb}, we are able to collect all possible particles whose {\tt vp}*{\tt dt}$<${\tt ncb} from physical+buffer region to its physical region.
+In order to keep particles ordered, we have to execute ${\tt xp}={\tt xp}+{\tt vp}*{\tt dt}$ twice, first time to obtain an updated density field on the tile, {\tt rho\_c\_new}, and second time to generate a new, local particle list {\tt xp\_new} and {\tt vp\_new} on the tile.
+The reason for this repeated execution is the dependence of {\tt x\_new} and {\tt v\_new}'s value and ordering on both the old {\tt xp}, {\tt vp} and the new {\tt rho\_c\_new}.
+However, ${\tt xp}={\tt xp}+{\tt vp}*{\tt dt}$ scales as $o(N)$ and is computational inexpensive.
+Although {\tt xp\_new} and {\tt vp\_new} arrays take extra memory, they do not appear simultaneously for multiple tiles, and compared to {\tt xp} and {\tt vp}, the extra memory overhead is by a factor of $1/{\tt nnt}^3$.
+This requires the ``gather'' algorithm to move particles -- as we synchronize the correct \{{\tt rho\_c}, {\tt xp}, {\tt vp}\} in the buffer region of each tile, and the particles in the extended (physical+buffer) region are a superset of particles locating in the physical region in the next time step, given that ${\tt vp}*{\tt dt}$ is not greater than the buffer depth.
+In {\tt update\_xp}, \{{\tt rho\_c\_new}, {\tt xp\_new}, {\tt vp\_new}\} can be set to collect only the physical region of the tile, with the rest of this subset discarded.
+At the end, \{{\tt rho\_c}, {\tt xp}, {\tt vp}\} is replaced by \{{\tt xp\_new}, {\tt vp\_new}, {\tt rho\_c\_new}, {\tt vfield}\}, and the memory of latter is freed up.
+After this step, the tile does not contain buffered regions, being a disjoint state.
 
-In each iteration of the main loop, we firstly call {\tt timestep}, where a increment of time {\tt dt} is controlled by particles' maximum velocities, accelerations, and cosmic expansions.
-According to {\tt dt}, subroutine {\tt update\_x} updates the positions of particles in a ``{\it gather}'' algorithm.
-${\tt x}={\tt x}+{\tt v}*{\tt dt}$ is executed twice, first time to determine a updated density field on the tile, {\tt rho\_c\_new}, and second time to generate a new particle list {\tt x\_new} and {\tt v\_new} on the tile.
-The reason for this repeated execution is the dependence of {\tt x\_new} and {\tt v\_new}'s value and ordering on both the old {\tt x}, {\tt v} and the new {\tt rho\_c\_new}.
-However, ${\tt x}={\tt x}+{\tt v}*{\tt dt}$ scales as $o(N)$ and is computational inexpensive.
-Although {\tt x\_new} and {\tt v\_new} arrays take extra memory, they do not appear simultaneously for multiple tiles, and compared to {\tt x} and {\tt v}, the extra memory overhead is by factor of $1/{\tt nnt}^3$.
-This requires the ``gather'' algorithm to move particles\footnote{In contrast, {\tt CUBEP3M} uses a ``{\it scatter}'' algorithm.} -- as we synchronize the correct \{{\tt rho\_c}, {\tt x}, {\tt v}\} in the buffer region of each tile, and the particles in the extended (physical+buffer) region are a superset of particles locating in the physical region in the next time step, given that ${\tt v}*{\tt dt}$ is not greater than the buffer depth.
-In {\tt update\_x}, \{{\tt rho\_c\_new}, {\tt x\_new}, {\tt v\_new}\} can be set to collect only the physical region of the tile, with the rest of this subset discarded.
-At the end, \{{\tt rho\_c}, {\tt x}, {\tt v}\} is replaced by \{{\tt rho\_c\_new}, {\tt x\_new}, {\tt v\_new}\}, and the memory of latter is freed up.
+{\tt update\_vp} -- 
+After drift $D$, we need to update velocities (kick $K$) of particles in physical region of each tile.
+Before this step, we need to call {\tt buffer\_density} and {\tt buffer\_xp} in order that particle positions are in buffered state.
+{\tt CUBE} uses a 2-level particle mesh scheme \citep{2013MNRAS.436..540H}.
+Local fine forces have a force cutoff ${\tt nforce\_cutoff}\leq{\tt ncb}$.
+So, if we apply a fine-grid particle-mesh on an extended tile with buffered-depth {\tt ncb} (see Figure \ref{fig.tile} regard it periodic), the force on physical regions does not depend on the false periodic boundary assumption, and the resulting fine force and velocity update on physical region is correct.
 
-Next, {\tt buffer\_density} and {\tt buffer\_x} are called again to synchronize the updated particle positions in order that tile-based local fine forces are computed correctly.
+The compensating coarse grid force is globally computed by using a coarser (usually by factor of 4) mesh by dimensional splitting -- a distributed-memory cubic decomposed 3D coarse density field is interpolated by particles, and we Fourier transform data in consecutive three dimensions with global transposition in between (known as the pencil decomposition).
+After the multiplication of force kernels, the inverse transform takes place to get the cubic distributed coarse force field, upon which velocities are updated again.
 
-A standard particle-mesh scheme is then performed in the {\tt PM} subroutine.
-First, by looping over tiles, fine force is calculated and velocities are updated on the physical region of each tile.
-The memory overhead of fine-mesh arrays are well controlled by $1/{\tt nnt}^3$.
-Next, global coarse force is solved by using a coarser (usually by factor of 4) mesh by dimensional splitting -- a distributed-memory cubic decomposed 3D coarse density field is interpolated by particles, and we Fourier transform data in consecutive three dimensions with global transposition in between (known as the pencil decomposition).
-After the multiplication of force kernels, the inverse transform takes place to get the cubic distributed coarse force field, upon which velocities are updated again.
-The memory usage of coarse arrays are negligible.
 An optional particle-particle (PP) force can be called to increase the force resolution and the velocities are updated last time according to this.
-As discussed in \S\ref{s.velocity}, updating velocities use model predicted velocity distribution, thus ${\tt v}={\tt v}+{\tt dv}$ is executed only once, scaling directly to the velocity distribution of the next time step ${\tt d}+{\tt dt}$.
+\tcb{v field}.
 During each of these force calculations, maximum accelerations are collected serving for the {\tt timestep} in the next iteration, controlling next {\tt dt}.
+
+After all possible velocity updates, {\tt vp} in physical regions are correctly updated.
+Particle locations in the buffer regions has updated before {\tt PM} and remained unchanged.
+So we simply call {\tt buffer\_v} again such that the {\tt update\_x} in the next iteration will be done correctly.
+We call {\tt update\_vp} to convert {\tt vp} into buffered state.
+
+{\tt checkpoint} -- 
+If a desired redshift is reached, we execute the last drift step in the $(DKKD)^n$ operation by {\tt update\_xp}, and call {\tt checkpoint} to save the disjoint state of \{{\tt xp\_new}, {\tt vp\_new}, {\tt rho\_c\_new}, {\tt vfield}\} on disk.
+Related operations, like run-time halo finder, projections are also done at this point.
+If final desired redshift is reached, {\tt final\_step} let us exit the main loop.
+These corresponding logical variables are controlled in {\tt timestep}.
+
+{\tt finalize} --
+Finally, in {\tt finalize} subroutine we destroy all the FFT plans and finish up any timing or statistics taken in the simulation.
 
@@
 
-By far, particle velocities in the physical regions of each tile are updated.
-Particle locations in the buffer regions has updated before {\tt PM} and remained unchanged.
-So we simply call {\tt buffer\_v} again, such that the {\tt update\_x} in the next iteration will be done correctly.
-
-If a desired redshift is reached, we call {\tt checkpoint} and/or exit the main loop.
-The corresponding logical variables are controlled in {\tt timestep}.
-Finally, in {\tt finalize} subroutine we destroy all the FFT plans and finish up any timing or statistics taken in the simulation.
-
 \begin{figure}[t]
@@
 
-\subsection{Memory layout}
+\subsection{Memory layout}\label{ss.memory}
+The memory overhead of fine force calculation is controlled by $(1/{\tt nnt})^3$. The memory usage of coarse arrays are negligible.
+
 Here we summarize the memory usage in unit of bpp (byte per particle).
 
-\section{Results}
+\section{Results}\label{s.results}
 We use the same initial conditions, and use {\tt CUBEP3M} and {\tt CUBE} to simulate the large scale structure formation separately and compare their results.
@@
 
-\section{Discussion and Conclusion}
+\section{Discussion and Conclusion}\label{s.discussion}
 Cosmological simulation, PM method.
@@
 \appendix
-\section{A. Velocity distribution function}
 
-\section{B. Initial condition generator}
+\section{A. Initial condition generator}
 
 \acknowledgements
-Thanks.
 

### 2017-09-24 ms_caf/ms_caf.tex
revised algorithm section.

@@
   \includegraphics[width=1.0\linewidth]{vdisp.pdf}
- \caption{\tcb{please check units}}
+ \caption{\tcb{Xin, could you please check the units of both $y$-axis?}}
 \label{fig.vdisp}
@@
 averaged velocity field on the same coarse grid $v_c$ and a residual $\Delta v$ 
-relative to this field: $v_d=v_c+\Delta v$. $v_c$ is always recorded and 
+relative to this field: 
+\begin{equation}
+	v_d=v_c+\Delta v.
+\end{equation}
+$v_c$ is always recorded and 
 kept updated, and occupies negligible memory. We then divide velocity space 
@@
 \begin{equation}\label{eq.vdisp}
-	\sigma_v(a,k) = \frac{1}{3}  (a H f D)^2 \int_0^kd^3q\frac{P_{L}(q)}{q^2},
+	\tcb{ \sigma_v(a,k) = \frac{1}{3}  (a H f D)^2 \int_0^kd^3q\frac{P_{L}(q)}{q^2}},
 \end{equation}
@@
 The simulation results are very insensitive if we manually tune the variance of the model $\sigma_\Delta$ within an order of magnitude.
-However, the method of using uneven bins gets much better results than simply using equal bins between minimum and maximum values $[\min(\Delta v),\max(\Delta v)]$.
-So, one can safely use a standard $\Lambda$CDM (Cold Dark Matter with a cosmological constant) for slightly different cosmological models, in equation (\ref{eq.vdisp}).
+However, in $n_\nu=1$ case, the method of using uneven bins gets much better results than simply using equal bins between minimum and maximum values $[\min(\Delta v),\max(\Delta v)]$.
+So, one can safely use a standard $\Lambda$CDM (cold dark matter with a cosmological constant as dark energy) for slightly different cosmological models, in equation (\ref{eq.vdisp}).
 In {\tt CUBE}, the velocity conversion takes the formula
 \begin{equation}\label{eq.nu}
-	\nu_d=\left\lfloor(2^{8n_\nu}-1)\pi^{-1}\tan^{-1}\left(v_d\sqrt{\pi/2\sigma_\Delta^2}\right)\right\rfloor,
+	\nu_d=\left\lfloor(2^{8n_\nu}-1)\pi^{-1}\tan^{-1}\left((v_d-v_c)\sqrt{\pi/2\sigma_\Delta^2}\right)\right\rfloor,
 \end{equation}
 \begin{equation}\label{eq.v}
-	v_d=\tan\left(\frac{\pi\nu_d}{2^{8n_\nu}-1}\right)\sqrt{2\sigma_\Delta^2/\pi},
+	v_d=v_c+\tan\left(\frac{\pi\nu_d}{2^{8n_\nu}-1}\right)\sqrt{2\sigma_\Delta^2/\pi},
 \end{equation}
@@
 Tangent functions are convenient and compute very fast. Compared to error 
-functions used in Gaussian case, they take the same variance at $v_d=0$ but 
+functions used in Gaussian case, they take the same variance at $\nu_d=0$ but 
 resolve high velocities relatively better.
@@
 The global simulation volume is decomposed into ${\tt nn}^3$ cubic sub-volumes, and each of these are assigned to a coarray {\it image}.
-Inside of an image, the sub-volume is further decomposed into ${\tt nnt}^3$ cubic {\it tiles}.
+Inside of an image, the sub-volume is further decomposed into ${\tt nnt}^3$ cubic {\it tiles}, which is a essential unit in the calculations.
 Each tile is surrounded by a {\it buffer} region which is ${\tt ncb}$ coarse cells thick.
-The buffer is designed for two reasons: (1) computing the fine mesh force, whose cut-off ${\tt nforce\_cutoff}\leq{\tt ncb}$, and (2) collecting all possible particles travelling from a tile's buffer region to its center, {\it physical} region.
-Thus, a integer-based coarse mesh particle number density ($\rho_c$) is defined as
-\begin{equation*}
-{\tt integer(1)\ rho\_c(nex,nex,nex,nnt,nnt,nnt)[nn,nn,*]}\\
-\end{equation*}
+The buffer is designed for two reasons: (1) computing the fine mesh force, whose cut-off length ${\tt nforce\_cutoff}\leq{\tt ncb}$, and (2) collecting all possible particles travelling from a tile's buffer region to its center, {\it physical} region.
+Figure \ref{fig.tile} shows the spacial decomposition in a 2-dimensional analogy, with ${\tt nn=2}$ and ${\tt nnt=2}$.
+According to this spatial decomposition, and as discussed in the last two subsections, we declare $\{\rho_c,\chi_d,v_c,\nu_d\}$, or in variable expression $\{{\tt rho\_c},{\tt xp},{\tt vfield},{\tt vp}\}$ as follows:
+\begin{eqnarray*}
+	&&{\tt integer(1)\ rho\_c(nex,nex,nex,nnt,nnt,nnt)[nn,nn,*]}\\
+	&&{\tt integer(}n_\chi{\tt )\ xp(npmax)[nn,nn,*]}\\
+	&&{\tt real(4)\ vfield(nex,nex,nex,nnt,nnt,nnt)[nn,nn,*]}\\
+	&&{\tt integer(}n_\nu{\tt )\ vp(npmax)[nn,nn,*]}
+\end{eqnarray*}
 where ${\tt nex}={\tt nt}+2\times{\tt ncb}$ covers the buffer region on both sides, {\tt nnt} is the tile dimemsions, and {\tt nn} is the image {\it codimensions}\footnote{Coarray Fortran concept.
 Codimensions can do communications between images.}.
-In the buffer region, phase space coordinates of particles {\tt xp} and {\tt vp} are copied from adjacent tiles and images, when necessary.
-Figure \ref{fig.tile} shows the spacial decomposition in a 2-dimensional analogy, with ${\tt nn=2}$ and ${\tt nnt=2}$.
-
+We denote actual number of particles in a given image {\tt nplocal}, and ${\tt npmax}>{\tt nplocal}$ is a value large enough to store buffered particles and should take into account the fact that {\tt nplocal} on some images (with densely clustered structures) could be large.
 The particle position and velocity arrays {\tt xp} and {\tt vp} (equivalent to $\chi_d$ and $\nu_d$ in \ref{ss.position} and \ref{ss.velocity} respectively) are required to be sorted according to the same memory layout of {\tt rho\_c}, such that $n_c$ and thus global positions of particles $x_d$ can be obtained by equation (\ref{eq.x}).
+$\{{\tt rho\_c},{\tt xp},{\tt vfield},{\tt vp}\}$ provides a complete information of positions and velocities of particles, and we call it a snapshot, or checkpoint.
 
@@
 Force kernels are also created or read in here.
-In {\tt read\_particles}, for each image, we read in all particles in {\it physical} regions of every tile (indicated in Figure \ref{fig.tile}), i.e., their initial positions {\tt xp}, velocities {\tt vp} in the compressed format, and their corresponding, physical part of the number density field {\tt rho\_c}.
+In {\tt read\_particles}, for each image, we read in all particles in {\it physical} regions of every tile (indicated in Figure \ref{fig.tile}), i.e., $\{{\tt rho\_c},{\tt xp},{\tt vfield},{\tt vp}\}$.
 These are obtained by the initial condition generator (see Appendix A).
@@
 We call it ``disjoint state''.
-At this stage, {\tt rho\_c}'s buffer regions of each tile are 0, and the elements of {\tt xp} and {\tt vp} beyond physical number can be arbitrary and are not used.
+At this stage, {\tt rho\_c}'s buffer regions of each tile are 0, and the elements of {\tt xp} and {\tt vp} beyond physical number -- {\tt xp(nplocal+1:)} and {\tt vp(nplocal+1:)} can be arbitrary and are not used.
 
@@
 We call this stage ``buffered state''.
-After {\tt particle\_initialize} is done by all images (synchronized by ``{\tt sync all}'', which is equivalent to a {\tt mpi\_barrier}), these three subroutines are called and particles are converted from disjoint state to buffered state.
+After {\tt particle\_initialize} is done by all images (synchronized by a ``{\tt sync all}'', which is equivalent to a {\tt mpi\_barrier}), these three subroutines are called and particles are converted from disjoint state to buffered state.
 
@@
       call read_particles
-      sync all
       call buffer_density
@@
         call update_xp
-sys
+        call buffer_density
         call buffer_xp
@@
 The actual simulation applies varied time steps.
-In each iteration of the main loop, we firstly call {\tt timestep}, where a increment of time {\tt dt} is controlled by particles' maximum velocities, accelerations, and cosmic expansions.
+In each iteration of the main loop, we firstly call {\tt timestep}, where a increment of time {\tt dt} is controlled by particles' maximum velocities, accelerations, cosmic expansion and any other desired conditions.
 
@@
 According to {\tt dt}, subroutine {\tt update\_xp} updates the positions of particles in a ``{\it gather}'' algorithm (in contrast, {\tt CUBEP3M} uses ``scatter'' algorithm) tile by tile.
-Because each tile is in the buffered state with buffer depth {\tt ncb}, we are able to collect all possible particles whose {\tt vp}*{\tt dt}$<${\tt ncb} from physical+buffer region to its physical region.
-In order to keep particles ordered, we have to execute ${\tt xp}={\tt xp}+{\tt vp}*{\tt dt}$ twice, first time to obtain an updated density field on the tile, {\tt rho\_c\_new}, and second time to generate a new, local particle list {\tt xp\_new} and {\tt vp\_new} on the tile.
-The reason for this repeated execution is the dependence of {\tt x\_new} and {\tt v\_new}'s value and ordering on both the old {\tt xp}, {\tt vp} and the new {\tt rho\_c\_new}.
-However, ${\tt xp}={\tt xp}+{\tt vp}*{\tt dt}$ scales as $o(N)$ and is computational inexpensive.
-Although {\tt xp\_new} and {\tt vp\_new} arrays take extra memory, they do not appear simultaneously for multiple tiles, and compared to {\tt xp} and {\tt vp}, the extra memory overhead is by a factor of $1/{\tt nnt}^3$.
-This requires the ``gather'' algorithm to move particles -- as we synchronize the correct \{{\tt rho\_c}, {\tt xp}, {\tt vp}\} in the buffer region of each tile, and the particles in the extended (physical+buffer) region are a superset of particles locating in the physical region in the next time step, given that ${\tt vp}*{\tt dt}$ is not greater than the buffer depth.
-In {\tt update\_xp}, \{{\tt rho\_c\_new}, {\tt xp\_new}, {\tt vp\_new}\} can be set to collect only the physical region of the tile, with the rest of this subset discarded.
-At the end, \{{\tt rho\_c}, {\tt xp}, {\tt vp}\} is replaced by \{{\tt xp\_new}, {\tt vp\_new}, {\tt rho\_c\_new}, {\tt vfield}\}, and the memory of latter is freed up.
-After this step, the tile does not contain buffered regions, being a disjoint state.
+For each particle, {\tt xp} and {\tt vp} are converted to $x_d$ and $v_d$ by Equations(\ref{eq.x},\ref{eq.v}).
+Because each tile is in the buffered state with buffer depth {\tt ncb}, we are able to collect all possible particles whose $v_d\times{\tt dt}<{\tt ncb}$ from physical+buffer region to its physical region.
+
+In order to keep particles ordered, for each tile, we first do $x_d=x_d+v_d\times{\tt dt}$ on all particles to obtain an updated density and velocity field on the tile, {\tt rho\_c\_new} and {\tt vfield\_new}.
+Then, this calculation in done on the same tile again\footnote{This repetition scales as $o(N)$ and is computational inexpensive.} to generate a new, local particle list {\tt xp\_new} and {\tt vp\_new} on the tile by Equations(\ref{eq.chi},\ref{eq.nu}).
+Here, the ordering of particles depends on {\tt rho\_c\_new}, and {\tt vfield\_new} is used as $v_c$ in Equation(\ref{eq.nu}).
+Then, another iteration is done on this tile to discard buffer regions of $\{{\tt rho\_c\_new},{\tt xp\_new},{\tt vfield\_new},{\tt vp\_new}\}$, converting buffer state to disjoint state, and it replaces corresponding part of $\{{\tt rho\_c},{\tt xp},{\tt vfield},{\tt vp}\}$.
+When all tiles are iterated, the entire image is in disjoint state, and {\tt nplocal} is updated.
+Then all images is synchronized by {\tt sync all} and we sum up {\tt nplocal} over images to check if total number of particles {\tt npglobal} is conserved.
 
 {\tt update\_vp} -- 
-After drift $D$, we need to update velocities (kick $K$) of particles in physical region of each tile.
-Before this step, we need to call {\tt buffer\_density} and {\tt buffer\_xp} in order that particle positions are in buffered state.
+After {\tt update\_xp}, or drift $D$, we call {\tt buffer\_density} and {\tt buffer\_xp} in order that particle positions are in buffered state, based on which, we update velocities (kick $K$) of particles in physical region of each tile.
 {\tt CUBE} uses a 2-level particle mesh scheme \citep{2013MNRAS.436..540H}.
-Local fine forces have a force cutoff ${\tt nforce\_cutoff}\leq{\tt ncb}$.
-So, if we apply a fine-grid particle-mesh on an extended tile with buffered-depth {\tt ncb} (see Figure \ref{fig.tile} regard it periodic), the force on physical regions does not depend on the false periodic boundary assumption, and the resulting fine force and velocity update on physical region is correct.
+Local fine forces have a force cutoff ${\tt nforce\_cutoff}\leq{\tt ncb}$, i.e., $F_{\rm fine}(r>{\tt ncb})=0$.
+So, if we apply a fine-grid particle-mesh on an extended tile with buffered-depth {\tt ncb} (see Figure \ref{fig.tile} and regard it periodic), the force on physical regions does not depend on the false periodic boundary assumption, and the resulting fine force $F_{\rm fine}$ and velocity update on physical region is correct.
 
-The compensating coarse grid force is globally computed by using a coarser (usually by factor of 4) mesh by dimensional splitting -- a distributed-memory cubic decomposed 3D coarse density field is interpolated by particles, and we Fourier transform data in consecutive three dimensions with global transposition in between (known as the pencil decomposition).
-After the multiplication of force kernels, the inverse transform takes place to get the cubic distributed coarse force field, upon which velocities are updated again.
+The compensating coarse grid force $F_{\rm coarse}$ is globally computed by using a coarser (usually by factor of 4) mesh by dimensional splitting -- a distributed-memory cubic decomposed 3D coarse density field is interpolated by particles, and we Fourier transform data in consecutive three dimensions with global transposition in between (known as the pencil decomposition).
+After the multiplication of force kernels, the inverse transform takes place to get the cubic distributed coarse force field $F_{\rm coarse}$, upon which velocities are updated again.
 
-An optional particle-particle (PP) force can be called to increase the force resolution and the velocities are updated last time according to this.
-\tcb{v field}.
-During each of these force calculations, maximum accelerations are collected serving for the {\tt timestep} in the next iteration, controlling next {\tt dt}.
+An optional particle-particle (PP) force $F_{\rm pp}$ can be called to increase the force resolution and the velocities are updated again.
+The collective operations of velocity update is Equation(\ref{eq.v}), $v_d=v_d+F_{\rm total}$ and Equation(\ref{eq.nu}).
 
-After all possible velocity updates, {\tt vp} in physical regions are correctly updated.
-Particle locations in the buffer regions has updated before {\tt PM} and remained unchanged.
+The maximum accelerations $\max(\dot{v}_{\rm fine})$, $\max(\dot{v}_{\rm coarse})$, $\max(\dot{v}_{\rm pp})$ and maximum of velocities $\max(v_d)$ are collected controlling {\tt dt} for the {\tt timestep} in the next iteration.
+We also update $\sigma^2_{\Delta}$ according to the new $v_d-v_c$.
+
+By far, {\tt vp} in physical regions are correctly updated.
+Remember that particle locations in the buffer regions has updated before {\tt PM} and remained unchanged.
 So we simply call {\tt buffer\_v} again such that the {\tt update\_x} in the next iteration will be done correctly.
@@
 {\tt checkpoint} -- 
-If a desired redshift is reached, we execute the last drift step in the $(DKKD)^n$ operation by {\tt update\_xp}, and call {\tt checkpoint} to save the disjoint state of \{{\tt xp\_new}, {\tt vp\_new}, {\tt rho\_c\_new}, {\tt vfield}\} on disk.
+If a desired redshift is reached, we execute the last drift step in the $(DKKD)^n$ operation by {\tt update\_xp}, and call {\tt checkpoint} to save the disjoint state of \{{\tt xp}, {\tt vp}, {\tt rho\_c}, {\tt vfield}\} on disk.
 Related operations, like run-time halo finder, projections are also done at this point.
@@
 The memory overhead of fine force calculation is controlled by $(1/{\tt nnt})^3$. The memory usage of coarse arrays are negligible. 
-
+Although {\tt xp\_new} and {\tt vp\_new} arrays take extra memory, they do not appear simultaneously for multiple tiles, and compared to {\tt xp} and {\tt vp}, the extra memory overhead is by a factor of $1/{\tt nnt}^3$.
 Here we summarize the memory usage in unit of bpp (byte per particle).

### 2017-09-29 ms_caf/ms_caf.tex
memory section

@@
 {\tt CUBE} uses cubic decomposition structures.
-The global simulation volume is decomposed into ${\tt nn}^3$ cubic sub-volumes, and each of these are assigned to a coarray {\it image}.
-Inside of an image, the sub-volume is further decomposed into ${\tt nnt}^3$ cubic {\it tiles}, which is a essential unit in the calculations.
+The global simulation volume is decomposed into ${\tt nn}^3$ cubic sub-volumes with {\tt nc} coarse grids per side, and each of these are assigned to a coarray {\it image}.
+Inside of an image, the sub-volume is further decomposed into ${\tt nnt}^3$ cubic {\it tiles}, with {\tt nt}={\tt nc}/{\tt nnt} coarse grids per side, which is a essential unit in the calculations.
 Each tile is surrounded by a {\it buffer} region which is ${\tt ncb}$ coarse cells thick.
@@
 Codimensions can do communications between images.}.
-We denote actual number of particles in a given image {\tt nplocal}, and ${\tt npmax}>{\tt nplocal}$ is a value large enough to store buffered particles and should take into account the fact that {\tt nplocal} on some images (with densely clustered structures) could be large.
+We denote actual number of particles in a given image {\tt nplocal}, and ${\tt npmax}>{\tt nplocal}$ is a value (discussed in Section \ref{ss.memory}) large enough to store particles.
 The particle position and velocity arrays {\tt xp} and {\tt vp} (equivalent to $\chi_d$ and $\nu_d$ in \ref{ss.position} and \ref{ss.velocity} respectively) are required to be sorted according to the same memory layout of {\tt rho\_c}, such that $n_c$ and thus global positions of particles $x_d$ can be obtained by equation (\ref{eq.x}).
@@
 {\tt initialize} and {\tt read\_particles} -- The subroutine {\tt initialize} creates necessary FFT plans and read in configuration files telling the program at which redshifts we need to do checkpoints, halofinds, or stop the simulation.
-Force kernels are also created or read in here.
+Force kernels {\tt kern\_c}, {\tt kern\_f} are also created or read in here.
 In {\tt read\_particles}, for each image, we read in all particles in {\it physical} regions of every tile (indicated in Figure \ref{fig.tile}), i.e., $\{{\tt rho\_c},{\tt xp},{\tt vfield},{\tt vp}\}$.
@@
 Then all images is synchronized by {\tt sync all} and we sum up {\tt nplocal} over images to check if total number of particles {\tt npglobal} is conserved.
+These steps are summarized in Figure \ref{fig.update_xp}.
+
+\begin{figure}[t]
+\begin{verbatim}
+
+subroutine update_xp
+  do (each tile)
+    do (each coarse grid)
+      do (each particle)
+        calculate particle's new coarse grid position
+        update rho_c_new
+        update vfield_new
+      enddo
+    enddo
+    do (each coarse grid)
+      do (each particle)
+        calculate particle's new accurate position
+        calculate particle's index
+        update xp_new(index) & vp_new(index)
+      enddo
+    enddo
+    do (each coarse grid)
+      discard buffer information
+      replace xp and vp for this tile
+    enddo
+  enddo
+  sync all
+  update velocity dispersion
+  check nplocal & npglobal
+end
+\end{verbatim}
+\caption{Pseudocode for subroutine {\tt update\_xp}.}
+\label{fig.update_xp}
+\end{figure}
 
@@
 So, if we apply a fine-grid particle-mesh on an extended tile with buffered-depth {\tt ncb} (see Figure \ref{fig.tile} and regard it periodic), the force on physical regions does not depend on the false periodic boundary assumption, and the resulting fine force $F_{\rm fine}$ and velocity update on physical region is correct.
+
+\begin{figure}[t]
+\begin{verbatim}
+
+subroutine update_vp
+  do (each tile)
+    calculate fine mesh density
+    calculate fine mesh force: local FFT
+    update fine mesh velocity
+    update maximum acceleration
+    if (PP force) then
+      PP calculation & velocity update
+      update maximum acceleration
+    endif
+  enddo
+  sync all
+  calculate coarse mesh density
+  calculate coarse mesh force: global pencil FFT
+  update coarse mesh velocity
+  update maximum acceleration
+  update maximum velocity
+end
+\end{verbatim}
+\caption{Pseudocode for subroutine {\tt update\_vp}.}
+\label{fig.update_vp}
+\end{figure}
 
@@
 We call {\tt update\_vp} to convert {\tt vp} into buffered state.
+These steps are summarized in Figure \ref{fig.update_vp}.
 
@@
 \subsection{Memory layout}\label{ss.memory}
-The memory overhead of fine force calculation is controlled by $(1/{\tt nnt})^3$. The memory usage of coarse arrays are negligible. 
-Although {\tt xp\_new} and {\tt vp\_new} arrays take extra memory, they do not appear simultaneously for multiple tiles, and compared to {\tt xp} and {\tt vp}, the extra memory overhead is by a factor of $1/{\tt nnt}^3$.
-Here we summarize the memory usage in unit of bpp (byte per particle).
+Here we list the memory-consuming arrays and how they scale with different configurations of the simulation.
+We classify them into (1) arrays of particles, (2) coarse mesh arrays and (3) fine mesh arrays.
+
+\subsubsection{Arrays of particles}
+The arrays of particles contains {\tt xvp(6,npmax)} and a temporary {\tt xvp\_new(6,nptile)}\footnote{We use ``{\tt xvp}'' to represent $\{{\tt xp},{\tt vp}\}$, and use ``{\tt xvp\_new}'' to represent $\{{\tt xp\_new},{\tt vp\_new}\}$.}.
+{\tt npmax} is the number of particles {\tt xvp} can store:
+\begin{equation}
+	{\tt npmax}=\left\langle {\tt nplocal} \right\rangle \left( 1+\frac{2\times {\tt ncb}}{\tt nt} \right)^3(1+\epsilon_{\rm image}),
+\end{equation}
+where $\left\langle {\tt nplocal} \right\rangle$ is the average number of particles per image.
+The second term above let us store particles in buffer regions, and the third term $1+\epsilon_{\rm image}$ takes into account the inhomogeneity of {\tt nplocal} on different images.
+When each image models smaller physical scales, $\epsilon_{\rm image}$ should be set larger.
+
+{\tt xvp\_new} stores temporary particles only on tiles:
+\begin{equation}
+	{\tt nptile}=\left\langle {\tt nplocal} \right\rangle \left( \frac{1}{\tt nnt} \right)^3(1+\epsilon_{\rm tile}),
+\end{equation}
+where, recall that {\tt nnt} is the number of tiles per image per dimension, and similarly $\epsilon_{\rm tile}$ controls the inhomogeneity on scale of tiles.
+Larger {\tt nnt} causes more inhomogeneity on smaller tiles, and $\epsilon_{\rm tile}$ can be much larger than $\epsilon_{\rm image}$, however the term ${\tt nnt}^{-3}$ decreases much faster.
+So, the majority memory is occupied by {\tt xvp}.
+
+\subsubsection{Coarse mesh arrays}
+On coarse mesh, {\tt rho\_c}, {\tt vfield} and force kernel {\tt kern\_c} should always be kept.
+They are usually configured to be 4 times coarser than fine grids and particle number density.
+In this case, each coarse grid, 3D {\tt vfield} takes only 12 bytes, and 64 particles will at least take 384 bytes ($n_\chi=n_\nu=1$ case).
+Similarly, pencil-FFT arrays, coarse force arrays etc.
+are memory-light and temporary.
+The total coarse mesh arrays take less than 10\% of the memory of {\tt xvp}.
+
+\subsubsection{Fine mesh arrays}
+On the local fine mesh, only a force kernel array needs to be kept.
+The fine mesh density arrays, force field arrays are temporary.
+Since {\tt xvp\_new}, coarse force arrays, pencil-FFT arrays are also temporary and are not used in any calculation simultaneously, they can be overlapped in memory by using equivalent statements.
+The total fine mesh arrays take less than 5\% of the memory of {\tt xvp}.
+
+\subsubsection{Compare with traditional algorithms}
+To show the improvement of memory usage, we give an example of TianNu simulation \citep{2017NatAs...1E.143Y} on the Tianhe-2 supercomputer, which used a traditional {\tt CUBEP3M} code.
+TianNu's particles number is limited by memory per computing node -- per computing node, an average of $576^3$ neutrino particles and $288^3$ CDM particles are used, and consumes about 40 GB of memory\footnote{Additional memory is used for OpenMP parallelization and for particle IDs to differentiate different particle spices.}, or about 186 bpp.
+A memory efficient case of {\tt CUBEP3M} by using large physical scales and at costs of speed, still uses about 40 bpp.
+
+By using {\tt CUBE}, with parameters $n_\chi=n_\nu=1$, ${\tt nc}=384$, ${\tt nnt}=4$, ${\tt ncb}=6$, $\epsilon_{\rm image}=10\%$, $\epsilon_{\rm tile}=100\%$ and $\left\langle {\tt nplocal} \right\rangle=1536^3$, we use about 38.3 GB of memory, corresponding to 10.6 bpp.
+This can be done on most of the supercomputers, even laptops.
+
+\begin{table}[]
+\centering
+\caption{Memory layout for a certain configuration}
+\label{t.memory}
+\begin{tabular}{llrrr}
+\hline
+& & \multicolumn{3}{c}{Memory usage}\\
+\cline{3-5}
+Type     & Array & /GB & /bpp & Percentage \\
+\hline
+Particles  & \tt xvp          & 29.9   & 8.24   & 87.4\%     \\
+           & \tt xvp\_new     & \underline{1.69}   & \underline{0.465}  & \underline{4.93\%}`     \\
+           & \tt PID          & 0      & 0      &    0\%     \\
+           & Subtotal         & 31.5   & 8.71   & 92.4\%     \\
+\hline
+Coarse mesh& \tt rho\_c       & 0.296  & 0.0818 & 0.870\%    \\
+           & \tt vfield       & 0.889  & 0.245  & 2.60\%     \\
+           & \tt kern\_c      & 0.340  & 0.0939 & 0.996\%    \\
+           & \tt force\_c     & (0.690)  & (0.190)  & (2.02\%)    \\
+           & \tt rho\_c\_new  & 0.0140 & 0.00388  & 0.0411\% \\
+           & Pencil-FFT       & (0.454)  & (0.125) & (1.33\%)      \\
+           & Subtotal         & 1.55   & 0.429 & 4.55\%     \\
+\hline
+Fine mesh  & \tt kern\_f      & 1.06   & 0.292 & 3.09\%      \\
+           & \tt force\_f     & (1.63) & (0.450) & (4.77\%)     \\
+           & Fine-FFT     &   (1.41) & (0.389) & (4.13\%)     \\
+           & Subtotal      & 1.06   & 0.292 & 3.09\%      \\
+\hline
+Total &                      & 34.2 & 9.43 & 100\%\\
+\hline
+\multicolumn{2}{l} {Optimal limit} & 21.7 & 6 & 63.6\%\\
+\hline
+\end{tabular}
+\end{table}
 
@@
 \centering
-  \includegraphics[width=0.95\linewidth]{diff_power.pdf}
+  \includegraphics[width=1.1\linewidth]{diff_power.pdf}
  \caption{power spectrum and cross correlation.}
@@
 
-\section{Discussion and Conclusion}\label{s.discussion}
  Cosmological simulation, PM method.

### 2017-09-29 ms_caf/ms_caf.tex
corrected numbers in table 1 -- fixed equivalence bugs.

@@
   \includegraphics[width=0.95\linewidth]{overplotparticles.pdf}
- \caption{particles.}
+ \caption{Difference between using {\tt CUBEP3M} (grey dots) and {\tt CUBE}-x1v1 (red dots). The box shows a projected $15.625^2\times 3.90625\ ({\rm Mpc}/h)^3$ region of the simulation. }
 \label{fig.particles}
@@
   \includegraphics[width=0.95\linewidth]{ddsp_pcdf_256_box80.pdf}
- \caption{Displacement.}
+ \caption{Error of the displacement of particles.}
 \label{fig.dsp}
@@
 are memory-light and temporary.
-The total coarse mesh arrays take less than 10\% of the memory of {\tt xvp}.
 
@@
 Since {\tt xvp\_new}, coarse force arrays, pencil-FFT arrays are also temporary and are not used in any calculation simultaneously, they can be overlapped in memory by using equivalent statements.
-The total fine mesh arrays take less than 5\% of the memory of {\tt xvp}.
 
@@
 
-By using {\tt CUBE}, with parameters $n_\chi=n_\nu=1$, ${\tt nc}=384$, ${\tt nnt}=4$, ${\tt ncb}=6$, $\epsilon_{\rm image}=10\%$, $\epsilon_{\rm tile}=100\%$ and $\left\langle {\tt nplocal} \right\rangle=1536^3$, we use about 38.3 GB of memory, corresponding to 10.6 bpp.
-This can be done on most of the supercomputers, even laptops.
+By using {\tt CUBE}, with parameters $n_\chi=n_\nu=1$, ${\tt nc}=384$, ${\tt nnt}=3$, ${\tt ncb}=6$, $\epsilon_{\rm image}=5\%$, $\epsilon_{\rm tile}=200\%$ and $\left\langle {\tt nplocal} \right\rangle=1536^3$, we use about 35.6 GB of memory, corresponding to 9.84 bpp.
+This can be done on most of the supercomputers, even modern laptops.
 
@@
 \hline
-Particles  & \tt xvp          & 29.9   & 8.24   & 87.4\%     \\
-           & \tt xvp\_new     & \underline{1.69}   & \underline{0.465}  & \underline{4.93\%}`     \\
+Particles  & \tt xvp          & 29.9   & 8.24   & 83.8\%     \\
+           & \tt xvp\_new     & \underline{3.16}   & \underline{0.872}  & \underline{8.87\%}     \\
            & \tt PID          & 0      & 0      &    0\%     \\
-           & Subtotal         & 31.5   & 8.71   & 92.4\%     \\
+           & Subtotal         & 33.0   & 9.12   & 92.7\%     \\
 \hline
-Coarse mesh& \tt rho\_c       & 0.296  & 0.0818 & 0.870\%    \\
-           & \tt vfield       & 0.889  & 0.245  & 2.60\%     \\
-           & \tt kern\_c      & 0.340  & 0.0939 & 0.996\%    \\
-           & \tt force\_c     & (0.690)  & (0.190)  & (2.02\%)    \\
-           & \tt rho\_c\_new  & 0.0140 & 0.00388  & 0.0411\% \\
-           & Pencil-FFT       & (0.454)  & (0.125) & (1.33\%)      \\
-           & Subtotal         & 1.55   & 0.429 & 4.55\%     \\
+Coarse mesh& \tt rho\_c       & 0.296  & 0.0818 & 0.831\%    \\
+           & \tt vfield       & 0.889  & 0.245  & 2.49\%     \\
+           & \tt kern\_c      & 0.340  & 0.0939 & 0.954\%    \\
+           & \tt force\_c     & (0.690)  & (0.190)  & (1.94\%)    \\
+           & \tt rho\_c\_new  & 0.0140 & 0.00388  & 0.0394\% \\
+           & Pencil-FFT       & (0.454)  & (0.125) & (1.27\%)      \\
+           & Subtotal         & 1.55   & 0.429 & 4.36\%     \\
 \hline
-Fine mesh  & \tt kern\_f      & 1.06   & 0.292 & 3.09\%      \\
-           & \tt force\_f     & (1.63) & (0.450) & (4.77\%)     \\
-           & Fine-FFT     &   (1.41) & (0.389) & (4.13\%)     \\
-           & Subtotal      & 1.06   & 0.292 & 3.09\%      \\
+Fine mesh  & \tt kern\_f      & 1.06   & 0.292 & 2.97\%      \\
+           & \tt force\_f     & (1.63) & (0.450) & (4.57\%)     \\
+           & Fine-FFT     &   (1.41) & (0.389) & (3.96\%)     \\
+           & Subtotal      & 1.06   & 0.292 & 2.97	\%      \\
 \hline
-Total &                      & 34.2 & 9.43 & 100\%\\
+Total &                      & 35.6 & 9.84 & 100\%\\
 \hline
-\multicolumn{2}{l} {Optimal limit} & 21.7 & 6 & 63.6\%\\
+\multicolumn{2}{l} {Optimal limit} & 21.7 & 6 & 61.0\%\\
 \hline
@@
 
-Results of {\tt izipx=1}
-
-Results of {\tt izipx=2}
-
 \begin{figure}
@@
   \includegraphics[width=0.95\linewidth]{power_universe1-8_z0and1.pdf}
- \caption{power spectrum.}
+ \caption{We don't need this figure.}
 \label{fig.power}
@@
   \includegraphics[width=1.1\linewidth]{diff_power.pdf}
- \caption{power spectrum and cross correlation.}
+ \caption{Fractional error on power spectrum and cross correlations.}
 \label{fig.ccc}

### 2017-09-30 ms_caf/ms_caf.tex
accuracy section -- displacement of particles

@@
 
-{\tt update\_vp} -- 
-After {\tt update\_xp}, or drift $D$, we call {\tt buffer\_density} and {\tt buffer\_xp} in order that particle positions are in buffered state, based on which, we update velocities (kick $K$) of particles in physical region of each tile.
-{\tt CUBE} uses a 2-level particle mesh scheme \citep{2013MNRAS.436..540H}.
-Local fine forces have a force cutoff ${\tt nforce\_cutoff}\leq{\tt ncb}$, i.e., $F_{\rm fine}(r>{\tt ncb})=0$.
-So, if we apply a fine-grid particle-mesh on an extended tile with buffered-depth {\tt ncb} (see Figure \ref{fig.tile} and regard it periodic), the force on physical regions does not depend on the false periodic boundary assumption, and the resulting fine force $F_{\rm fine}$ and velocity update on physical region is correct.
-
 \begin{figure}[t]
@@
 
+{\tt update\_vp} -- 
+After {\tt update\_xp}, or drift $D$, we call {\tt buffer\_density} and {\tt buffer\_xp} in order that particle positions are in buffered state, based on which, we update velocities (kick $K$) of particles in physical region of each tile.
+{\tt CUBE} uses a 2-level particle mesh scheme \citep{2013MNRAS.436..540H}.
+Local fine forces have a force cutoff ${\tt nforce\_cutoff}\leq{\tt ncb}$, i.e., $F_{\rm fine}(r>{\tt ncb})=0$.
+So, if we apply a fine-grid particle-mesh on an extended tile with buffered-depth {\tt ncb} (see Figure \ref{fig.tile} and regard it periodic), the force on physical regions does not depend on the false periodic boundary assumption, and the resulting fine force $F_{\rm fine}$ and velocity update on physical region is correct.
+
 The compensating coarse grid force $F_{\rm coarse}$ is globally computed by using a coarser (usually by factor of 4) mesh by dimensional splitting -- a distributed-memory cubic decomposed 3D coarse density field is interpolated by particles, and we Fourier transform data in consecutive three dimensions with global transposition in between (known as the pencil decomposition).
@@
 
-\begin{figure}[t]
-\centering
-  \includegraphics[width=0.95\linewidth]{overplotparticles.pdf}
- \caption{Difference between using {\tt CUBEP3M} (grey dots) and {\tt CUBE}-x1v1 (red dots). The box shows a projected $15.625^2\times 3.90625\ ({\rm Mpc}/h)^3$ region of the simulation. }
-\label{fig.particles}
-\end{figure}
-
-\begin{figure}[t]
-\centering
-  \includegraphics[width=0.95\linewidth]{ddsp_pcdf_256_box80.pdf}
- \caption{Error of the displacement of particles.}
-\label{fig.dsp}
-\end{figure}
-
 \subsection{Memory layout}\label{ss.memory}
@@
 {\tt npmax} is the number of particles {\tt xvp} can store:
-\begin{equation}
+\begin{equation}\label{eq.npmax}
 	{\tt npmax}=\left\langle {\tt nplocal} \right\rangle \left( 1+\frac{2\times {\tt ncb}}{\tt nt} \right)^3(1+\epsilon_{\rm image}),
@@
 {\tt xvp\_new} stores temporary particles only on tiles:
-\begin{equation}
+\begin{equation}\label{eq.nptile}
 	{\tt nptile}=\left\langle {\tt nplocal} \right\rangle \left( \frac{1}{\tt nnt} \right)^3(1+\epsilon_{\rm tile}),
@@
 Similarly, pencil-FFT arrays, coarse force arrays etc.
-are memory-light and temporary.
+are memory-light and/or temporary.
 
@@
 Since {\tt xvp\_new}, coarse force arrays, pencil-FFT arrays are also temporary and are not used in any calculation simultaneously, they can be overlapped in memory by using equivalent statements.
-
-\subsubsection{Compare with traditional algorithms}
-To show the improvement of memory usage, we give an example of TianNu simulation \citep{2017NatAs...1E.143Y} on the Tianhe-2 supercomputer, which used a traditional {\tt CUBEP3M} code.
-TianNu's particles number is limited by memory per computing node -- per computing node, an average of $576^3$ neutrino particles and $288^3$ CDM particles are used, and consumes about 40 GB of memory\footnote{Additional memory is used for OpenMP parallelization and for particle IDs to differentiate different particle spices.}, or about 186 bpp.
-A memory efficient case of {\tt CUBEP3M} by using large physical scales and at costs of speed, still uses about 40 bpp.
-
-By using {\tt CUBE}, with parameters $n_\chi=n_\nu=1$, ${\tt nc}=384$, ${\tt nnt}=3$, ${\tt ncb}=6$, $\epsilon_{\rm image}=5\%$, $\epsilon_{\rm tile}=200\%$ and $\left\langle {\tt nplocal} \right\rangle=1536^3$, we use about 35.6 GB of memory, corresponding to 9.84 bpp.
-This can be done on most of the supercomputers, even modern laptops.
 
@@
 
-\section{Results}\label{s.results}
-We use the same initial conditions, and use {\tt CUBEP3M} and {\tt CUBE} to simulate the large scale structure formation separately and compare their results.
+\subsubsection{Compare with traditional algorithms}
+To illustrate the improvement of memory usage, we give an example of TianNu simulation \citep{2017NatAs...1E.143Y} on the Tianhe-2 supercomputer, which used a traditional {\tt CUBEP3M} code.
+TianNu's particles number is limited by memory per computing node -- per computing node, an average of $576^3$ neutrino particles and $288^3$ CDM particles are used, and consumes about 40 GB of memory\footnote{Additional memory is used for OpenMP parallelization and for particle IDs to differentiate different particle spices.}, or about 186 bpp.
+A memory efficient case of {\tt CUBEP3M} by using large physical scales and at costs of speed, still uses about 40 bpp.
+
+By using {\tt CUBE}, with parameters $n_\chi=n_\nu=1$, ${\tt nc}=384$, ${\tt nnt}=3$, ${\tt ncb}=6$, $\epsilon_{\rm image}=5\%$, $\epsilon_{\rm tile}=200\%$ and $\left\langle {\tt nplocal} \right\rangle=1536^3$, we use about 35.6 GB of memory, corresponding to 9.84 bpp.
+This can be done on most of the supercomputers, even modern laptops.
+
+Table \ref{t.memory} shows the memory consumption for this test simulation.
+The memory-consuming arrays are listed and classified into the three types above mentioned, and their memory usages are in unit of GB ($10^9$ bytes), byte per particle (bpp), and their percentage to the total memory usage.
+The parenthesized numbers show overlapped memory, which is saved by equivalencing them with the underscored numbers.
+There are other, unlisted variables which are memory-light, and can also be equivalenced with listed variables.
+For this $1536^3$ per image simulation, using ${\tt nnt}^3=27$ tiles per image optimizes the memory usage, because it well controls the balance between the second terms in Equations (\ref{eq.npmax},\ref{eq.nptile}).
+A great amount of temporary memory in coarse and fine mesh force calculations is equivalenced to {\tt xvp\_new}.
+
+In the bottom of Table \ref{t.memory} we stress that the optimal memory usage is 6 bpp, or 21.7 GB, 61\% of the actual memory usage in this test simulation.
+The most prominent departure from this limit is that {\tt xvp} already uses 8.24 bpp, and all other variables occupy only additional 8\%.
+Of coarse, on modern super computers, the memory per computing node is usually much larger, by scaling up the number of particles per node, the buffer ratio ${\tt ncb}/{\tt nt}$ will be lowered and we can approach closer to the 6 bpp limit.
+
+\section{Accuracy}\label{s.results}
+We run a group of simulations to test the accuracy of {\tt CUBE}.
+We use same seeds to generate same Gaussian random fields in the initial condition generators of {\tt CUBEP3M} and {\tt CUBE}, and then they produce initial conditions of their own format respectively.
+Then the main $N$-body codes run their own initial conditions to redshift $z=0$.
+
+First, by using different configurations -- different number of computing nodes, box sizes, particle resolutions, different number of tiles per node/image etc., we find that by using 2-byte integers for both positions and velocities ($n_\chi=n_\nu=2$, or x2v2) {\tt CUBE} gives exact results compared to {\tt CUBEP3M}.
+So, if memory usage is not a problem, one can always double the memory usage to get exact results, and the optimal memory limit of this case is 12 bpp, still much lower than traditional methods.
+Next, we focus on the accuracy of the other three cases -- x1v2, x2v1 and x1v1.
+
+We list the names and configurations of simulations used in Table \ref{t.sim}, where $N_{\rm image}$, $L$, $z_i$, $z_{\rm checkpoint}$ and $N_p$ are respectively the number of computing nodes used, length of the side of the box, initial redshift, redshift where we do checkpoint and total number of particles.
+These configurations are run by {\tt CUBEP3M}, x2v2, x1v2, x2v1 and x1v1 versions of {\tt CUBE}.
+Using different number of tiles per image gives exact same results.
+
+\begin{table}[]
+\centering
+\caption{Simulation configurations}
+\label{t.sim}
+\begin{tabular}{lrrrrr}
+\hline
+& \multicolumn{5}{c}{Configurations}\\
+\cline{2-6}
+Name  & $N_{\rm image}$ & $L/(h^{-1}{\rm Mpc})$ & $z_i$ & $z_{\rm checkpoint}$ & $N_p$ \\
+\hline
+S512   & 8     & 200.0   & 49.0 & 0      & $512^3$     \\
+S256   & 1     & 80.0    & 49.0 & 0      & $256^3$     \\
+S2048S & 64    & 300.0   & 49.0 & 1.0, 0  & $2048^3$     \\
+S2048L & 64    & 1200.0  & 49.0 & 1.0, 0  & $2048^3$     \\
+\hline
+\end{tabular}
+\end{table}
+
+\begin{figure}[t]
+\centering
+  \includegraphics[width=0.95\linewidth]{overplotparticles.pdf}
+ \caption{Difference between using {\tt CUBEP3M} (grey dots) and {\tt CUBE}-x1v1 (red dots). The box shows a projected $15.625^2\times 3.90625\ ({\rm Mpc}/h)^3$ region of the simulation. }
+\label{fig.particles}
+\end{figure}
+
+\subsection{Displacement of particles}
+In S512, we zoom into a small region of $15.625^2\times 3.90625\ ({\rm Mpc}/h)^3$ and compare the particle distribution between {\tt CUBEP3M} and {\tt CUBE}-x1v1 in Figure \ref{fig.particles}.
+For clarity, {\tt CUBEP3M} particles are marked with bigger, grey dots, whereas smaller red dots are {\tt CUBE}-x1v1 particles over-plotted onto them.
+One can see the difference between them.
+1 ${\rm Mpc}/h$, fine grid and coarse grid scales are shown in the figure.
+The position resolution of particles in {\tt CUBE}-x1v1 is 1/256 of a coarse grid, or 1/64 of a fine grid.
+
+To quantify the offset in the final particle distributions, we use PIDs to track the displacement ${\bs \Psi}({\bs q})\equiv{\bs x}-{\bs q}$ of every particle \citep{2017PhRvD..95d3501Y}, where ${\bs x}$ and ${\bs q}$ are Eulerian and Lagrangian coordinates of the particle.
+Then we calculate the absolute value of the offset vector
+\begin{equation}\label{eq.offset}
+	\Delta\Psi\equiv|{\bs\Psi}_i-{\bs\Psi}_0|.
+\end{equation}
+Here, ${\bs\Psi}_0$ stands for {\tt CUBEP3M} and subscript $_i$ can stand for x2v2, x1v2, x2v1 or x1v1.
+The probability distribution functions (PDFs) and cumulative distribution functions (CDFs) of $\Delta\Psi$ in S256 are shown in Figure \ref{fig.dsp}.
+Results from x2v2, x1v2, x2v1 or x1v1 are in black, red, blue and green respectively.
+The results from absolute displacement of particles (by replacing ${\bs\Psi}_i$ with ${\bs q}$ in Equation (\ref{eq.offset})) are shown in grey for comparison.
+
+For x2v2, almost all particles are accurate up to 1/100 of a fine grid, and the worst particle is $\sim 1/10$ of a fine grid away from its counterpart in {\tt CUBEP3M}.
+The difference is caused by truncation errors and is negligible in physical and cosmological applications.
+The accuracy of x1v2 is between x2v2 and x1v1, and x2v1 gives only minor improvement from x1v1.
+
+\subsection{Power spectrum}
+
+\begin{figure}[t]
+\centering
+  \includegraphics[width=0.95\linewidth]{ddsp_pcdf_256_box80.pdf}
+ \caption{Error of the displacement of particles.}
+\label{fig.dsp}
+\end{figure}
 

### 2017-10-03 ms_caf/ms_caf.tex
discussion section

@@
 \author{
-Hao-Ran Yu$^{1,2}$,
-Ue-Li Pen$^{2,3,4,5}$,
+Hao-Ran Yu$^{1,2,3}$,
+Ue-Li Pen$^{2,4,5,6}$,
 Xin Wang$^{2}$
@@
 $^2$Canadian Institute for Theoretical Astrophysics, University of Toronto, Toronto, ON M5H 3H8 Canada\\
-$^3$Dunlap Institute for Astronomy and Astrophysics, University of Toronto, Toronto, ON M5S 3H4, Canada\\
-$^4$Canadian Institute for Advanced Research, CIFAR Program in Gravitation and Cosmology, Toronto, ON M5G 1Z8, Canada\\
-$^5$Perimeter Institute for Theoretical Physics, Waterloo, ON N2L 2Y5, Canada
+$^3$Department of Physics and Astronomy, Shanghai Jiao Tong University, Shanghai, 200240, China\\
+$^4$Dunlap Institute for Astronomy and Astrophysics, University of Toronto, Toronto, ON M5S 3H4, Canada\\
+$^5$Canadian Institute for Advanced Research, CIFAR Program in Gravitation and Cosmology, Toronto, ON M5G 1Z8, Canada\\
+$^6$Perimeter Institute for Theoretical Physics, Waterloo, ON N2L 2Y5, Canada
 }
@@
 
+\subsubsection{Initial conditions}
+The cosmological initial condition generator is compiled and run separately from the main $N$-body code.
+Here we briefly describe it for completeness, and declare that it uses less memory than the main $N$-body code.
+
 \subsubsection{Algorithm}\label{sss.algorithm}
@@
 \end{verbatim}
-\caption{Overall structure of {\tt CUBE}.}
+\caption{Overall structure of {\tt CUBE}. Sections of the code are grouped into Fortran subroutines, which are described in paragraphs of Section \ref{sss.algorithm}.}
 \label{fig.code}
@@
 
+\begin{figure*}[t]
+\centering
+  \includegraphics[width=0.75\linewidth]{overplotparticles.pdf}
+ \caption{Offset in the particle distribution induced by integer-1 based algorithm (x1v1).
+In S512's subregion of volume $15.625^2\times 3.90625\ ({\rm Mpc}/h)^3$, particles from {\tt CUBEP3M} (bigger grey dots in the background) and {\tt CUBE}-x1v1 (smaller red dots) are projected onto the plane of $(15.625\ {\rm Mpc}/h)^2$.
+The comparing rules show 1 ${\rm Mpc}/h$, fine grid and coarse grid respectively, and position resolution of x1v1 is 1/64 of a fine grid.}
+\label{fig.particles}
+\end{figure*}
+
 \subsubsection{Compare with traditional algorithms}
@@
 
-We list the names and configurations of simulations used in Table \ref{t.sim}, where $N_{\rm image}$, $L$, $z_i$, $z_{\rm checkpoint}$ and $N_p$ are respectively the number of computing nodes used, length of the side of the box, initial redshift, redshift where we do checkpoint and total number of particles.
+We list the names and configurations of simulations used in Table \ref{t.sim}, where $N_{\rm node}$, $L$, $z_i$, $z_{\rm ckpt}$, $N_p$ and $m_p$ are respectively the number of computing nodes used, length of the side of the box, initial redshift, redshift where we do checkpoint, total number of particles and particle mass.
 These configurations are run by {\tt CUBEP3M}, x2v2, x1v2, x2v1 and x1v1 versions of {\tt CUBE}.
@@
 \label{t.sim}
-\begin{tabular}{lrrrrr}
+\begin{tabular}{lrrrrrr}
 \hline
-& \multicolumn{5}{c}{Configurations}\\
-\cline{2-6}
-Name  & $N_{\rm image}$ & $L/(h^{-1}{\rm Mpc})$ & $z_i$ & $z_{\rm checkpoint}$ & $N_p$ \\
+& \multicolumn{6}{c}{Configurations}\\
+\cline{2-7}
+Name  & $N_{\rm node}$ & $L/({\rm Mpc}/h)$ & $z_i$ & $z_{\rm ckpt}$ & $N_p$ & $m_p/\Msun$ \\
 \hline
-S512   & 8     & 200.0   & 49.0 & 0      & $512^3$     \\
-S256   & 1     & 80.0    & 49.0 & 0      & $256^3$     \\
-S2048S & 64    & 300.0   & 49.0 & 1.0, 0  & $2048^3$     \\
-S2048L & 64    & 1200.0  & 49.0 & 1.0, 0  & $2048^3$     \\
+S512   & 8     & 200   & 49 & 0    & $512^3$  & $7.5\times 10^9$    \\
+S256   & 1     & 80    & 49 & 0    & $256^3$  & $3.8\times 10^9$    \\
+S2048S & 64    & 300   & 49 & 1, 0  & $2048^3$ & $4.0\times 10^8$    \\
+S2048L & 64    & 1200  & 49 & 1, 0  & $2048^3$ & $2.5\times 10^{10}$    \\
 \hline
@@
 \end{table}
-
-\begin{figure}[t]
-\centering
-  \includegraphics[width=0.95\linewidth]{overplotparticles.pdf}
- \caption{Difference between using {\tt CUBEP3M} (grey dots) and {\tt CUBE}-x1v1 (red dots). The box shows a projected $15.625^2\times 3.90625\ ({\rm Mpc}/h)^3$ region of the simulation. }
-\label{fig.particles}
-\end{figure}
 
@@
 One can see the difference between them.
-1 ${\rm Mpc}/h$, fine grid and coarse grid scales are shown in the figure.
+The 1 ${\rm Mpc}/h$, fine grid and coarse grid scales are shown in the figure.
 The position resolution of particles in {\tt CUBE}-x1v1 is 1/256 of a coarse grid, or 1/64 of a fine grid.
+\begin{figure}[]
+\centering
+  \includegraphics[width=0.95\linewidth]{ddsp_pcdf_256_box80.pdf}
+ \caption{Statistics of the error in particle displacement $\Delta{\bs\Psi}$ induced by integer based algorithms in {\tt CUBE}.
+The PDF and CDF of $|\Delta{\bs\Psi}|$ from S256 are shown in the upper and lower axes, as functions of fine grids.
+Black, red, blue and green correspond to x2v2, x1v2, x2v1 and x1v1 respectively.
+The grey lines, marked with ``initial'', shows the distribution of actual displacement of particles in {\tt CUBEP3M}, $|{\bs\Psi}_0|$, which is orders of magnitudes larger than $|\Delta{\bs\Psi}|$.
+}
+\label{fig.dsp}
+\end{figure}
 
@@
 Here, ${\bs\Psi}_0$ stands for {\tt CUBEP3M} and subscript $_i$ can stand for x2v2, x1v2, x2v1 or x1v1.
-The probability distribution functions (PDFs) and cumulative distribution functions (CDFs) of $\Delta\Psi$ in S256 are shown in Figure \ref{fig.dsp}.
+The probability distribution functions (PDFs) and cumulative distribution functions (CDFs) of $\Delta\bs\Psi$ in S256 are shown in Figure \ref{fig.dsp}.
 Results from x2v2, x1v2, x2v1 or x1v1 are in black, red, blue and green respectively.
@@
 The accuracy of x1v2 is between x2v2 and x1v1, and x2v1 gives only minor improvement from x1v1.
+We also run a simulation with same number of particles but with $L=600\ {\rm Mpc}/h$ ($m_p=1.2\times 10^{12}\Msun$), and find that the accuracy of x1v2 is in turn between x2v1 and x1v1.
+We interpret that, in this latter case, particles have lower mass resolution, so they move slower and need higher position resolution but need lower velocity resolution, then x2v1 outperforms x1v2.
 
+\begin{figure*}
+\centering
+  \includegraphics[width=1.0\linewidth]{diff_power.pdf}
+ \caption{Fractional deviations in the power spectrum $\Delta^2(k)/\Delta^2_0(k)-1$ (upper axes) and decorrelation $\xi$ (lower axes) induced by the integer-based algorithms in {\tt CUBE}.
+The four panels show the results of two redshifts in S2048S and S2048L: $\{ z=1, z=0 \} \times \{ L=300{\rm Mpc}/h, L=1200{\rm Mpc}/h \}$.
+The four solid lines show the results from \{x2v2,x1v2,x2v1,x1v1\}, whose colors are shown in the legends in the last panel.
+The orange dashed lines show the scale $k=0.2\ k_{\rm Nyquist}$, and smaller scales are beyond our interest.}
+\label{fig.ccc}
+\end{figure*}
 \subsection{Power spectrum}
+In S2048S and S2048L are two simulations with $2048^3$ particles on a small ($L=300\ {\rm Mpc}/h$) and large ($L=1200\ {\rm Mpc}/h$) box sizes.
+We compare their accuracy by their power spectra and their cross correlations with {\tt CUBEP3M} at two checkpoints $z=1$ and $z=0$.
+For each simulation the particles are firstly cloud-in-cell (CIC) interpolated onto the fine mesh grid of simulation, and from the density field $\rho$ we define the density contrast $\delta\equiv \rho/\left\langle \rho \right\rangle-1$.
+More generally we define the cross power spectrum $P_{\alpha\beta}(k)$ between two fields $\delta_\alpha$ and $\delta_\beta$ ($\delta_\alpha=\delta_\beta$ for auto power spectrum) in Fourier space as
+\begin{equation}\label{eq.cps}
+	\left\langle \delta_\alpha^\dagger({\bs k})\delta_\beta({\bs k}') \right\rangle=
+    (2\pi)^3 P_{\alpha\beta}(k){\bs \delta}_{\rm 3D}({\bs k}-{\bs k}'),
+\end{equation}
+where ${\bs \delta}_{\rm 3D}$ is the three-dimensional Dirac delta function.
+In cosmology we usually consider the dimensionless power spectrum $\Delta^2_{\alpha\beta}(k)\equiv k^3 P_{\alpha\beta}(k)/(2\pi^2)$.
+The cross correlation coefficient is defined as
+\begin{equation}\label{eq.ccc}
+	\xi(k)\equiv P_{\alpha\beta}/\sqrt{P_{\alpha\alpha}P_{\beta\beta}}.
+\end{equation}
 
-\begin{figure}[t]
-\centering
-  \includegraphics[width=0.95\linewidth]{ddsp_pcdf_256_box80.pdf}
- \caption{Error of the displacement of particles.}
-\label{fig.dsp}
-\end{figure}
+We study the fractional power spectrum deviations $\Delta^2(k)/\Delta^2_0(k)-1$ (where $\Delta^2_0(k)$ denotes the {\tt CUBEP3M} power spectrum) and the cross correlation coefficient functions $\xi(k)$ between {\tt CUBE} and {\tt CUBEP3M}, and show the results in Figure \ref{fig.ccc}.
+For the four panels, the left shows S2048S and right shows S2048L; the upper shows $z=1$ and lower shows $z=0$.
+We label $k=0.2\ k_{\rm Nyquist}$ as dashed lines, where $k_{\rm Nyquist}$ is the fine mesh scale, and the average particle separation scale.
+On the scale $k=0.2\ k_{\rm Nyquist}$, the power spectrum is suppressed by $\sim 20\%$ compared to higher resolution simulations by Poisson noise of particles and we do not consider $k>0.2\ k_{\rm Nyquist}$.
 
-\begin{figure}
-\centering
-  \includegraphics[width=0.95\linewidth]{power_universe1-8_z0and1.pdf}
- \caption{We don't need this figure.}
-\label{fig.power}
-\end{figure}
+The upper parts of the four panels show the fractional power spectrum deviations.
+Because x2v2 gives exact results as {\tt CUBEP3M}, $\Delta^2_{\rm x2v2}(k)/\Delta^2_0(k)-1=0$ for both simulations and redshifts, shown by the four black straight lines.
+Similar to the results form Figure \ref{fig.dsp}, x1v2 (red) curves give better results than x2v1 (blue) and x1v1 (green), and x2v1 and x1v1 give nearly same results.
+Another trend is that higher redshifts and/or larger box sizes gives much better results.
+The worst case is S2048S, $z=0$ case -- There is a $<2\%$ deviation in the power spectrum.
 
-\begin{figure}
-\centering
-  \includegraphics[width=1.1\linewidth]{diff_power.pdf}
- \caption{Fractional error on power spectrum and cross correlations.}
-\label{fig.ccc}
-\end{figure}
+The lower parts of the four panels show the cross correlation coefficient function between {\tt CUBE} and {\tt CUBEP3M}.
+Again, x2v2 performs perfect in all the four cases.
+For x1v2, $\xi(k)>99.5\%$ for all cases within our scale of interest.
+For x1v2 and x1v1, $\xi(k)$ are very close, and they perform worst in S2048S, $z=0$ case -- on scale $k=0.2\ k_{\rm Nyquist}$, the correlation drops to $90\%$.
+
+\section{Discussion and conclusion}\label{s.discussion}
+We present an open-source parallel integer-based $N$-body algorithm that needs use very low memory usage, approaching 6 byte per particle (6 bpp).
+The accuracy of this 6pp configuration is acceptable on most scales of our interest.
+The accuracy is also adjustable in that we can choose 1-byte/2-byte integers separately for positions and velocities of particles.
+In the case of using 2-byte integers for both positions and velocities (``x2v2'', and memory can be 12 bpp), the algorithm gives exact results given by traditional $N$-body algorithms.
+
+We analyze the deviation of power spectrum and de-correlation caused by integer-1 based algorithms in two simulations S2048S and S2048L.
+They have comparable number of particles per computing as the TianNu simulation.
+S2048S focuses on mass resolution -- it simulates a very small physical volume and its mass resolution $m_p=4\times 10^8\Msun$ is even higher than TianNu ($7\times 10^8\Msun$ for its CDM particles), whereas S2048L focuses on simulation volume and have higher particle mass.
+In the power spectra and cross correlations, S2048L performs better, and we conclude that in high-$m_p$ cases, we can safely use x1v1 in cosmological simulations.
+
+Another benefit for this algorithm is cosmological simulations with neutrinos, although the neutrino modules of {\tt CUBE} is in development.
+Neutrinos have a high velocity dispersion and move much faster than CDM, and their small scale errors are dominated by their Poisson noise.
+We expect that, compared to CDM, x1v1 gives less power spectrum deviation to neutrinos, as they behave more Gaussian and less clustered.
+A cosmological $N$-body simulation with neutrinos, like TianNu, contains much more (for TianNu, 8 times more) neutrino $N$-body particles than CDM, which dominates the memory.
+For a TianNu-like simulation, one can safely use x1v2 or x2v2 for CDM and x1v1 for neutrinos, and the memory usage can approach 6 bpp.
+This allows much more particles in the simulation and can lower the Poisson noise of neutrinos prominently.
 
@@
 
-\clearpage
+\begin{figure}
+\centering
+  \includegraphics[width=0.95\linewidth]{power_universe1-8_z0and1.pdf}
+ \caption{\tcb{We don't need this figure in submission.} The solid lines show the power spectra of S2048L and S2048S at $z=0$, and dotted lines show the $z=1$ counterparts.
+For $z=0$, the two data tips are at $k=0.2\ k_{\rm Nyquist}$, where the power spectrum is suppressed by more than 20\% compared to higher resolution simulations.}
+\label{fig.power}
+\end{figure}
 
-\appendix
+%\clearpage
 
-\section{A. Initial condition generator}
+%\appendix
+
+%\section{A.}
 

### 2017-10-03 ms_caf/ms_caf.tex
minor changes

@@
 
+At a sufficient high initial redshift $z_i$, we firstly generate a linear density fluctuation $\delta_L$ by multiplying a Gaussian random field with the transfer function $T(k)$ in Fourier space.
+Then we solve for a curl-less displacement field to displace particles from their initial regular positions with Zel'dovich approximation (ZA) \citep{1970A&A.....5...84Z}.
+We solve the potential $\Phi$ of this displacement field by Poisson equation, and then the particles are generated and displaced by locally differentiating $\Phi$ in real space.
+At the same time the particle is assigned with a initial velocity.
+Before the particles are generated, a global Pencil FFTs are applied on the fine mesh to transform the initial random Gaussian field into the final displacement potential $\Phi$.
+The initial condition generator is applied only once, and in principle, at some costs of speed, one can use only one scalar field on the fine mesh, and FFTs can be done in-place.
+If the number of particles equals the number of fine grids, then the memory consumption is dominated by the single precision fine mesh field, or 4 bpp.
+
+The particles can be generated and saved to disk tile by tile, locally.
+Assuming that we initially put particles at the center of every fine grid, we first generate these regularly placed particles in an extended tile, and then according to their new positions (coarse grid positions) given by ZA we get a coarse density field on tile.
+Secondly, according to the existing coarse density field, we apply ZA again to get their accurate positions and initial velocities, and we generate their integer based phase space coordinates and put them in correct orders according to the coarse density field.
+Lastly, we discard those particles out of the physical region of the tile, and write only particles in physical region on to the disk.
+As we will see in Section \ref{sss.algorithm}, {\tt update\_xp} paragraph, it is very similar to the ``drift'' step in the main $N$-body code, where an ``buffered'' state of particles is firstly created by two iterations of particles, and the buffered particles are discarded to achieve a ``disjoint'' state.
+After working on this tile, the arrays for positions and velocities are freed up and can be reused for the next tile.
+During this step, the only major memory usage is given by the fine grid potential $\Phi$.
+
 \subsubsection{Algorithm}\label{sss.algorithm}
@@
 In {\tt read\_particles}, for each image, we read in all particles in {\it physical} regions of every tile (indicated in Figure \ref{fig.tile}), i.e., $\{{\tt rho\_c},{\tt xp},{\tt vfield},{\tt vp}\}$.
-These are obtained by the initial condition generator (see Appendix A).
+These are obtained by the initial condition generator.
 Because physical regions of tile are {\it complete} and {\it disjoint} in space, particles at this stage are also complete and disjoint.
@@
 
-Another benefit for this algorithm is cosmological simulations with neutrinos, although the neutrino modules of {\tt CUBE} is in development.
+Another benefit for this algorithm is LSS simulations with neutrinos, although the neutrino modules of {\tt CUBE} is in development.
 Neutrinos have a high velocity dispersion and move much faster than CDM, and their small scale errors are dominated by their Poisson noise.
 We expect that, compared to CDM, x1v1 gives less power spectrum deviation to neutrinos, as they behave more Gaussian and less clustered.
-A cosmological $N$-body simulation with neutrinos, like TianNu, contains much more (for TianNu, 8 times more) neutrino $N$-body particles than CDM, which dominates the memory.
+LSS-neutrino simulations, like TianNu, can contain much more (for TianNu, 8 times more) neutrino $N$-body particles than CDM, which dominates the memory.
 For a TianNu-like simulation, one can safely use x1v2 or x2v2 for CDM and x1v1 for neutrinos, and the memory usage can approach 6 bpp.
 This allows much more particles in the simulation and can lower the Poisson noise of neutrinos prominently.
+A 1-byte PID per particle increases minor memory usage and can differentiate 8 kinds of particles, or we can store different particles in different arrays without using PID.
 
- Cosmological simulation, PM method.
+We did not include PP force in {\tt CUBE} and {\tt CUBEP3M} in this paper.
+If one wants to focus on smaller scales, like halo masses and profiles, an extended PP forces, which act up to adjacent fine cells, should take into account.
+The memory consumption for PP force is only local and is negligible compared to particles.
+In these cases, even x2v2 is used, a 12 bpp memory usage is still much lower than traditional $N$-body algorithms.
 
-PID.
-
-Summarize memory.
-
-Phi, GPU.
+Traditional $N$-body codes consumes considerable memory while doing a relatively light computations.
+{\tt CUBE} prominently reduces the memory usage in $N$-body simulations and approaches toward a better balance between memory and computation.
+The next steps are optimization of the code and adapting it for various kinds of heterogeneous computing systems, e.g.
+MIC and GPUs.
+Optimizing the velocity storage may further improve the accuracy of x1v1, and whose effects on neutrino-LSS simulation is going to be discovered.
 

### 2017-10-05 ms_caf/ms_caf.tex
IC section

@@
 \centering
-  \includegraphics[width=0.95\linewidth]{fig_tile.pdf}
+  \includegraphics[width=0.99\linewidth]{fig_tile.pdf}
  \caption{Spacial decomposition in {\tt CUBE} in a 2D analogy.
 In this example, there are 2 images per dimension (${\tt nn}=2$), and 2 tiles per image per dimension (${\tt nnt}=2$).
-The orange boxes show the overlapped physical+buffer regions, inside of which one physical regions is indicated with green.}
+The orange boxes show the overlapped physical+buffer=extended tile regions, inside of which one physical regions is indicated with green.}
 \label{fig.tile}
@@
 Figure \ref{fig.tile} shows the spacial decomposition in a 2-dimensional analogy, with ${\tt nn=2}$ and ${\tt nnt=2}$.
-According to this spatial decomposition, and as discussed in the last two subsections, we declare $\{\rho_c,\chi_d,v_c,\nu_d\}$, or in variable expression $\{{\tt rho\_c},{\tt xp},{\tt vfield},{\tt vp}\}$ as follows:
+
+According to this spatial decomposition, and as discussed in the last two subsections, we declare $\{\rho_c,\chi_d,v_c,\nu_d\}$, or in variable expression $\{{\tt rho\_c},{\tt xp},{\tt vfield},{\tt vp}\}$, as follows:
 \begin{eqnarray*}
@@
 
+An additional particle-ID (PID) array, {\tt PID}, can also be declared to differentiate particle types (e.g.
+CDM and neutrino particles) or differentiate every particle, by using 1-byte, 4-byte or 8-byte integer per particle.
+If PID is turned on, the array {\tt PID} is also included in the checkpoints in an same order as {\tt xp} and {\tt vp}.
+
 \subsubsection{Initial conditions}
-The cosmological initial condition generator is compiled and run separately from the main $N$-body code.
-Here we briefly describe it for completeness, and declare that it uses less memory than the main $N$-body code.
+The cosmological initial condition generator is compiled and run separately from the main $N$-body code. Here we briefly describe it for completeness.
 
-At a sufficient high initial redshift $z_i$, we firstly generate a linear density fluctuation $\delta_L$ by multiplying a Gaussian random field with the transfer function $T(k)$ in Fourier space.
-Then we solve for a curl-less displacement field to displace particles from their initial regular positions with Zel'dovich approximation (ZA) \citep{1970A&A.....5...84Z}.
-We solve the potential $\Phi$ of this displacement field by Poisson equation, and then the particles are generated and displaced by locally differentiating $\Phi$ in real space.
-At the same time the particle is assigned with a initial velocity.
-Before the particles are generated, a global Pencil FFTs are applied on the fine mesh to transform the initial random Gaussian field into the final displacement potential $\Phi$.
+The first step is the calculation of the potential field.
+At a sufficient high initial redshift $z_i$, we firstly generate a linear density fluctuation $\delta_L({\bs q})$ on Lagrangian grid ${\bs q}$ by multiplying a Gaussian random field with the transfer function $T(k)$ in Fourier space, where $T(k)$ is given by the assumed cosmological model.
+Then we solve for the potential ${\Phi({\bs q}, z_i)}$ of a curl-less displacement field ${\bs\Psi({\bs q})}$ by a Poisson equation $-\nabla^2\Phi=\delta_L$.
+At this step, a global Pencil FFTs are applied on the fine mesh to transform the initial random Gaussian field into the final displacement potential $\Phi$.
 The initial condition generator is applied only once, and in principle, at some costs of speed, one can use only one scalar field on the fine mesh, and FFTs can be done in-place.
 If the number of particles equals the number of fine grids, then the memory consumption is dominated by the single precision fine mesh field, or 4 bpp.
+
+The second step is to generate particles on Lagrangian coordinates and displace them with Zel'dovich approximation (ZA) \citep{1970A&A.....5...84Z}.
+The particles are displaced by locally differentiating $\Phi$ in real space meanwhile their initial velocities are assigned and immediately converted into integer-based velocity by Equation (\ref{eq.nu}).
+PID.
 

### 2017-10-06 ms_caf/ms_caf.tex
initial conditions rewritten

@@
 Then we solve for the potential ${\Phi({\bs q}, z_i)}$ of a curl-less displacement field ${\bs\Psi({\bs q})}$ by a Poisson equation $-\nabla^2\Phi=\delta_L$.
-At this step, a global Pencil FFTs are applied on the fine mesh to transform the initial random Gaussian field into the final displacement potential $\Phi$.
+At this step, global Pencil FFTs are applied on the fine mesh to transform the initial random Gaussian field into the final displacement potential $\Phi$.
 The initial condition generator is applied only once, and in principle, at some costs of speed, one can use only one scalar field on the fine mesh, and FFTs can be done in-place.
@@
 
-The second step is to generate particles on Lagrangian coordinates and displace them with Zel'dovich approximation (ZA) \citep{1970A&A.....5...84Z}.
-The particles are displaced by locally differentiating $\Phi$ in real space meanwhile their initial velocities are assigned and immediately converted into integer-based velocity by Equation (\ref{eq.nu}).
-PID.
-
-The particles can be generated and saved to disk tile by tile, locally.
-Assuming that we initially put particles at the center of every fine grid, we first generate these regularly placed particles in an extended tile, and then according to their new positions (coarse grid positions) given by ZA we get a coarse density field on tile.
-Secondly, according to the existing coarse density field, we apply ZA again to get their accurate positions and initial velocities, and we generate their integer based phase space coordinates and put them in correct orders according to the coarse density field.
+The second step is to generate particles on Lagrangian coordinates ${\bs q}$ and displace them with Zel'dovich approximation (ZA) \citep{1970A&A.....5...84Z}, by locally differentiating $\Phi$ in real space.
+There are two iterations over particles' Lagrangian coordinates ${\bs q}$ (usually at the center of every fine grid).
+The first iteration calculates particles' new, coarse grid positions by ZA and get a coarse density field on tile.
+The second iteration, ZA is applied again to calculate the accurate positions of particles and their velocities, and convert them into integer based ones by Equations (\ref{eq.chi},\ref{eq.nu}) and place them in a certain order according to the coarse density field generated by the first iteration.
 Lastly, we discard those particles out of the physical region of the tile, and write only particles in physical region on to the disk.
 As we will see in Section \ref{sss.algorithm}, {\tt update\_xp} paragraph, it is very similar to the ``drift'' step in the main $N$-body code, where an ``buffered'' state of particles is firstly created by two iterations of particles, and the buffered particles are discarded to achieve a ``disjoint'' state.
-After working on this tile, the arrays for positions and velocities are freed up and can be reused for the next tile.
-During this step, the only major memory usage is given by the fine grid potential $\Phi$.
+If PIDs are needed, they are generated at the second iteration and are manipulated similarly as integer based velocities.
+Because particles are generated and saved to disk tile by tile, locally, after working on certain tile, the arrays for positions, velocities and PIDs (if any) are freed up and can be reused for the next tile.
+During this step, the only major memory usage is still dominated by the fine grid potential $\Phi$.
+
+After working on all tiles, we sum up the total particle number on this image, {\tt nplocal}, and write it in the header of the checkpoint.
+Finally, $\{{\tt rho\_c},{\tt xp},{\tt vfield},{\tt vp},{\tt PID}({\rm optional})\}$ is rewritten to the disk as the checkpoint of $z=z_i$.
 

### 2017-10-08 ms_caf/ms_caf.tex
change S2048S to L=400

@@
 CDM and neutrino particles) or differentiate every particle, by using 1-byte, 4-byte or 8-byte integer per particle.
-If PID is turned on, the array {\tt PID} is also included in the checkpoints in an same order as {\tt xp} and {\tt vp}.
+If PID is turned on, the array {\tt PID} is also included in the checkpoints, and the ordering of PID is same as {\tt xp} and {\tt vp}.
 
-\subsubsection{Initial conditions}
+\subsubsection{Initial conditions}\label{sss.ic}
 The cosmological initial condition generator is compiled and run separately from the main $N$-body code. Here we briefly describe it for completeness.
@@
 The first step is the calculation of the potential field.
-At a sufficient high initial redshift $z_i$, we firstly generate a linear density fluctuation $\delta_L({\bs q})$ on Lagrangian grid ${\bs q}$ by multiplying a Gaussian random field with the transfer function $T(k)$ in Fourier space, where $T(k)$ is given by the assumed cosmological model.
-Then we solve for the potential ${\Phi({\bs q}, z_i)}$ of a curl-less displacement field ${\bs\Psi({\bs q})}$ by a Poisson equation $-\nabla^2\Phi=\delta_L$.
-At this step, global Pencil FFTs are applied on the fine mesh to transform the initial random Gaussian field into the final displacement potential $\Phi$.
+At a sufficient high initial redshift $z_i$, we first generate a linear density fluctuation $\delta_L({\bs q})$ on Lagrangian grid ${\bs q}$ by multiplying a Gaussian random field with the transfer function $T(k)$ in Fourier space, where $T(k)$ is given by the assumed cosmological model.
+Then we solve for the potential ${\Phi({\bs q}, z_i)}$ of a curl-less displacement field ${\bs\Psi({\bs q})}$ by Poisson equation $-\nabla^2\Phi=\delta_L$.
+At this step, global Pencil FFTs are applied on the fine mesh to transform the initial random Gaussian field to the displacement potential $\Phi$.
 The initial condition generator is applied only once, and in principle, at some costs of speed, one can use only one scalar field on the fine mesh, and FFTs can be done in-place.
@@
 The second step is to generate particles on Lagrangian coordinates ${\bs q}$ and displace them with Zel'dovich approximation (ZA) \citep{1970A&A.....5...84Z}, by locally differentiating $\Phi$ in real space.
-There are two iterations over particles' Lagrangian coordinates ${\bs q}$ (usually at the center of every fine grid).
+This step is done tile by tile.
+For each tile, we iterate twice over particles' Lagrangian coordinates ${\bs q}$ (usually at the center of every fine grid).
 The first iteration calculates particles' new, coarse grid positions by ZA and get a coarse density field on tile.
@@
 Lastly, we discard those particles out of the physical region of the tile, and write only particles in physical region on to the disk.
-As we will see in Section \ref{sss.algorithm}, {\tt update\_xp} paragraph, it is very similar to the ``drift'' step in the main $N$-body code, where an ``buffered'' state of particles is firstly created by two iterations of particles, and the buffered particles are discarded to achieve a ``disjoint'' state.
-If PIDs are needed, they are generated at the second iteration and are manipulated similarly as integer based velocities.
+As we will see in Section \ref{sss.algorithm}, {\tt update\_xp} paragraph, it is very similar to the ``drift'' step in the main $N$-body code, where an ``buffered'' state of particles is firstly created by two iterations of particles, and the particles in the buffer region are discarded to achieve a ``disjoint'' state.
+If PIDs are needed, they are also generated at the second iteration and are manipulated similarly as integer based velocities.
 Because particles are generated and saved to disk tile by tile, locally, after working on certain tile, the arrays for positions, velocities and PIDs (if any) are freed up and can be reused for the next tile.
@@
 Finally, $\{{\tt rho\_c},{\tt xp},{\tt vfield},{\tt vp},{\tt PID}({\rm optional})\}$ is rewritten to the disk as the checkpoint of $z=z_i$.
+We summarize the above steps into a pseudocode in Figure \ref{fig.ic}, where the ``{\tt do}'' loop to create particles per tile is very similar to the following {\tt update\_xp} subroutine in the main $N$-body code.
+The corresponding arrays are also discussed below.
 
@@
 
+\begin{figure}[t]
+\begin{verbatim}
+
+program Initial_Condition_Generator_for_CUBE
+  Get random seed
+  Generate Gaussian random field
+  Multiply by power spectrum in Fourier space
+  Get linear density field
+  Solve Poisson equation
+  Get displacement potential  
+  do (each tile) ! create particles per tile
+    do (each fine grid in the extended tile region)
+      calculate particle's coarse grid position
+      update rho_c_new
+    enddo
+    do (each fine grid in the extended tile region)
+      calculate particle's new accurate position
+      calculate praticle's initial velocity
+      calculate particle's index
+      create xp_new(index) & vp_new(index)
+      if (PID_flag) create PID(index)
+    enddo
+    do (each coarse grid)
+      discard buffer information
+    enddo
+    write to disk
+    sum up nplocal
+  enddo
+  sync all
+  sum up npglobal
+  write to headers of the checkpoint
+end
+\end{verbatim}
+\caption{Pseudocode for the initial condition generator.}
+\label{fig.ic}
+\end{figure}
+
 {\tt initialize} and {\tt read\_particles} -- The subroutine {\tt initialize} creates necessary FFT plans and read in configuration files telling the program at which redshifts we need to do checkpoints, halofinds, or stop the simulation.
 Force kernels {\tt kern\_c}, {\tt kern\_f} are also created or read in here.
-In {\tt read\_particles}, for each image, we read in all particles in {\it physical} regions of every tile (indicated in Figure \ref{fig.tile}), i.e., $\{{\tt rho\_c},{\tt xp},{\tt vfield},{\tt vp}\}$.
-These are obtained by the initial condition generator.
-Because physical regions of tile are {\it complete} and {\it disjoint} in space, particles at this stage are also complete and disjoint.
+In {\tt read\_particles}, for each image, we read in all particles in {\it physical} regions of every tile (indicated in Figure \ref{fig.tile}), i.e., $\{{\tt rho\_c},{\tt xp},{\tt vfield},{\tt vp},{\tt PID}({\rm if\ any})\}$.
+These are obtained by the initial condition generator (Section \ref{sss.ic}).
+Because physical regions of tiles are {\it complete} and {\it disjoint} in space, particles at this stage are also complete and disjoint.
 We call it ``disjoint state''.
-At this stage, {\tt rho\_c}'s buffer regions of each tile are 0, and the elements of {\tt xp} and {\tt vp} beyond physical number -- {\tt xp(nplocal+1:)} and {\tt vp(nplocal+1:)} can be arbitrary and are not used.
+At this stage, {\tt rho\_c}'s values in buffer regions are 0's, and the elements of {\tt xp} and {\tt vp} beyond physical number -- {\tt xp(nplocal+1:)} and {\tt vp(nplocal+1:)} can be arbitrary and are not used.
 
 {\tt buffer\_density}, {\tt buffer\_x} and {\tt buffer\_v} -- In order to use the integer based format, {\tt xp} and {\tt vp} must always be ordered, and their number density field {\tt rho\_c} must always be present.
-In updating arrays of {\tt xp} and {\tt vp} (not simultaneously) of a local tile, a vicinity (buffer) region of the physical region is also needed.
+In updating arrays of {\tt xp} and {\tt vp} (not simultaneously) of a local tile, the buffer region of the tile is also needed.
 First, buffer regions of {\tt rho\_c} is synchronized between tiles and images by subroutine {\tt buffer\_density}.
 Then, by subroutines {\tt buffer\_x} and {\tt buffer\_v} respectively, {\tt xp} and {\tt vp} are updated to contain common, buffered particles, in an order according to the new, buffered {\tt rho\_c}.
+If there are PIDs used, then {\tt PID} is also updated in subroutine {\tt update\_v}.
 We call this stage ``buffered state''.
@@
 
-\begin{figure}[t]
+\begin{figure}[]
 \begin{verbatim}
@@
 In order to keep particles ordered, for each tile, we first do $x_d=x_d+v_d\times{\tt dt}$ on all particles to obtain an updated density and velocity field on the tile, {\tt rho\_c\_new} and {\tt vfield\_new}.
-Then, this calculation in done on the same tile again\footnote{This repetition scales as $o(N)$ and is computational inexpensive.} to generate a new, local particle list {\tt xp\_new} and {\tt vp\_new} on the tile by Equations(\ref{eq.chi},\ref{eq.nu}).
+Then, this calculation in done on the same tile again\footnote{This repetition scales as $o(N)$ and is computational inexpensive.} to generate a new, local particle list {\tt xp\_new} and {\tt vp\_new} (also {\tt PID\_new} if needed) on the tile by Equations(\ref{eq.chi},\ref{eq.nu}).
 Here, the ordering of particles depends on {\tt rho\_c\_new}, and {\tt vfield\_new} is used as $v_c$ in Equation(\ref{eq.nu}).
-Then, another iteration is done on this tile to discard buffer regions of $\{{\tt rho\_c\_new},{\tt xp\_new},{\tt vfield\_new},{\tt vp\_new}\}$, converting buffer state to disjoint state, and it replaces corresponding part of $\{{\tt rho\_c},{\tt xp},{\tt vfield},{\tt vp}\}$.
+Then, another iteration is done on this tile to discard buffer regions of $\{{\tt rho\_c\_new},{\tt xp\_new},{\tt vfield\_new},{\tt vp\_new},{\tt PID\_new}({\rm if\ any})\}$, converting buffer state to disjoint state, and it replaces corresponding part of $\{{\tt rho\_c},{\tt xp},{\tt vfield},{\tt vp},{\tt PID}({\rm if\ any})\}$.
 When all tiles are iterated, the entire image is in disjoint state, and {\tt nplocal} is updated.
@@
 {\tt update\_vp} -- 
+The particle-mesh (PM) or particle-particle particle-mesh (P$^3$M) algorithm is applied in this subroutine to update particles's velocities.
 After {\tt update\_xp}, or drift $D$, we call {\tt buffer\_density} and {\tt buffer\_xp} in order that particle positions are in buffered state, based on which, we update velocities (kick $K$) of particles in physical region of each tile.
@@
 We also update $\sigma^2_{\Delta}$ according to the new $v_d-v_c$.
+These steps are summarized in Figure \ref{fig.update_vp}.
 
@@
 Remember that particle locations in the buffer regions has updated before {\tt PM} and remained unchanged.
-So we simply call {\tt buffer\_v} again such that the {\tt update\_x} in the next iteration will be done correctly.
-We call {\tt update\_vp} to convert {\tt vp} into buffered state.
-These steps are summarized in Figure \ref{fig.update_vp}.
+So, in the main code we simply call {\tt buffer\_v} again to bring {\tt vp} and {\tt PID} (if any) into buffered state, such that the {\tt update\_x} in the next iteration will be done correctly.
 
@@
 S256   & 1     & 80    & 49 & 0    & $256^3$  & $3.8\times 10^9$    \\
-S2048S & 64    & 300   & 49 & 1, 0  & $2048^3$ & $4.0\times 10^8$    \\
-S2048L & 64    & 1200  & 49 & 1, 0  & $2048^3$ & $2.5\times 10^{10}$    \\
+S2048S & 64    & 400   & 49 & 1,0  & $2048^3$ & $9.4\times 10^8$    \\
+S2048L & 64    & 1200  & 49 & 1,0  & $2048^3$ & $2.5\times 10^{10}$ \\
+\hline
+TianNu & 13824 & 1200  & 100& 0.01 & $6912^3$ & $7.0\times 10^8$\\
+       &       &       & 5  & 0.01 & $13824^3$& $3.0\times 10^5$\\
 \hline
@@
  \caption{Fractional deviations in the power spectrum $\Delta^2(k)/\Delta^2_0(k)-1$ (upper axes) and decorrelation $\xi$ (lower axes) induced by the integer-based algorithms in {\tt CUBE}.
-The four panels show the results of two redshifts in S2048S and S2048L: $\{ z=1, z=0 \} \times \{ L=300{\rm Mpc}/h, L=1200{\rm Mpc}/h \}$.
+The four panels show the results of two redshifts in S2048S and S2048L: $\{ z=1, z=0 \} \times \{ L=400{\rm Mpc}/h, L=1200{\rm Mpc}/h \}$.
 The four solid lines show the results from \{x2v2,x1v2,x2v1,x1v1\}, whose colors are shown in the legends in the last panel.
@@
 \subsection{Power spectrum}
-In S2048S and S2048L are two simulations with $2048^3$ particles on a small ($L=300\ {\rm Mpc}/h$) and large ($L=1200\ {\rm Mpc}/h$) box sizes.
+In S2048S and S2048L are two simulations with $2048^3$ particles on small ($L=400\ {\rm Mpc}/h$) and large ($L=1200\ {\rm Mpc}/h$) box sizes.
 We compare their accuracy by their power spectra and their cross correlations with {\tt CUBEP3M} at two checkpoints $z=1$ and $z=0$.

### 2017-10-09 ms_caf/ms_caf.tex
add TianNu and TianZero configs in Table 2.

@@
 
-As a $n_\chi=1$, 1D ($d=1$), 4-coarse-cell ($N_c=4$) example, if $$\chi_1=(-128,127,0,60),$$ particle number density $${\rho_c^{\rm 1D}}=(1,0,2,1),$$ then in unit of coarse cells, $$x_1=(0.001953125, 2.998046875, 2.501953125, 3.736328125).$$
+As a $n_\chi=1$, 1D ($d=1$), 4-coarse-cell ($N_c=4$) example, if $$\chi_1=(-128,127,0,60),$$ particle number density $${\rho_c^{\rm 1D}}=(1,0,2,1),$$ then in unit of coarse cells, the accurate positions of these four particles are $$x_1=(0.001953125, 2.998046875, 2.501953125, 3.736328125).$$
 
@@
 
-\subsubsection{Initial conditions}\label{sss.ic}
-The cosmological initial condition generator is compiled and run separately from the main $N$-body code. Here we briefly describe it for completeness.
-
-The first step is the calculation of the potential field.
-At a sufficient high initial redshift $z_i$, we first generate a linear density fluctuation $\delta_L({\bs q})$ on Lagrangian grid ${\bs q}$ by multiplying a Gaussian random field with the transfer function $T(k)$ in Fourier space, where $T(k)$ is given by the assumed cosmological model.
-Then we solve for the potential ${\Phi({\bs q}, z_i)}$ of a curl-less displacement field ${\bs\Psi({\bs q})}$ by Poisson equation $-\nabla^2\Phi=\delta_L$.
-At this step, global Pencil FFTs are applied on the fine mesh to transform the initial random Gaussian field to the displacement potential $\Phi$.
-The initial condition generator is applied only once, and in principle, at some costs of speed, one can use only one scalar field on the fine mesh, and FFTs can be done in-place.
-If the number of particles equals the number of fine grids, then the memory consumption is dominated by the single precision fine mesh field, or 4 bpp.
-
-The second step is to generate particles on Lagrangian coordinates ${\bs q}$ and displace them with Zel'dovich approximation (ZA) \citep{1970A&A.....5...84Z}, by locally differentiating $\Phi$ in real space.
-This step is done tile by tile.
-For each tile, we iterate twice over particles' Lagrangian coordinates ${\bs q}$ (usually at the center of every fine grid).
-The first iteration calculates particles' new, coarse grid positions by ZA and get a coarse density field on tile.
-The second iteration, ZA is applied again to calculate the accurate positions of particles and their velocities, and convert them into integer based ones by Equations (\ref{eq.chi},\ref{eq.nu}) and place them in a certain order according to the coarse density field generated by the first iteration.
-Lastly, we discard those particles out of the physical region of the tile, and write only particles in physical region on to the disk.
-As we will see in Section \ref{sss.algorithm}, {\tt update\_xp} paragraph, it is very similar to the ``drift'' step in the main $N$-body code, where an ``buffered'' state of particles is firstly created by two iterations of particles, and the particles in the buffer region are discarded to achieve a ``disjoint'' state.
-If PIDs are needed, they are also generated at the second iteration and are manipulated similarly as integer based velocities.
-Because particles are generated and saved to disk tile by tile, locally, after working on certain tile, the arrays for positions, velocities and PIDs (if any) are freed up and can be reused for the next tile.
-During this step, the only major memory usage is still dominated by the fine grid potential $\Phi$.
-
-After working on all tiles, we sum up the total particle number on this image, {\tt nplocal}, and write it in the header of the checkpoint.
-Finally, $\{{\tt rho\_c},{\tt xp},{\tt vfield},{\tt vp},{\tt PID}({\rm optional})\}$ is rewritten to the disk as the checkpoint of $z=z_i$.
-We summarize the above steps into a pseudocode in Figure \ref{fig.ic}, where the ``{\tt do}'' loop to create particles per tile is very similar to the following {\tt update\_xp} subroutine in the main $N$-body code.
-The corresponding arrays are also discussed below.
-
-\subsubsection{Algorithm}\label{sss.algorithm}
-Figure \ref{fig.code} shows the overall structure of the code and these subroutines are described in following paragraphs.
-
 \begin{figure}[t]
@@
 
-{\tt initialize} and {\tt read\_particles} -- The subroutine {\tt initialize} creates necessary FFT plans and read in configuration files telling the program at which redshifts we need to do checkpoints, halofinds, or stop the simulation.
-Force kernels {\tt kern\_c}, {\tt kern\_f} are also created or read in here.
-In {\tt read\_particles}, for each image, we read in all particles in {\it physical} regions of every tile (indicated in Figure \ref{fig.tile}), i.e., $\{{\tt rho\_c},{\tt xp},{\tt vfield},{\tt vp},{\tt PID}({\rm if\ any})\}$.
-These are obtained by the initial condition generator (Section \ref{sss.ic}).
-Because physical regions of tiles are {\it complete} and {\it disjoint} in space, particles at this stage are also complete and disjoint.
-We call it ``disjoint state''.
-At this stage, {\tt rho\_c}'s values in buffer regions are 0's, and the elements of {\tt xp} and {\tt vp} beyond physical number -- {\tt xp(nplocal+1:)} and {\tt vp(nplocal+1:)} can be arbitrary and are not used.
+\subsubsection{Initial conditions}\label{sss.ic}
+The cosmological initial condition generator is compiled and run separately from the main $N$-body code. Here we briefly describe it for completeness.
 
-{\tt buffer\_density}, {\tt buffer\_x} and {\tt buffer\_v} -- In order to use the integer based format, {\tt xp} and {\tt vp} must always be ordered, and their number density field {\tt rho\_c} must always be present.
-In updating arrays of {\tt xp} and {\tt vp} (not simultaneously) of a local tile, the buffer region of the tile is also needed.
-First, buffer regions of {\tt rho\_c} is synchronized between tiles and images by subroutine {\tt buffer\_density}.
-Then, by subroutines {\tt buffer\_x} and {\tt buffer\_v} respectively, {\tt xp} and {\tt vp} are updated to contain common, buffered particles, in an order according to the new, buffered {\tt rho\_c}.
-If there are PIDs used, then {\tt PID} is also updated in subroutine {\tt update\_v}.
-We call this stage ``buffered state''.
-After {\tt particle\_initialize} is done by all images (synchronized by a ``{\tt sync all}'', which is equivalent to a {\tt mpi\_barrier}), these three subroutines are called and particles are converted from disjoint state to buffered state.
+The first step is the calculation of the potential field.
+At a sufficient high initial redshift $z_i$, we first generate a linear density fluctuation $\delta_L({\bs q})$ on Lagrangian grid ${\bs q}$ by multiplying a Gaussian random field with the transfer function $T(k)$ in Fourier space, where $T(k)$ is given by the assumed cosmological model.
+Then we solve for the potential ${\Phi({\bs q}, z_i)}$ of a curl-less displacement field ${\bs\Psi({\bs q})}$ by Poisson equation $-\nabla^2\Phi=\delta_L$.
+At this step, global Pencil FFTs are applied on the fine mesh to transform the initial random Gaussian field to the displacement potential $\Phi$.
+The initial condition generator is applied only once, and in principle, at some costs of speed, one can use only one scalar field on the fine mesh, and FFTs can be done in-place.
+If the number of particles equals the number of fine grids, then the memory consumption is dominated by the single precision fine mesh field, or 4 bpp.
+
+The second step is to generate particles on Lagrangian coordinates ${\bs q}$ and displace them with Zel'dovich approximation (ZA) \citep{1970A&A.....5...84Z}, by locally differentiating $\Phi$ in real space.
+This step is done tile by tile.
+For each tile, we iterate twice over particles' Lagrangian coordinates ${\bs q}$ (usually at the center of every fine grid).
+The first iteration calculates particles' new, coarse grid positions by ZA and get a coarse density field on tile.
+The second iteration, ZA is applied again to calculate the accurate positions of particles and their velocities, and convert them into integer based ones by Equations (\ref{eq.chi},\ref{eq.nu}) and place them in a certain order according to the coarse density field generated by the first iteration.
+Lastly, we discard those particles out of the physical region of the tile, and write only particles in physical region on to the disk.
+As we will see in Section \ref{sss.algorithm}, {\tt update\_xp} paragraph, it is very similar to the ``drift'' step in the main $N$-body code, where an ``buffered'' state of particles is firstly created by two iterations of particles, and the particles in the buffer region are discarded to achieve a ``disjoint'' state.
+If PIDs are needed, they are also generated at the second iteration and are manipulated similarly as integer based velocities.
+Because particles are generated and saved to disk tile by tile, locally, after working on certain tile, the arrays for positions, velocities and PIDs (if any) are freed up and can be reused for the next tile.
+During this step, the only major memory usage is still dominated by the fine grid potential $\Phi$.
+
+After working on all tiles, we sum up the total particle number on this image, {\tt nplocal}, and write it in the header of the checkpoint.
+Finally, $\{{\tt rho\_c},{\tt xp},{\tt vfield},{\tt vp},{\tt PID}({\rm optional})\}$ is rewritten to the disk as the checkpoint of $z=z_i$.
+We summarize the above steps into a pseudocode in Figure \ref{fig.ic}, where the ``{\tt do}'' loop to create particles per tile is very similar to the following {\tt update\_xp} subroutine in the main $N$-body code.
+The corresponding arrays are also discussed below.
 
@@
 \end{figure}
+
+\subsubsection{Algorithm}\label{sss.algorithm}
+Figure \ref{fig.code} shows the overall structure of the code and these subroutines are described in following paragraphs.
+
+{\tt initialize} and {\tt read\_particles} -- The subroutine {\tt initialize} creates necessary FFT plans and read in configuration files telling the program at which redshifts we need to do checkpoints, halofinds, or stop the simulation.
+Force kernels {\tt kern\_c}, {\tt kern\_f} are also created or read in here.
+In {\tt read\_particles}, for each image, we read in all particles in {\it physical} regions of every tile (indicated in Figure \ref{fig.tile}), i.e., $\{{\tt rho\_c},{\tt xp},{\tt vfield},{\tt vp},{\tt PID}({\rm if\ any})\}$.
+These are obtained by the initial condition generator (Section \ref{sss.ic}).
+Because physical regions of tiles are {\it complete} and {\it disjoint} in space, particles at this stage are also complete and disjoint.
+We call it ``disjoint state''.
+At this stage, {\tt rho\_c}'s values in buffer regions are 0's, and the elements of {\tt xp} and {\tt vp} beyond physical number -- {\tt xp(nplocal+1:)} and {\tt vp(nplocal+1:)} can be arbitrary and are not used.
+
+{\tt buffer\_density}, {\tt buffer\_x} and {\tt buffer\_v} -- In order to use the integer based format, {\tt xp} and {\tt vp} must always be ordered, and their number density field {\tt rho\_c} must always be present.
+In updating arrays of {\tt xp} and {\tt vp} (not simultaneously) of a local tile, the buffer region of the tile is also needed.
+First, buffer regions of {\tt rho\_c} is synchronized between tiles and images by subroutine {\tt buffer\_density}.
+Then, by subroutines {\tt buffer\_x} and {\tt buffer\_v} respectively, {\tt xp} and {\tt vp} are updated to contain common, buffered particles, in an order according to the new, buffered {\tt rho\_c}.
+If there are PIDs used, then {\tt PID} is also updated in subroutine {\tt update\_v}.
+We call this stage ``buffered state''.
+After {\tt particle\_initialize} is done by all images (synchronized by a ``{\tt sync all}'', which is equivalent to a {\tt mpi\_barrier}), these three subroutines are called and particles are converted from disjoint state to buffered state.
 
@@
 Using different number of tiles per image gives exact same results.
+We also list the configurations for TianNu and TianZero \citep{2017NatAs...1E.143Y,2017RAA....17...85E} simulations as a reference.
 
@@
 \hline
-TianNu & 13824 & 1200  & 100& 0.01 & $6912^3$ & $7.0\times 10^8$\\
-       &       &       & 5  & 0.01 & $13824^3$& $3.0\times 10^5$\\
+TianNu & 13824 & 1200  & 100& 0.01 & $6912^3$ & $6.9\times 10^8$\\
+       &       &       & 5  & 0.01 & $13824^3$& $3.2\times 10^5$\\
+TianZero & 13824 & 1200 & 100& 0.01 & $6912^3$ & $7.0\times 10^8$\\
 \hline
@@
 We compare their accuracy by their power spectra and their cross correlations with {\tt CUBEP3M} at two checkpoints $z=1$ and $z=0$.
+The physical scale of S2048S is designed such that the particle mass resolution $m_p$ is comparable to TianNu and TianZero simulations (their parameters are also listed in Table \ref{t.sim}) to study halo properties.
+On the other hand, S2048L focuses on larger structures, on which scale one can study weak gravitational lensing, reconstruction of baryonic acoustic oscillation (BAO) \citep{2017ApJ...841L..29W} etc.
+
 For each simulation the particles are firstly cloud-in-cell (CIC) interpolated onto the fine mesh grid of simulation, and from the density field $\rho$ we define the density contrast $\delta\equiv \rho/\left\langle \rho \right\rangle-1$.
@@
 \section{Discussion and conclusion}\label{s.discussion}
-We present an open-source parallel integer-based $N$-body algorithm that needs use very low memory usage, approaching 6 byte per particle (6 bpp).
-The accuracy of this 6pp configuration is acceptable on most scales of our interest.
+We present an parallel integer-based $N$-body algorithm.
+This open-source code {\tt CUBE}, has recently been used in many studies of LSS, e.g.
+\cite{2017PhRvD..95d3501Y,2017ApJ...841L..29W,2017MNRAS.469.1968P}.
+It needs very low memory usage, approaching 6 byte per particle (6 bpp).
+The accuracy of this 6pp configuration is acceptable on most scales of our interest, e.g.
+BAO.
 The accuracy is also adjustable in that we can choose 1-byte/2-byte integers separately for positions and velocities of particles.
@@
 They have comparable number of particles per computing as the TianNu simulation.
-S2048S focuses on mass resolution -- it simulates a very small physical volume and its mass resolution $m_p=4\times 10^8\Msun$ is even higher than TianNu ($7\times 10^8\Msun$ for its CDM particles), whereas S2048L focuses on simulation volume and have higher particle mass.
+S2048S focuses on mass resolution -- comparable to TianNu and TianZero simulations, whereas S2048L focuses on simulation volume and have higher particle mass.
 In the power spectra and cross correlations, S2048L performs better, and we conclude that in high-$m_p$ cases, we can safely use x1v1 in cosmological simulations.
@@
   \includegraphics[width=0.95\linewidth]{power_universe1-8_z0and1.pdf}
- \caption{\tcb{We don't need this figure in submission.} The solid lines show the power spectra of S2048L and S2048S at $z=0$, and dotted lines show the $z=1$ counterparts.
+ \caption{\tcb{We don't need this figure in submission.
+However we need to compare the results with nonlinear power spectrum by {\tt HALOFIT}.} The solid lines show the power spectra of S2048L and S2048XS ($L=300\ {\rm Mpc}/h$) at $z=0$, and dotted lines show the $z=1$ counterparts.
 For $z=0$, the two data tips are at $k=0.2\ k_{\rm Nyquist}$, where the power spectrum is suppressed by more than 20\% compared to higher resolution simulations.}

### 2017-10-16 ms_caf/ms_caf.tex
revised fig.1, introduction, and more citations

@@
 \usepackage[english]{babel}
-\usepackage{amsmath,color}
+\usepackage{amsmath,color,aas_macros}
 
@@
 precision floating point numbers keep track phase space coordinates of each particle.
-Here we present the algorithm and accuracy of a new parallel, memory-lite, Particle-
-Mesh based $N$-body code, where each particle can occupy as low as 6-byte memory. This 
+Here we present the algorithm and accuracy of a new parallel, memory-lite, Particle-Mesh
+based $N$-body code, where each particle can occupy as low as 6-byte memory. This 
 is accomplished by storing relative position and relative velocity of each particle, 
@@
 \section{Introduction}\label{s.intro}
-$N$-body simulation is a powerful tool to solve the highly nonlinear dynamic problems 
-\citep{1988csup.book.....H}. It is widely used in cosmology to model the formation and 
-evolution of the large scale structure (LSS). With the fast development of parallel 
+The $N$-body simulation, a dynamical simulation of a group of particles, is a powerful tool in physics and astronomy \citep{1988csup.book.....H}.
+It is widely used in cosmology to model the large scale structure (LSS) of the Universe \citep{1985ApJ...292..371D}.
+Current percent and sub-percent level LSS measurements of cosmological parameters, via the matter power spectrum \citep{2005MNRAS.360L..82R,2011ApJ...726....7T}, baryonic acoustic oscillations (BAO) \citep{2005ApJ...633..560E,2009ApJ...700..479T}, weak gravitational lensing \citep{2003ApJ...592..699V,2009A&A...499...31H,2009ApJ...701..945S} etc., require understandings of nonlinear dynamics of the cosmic structure, and rely on high-resolution and high dynamic range $N$-body simulations.
+
+When $N$ is large, the brute force pairwise particle-particle (PP) force brings unaffordable $o(N^2)$ computations.
+The workload can be alleviated by many algorithms, for example, the ``tree'' algorithms like {\tt GADGET} \citep{2001NewA....6...79S,2005MNRAS.364.1105S} and its simulation ``Millennium'' \citep{2005Natur.435..629S,2012MNRAS.426.2046A}, {\tt TPM} \citep{1995ApJS...98..355X} and {\tt GOTPM} \citep{2004NewA....9..111D}, adaptive grid algorithm including {\tt HYDRA} \citep{1995ApJ...452..797C} and {\tt RAMSES} \citep{2010ascl.soft11007T}, as well as mesh-refined codes \citep{1991ApJ...368L..23C} and moving adaptive particle-mesh codes \citep{1995ApJS..100..269P}.
+Among these, the standard particle-mesh (PM) algorithm \citep{1988csup.book.....H} is most memory and computational efficient if we focus on large cosmological scales.
+The load balancing problem is minimized as the matter distribution is rather homogeneous, and the speed benefits from the fast Fourier transform (FFT) libraries, such as {\tt FFTW3} \citep{Frigo05thedesign}.
+Based on PM, the parallel code {\tt PMFAST} \citep{2005NewA...10..393M} uses a two-level PM algorithm to reduce the global communication and memory.
+After subsequent developments, {\tt CUBEP3M} \citep{2013MNRAS.436..540H} uses cubic spatial decomposition, and adds in-cell or adjacent-cell PP force and many other features.
+
+With the fast development of parallel 
 supercomputers, we are able to simulate a system of more than a trillion ($10^{12}$) 
@@
 
-The $N$-body simulations use considerable amount of memory, because the phase space 
+$N$-body simulations use considerable amount of memory, because the phase space 
 coordinates $(x,y,z,v_x,v_y,v_z)$ of each $N$-body particle must be stored as, at 
@@
   \includegraphics[width=1.0\linewidth]{vdisp.pdf}
- \caption{\tcb{Xin, could you please check the units of both $y$-axis?}}
+ \caption{Variance of the velocity distribution given by Equation \ref{eq.vdisp}.
+The $x$-axis is the scale factor characterizing the cosmic evolution and the $y$-axis shows the $\sigma_v$ in unit of km/s.
+The 5 curves, from top to bottom, shows the $\sigma_v(a)$ with increasing smoothing scale $r$ from 0 to 20 Mpc$/h$.}
 \label{fig.vdisp}
@@
 \begin{equation}\label{eq.vdisp}
-	\tcb{ \sigma_v(a,k) = \frac{1}{3}  (a H f D)^2 \int_0^kd^3q\frac{P_{L}(q)}{q^2}},
+	\sigma_v(a,k) = \frac{1}{3}  (a H f D)^2 \int_0^kd^3q\frac{P_{L}(q)}{q^2},
 \end{equation}
@@
 
-\section{Accuracy}\label{s.results}
-We run a group of simulations to test the accuracy of {\tt CUBE}.
-We use same seeds to generate same Gaussian random fields in the initial condition generators of {\tt CUBEP3M} and {\tt CUBE}, and then they produce initial conditions of their own format respectively.
-Then the main $N$-body codes run their own initial conditions to redshift $z=0$.
-
-First, by using different configurations -- different number of computing nodes, box sizes, particle resolutions, different number of tiles per node/image etc., we find that by using 2-byte integers for both positions and velocities ($n_\chi=n_\nu=2$, or x2v2) {\tt CUBE} gives exact results compared to {\tt CUBEP3M}.
-So, if memory usage is not a problem, one can always double the memory usage to get exact results, and the optimal memory limit of this case is 12 bpp, still much lower than traditional methods.
-Next, we focus on the accuracy of the other three cases -- x1v2, x2v1 and x1v1.
-
-We list the names and configurations of simulations used in Table \ref{t.sim}, where $N_{\rm node}$, $L$, $z_i$, $z_{\rm ckpt}$, $N_p$ and $m_p$ are respectively the number of computing nodes used, length of the side of the box, initial redshift, redshift where we do checkpoint, total number of particles and particle mass.
-These configurations are run by {\tt CUBEP3M}, x2v2, x1v2, x2v1 and x1v1 versions of {\tt CUBE}.
-Using different number of tiles per image gives exact same results.
-We also list the configurations for TianNu and TianZero \citep{2017NatAs...1E.143Y,2017RAA....17...85E} simulations as a reference.
-
 \begin{table}[]
@@
 \end{table}
+
+\section{Accuracy}\label{s.results}
+We run a group of simulations to test the accuracy of {\tt CUBE}.
+We use same seeds to generate same Gaussian random fields in the initial condition generators of {\tt CUBEP3M} and {\tt CUBE}, and then they produce initial conditions of their own format respectively.
+Then the main $N$-body codes run their own initial conditions to redshift $z=0$.
+
+First, by using different configurations -- different number of computing nodes, box sizes, particle resolutions, different number of tiles per node/image etc., we find that by using 2-byte integers for both positions and velocities ($n_\chi=n_\nu=2$, or x2v2) {\tt CUBE} gives exact results compared to {\tt CUBEP3M}.
+So, if memory usage is not a problem, one can always double the memory usage to get exact results, and the optimal memory limit of this case is 12 bpp, still much lower than traditional methods.
+Next, we focus on the accuracy of the other three cases -- x1v2, x2v1 and x1v1.
+
+We list the names and configurations of simulations used in Table \ref{t.sim}, where $N_{\rm node}$, $L$, $z_i$, $z_{\rm ckpt}$, $N_p$ and $m_p$ are respectively the number of computing nodes used, length of the side of the box, initial redshift, redshift where we do checkpoint, total number of particles and particle mass.
+These configurations are run by {\tt CUBEP3M}, x2v2, x1v2, x2v1 and x1v1 versions of {\tt CUBE}.
+Using different number of tiles per image gives exact same results.
+We also list the configurations for TianNu and TianZero \citep{2017NatAs...1E.143Y,2017RAA....17...85E} simulations as a reference.
 
@@
 The physical scale of S2048S is designed such that the particle mass resolution $m_p$ is comparable to TianNu and TianZero simulations (their parameters are also listed in Table \ref{t.sim}) to study halo properties.
-On the other hand, S2048L focuses on larger structures, on which scale one can study weak gravitational lensing, reconstruction of baryonic acoustic oscillation (BAO) \citep{2017ApJ...841L..29W} etc.
+On the other hand, S2048L focuses on larger structures, on which scale one can study weak gravitational lensing, baryonic acoustic oscillation (BAO) \citep{2005ApJ...633..560E} and its reconstruction \citep{2007ApJ...664..675E,2017ApJ...841L..29W} etc.
 
@@
 They have comparable number of particles per computing as the TianNu simulation.
-S2048S focuses on mass resolution -- comparable to TianNu and TianZero simulations, whereas S2048L focuses on simulation volume and have higher particle mass.
-In the power spectra and cross correlations, S2048L performs better, and we conclude that in high-$m_p$ cases, we can safely use x1v1 in cosmological simulations.
+S2048S focuses on mass resolution -- comparable to TianNu and TianZero simulations, whereas S2048L focuses on simulation volume and have high
+er particle mass. In the power spectra and cross correlations, S2048L performs better, and we conclude that in high-$m_p$ cases, we can safely use x1v1 in cosmological simulations.
 
@@
 
-\begin{figure}
-\centering
-  \includegraphics[width=0.95\linewidth]{power_universe1-8_z0and1.pdf}
- \caption{\tcb{We don't need this figure in submission.
+%\begin{figure}
+%\centering
+%  \includegraphics[width=0.95\linewidth]{power_universe1-8_z0and1.pdf}
+% \caption{\tcb{We don't need this figure in submission.
 However we need to compare the results with nonlinear power spectrum by {\tt HALOFIT}.} The solid lines show the power spectra of S2048L and S2048XS ($L=300\ {\rm Mpc}/h$) at $z=0$, and dotted lines show the $z=1$ counterparts.
 For $z=0$, the two data tips are at $k=0.2\ k_{\rm Nyquist}$, where the power spectrum is suppressed by more than 20\% compared to higher resolution simulations.}
-\label{fig.power}
-\end{figure}
+%\label{fig.power}
+%\end{figure}
 

### 2017-10-19 ms_caf/ms_caf.tex
draft v1.0

@@
 \begin{document}
-\title{CUBE: A Memory-lite Cosmological $N$-body Algorithm}
+\title{CUBE: A Memory-optimized Cosmological $N$-body Algorithm}
 \author{
@@
 \begin{abstract}
-Cosmological large scale structure $N$-body simulations are computation-light,
-memory-heavy problems in supercomputing. Traditional $N$-body simulation
-algorithms use at least 24-byte memory per particle, of which six 4-byte single 
-precision floating point numbers keep track phase space coordinates of each particle.
-Here we present the algorithm and accuracy of a new parallel, memory-lite, Particle-Mesh
-based $N$-body code, where each particle can occupy as low as 6-byte memory. This 
-is accomplished by storing relative position and relative velocity of each particle, 
-in the format of 1-byte-integer, respect to their averaged value of a mesh-grid. The 
-remaining information is given by complimentary density and velocity fields, which are 
-negligible in memory space, and proper ordering of particles, which gives no extra 
-memory. Our numerical experiments show that this integer based $N$-body algorithm 
-provides acceptable accuracy compared to traditional algorithm in cosmological 
-simulations. This significant lowering of memory-to-computation ratio breaks the 
-bottleneck of scaling up and speeding up large cosmological $N$-body simulations on 
-multi-core and heterogenous computing systems.
+Cosmological large scale structure $N$-body simulations are computation-light, memory-heavy problems in supercomputing.
+Traditional $N$-body simulation algorithms use at least 24-byte memory per particle, of which six 4-byte single precision floating point numbers keep track phase space coordinates of each particle.
+Here we present the algorithm and accuracy of a new parallel, memory-lite, Particle-Mesh based $N$-body code, where each particle can occupy as low as 6-byte memory.
+This is accomplished by storing relative position and relative velocity of each particle, in the format of 1-byte-integer, respect to their averaged value of a mesh-grid.
+The remaining information is given by complimentary density and velocity fields, which are negligible in memory space, and proper ordering of particles, which gives no extra memory.
+Our numerical experiments show that this integer based $N$-body algorithm provides acceptable accuracy compared to traditional algorithm in cosmological simulations.
+This significant lowering of memory-to-computation ratio breaks the bottleneck of scaling up and speeding up large cosmological $N$-body simulations on multi-core and heterogenous computing systems.
 
@@
 When $N$ is large, the brute force pairwise particle-particle (PP) force brings unaffordable $o(N^2)$ computations.
-The workload can be alleviated by many algorithms, for example, the ``tree'' algorithms like {\tt GADGET} \citep{2001NewA....6...79S,2005MNRAS.364.1105S} and its simulation ``Millennium'' \citep{2005Natur.435..629S,2012MNRAS.426.2046A}, {\tt TPM} \citep{1995ApJS...98..355X} and {\tt GOTPM} \citep{2004NewA....9..111D}, adaptive grid algorithm including {\tt HYDRA} \citep{1995ApJ...452..797C} and {\tt RAMSES} \citep{2010ascl.soft11007T}, as well as mesh-refined codes \citep{1991ApJ...368L..23C} and moving adaptive particle-mesh codes \citep{1995ApJS..100..269P}.
+However the workload can be alleviated to $o(N\log N)$ by many algorithms.
+For example, the ``tree'' algorithms like {\tt GADGET} \citep{2001NewA....6...79S,2005MNRAS.364.1105S} and its simulation ``Millennium'' \citep{2005Natur.435..629S,2012MNRAS.426.2046A}, {\tt TPM} \citep{1995ApJS...98..355X} and {\tt GOTPM} \citep{2004NewA....9..111D}, adaptive grid algorithm including {\tt HYDRA} \citep{1995ApJ...452..797C} and {\tt RAMSES} \citep{2010ascl.soft11007T}, as well as mesh-refined codes \citep{1991ApJ...368L..23C} and moving adaptive particle-mesh codes \citep{1995ApJS..100..269P}.
 Among these, the standard particle-mesh (PM) algorithm \citep{1988csup.book.....H} is most memory and computational efficient if we focus on large cosmological scales.
 The load balancing problem is minimized as the matter distribution is rather homogeneous, and the speed benefits from the fast Fourier transform (FFT) libraries, such as {\tt FFTW3} \citep{Frigo05thedesign}.
-Based on PM, the parallel code {\tt PMFAST} \citep{2005NewA...10..393M} uses a two-level PM algorithm to reduce the global communication and memory.
-After subsequent developments, {\tt CUBEP3M} \citep{2013MNRAS.436..540H} uses cubic spatial decomposition, and adds in-cell or adjacent-cell PP force and many other features.
+{\tt PMFAST} \citep{2005NewA...10..393M} is first codes to use a 2-level PM algorithm, aiming to put PM codes toward speed, memory compactness and scalability.
+After subsequent developments on {\tt PMFAST}, {\tt CUBEP3M} \citep{2013MNRAS.436..540H} uses cubic spatial decomposition, and adds PP force and many other features.
 
-With the fast development of parallel 
-supercomputers, we are able to simulate a system of more than a trillion ($10^{12}$) 
-$N$-body particles. To date the largest $N$-body simulation in application is the 
-``TianNu'' simulation \citep{2017NatAs...1E.143Y,2017RAA....17...85E} run on the 
-TianHe-2 supercomputer by cosmological simulation code 
-{\tt CUBEP3M} \citep{2013MNRAS.436..540H}. It uses nearly $3\times 10^{12}$ particles 
-to simulate the cold dark matter (CDM) and cosmic neutrino evolution through the 
-cosmic age.
+Beside the methodology, the fast development of parallel supercomputers enable us to simulate a system of more than a trillion ($10^{12}$) $N$-body particles.
+To date the largest $N$-body simulation in application is the ``TianNu'' \citep{2017NatAs...1E.143Y,2017RAA....17...85E} run on the TianHe-2 supercomputer.
+With the code {\tt CUBEP3M} adding neutrino modules, it uses $3\times 10^{12}$ particles to simulate the cold dark matter (CDM) and cosmic neutrino evolution through the cosmic age.
 
-$N$-body simulations use considerable amount of memory, because the phase space 
-coordinates $(x,y,z,v_x,v_y,v_z)$ of each $N$-body particle must be stored as, at 
-least, six single precision floating point numbers (24 bytes). Contrarily, their 
-computing workload can be alleviated by many algorithms, like Particle-Mesh [cite] and 
-Tree [cite], scale as $o(N\log N)$ or even $o(N)$. On the other hand, modern 
-supercomputer systems use multi cores, many integrated cores (MIC) and even densely 
-parallelized GPUs, bringing orders of magnitude higher computing power, whereas these 
-architectures usually have limited memory allocation. Thus, the computation-light but 
-memory-heavy applications, compared to matrix multiplication and decomposition 
-calculations, are less suitable for fully usage of the computing power of modern 
-supercomputers. For example, although native and offload modes of {\tt CUBEP3M} are 
-able to run on the Intel Xeon-PHI MIC architectures, with the require ment of enough 
-memory, TianNu simulation were done on TianHe-2 with only its CPUs -- 73\% of the 
-total memory but only 13\% of the total computing power.
+Relative to the optimized computation optimizations, $N$-body simulations use considerable amount of memory, because the phase space coordinates $(x,y,z,v_x,v_y,v_z)$ of each $N$-body particle must be stored as at least six single precision floating point numbers (24 bytes).
+On the other hand, modern supercomputer systems use multi cores, many integrated cores (MIC) and even densely parallelized GPUs, bringing orders of magnitude higher computing power, whereas these architectures usually have limited memory allocation.
+Thus, the computation-light but memory-heavy applications, compared to matrix multiplication and decomposition calculations, are less suitable for fully usage of the computing power of modern supercomputers.
+For example, although native and offload modes of {\tt CUBEP3M} are able to run on the Intel Xeon-PHI MIC architectures, with the require ment of enough memory, TianNu simulation were done on TianHe-2 with only its CPUs -- 73\% of the total memory but only 13\% of the total computing power.
 
-We present a new $N$-body simulation code {\tt CUBE}, using as low as 6 byte
-per particle (here after we use ``bpp'' referring to ``byte per particle'').
+We present a new $N$-body simulation code {\tt CUBE}, using as low as 6 byte per particle (here after we use ``bpp'' referring to ``byte per particle'').
 We show that it gives accurate results in cosmological LSS simulations.
-The method is presented in \S\ref{s.method}, and a comparison between this method
-and traditional method is shown in \S\ref{s.results}. Discussions and 
-conclusions are in \S\ref{s.discussion}.
+In \S\ref{s.method} we show how the memory can be saved by using a ``integer-based storage'' and how the PM $N$-body algorithm is adapted with this storage format.
+In \S\ref{s.results} we quantify the accuracy of this algorithm by groups of simulations by {\tt CUBEP3M} and {\tt CUBE}.
+Discussions and conclusions are in \S\ref{s.discussion}.
 
 \section{Method}\label{s.method}
-The most memory consuming part of a $N$-body simulation is usually the phase space 
-coordinates of $N$-body particles -- 24 bpp (4 single precision floating numbers) must 
-be used to store each particles' 3D position and velocity vectors. {\tt CUBEP3M}, an 
-example of a memory-efficient parallel $N$-body code, can use as low as 40 bpp in 
-sacrificing computing speed \citep{2013MNRAS.436..540H}. This includes the phase 
-coordinates (24 bpp) for particles in physical domain and buffered region, a linked 
-list (4 bpp), and a global coarse mesh and local fine mesh. 4-byte real numbers are 
-not necessarily adequate in representing the {\it global} coordinates in simulations.
-If the box size is many orders of magnitude larger than interactive distance
-between particles, especially in the field of re-simulation of dense subregions,
-double precision (8-byte) coordinates are needed to avoid truncation errors.
+The most memory consuming part of a $N$-body simulation is usually the phase space coordinates of $N$-body particles -- 24 bpp (4 single precision floating numbers) must be used to store each particles' 3D position and velocity vectors.
+{\tt CUBEP3M}, an example of a memory-efficient parallel $N$-body code, can use as low as 40 bpp in sacrificing computing speed \citep{2013MNRAS.436..540H}.
+This includes the phase coordinates (24 bpp) for particles in physical domain and buffered region, a linked list (4 bpp), and a global coarse mesh and local fine mesh.
+4-byte real numbers are not necessarily adequate in representing the {\it global} coordinates in simulations.If the box size is many orders of magnitude larger than interactive distance between particles, especially in the field of re-simulation of dense subregions, double precision (8-byte) coordinates are needed to avoid truncation errors.
 Another solution is to record {\it relative} coordinates for both position and velocity.
-{\tt CUBE} replaces the coordinates and linked list 24+4=28 bpp memory usage
-with an integer based storage, reduces the basic memory usage from 28 bpp down to
-6 bpp, described as following \ref{ss.position} and \ref{ss.velocity}. The algorithm is described in \ref{ss.overview}.
+{\tt CUBE} replaces the coordinates and linked list 24+4=28 bpp memory usage with an integer based storage, reduces the basic memory usage from 28 bpp down to 6 bpp, described as following \ref{ss.position} and \ref{ss.velocity}.
+The algorithm is described in \ref{ss.overview}.
 
 \subsection{Particle position storage}\label{ss.position}
-We construct a uniform mesh throughout the space and each particle belongs 
-to its parent cell of the mesh. Instead of storing global coordinates of 
-each particle, we store its offset relative to its parent cell which 
-contains the particle. We divide the cell, in each dimension $d$, evenly into 
-$2^8=256$ bins, and use a 1-byte (8 bits) integer $\chi_d \in \{-128,-127,...,127\}$ to 
-indicate which bin it locates in this dimension. The global locations of 
-particles are given by cell-ordered format in memory space, and a 
-complimentary number count of particle number in this mesh (density field) 
-will give complete information of particle distribution in the mesh. Then 
-the global coordinate in $d$th dimension $x_d$ is given by $x_d=(n_c-1)+
-(\chi_d+128+1/2)/256$, where $n_c=1,2,...,N_c$ 
-is the index of the coarse grid. The mesh 
-is chosen to be coarse enough such that the density field takes negligible 
-memory. This coarse density field can be further compressed into 1-byte 
-integer format, such that a 1-byte integer show the particle number in this 
-coarse cell in range 0 to 255. In the densest cells (rarely happened) where 
-there are $\ge 255$ particles, we can just write 255, and write the actual 
-number as a 4-byte integer in another file.
+We construct a uniform mesh throughout the space and each particle belongs to its parent cell of the mesh.
+Instead of storing global coordinates of each particle, we store its offset relative to its parent cell which contains the particle.
+We divide the cell, in each dimension $d$, evenly into $2^8=256$ bins, and use a 1-byte (8 bits) integer $\chi_d \in \{-128,-127,...,127\}$ to indicate which bin it locates in this dimension.
+The global locations of particles are given by cell-ordered format in memory space, and a complimentary number count of particle number in this mesh (density field) will give complete information of particle distribution in the mesh.
+Then the global coordinate in $d$th dimension $x_d$ is given by $x_d=(n_c-1)+(\chi_d+128+1/2)/256$, where $n_c=1,2,...,N_c$ is the index of the coarse grid.
+The mesh is chosen to be coarse enough such that the density field takes negligible memory.
+This coarse density field can be further compressed into 1-byte integer format, such that a 1-byte integer show the particle number in this coarse cell in range 0 to 255.
+In the densest cells (rarely happened) where there are $\ge 255$ particles, we can just write 255, and write the actual number as a 4-byte integer in another file.
 
-In a simulation with volume $L^3$ and $N_c^3$ coarse cells, particle 
-positions are stored with a precision of $L/(256N_c)$. The force calculation 
-(e.g. softening length) should be configured much finer than this 
-resolution, discussed in later sections. On the other hand, particle 
-position can also be stored as 2-byte (16 bits) integers to increase the resolution. 
-In this case, each coarse cell is divided into $2^{16}=65536$ bins and the 
-position precision is $L/(65536N_c)$, precise enough compared to using
-4-byte global coordinates, see later results. We denote this case ``x2'' and 
-denote using 1-byte integers for positions ``x1''.
+In a simulation with volume $L^3$ and $N_c^3$ coarse cells, particle positions are stored with a precision of $L/(256N_c)$.
+The force calculation (e.g.
+softening length) should be configured much finer than this resolution, discussed in later sections.
+On the other hand, particle position can also be stored as 2-byte (16 bits) integers to increase the resolution.
+In this case, each coarse cell is divided into $2^{16}=65536$ bins and the position precision is $L/(65536N_c)$, precise enough compared to using4-byte global coordinates, see later results.
+We denote this case ``x2'' and denote using 1-byte integers for positions ``x1''.
 
@@
 \centering
-  \includegraphics[width=1.0\linewidth]{vdisp.pdf}
+  \includegraphics[width=0.88\linewidth]{vdisp.pdf}
  \caption{Variance of the velocity distribution given by Equation \ref{eq.vdisp}.
@@
 \subsection{Particle velocity storage}\label{ss.velocity}
-Similarly, actual velocity in $d$th dimension $v_d$ is decomposed into an 
-averaged velocity field on the same coarse grid $v_c$ and a residual $\Delta v$ 
-relative to this field: 
+Similarly, actual velocity in $d$th dimension $v_d$ is decomposed into an averaged velocity field on the same coarse grid $v_c$ and a residual $\Delta v$ relative to this field: 
 \begin{equation}
@@
 \end{equation}
-$v_c$ is always recorded and 
-kept updated, and occupies negligible memory. We then divide velocity space 
-$\Delta v$ into uneven bins, and use a $n_\nu$-byte integer to indicate
-which $\Delta v$ bin the particle is located.
+$v_c$ is always recorded and kept updated, and occupies negligible memory.
+We then divide velocity space $\Delta v$ into uneven bins, and use a $n_\nu$-byte integer to indicate which $\Delta v$ bin the particle is located.
 
-The reason why we use uneven bins is that, slower particles are more 
-abundant compared to faster ones, and one should better resolve slower 
-particles tracing at least linear evolution. On the other hand, there could 
-be extreme scattering particles (in case of particle-particle force), and we 
-can safely ignore or less resolve those nonphysical particles. One of the 
-solution is that, if we know the probability distribution function (PDF) 
-$f(\Delta v)$ we divide its cumulative distribution function (CDF) $F(\Delta 
-v)\in(0,1)$ into $2^{8n_\nu}$ bins to determine the boundary of $\Delta v$ 
-bins, and particles should evenly distribute in the corresponding uneven $
-\Delta v$ bins. Practically we find that either $f(v_d)$ or $f(\Delta v)$ is 
-close to Gaussian, so we can use Gaussian CDF, or any convenient analytic 
-functions which are close to Gaussian, to convert velocity between real 
-numbers and integers.
+The reason why we use uneven bins is that, slower particles are more abundant compared to faster ones, and one should better resolve slower particles tracing at least linear evolution.
+On the other hand, there could be extreme scattering particles (in case of particle-particle force), and we can safely ignore or less resolve those nonphysical particles.
+One of the solution is that, if we know the probability distribution function (PDF) $f(\Delta v)$ we divide its cumulative distribution function (CDF) $F(\Delta v)\in(0,1)$ into $2^{8n_\nu}$ bins to determine the boundary of $\Delta v$ bins, and particles should evenly distribute in the corresponding uneven $\Delta v$ bins.
+Practically we find that either $f(v_d)$ or $f(\Delta v)$ is close to Gaussian, so we can use Gaussian CDF, or any convenient analytic functions which are close to Gaussian, to convert velocity between real numbers and integers.
 
-The essential parameter of the velocity distribution is its variance. On 
-non-linear scale, the velocity distribution function is non-Gaussian. 
-However, to the first order approximation, we simply assume it as Gaussian 
-and characterized by the variance
+The essential parameter of the velocity distribution is its variance.
+On non-linear scale, the velocity distribution function is non-Gaussian.
+However, to the first order approximation, we simply assume it as Gaussian and characterized by the variance
 \begin{equation}\label{eq.vdisp}
@@
 \end{equation}
-where $a(z)$ is the scale factor, $H(z)$ the Hubble parameter, $D$ is the 
-linear growth factor, $f=d \ln D/d\ln a$, and $P_L(k)$ is the linear power 
-spectrum of density contrast at redshift zero.
-$\sigma_v(a,k)$ is a function of cosmic evolution $a$ and a smoothing scale 
-$k$, or $r$ (see Figure \ref{fig.vdisp}). $\Delta v$
-is the velocity dispersion relative to the coarse grid, so we approximate 
-its variance as
+where $a(z)$ is the scale factor, $H(z)$ the Hubble parameter, $D$ is the linear growth factor, $f=d \ln D/d\ln a$, and $P_L(k)$ is the linear power spectrum of density contrast at redshift zero.
+$\sigma_v(a,k)$ is a function of cosmic evolution $a$ and a smoothing scale $k$, or $r$ (see Figure \ref{fig.vdisp}).
+$\Delta v$is the velocity dispersion relative to the coarse grid, so we approximate its variance as
 \begin{equation}\label{eq.vdelta}
@@
 \end{equation}
-where $r_c$ is the scale of
-coarse grid, and $r_p$ is the scale of average particle separation. In each
-dimension of 3D velocity field, we use $\sigma^2_{\Delta}(a)/3$ according 
-to the equipartition therorem. On different scales, we measure the 
-statistics of $v_d$, $v_c$ and $\Delta v$ and find good agreement with the 
-above model.
+where $r_c$ is the scale of coarse grid, and $r_p$ is the scale of average particle separation.
+In each dimension of 3D velocity field, we use $\sigma^2_{\Delta}(a)/3$ according to the equipartition therorem.
+On different scales, we measure the statistics of $v_d$, $v_c$ and $\Delta v$ and find good agreement with the above model.
 
@@
 \centering
-  \includegraphics[width=0.99\linewidth]{fig_tile.pdf}
+  \includegraphics[width=0.90\linewidth]{fig_tile.pdf}
  \caption{Spacial decomposition in {\tt CUBE} in a 2D analogy.
@@
 where $\lfloor\ \rfloor$ is the operator to take the nearest integer.
-Tangent functions are convenient and compute very fast. Compared to error 
-functions used in Gaussian case, they take the same variance at $\nu_d=0$ but 
-resolve high velocities relatively better.
-Note again that proper choice of conversion formulae and $\sigma_\Delta$ 
-only optimizes the velocity space sampling, but does not affect the physics.
+Tangent functions are convenient and compute very fast.
+Compared to error functions used in Gaussian case, they take the same variance at $\nu_d=0$ but resolve high velocities relatively better.
+Note again that proper choice of conversion formulae and $\sigma_\Delta$ only optimizes the velocity space sampling, but does not affect the physics.
 
@@
 
-{\tt update\_xp} --
-According to {\tt dt}, subroutine {\tt update\_xp} updates the positions of particles in a ``{\it gather}'' algorithm (in contrast, {\tt CUBEP3M} uses ``scatter'' algorithm) tile by tile.
-For each particle, {\tt xp} and {\tt vp} are converted to $x_d$ and $v_d$ by Equations(\ref{eq.x},\ref{eq.v}).
-Because each tile is in the buffered state with buffer depth {\tt ncb}, we are able to collect all possible particles whose $v_d\times{\tt dt}<{\tt ncb}$ from physical+buffer region to its physical region.
-
-In order to keep particles ordered, for each tile, we first do $x_d=x_d+v_d\times{\tt dt}$ on all particles to obtain an updated density and velocity field on the tile, {\tt rho\_c\_new} and {\tt vfield\_new}.
-Then, this calculation in done on the same tile again\footnote{This repetition scales as $o(N)$ and is computational inexpensive.} to generate a new, local particle list {\tt xp\_new} and {\tt vp\_new} (also {\tt PID\_new} if needed) on the tile by Equations(\ref{eq.chi},\ref{eq.nu}).
-Here, the ordering of particles depends on {\tt rho\_c\_new}, and {\tt vfield\_new} is used as $v_c$ in Equation(\ref{eq.nu}).
-Then, another iteration is done on this tile to discard buffer regions of $\{{\tt rho\_c\_new},{\tt xp\_new},{\tt vfield\_new},{\tt vp\_new},{\tt PID\_new}({\rm if\ any})\}$, converting buffer state to disjoint state, and it replaces corresponding part of $\{{\tt rho\_c},{\tt xp},{\tt vfield},{\tt vp},{\tt PID}({\rm if\ any})\}$.
-When all tiles are iterated, the entire image is in disjoint state, and {\tt nplocal} is updated.
-Then all images is synchronized by {\tt sync all} and we sum up {\tt nplocal} over images to check if total number of particles {\tt npglobal} is conserved.
-These steps are summarized in Figure \ref{fig.update_xp}.
-
 \begin{figure}[]
@@
 
+{\tt update\_xp} --
+According to {\tt dt}, subroutine {\tt update\_xp} updates the positions of particles in a ``{\it gather}'' algorithm (in contrast, {\tt CUBEP3M} uses ``scatter'' algorithm) tile by tile.
+For each particle, {\tt xp} and {\tt vp} are converted to $x_d$ and $v_d$ by Equations(\ref{eq.x},\ref{eq.v}).
+Because each tile is in the buffered state with buffer depth {\tt ncb}, we are able to collect all possible particles whose $v_d\times{\tt dt}<{\tt ncb}$ from physical+buffer region to its physical region.
+
+In order to keep particles ordered, for each tile, we first do $x_d=x_d+v_d\times{\tt dt}$ on all particles to obtain an updated density and velocity field on the tile, {\tt rho\_c\_new} and {\tt vfield\_new}.
+Then, this calculation in done on the same tile again\footnote{This repetition scales as $o(N)$ and is computational inexpensive.} to generate a new, local particle list {\tt xp\_new} and {\tt vp\_new} (also {\tt PID\_new} if needed) on the tile by Equations(\ref{eq.chi},\ref{eq.nu}).
+Here, the ordering of particles depends on {\tt rho\_c\_new}, and {\tt vfield\_new} is used as $v_c$ in Equation(\ref{eq.nu}).
+Then, another iteration is done on this tile to discard buffer regions of $\{{\tt rho\_c\_new},{\tt xp\_new},{\tt vfield\_new},{\tt vp\_new},{\tt PID\_new}({\rm if\ any})\}$, converting buffer state to disjoint state, and it replaces corresponding part of $\{{\tt rho\_c},{\tt xp},{\tt vfield},{\tt vp},{\tt PID}({\rm if\ any})\}$.
+When all tiles are iterated, the entire image is in disjoint state, and {\tt nplocal} is updated.
+Then all images is synchronized by {\tt sync all} and we sum up {\tt nplocal} over images to check if total number of particles {\tt npglobal} is conserved.
+These steps are summarized in Figure \ref{fig.update_xp}.
+
+{\tt update\_vp} -- 
+The particle-mesh (PM) or particle-particle particle-mesh (P$^3$M) algorithm is applied in this subroutine to update particles's velocities.
+After {\tt update\_xp}, or drift $D$, we call {\tt buffer\_density} and {\tt buffer\_xp} in order that particle positions are in buffered state, based on which, we update velocities (kick $K$) of particles in physical region of each tile.
+{\tt CUBE} uses a 2-level particle mesh scheme \citep{2005NewA...10..393M,2013MNRAS.436..540H}.
+Local fine forces have a force cutoff ${\tt nforce\_cutoff}\leq{\tt ncb}$, i.e., $F_{\rm fine}(r>{\tt ncb})=0$.
+So, if we apply a fine-grid particle-mesh on an extended tile with buffered-depth {\tt ncb} (see Figure \ref{fig.tile} and regard it periodic), the force on physical regions does not depend on the false periodic boundary assumption, and the resulting fine force $F_{\rm fine}$ and velocity update on physical region is correct.
+
 \begin{figure}[]
@@
 \end{figure}
-
-{\tt update\_vp} -- 
-The particle-mesh (PM) or particle-particle particle-mesh (P$^3$M) algorithm is applied in this subroutine to update particles's velocities.
-After {\tt update\_xp}, or drift $D$, we call {\tt buffer\_density} and {\tt buffer\_xp} in order that particle positions are in buffered state, based on which, we update velocities (kick $K$) of particles in physical region of each tile.
-{\tt CUBE} uses a 2-level particle mesh scheme \citep{2013MNRAS.436..540H}.
-Local fine forces have a force cutoff ${\tt nforce\_cutoff}\leq{\tt ncb}$, i.e., $F_{\rm fine}(r>{\tt ncb})=0$.
-So, if we apply a fine-grid particle-mesh on an extended tile with buffered-depth {\tt ncb} (see Figure \ref{fig.tile} and regard it periodic), the force on physical regions does not depend on the false periodic boundary assumption, and the resulting fine force $F_{\rm fine}$ and velocity update on physical region is correct.
 
@@
 
-\begin{figure*}[t]
+\begin{figure*}[]
 \centering
@@
 \subsubsection{Compare with traditional algorithms}
-To illustrate the improvement of memory usage, we give an example of TianNu simulation \citep{2017NatAs...1E.143Y} on the Tianhe-2 supercomputer, which used a traditional {\tt CUBEP3M} code.
-TianNu's particles number is limited by memory per computing node -- per computing node, an average of $576^3$ neutrino particles and $288^3$ CDM particles are used, and consumes about 40 GB of memory\footnote{Additional memory is used for OpenMP parallelization and for particle IDs to differentiate different particle spices.}, or about 186 bpp.
+To illustrate the improvement of memory usage, we refer to the TianNu simulation \citep{2017NatAs...1E.143Y} on the Tianhe-2 supercomputer, which used a traditional $N$-body code {\tt CUBEP3M}.
+TianNu's particles number is limited by memory per computing node -- for each computing node, an average of $576^3$ neutrino particles and $288^3$ CDM particles are used, and consumes about 40 GB of memory\footnote{Additional memory is used for OpenMP parallelization and for particle IDs to differentiate different particle spices.}, or about 186 bpp.
 A memory efficient case of {\tt CUBEP3M} by using large physical scales and at costs of speed, still uses about 40 bpp.
 
-By using {\tt CUBE}, with parameters $n_\chi=n_\nu=1$, ${\tt nc}=384$, ${\tt nnt}=3$, ${\tt ncb}=6$, $\epsilon_{\rm image}=5\%$, $\epsilon_{\rm tile}=200\%$ and $\left\langle {\tt nplocal} \right\rangle=1536^3$, we use about 35.6 GB of memory, corresponding to 9.84 bpp.
+If the same amount of memory is allocated to {\tt CUBE}, we can set parameters as $n_\chi=n_\nu=1$, ${\tt nc}=384$, ${\tt nnt}=3$, ${\tt ncb}=6$, $\epsilon_{\rm image}=5\%$, $\epsilon_{\rm tile}=200\%$.
+This gives $\left\langle {\tt nplocal} \right\rangle=1536^3$, and uses about 35.6 GB of memory, corresponding to 9.84 bpp.
 This can be done on most of the supercomputers, even modern laptops.
@@
 The position resolution of particles in {\tt CUBE}-x1v1 is 1/256 of a coarse grid, or 1/64 of a fine grid.
+
 \begin{figure}[]
 \centering
-  \includegraphics[width=0.95\linewidth]{ddsp_pcdf_256_box80.pdf}
+  \includegraphics[width=0.99\linewidth]{ddsp_pcdf_256_box80.pdf}
  \caption{Statistics of the error in particle displacement $\Delta{\bs\Psi}$ induced by integer based algorithms in {\tt CUBE}.
@@
 \end{figure}
+
+\begin{figure*}
+\centering
+  \includegraphics[width=1.0\linewidth]{diff_power.pdf}
+ \caption{Fractional deviations in the power spectrum $\Delta^2(k)/\Delta^2_0(k)-1$ (upper axes) and decorrelation $\xi$ (lower axes) induced by the integer-based algorithms in {\tt CUBE}.
+The four panels show the results of two redshifts in S2048S and S2048L: $\{ z=1, z=0 \} \times \{ L=400{\rm Mpc}/h, L=1200{\rm Mpc}/h \}$.
+The four solid lines show the results from \{x2v2,x1v2,x2v1,x1v1\}, whose colors are shown in the legends in the last panel.
+The orange dashed lines show the scale $k=0.2\ k_{\rm Nyquist}$, and smaller scales are beyond our interest.}
+\label{fig.ccc}
+\end{figure*}
+\subsection{Power spectrum}
 
@@
 
-\begin{figure*}
-\centering
-  \includegraphics[width=1.0\linewidth]{diff_power.pdf}
- \caption{Fractional deviations in the power spectrum $\Delta^2(k)/\Delta^2_0(k)-1$ (upper axes) and decorrelation $\xi$ (lower axes) induced by the integer-based algorithms in {\tt CUBE}.
-The four panels show the results of two redshifts in S2048S and S2048L: $\{ z=1, z=0 \} \times \{ L=400{\rm Mpc}/h, L=1200{\rm Mpc}/h \}$.
-The four solid lines show the results from \{x2v2,x1v2,x2v1,x1v1\}, whose colors are shown in the legends in the last panel.
-The orange dashed lines show the scale $k=0.2\ k_{\rm Nyquist}$, and smaller scales are beyond our interest.}
-\label{fig.ccc}
-\end{figure*}
-\subsection{Power spectrum}
 In S2048S and S2048L are two simulations with $2048^3$ particles on small ($L=400\ {\rm Mpc}/h$) and large ($L=1200\ {\rm Mpc}/h$) box sizes.

### 2017-11-06 ms_caf/ms_caf.tex
Ue-Li's revision

@@
 \begin{document}
-\title{CUBE: A Memory-optimized Cosmological $N$-body Algorithm}
+\title{CUBE: An Information-optimized parallel Cosmological $N$-body Algorithm}
 \author{
@@
 Cosmological large scale structure $N$-body simulations are computation-light, memory-heavy problems in supercomputing.
-Traditional $N$-body simulation algorithms use at least 24-byte memory per particle, of which six 4-byte single precision floating point numbers keep track phase space coordinates of each particle.
-Here we present the algorithm and accuracy of a new parallel, memory-lite, Particle-Mesh based $N$-body code, where each particle can occupy as low as 6-byte memory.
-This is accomplished by storing relative position and relative velocity of each particle, in the format of 1-byte-integer, respect to their averaged value of a mesh-grid.
-The remaining information is given by complimentary density and velocity fields, which are negligible in memory space, and proper ordering of particles, which gives no extra memory.
-Our numerical experiments show that this integer based $N$-body algorithm provides acceptable accuracy compared to traditional algorithm in cosmological simulations.
-This significant lowering of memory-to-computation ratio breaks the bottleneck of scaling up and speeding up large cosmological $N$-body simulations on multi-core and heterogenous computing systems.
+\tcb{This considerable amount of memory is occupied by the phase space information of particles.
+We show that keeping only a small fraction of the information gives negligible impact on the accuracy of the simulation.} This algorithm is applied in a new parallel, \tcb{information-optimized}, Particle-Mesh based $N$-body code CUBE, where only 6-byte memory is used to store the information of each particle.
+\tcb{This is accomplished by storing particle's relative phase space coordinates instead of global values, and thus can use a lower accuracy -- in the format of fixed point as light as 1-byte.
+The remaining information is given by complimentary density and velocity fields (negligible in memory space) and proper ordering of particles (no extra memory).} Our numerical experiments show that this \tcb{information-optimized $N$-body algorithm provides accurate results within the error of the Particle-Mesh algorithm.} This significant lowering of memory-to-computation ratio breaks the bottleneck of scaling up and speeding up large cosmological $N$-body simulations on multi-core and heterogenous computing systems.
 
@@
 
-Relative to the optimized computation optimizations, $N$-body simulations use considerable amount of memory, because the phase space coordinates $(x,y,z,v_x,v_y,v_z)$ of each $N$-body particle must be stored as at least six single precision floating point numbers (24 bytes).
+Relative to the optimized computation optimizations, $N$-body simulations use considerable amount of memory to store the \tcb{information of particles}.
+Their phase space coordinates $(x,y,z,v_x,v_y,v_z)$ are stored as at least 6 single-precision floating point numbers (total of 24 bytes).
 On the other hand, modern supercomputer systems use multi cores, many integrated cores (MIC) and even densely parallelized GPUs, bringing orders of magnitude higher computing power, whereas these architectures usually have limited memory allocation.
-Thus, the computation-light but memory-heavy applications, compared to matrix multiplication and decomposition calculations, are less suitable for fully usage of the computing power of modern supercomputers.
-For example, although native and offload modes of {\tt CUBEP3M} are able to run on the Intel Xeon-PHI MIC architectures, with the require ment of enough memory, TianNu simulation were done on TianHe-2 with only its CPUs -- 73\% of the total memory but only 13\% of the total computing power.
+Thus, these computation-light but memory-heavy applications, compared to matrix multiplication and decomposition calculations, are currently less suitable for fully usage of the computing power of modern supercomputers.
+For example, although native and offload modes of {\tt CUBEP3M} are able to run on the Intel Xeon-PHI MIC architectures, with the requirement of enough memory, TianNu simulation were done on TianHe-2 with only its CPUs -- 73\% of the total memory but only 13\% of the total computing power.
+\tcb{We investigate how most amount of information of particles can be deducted whereas still preserves the accuracy of $N$-body simulations, in order that the total memory usage for given $N$ is optimized.}
 
-We present a new $N$-body simulation code {\tt CUBE}, using as low as 6 byte per particle (here after we use ``bpp'' referring to ``byte per particle'').
-We show that it gives accurate results in cosmological LSS simulations.
+In the following sections we present a new, \tcb{information-optimized} parallel cosmological $N$-body simulation code {\tt CUBE}, using as low as 6 byte per particle.
+It gives accurate results in cosmological LSS simulations -- \tcb{the error induced by information optimization is below the error from the PM algorithm}.
 In \S\ref{s.method} we show how the memory can be saved by using a ``integer-based storage'' and how the PM $N$-body algorithm is adapted with this storage format.
@@
 \section{Method}\label{s.method}
-The most memory consuming part of a $N$-body simulation is usually the phase space coordinates of $N$-body particles -- 24 bpp (4 single precision floating numbers) must be used to store each particles' 3D position and velocity vectors.
+The most memory consuming part of a $N$-body simulation is usually the phase space coordinates of $N$-body particles -- 24 byte per particle \tcb{(here after we use ``bpp'' referring to ``byte per particle'')}, which contains 4 single precision floating numbers, must be used to store each particles' 3-dimensional position and velocity vectors.
 {\tt CUBEP3M}, an example of a memory-efficient parallel $N$-body code, can use as low as 40 bpp in sacrificing computing speed \citep{2013MNRAS.436..540H}.
 This includes the phase coordinates (24 bpp) for particles in physical domain and buffered region, a linked list (4 bpp), and a global coarse mesh and local fine mesh.
-4-byte real numbers are not necessarily adequate in representing the {\it global} coordinates in simulations.If the box size is many orders of magnitude larger than interactive distance between particles, especially in the field of re-simulation of dense subregions, double precision (8-byte) coordinates are needed to avoid truncation errors.
+Sometimes 4-byte real numbers are not necessarily adequate in representing the {\it global} coordinates in simulations.
+If the box size is many orders of magnitude larger than interactive distance between particles, especially in the field of re-simulation of dense subregions, double precision (8-byte) coordinates are needed to avoid truncation errors.
 Another solution is to record {\it relative} coordinates for both position and velocity.
@@
 $\sigma_v(a,k)$ is a function of cosmic evolution $a$ and a smoothing scale $k$, or $r$ (see Figure \ref{fig.vdisp}).
-$\Delta v$is the velocity dispersion relative to the coarse grid, so we approximate its variance as
+$\Delta v$ is the velocity dispersion relative to the coarse grid, so we approximate its variance as
 \begin{equation}\label{eq.vdelta}
@@
 {\tt CUBE} uses a 2-level PM force calculation, same as {\tt CUBEP3M}, However, in order to apply the integer based format to the $N$-body simulation, substantial structural changes need to be done.
-{\tt CUBE} is written in Coarray Fortran, where Coarray features implement MPI communication between computation nodes/images\footnote{Images are the concept of computing nodes or MPI tasks in Coarray Fortran.
+{\tt CUBE} is written in Coarray Fortran, where Coarray features \tcb{replace} MPI communications between computation nodes/images\footnote{Images are the concept of computing nodes or MPI tasks in Coarray Fortran.
 We use this terminology in this paper.}.
@@
 We operate a Runge-Kutta 2 method for time integration.
-i.e., we update position ($D$=drift) and velocity ($K$=kick) every half time step.
+i.e., updating positions ($D$=drift) and velocities ($K$=kick) are at every half time step by \tcb{operator splitting}.
 For $n$ time steps, the operation would be $(DKKD)^n$ which is 2nd order accurate.
@@
 \centering
-\caption{Memory layout for a certain configuration}
+\caption{Memory layout for a \tcb{certain} configuration}
 \label{t.memory}
@@
 We study the fractional power spectrum deviations $\Delta^2(k)/\Delta^2_0(k)-1$ (where $\Delta^2_0(k)$ denotes the {\tt CUBEP3M} power spectrum) and the cross correlation coefficient functions $\xi(k)$ between {\tt CUBE} and {\tt CUBEP3M}, and show the results in Figure \ref{fig.ccc}.
-For the four panels, the left shows S2048S and right shows S2048L; the upper shows $z=1$ and lower shows $z=0$.
-We label $k=0.2\ k_{\rm Nyquist}$ as dashed lines, where $k_{\rm Nyquist}$ is the fine mesh scale, and the average particle separation scale.
-On the scale $k=0.2\ k_{\rm Nyquist}$, the power spectrum is suppressed by $\sim 20\%$ compared to higher resolution simulations by Poisson noise of particles and we do not consider $k>0.2\ k_{\rm Nyquist}$.
+For the four panels, the left two show S2048S and right two shows S2048L; the upper two show $z=1$ and lower two show $z=0$.
+\tcb{We label $k=0.2\ k_{\rm Nyquist}$ as dashed lines, where $k_{\rm Nyquist}$ is the scale of fine mesh grids, and the scale of average particle separations.
+From many simulations we observe that, on the scale $k=0.2\ k_{\rm Nyquist}$, the power spectrum is suppressed by $\sim 20\%$ compared to higher resolution simulations.
+This error is caused by the limited resolution of PM algorithms.
+Our scales of interest should be $k\lesssim 0.2\ k_{\rm Nyquist}$ where PM do not dominate the error, however in Figure \ref{fig.ccc} we still plot $k> 0.2\ k_{\rm Nyquist}$.}
 
 The upper parts of the four panels show the fractional power spectrum deviations.
-Because x2v2 gives exact results as {\tt CUBEP3M}, $\Delta^2_{\rm x2v2}(k)/\Delta^2_0(k)-1=0$ for both simulations and redshifts, shown by the four black straight lines.
+\tcb{For both simulations and redshifts, the fractional power spectrum deviation is less than 1\% on scales of interest, where PM error do not dominate.
+If we just simply observe the scales where PM already contributes significant errors, there are differences between simulations.} Because x2v2 gives exact results as {\tt CUBEP3M} on all scales \tcb{(even including the scale where PM contribute significant error)} $\Delta^2_{\rm x2v2}(k)/\Delta^2_0(k)-1=0$ for both simulations and redshifts, shown by the four black straight lines.
 Similar to the results form Figure \ref{fig.dsp}, x1v2 (red) curves give better results than x2v1 (blue) and x1v1 (green), and x2v1 and x1v1 give nearly same results.
 Another trend is that higher redshifts and/or larger box sizes gives much better results.
-The worst case is S2048S, $z=0$ case -- There is a $<2\%$ deviation in the power spectrum.
+In the case of S2048S $z=0$ it shows more deviations near Nyquist frequency.
 
 The lower parts of the four panels show the cross correlation coefficient function between {\tt CUBE} and {\tt CUBEP3M}.
-Again, x2v2 performs perfect in all the four cases.
-For x1v2, $\xi(k)>99.5\%$ for all cases within our scale of interest.
-For x1v2 and x1v1, $\xi(k)$ are very close, and they perform worst in S2048S, $z=0$ case -- on scale $k=0.2\ k_{\rm Nyquist}$, the correlation drops to $90\%$.
+\tcb{For both redshifts of S2048L and $z=1$ of S2048S, the results show perfect cross correlation with {\tt CUBEP3M}, even for x1v1.
+For $z=0$ of S2048S, only x2v1 and x1v1 cases decorrelate from {\tt CUBEP3M}, on the scale of $k=0.2\ k_{\rm Nyquist}$, where the PM error is $\sim 20\%$, the decorrelation is only by $\sim 5\%$.}
 

### 2017-11-06 ms_caf/ms_caf.tex
conclusion

@@
 {\tt CUBE} uses a 2-level PM force calculation, same as {\tt CUBEP3M}, However, in order to apply the integer based format to the $N$-body simulation, substantial structural changes need to be done.
-{\tt CUBE} is written in Coarray Fortran, where Coarray features \tcb{replace} MPI communications between computation nodes/images\footnote{Images are the concept of computing nodes or MPI tasks in Coarray Fortran.
+{\tt CUBE} is written in Coarray Fortran, where Coarray features \tcb{replace} MPI (Message Passing Interface) communications between computation nodes/images\footnote{Images are the concept of computing nodes or MPI tasks in Coarray Fortran.
 We use this terminology in this paper.}.
@@
 \tcb{For both simulations and redshifts, the fractional power spectrum deviation is less than 1\% on scales of interest, where PM error do not dominate.
-If we just simply observe the scales where PM already contributes significant errors, there are differences between simulations.} Because x2v2 gives exact results as {\tt CUBEP3M} on all scales \tcb{(even including the scale where PM contribute significant error)} $\Delta^2_{\rm x2v2}(k)/\Delta^2_0(k)-1=0$ for both simulations and redshifts, shown by the four black straight lines.
+If we just simply observe the scales where PM already contributes significant errors, there are differences between simulations.} For x2v2, it gives exact results as {\tt CUBEP3M} on all scales \tcb{(even including the scale where PM contribute significant error)} $\Delta^2_{\rm x2v2}(k)/\Delta^2_0(k)-1=0$ for both simulations and redshifts, shown by the four black straight lines.
 Similar to the results form Figure \ref{fig.dsp}, x1v2 (red) curves give better results than x2v1 (blue) and x1v1 (green), and x2v1 and x1v1 give nearly same results.
@@
 
+\tcb{To summarize from Figure \ref{fig.ccc}, in terms of either power spectrum deviation or cross correlation, the error induced by information optimization (even for x1v1) is lower than the error from the PM algorithm, and we can safely use x1v1 for most of the LSS studies.
+If one still wants to compare the results when PM already contributes significant error ($k\gtrsim 0.2\ k_{\rm Nyquist}$), errors are from nonlinear (lower $m_p$ and/or lower redshift) and are primarily from compressing the velocity information (x2v1 and x1v1).}
+
 \section{Discussion and conclusion}\label{s.discussion}
-We present an parallel integer-based $N$-body algorithm.
+We present an parallel, \tcb{information-optimized} $N$-body algorithm.
 This open-source code {\tt CUBE}, has recently been used in many studies of LSS, e.g.
@@
 It needs very low memory usage, approaching 6 byte per particle (6 bpp).
-The accuracy of this 6pp configuration is acceptable on most scales of our interest, e.g.
-BAO.
-The accuracy is also adjustable in that we can choose 1-byte/2-byte integers separately for positions and velocities of particles.
+
+The accuracy of this code is adjustable in that we can choose 1-byte/2-byte integers separately for positions and velocities of particles.
 In the case of using 2-byte integers for both positions and velocities (``x2v2'', and memory can be 12 bpp), the algorithm gives exact results given by traditional $N$-body algorithms.
-
-We analyze the deviation of power spectrum and de-correlation caused by integer-1 based algorithms in two simulations S2048S and S2048L.
-They have comparable number of particles per computing as the TianNu simulation.
-S2048S focuses on mass resolution -- comparable to TianNu and TianZero simulations, whereas S2048L focuses on simulation volume and have high
-er particle mass. In the power spectra and cross correlations, S2048L performs better, and we conclude that in high-$m_p$ cases, we can safely use x1v1 in cosmological simulations.
+\tcb{Note that the results are exactly same in that they not only produce same physical statistics of LSS, but also the same error (not physical) from the PM algorithm near Nyquist frequencies.
+In another word, the positions and velocities of each particle are exact.
+In practice, we only require that the errors from information optimization is much lower than the errors from PM algorithm.
+In Figure \ref{fig.ccc} we see that it is the case even for the most memory efficient configuration x1v1.
+This shows that in most LSS studies, when our scales of interest are smaller than $k\simeq 0.2\ k_{\rm Nyquist}$, six 1-byte fixed point numbers contains sufficient information of every $N$-body particle.}
 
@@
 Traditional $N$-body codes consumes considerable memory while doing a relatively light computations.
-{\tt CUBE} prominently reduces the memory usage in $N$-body simulations and approaches toward a better balance between memory and computation.
-The next steps are optimization of the code and adapting it for various kinds of heterogeneous computing systems, e.g.
+\tcb{{\tt CUBE} is designed to optimize the efficiency of information and the memory usage in $N$-body simulations.
+{\tt CUBE} is written in Coarray Fortran -- the concise Coarray features are used instead of complicated MPI, and the code itself is much more concise than {\tt CUBEP3M} for future maintenance and development.} The next steps are optimization of the code and adapting it for various kinds of heterogeneous computing systems, e.g.
 MIC and GPUs.

### 2017-11-07 ms_caf/ms_caf.tex
minor changes

@@
 \begin{document}
-\title{CUBE: An Information-optimized parallel Cosmological $N$-body Algorithm}
+\title{CUBE: An \tcb{Information-optimized parallel} Cosmological $N$-body Algorithm}
 \author{
@@
 Cosmological large scale structure $N$-body simulations are computation-light, memory-heavy problems in supercomputing.
-\tcb{This considerable amount of memory is occupied by the phase space information of particles.
-We show that keeping only a small fraction of the information gives negligible impact on the accuracy of the simulation.} This algorithm is applied in a new parallel, \tcb{information-optimized}, Particle-Mesh based $N$-body code CUBE, where only 6-byte memory is used to store the information of each particle.
-\tcb{This is accomplished by storing particle's relative phase space coordinates instead of global values, and thus can use a lower accuracy -- in the format of fixed point as light as 1-byte.
-The remaining information is given by complimentary density and velocity fields (negligible in memory space) and proper ordering of particles (no extra memory).} Our numerical experiments show that this \tcb{information-optimized $N$-body algorithm provides accurate results within the error of the Particle-Mesh algorithm.} This significant lowering of memory-to-computation ratio breaks the bottleneck of scaling up and speeding up large cosmological $N$-body simulations on multi-core and heterogenous computing systems.
+\tcb{The considerable amount of memory is usually dominated by an inefficient way of storing more than sufficient phase space information of particles.} We present a new \tcb{parallel, information-optimized}, particle-mesh based $N$-body code CUBE, \tcb{in which information- and memory-efficiency is increased by nearly an order of magnitude.} This is accomplished by storing particle's relative phase space coordinates \tcb{instead of global values, and in the format of fixed point as light as 1-byte.
+The remaining information is given by complimentary density and velocity fields (negligible in memory space) and proper ordering of particles (no extra memory).} Our numerical experiments show that this \tcb{information-optimized $N$-body algorithm provides accurate results within the error of the particle-mesh algorithm.} This significant lowering of memory-to-computation ratio breaks the bottleneck of scaling up and speeding up large cosmological $N$-body simulations on multi-core and heterogenous computing systems.
 
@@
 However the workload can be alleviated to $o(N\log N)$ by many algorithms.
-For example, the ``tree'' algorithms like {\tt GADGET} \citep{2001NewA....6...79S,2005MNRAS.364.1105S} and its simulation ``Millennium'' \citep{2005Natur.435..629S,2012MNRAS.426.2046A}, {\tt TPM} \citep{1995ApJS...98..355X} and {\tt GOTPM} \citep{2004NewA....9..111D}, adaptive grid algorithm including {\tt HYDRA} \citep{1995ApJ...452..797C} and {\tt RAMSES} \citep{2010ascl.soft11007T}, as well as mesh-refined codes \citep{1991ApJ...368L..23C} and moving adaptive particle-mesh codes \citep{1995ApJS..100..269P}.
-Among these, the standard particle-mesh (PM) algorithm \citep{1988csup.book.....H} is most memory and computational efficient if we focus on large cosmological scales.
+For example, the ``tree'' algorithms like {\tt GADGET} \citep{2001NewA....6...79S,2005MNRAS.364.1105S} and its simulation ``Millennium'' \citep{2005Natur.435..629S,2012MNRAS.426.2046A}, {\tt TPM} \citep{1995ApJS...98..355X} and {\tt GOTPM} \citep{2004NewA....9..111D}, adaptive grid algorithm including {\tt HYDRA} \citep{1995ApJ...452..797C} and {\tt RAMSES} \citep{2010ascl.soft11007T}, as well as mesh-refined codes \citep{1991ApJ...368L..23C} and moving adaptive particle-mesh (PM) codes \citep{1995ApJS..100..269P}.
+Among these, the standard PM algorithm \citep{1988csup.book.....H} is most memory and computational efficient if we focus on large cosmological scales.
 The load balancing problem is minimized as the matter distribution is rather homogeneous, and the speed benefits from the fast Fourier transform (FFT) libraries, such as {\tt FFTW3} \citep{Frigo05thedesign}.

### 2017-11-29 ms_caf/ms_caf.tex
Paper: compared with CLASS power, added PM decorrelation

@@
 For example, although native and offload modes of {\tt CUBEP3M} are able to run on the Intel Xeon-PHI MIC architectures, with the requirement of enough memory, TianNu simulation were done on TianHe-2 with only its CPUs -- 73\% of the total memory but only 13\% of the total computing power.
-\tcb{We investigate how most amount of information of particles can be deducted whereas still preserves the accuracy of $N$-body simulations, in order that the total memory usage for given $N$ is optimized.}
+\tcb{We investigate how most amount of information of particles can be deducted whereas it still preserves the accuracy of $N$-body simulations, in order that the total memory usage for given $N$ is optimized.}
 
@@
 First, by using different configurations -- different number of computing nodes, box sizes, particle resolutions, different number of tiles per node/image etc., we find that by using 2-byte integers for both positions and velocities ($n_\chi=n_\nu=2$, or x2v2) {\tt CUBE} gives exact results compared to {\tt CUBEP3M}.
-So, if memory usage is not a problem, one can always double the memory usage to get exact results, and the optimal memory limit of this case is 12 bpp, still much lower than traditional methods.
+So, if memory usage is not a problem, one can always double the memory usage to get exact results as {\tt CUBEP3M}, and the optimal memory limit of this case is 12 bpp, still much lower than traditional methods.
 Next, we focus on the accuracy of the other three cases -- x1v2, x2v1 and x1v1.
@@
 We list the names and configurations of simulations used in Table \ref{t.sim}, where $N_{\rm node}$, $L$, $z_i$, $z_{\rm ckpt}$, $N_p$ and $m_p$ are respectively the number of computing nodes used, length of the side of the box, initial redshift, redshift where we do checkpoint, total number of particles and particle mass.
-These configurations are run by {\tt CUBEP3M}, x2v2, x1v2, x2v1 and x1v1 versions of {\tt CUBE}.
+These configurations are run by {\tt CUBEP3M}, x2v2, x1v2, x2v1 and x1v1 versions of {\tt CUBE} with same initial seeds.
 Using different number of tiles per image gives exact same results.
@@
 The position resolution of particles in {\tt CUBE}-x1v1 is 1/256 of a coarse grid, or 1/64 of a fine grid.
-
-\begin{figure}[]
-\centering
-  \includegraphics[width=0.99\linewidth]{ddsp_pcdf_256_box80.pdf}
- \caption{Statistics of the error in particle displacement $\Delta{\bs\Psi}$ induced by integer based algorithms in {\tt CUBE}.
-The PDF and CDF of $|\Delta{\bs\Psi}|$ from S256 are shown in the upper and lower axes, as functions of fine grids.
-Black, red, blue and green correspond to x2v2, x1v2, x2v1 and x1v1 respectively.
-The grey lines, marked with ``initial'', shows the distribution of actual displacement of particles in {\tt CUBEP3M}, $|{\bs\Psi}_0|$, which is orders of magnitudes larger than $|\Delta{\bs\Psi}|$.
-}
-\label{fig.dsp}
-\end{figure}
-
-\begin{figure*}
-\centering
-  \includegraphics[width=1.0\linewidth]{diff_power.pdf}
- \caption{Fractional deviations in the power spectrum $\Delta^2(k)/\Delta^2_0(k)-1$ (upper axes) and decorrelation $\xi$ (lower axes) induced by the integer-based algorithms in {\tt CUBE}.
-The four panels show the results of two redshifts in S2048S and S2048L: $\{ z=1, z=0 \} \times \{ L=400{\rm Mpc}/h, L=1200{\rm Mpc}/h \}$.
-The four solid lines show the results from \{x2v2,x1v2,x2v1,x1v1\}, whose colors are shown in the legends in the last panel.
-The orange dashed lines show the scale $k=0.2\ k_{\rm Nyquist}$, and smaller scales are beyond our interest.}
-\label{fig.ccc}
-\end{figure*}
-\subsection{Power spectrum}
 
@@
 
-In S2048S and S2048L are two simulations with $2048^3$ particles on small ($L=400\ {\rm Mpc}/h$) and large ($L=1200\ {\rm Mpc}/h$) box sizes.
+\begin{figure}[]
+\centering
+  \includegraphics[width=0.99\linewidth]{ddsp_pcdf_256_box80.pdf}
+ \caption{Statistics of the error in particle displacement $\Delta{\bs\Psi}$ induced by integer based algorithms in {\tt CUBE}.
+The PDF and CDF of $|\Delta{\bs\Psi}|$ from S256 are shown in the upper and lower axes, as functions of fine grids.
+Black, red, blue and green correspond to x2v2, x1v2, x2v1 and x1v1 respectively.
+The grey lines, marked with ``initial'', shows the distribution of actual displacement of particles in {\tt CUBEP3M}, $|{\bs\Psi}_0|$, which is orders of magnitudes larger than $|\Delta{\bs\Psi}|$.
+}
+\label{fig.dsp}
+\end{figure}
+
+\begin{figure*}
+\centering
+  \includegraphics[width=1.0\linewidth]{ccc.pdf}
+ \caption{\tcb{Fractional deviations $\Delta^2(k)/\Delta^2_0(k)-1$ in the power spectra relative to nonlinear predictions $\Delta^2_0(k)$ (upper axes), and {\tt CUBE}'s decorrelations $\xi(k)$ relative to {\tt CUBEP3M} (lower axes) induced by the integer-based algorithms.
+The four panels show the results of two redshifts in S2048S and S2048L: $\{ z=1, z=0 \} \times \{ L=400{\rm Mpc}/h, L=1200{\rm Mpc}/h \}$.
+The four solid lines show the results from \{x2v2,x1v2,x2v1,x1v1\}.
+In lower panels, the dashed grey curves show the decorrelations given by the PM algorithm.
+The orange dashed lines show the scale $k=0.2\ k_{\rm Nyquist}$.}}
+\label{fig.ccc}
+\end{figure*}
+\subsection{Power spectrum}
+
+S2048S and S2048L are two simulations with $2048^3$ particles in small ($L=400\ {\rm Mpc}/h$) and large ($L=1200\ {\rm Mpc}/h$) box sizes.
 We compare their accuracy by their power spectra and their cross correlations with {\tt CUBEP3M} at two checkpoints $z=1$ and $z=0$.
@@
 
-For each simulation the particles are firstly cloud-in-cell (CIC) interpolated onto the fine mesh grid of simulation, and from the density field $\rho$ we define the density contrast $\delta\equiv \rho/\left\langle \rho \right\rangle-1$.
+For each simulation, at redshifts $z=1$ and $z=0$, the particles are firstly cloud-in-cell (CIC) interpolated onto the fine mesh grid of simulation, and from the density field $\rho$ we define the density contrast $\delta\equiv \rho/\left\langle \rho \right\rangle-1$.
 More generally we define the cross power spectrum $P_{\alpha\beta}(k)$ between two fields $\delta_\alpha$ and $\delta_\beta$ ($\delta_\alpha=\delta_\beta$ for auto power spectrum) in Fourier space as
@@
 
-We study the fractional power spectrum deviations $\Delta^2(k)/\Delta^2_0(k)-1$ (where $\Delta^2_0(k)$ denotes the {\tt CUBEP3M} power spectrum) and the cross correlation coefficient functions $\xi(k)$ between {\tt CUBE} and {\tt CUBEP3M}, and show the results in Figure \ref{fig.ccc}.
-For the four panels, the left two show S2048S and right two shows S2048L; the upper two show $z=1$ and lower two show $z=0$.
-\tcb{We label $k=0.2\ k_{\rm Nyquist}$ as dashed lines, where $k_{\rm Nyquist}$ is the scale of fine mesh grids, and the scale of average particle separations.
-From many simulations we observe that, on the scale $k=0.2\ k_{\rm Nyquist}$, the power spectrum is suppressed by $\sim 20\%$ compared to higher resolution simulations.
-This error is caused by the limited resolution of PM algorithms.
-Our scales of interest should be $k\lesssim 0.2\ k_{\rm Nyquist}$ where PM do not dominate the error, however in Figure \ref{fig.ccc} we still plot $k> 0.2\ k_{\rm Nyquist}$.}
+\tcb{We compute the nonlinear prediction of the power spectrum by CLASS \citep{2011JCAP...07..034B}, denoted $\Delta^2_0(k)$, at $z=0,1$.
+Then we study the fractional power spectrum deviations $\Delta^2(k)/\Delta^2_0(k)-1$.
+In the four panels of figure \ref{fig.ccc} we denote the box sizes of the simulation and redshifts at the upper left corners.
+The upper parts of these four panels show $\Delta^2(k)/\Delta^2_0(k)-1$, and the four curves of different colors show the results of x2v2, x1v2, x2v1 and x1v1.
+They almost overlap with each other, and in general agree with the nonlinear predictions.}
 
-The upper parts of the four panels show the fractional power spectrum deviations.
-\tcb{For both simulations and redshifts, the fractional power spectrum deviation is less than 1\% on scales of interest, where PM error do not dominate.
-If we just simply observe the scales where PM already contributes significant errors, there are differences between simulations.} For x2v2, it gives exact results as {\tt CUBEP3M} on all scales \tcb{(even including the scale where PM contribute significant error)} $\Delta^2_{\rm x2v2}(k)/\Delta^2_0(k)-1=0$ for both simulations and redshifts, shown by the four black straight lines.
-Similar to the results form Figure \ref{fig.dsp}, x1v2 (red) curves give better results than x2v1 (blue) and x1v1 (green), and x2v1 and x1v1 give nearly same results.
-Another trend is that higher redshifts and/or larger box sizes gives much better results.
-In the case of S2048S $z=0$ it shows more deviations near Nyquist frequency.
+\tcb{We label $k=0.2k_{\rm Nyquist}$ as vertical dashed lines, where $k_{\rm Nyquist}$ is the scale of fine mesh grids, and the scale of average particle separations.
+On this scale, the power spectra are offset from nonlinear predictions by at least 20\%.
+This error is from the PM algorithm and one has to increase the resolution of the simulation to correct these offsets.
+We do not plot {\tt CUBEP3M} because we examined that {\tt CUBEP3M} and x2v2 of {\tt CUBE} produce same results.
+Thus, the differences between {\tt CUBEP3M} and different integer formats of {\tt CUBE} are negligible compared to the error of the PM algorithm.}
 
-The lower parts of the four panels show the cross correlation coefficient function between {\tt CUBE} and {\tt CUBEP3M}.
-\tcb{For both redshifts of S2048L and $z=1$ of S2048S, the results show perfect cross correlation with {\tt CUBEP3M}, even for x1v1.
-For $z=0$ of S2048S, only x2v1 and x1v1 cases decorrelate from {\tt CUBEP3M}, on the scale of $k=0.2\ k_{\rm Nyquist}$, where the PM error is $\sim 20\%$, the decorrelation is only by $\sim 5\%$.}
+\tcb{In the lower parts of these four panels study the cross correlations.
+This time we in turn compare everything with {\tt CUBEP3M}, to emphasize the decorrelation by different integer formats.
+Firstly, x2v2 is perfectly correlated and cross correlation of x1v2 is nearly 100\% in all cases.
+In order to quantify the PM error in terms of decorrelation, we run x2v2 simulations with same initial seeds, but 8 times more particles, and measure the decorrelation caused by coarser resolutions.
+These results are dashed curves (labeled ``PM'') in the lower parts of these four panels.
+We conclude that cross correlation of x1v1, in all cases, is not worse than the PM errors.}
 

### 2017-11-29 CUBE/utilities/Jmatrix_qspace.f90

@@ -5,6 +5,9 @@ program Jmatrix_qspace
   use powerspectrum
   implicit none
   save
+
+  real,parameter :: l_smooth=30 ! Mpc/h
+  real,parameter :: k_smooth=box/l_smooth
   integer i_redshift,i_dim
   real cube3(3,ng,ng,ng),cube(ng,ng,ng),delta_L(ng,ng,ng)
   real xi(10,nbin)[*]
@@ -71,7 +74,7 @@ contains
         dim_2=mod(dim_1,3)+1
         dim_3=mod(dim_2,3)+1
         pdim=(ekx(dim_1)-1)*(ekx(dim_2)+1)*(ekx(dim_3)+1)/4
-        cgrad(j_dim,i,j,k)=cgrad(j_dim,i,j,k)+pdim*cxyz(i,j,k)
+        cgrad(j_dim,i,j,k) = pdim*cxyz(i,j,k)*exp(-(kr/k_smooth)**2)
       enddo
     enddo
     enddo

### 2017-11-29 CUBE/utilities/cicpower.f90

@@ -68,6 +68,7 @@ program cicpower
     close(10)
 
     rho0=0
+    rho_grid=0
     nlast=0
     do itz=1,nnt
     do ity=1,nnt
@@ -155,7 +156,7 @@ program cicpower
     write(str_i,'(i6)') image
     write(str_z,'(f7.3)') z_checkpoint(cur_checkpoint)
     !open(15,file=output_name('delta_N'),access='stream')
-    open(15,file='../output/universe5/image'//&
+    open(15,file='../output/universe1/image'//&
                   trim(adjustl(str_i))//'/'//trim(adjustl(str_z))//'delta_N'//&
                   output_suffix(),access='stream')
     read(15) rho0

### 2017-11-29 CUBE/utilities/detJ.f90

@@ -81,7 +81,8 @@ program detJmatrix
     enddo
     enddo
     enddo
-
+    
+    print*, output_name('gradphiqinvJ')
     open(10,file=output_name('gradphiqinvJ'),status='replace',access='stream')
       write(10) rgrad(1,:,:,:)
       write(10) rgrad(2,:,:,:)

### 2017-11-29 CUBE/utilities/initial_conditions.f90

@@ -155,9 +155,9 @@ program initial_conditions
   norm=1
   if (head) print*, 'Normalization factor: norm =', norm
   !Delta^2
-  tf(2,:)=tf(2,:)**2.0 * tf(1,:)**(3+n_s) * norm / (2.0*pi**2)
-  tf(3,:)=tf(3,:)**2.0 * tf(1,:)**(3+n_s) * norm / (2.0*pi**2)
-  tf(6,:)=tf(6,:)**2.0 * tf(1,:)**(3+n_s) * norm / (2.0*pi**2)
+  tf(2,:)=tf(2,:)**2.0 * tf(1,:)**(3*0+n_s) * norm! / (2.0*pi**2)
+  tf(3,:)=tf(3,:)**2.0 * tf(1,:)**(3*0+n_s) * norm! / (2.0*pi**2)
+  tf(6,:)=tf(6,:)**2.0 * tf(1,:)**(3*0+n_s) * norm! / (2.0*pi**2)
   !dk
   tf(4,1)=tf(1,2)/2
   do k=2,nk-1
@@ -170,11 +170,18 @@ program initial_conditions
     if (tf(1,k)>kmax) exit
     v8=v8+tf(2,k)*tophat(tf(1,k)*8)**2*tf(4,k)/tf(1,k)
   enddo
+
 !  if (head) print*, 's8**2/v8:', v8, s8**2/v8,nyquest ;stop
   tf(2:3,:)=tf(2:3,:)*(s8**2/v8)*Dgrow(a)**2
 !  tf(2:3,:)= scalar_amp*tf(2:3,:)*Dgrow(a)**2 ! for Xin
   sync all
 
+print*, Dgrow(a)
+  open(11,file='tf.bin',access='stream')
+  write(11) tf(1,:)
+  write(11) tf(2,:)/Dgrow(a)**2/Dgrow(1./6)**2
+  close(11)
+  stop
 
   ! noisemap -------------------------------------
   if (head) print*,'Generating random noise'

### 2017-12-14 ms_caf/ms_caf.tex
resolution analysis and bug fix in IC.

@@
 Then the main $N$-body codes run their own initial conditions to redshift $z=0$.
+\tcb{We use the same force kernels as {\tt CUBEP3M} without PP force.
+Note that near find grid scales there are possibilities to enhance the force kernel to better match the nonlinear power spectrum predictions, however we use the conservative mode of {\tt CUBEP3M} in this paper.
+An extended PP force and an unbiased force matching algorithm will be added to {\tt CUBE}.
+The power spectrum studies are presented in \citet{2013MNRAS.436..540H}, and here we focus on the cross correlations between different integer based methods.}
 
@@
 
-\tcb{We label $k=0.2k_{\rm Nyquist}$ as vertical dashed lines, where $k_{\rm Nyquist}$ is the scale of fine mesh grids, and the scale of average particle separations.
+\tcb{We label $k=0.2\,k_{\rm Nyquist}$ as vertical dashed lines, where $k_{\rm Nyquist}$ is the scale of fine mesh grids, and the scale of average particle separations.
 On this scale, the power spectra are offset from nonlinear predictions by at least 20\%.
@@
 Firstly, x2v2 is perfectly correlated and cross correlation of x1v2 is nearly 100\% in all cases.
-In order to quantify the PM error in terms of decorrelation, we run x2v2 simulations with same initial seeds, but 8 times more particles, and measure the decorrelation caused by coarser resolutions.
+In order to quantify the PM error in terms of decorrelation, we run x2v2 simulations with same initial conditions, but 8 times more particles and cells, and measure the decorrelation caused by coarser resolutions.
 These results are dashed curves (labeled ``PM'') in the lower parts of these four panels.
@@
 \tcb{To summarize from Figure \ref{fig.ccc}, in terms of either power spectrum deviation or cross correlation, the error induced by information optimization (even for x1v1) is lower than the error from the PM algorithm, and we can safely use x1v1 for most of the LSS studies.
-If one still wants to compare the results when PM already contributes significant error ($k\gtrsim 0.2\ k_{\rm Nyquist}$), errors are from nonlinear (lower $m_p$ and/or lower redshift) and are primarily from compressing the velocity information (x2v1 and x1v1).}
+If one still wants to compare the results when PM already contributes significant error ($k\gtrsim 0.2\,k_{\rm Nyquist}$), errors are from nonlinear (lower $m_p$ and/or lower redshift) and are primarily from compressing the velocity information (x2v1 and x1v1).}
 
@@
 In Figure \ref{fig.ccc} we see that it is the case even for the most memory efficient configuration x1v1.
-This shows that in most LSS studies, when our scales of interest are smaller than $k\simeq 0.2\ k_{\rm Nyquist}$, six 1-byte fixed point numbers contains sufficient information of every $N$-body particle.}
+This shows that in most LSS studies, when our scales of interest are smaller than $k\simeq 0.2\,k_{\rm Nyquist}$, six 1-byte fixed point numbers contains sufficient information of every $N$-body particle.}
 

### 2017-12-14 CUBE/main/kernel_f.f90

@@ -1,4 +1,4 @@
-#define FPPKCORR
+!#define FPPKCORR
 subroutine kernel_f
   use variables
   implicit none
@@ -27,11 +27,6 @@ subroutine kernel_f
   fk_table(2,1,1,:)=fk_table(2,1,1,:)*1.6
   fk_table(1,2,1,:)=fk_table(1,2,1,:)*1.6
   fk_table(1,1,2,:)=fk_table(1,1,2,:)*1.6
-  print*, fk_table(1,1,1,:)
-  print*, fk_table(2,1,1,:)
-  print*, fk_table(1,2,1,:)
-  print*, fk_table(1,1,2,:)
-  stop
 #endif
 
   do i_dim=1,3

### 2017-12-14 CUBE/main/parameters.f90

@@ -2,7 +2,7 @@ module parameters
   implicit none
   save
 
-  include 'universe1.fh'
+  include 'universe3.fh'
   !! output directory for both IC and snapshots
   !character(*),parameter :: opath='../output/universe1/'
   !
@@ -31,6 +31,8 @@ module parameters
   ! ngrid /image/dim for pencil-fft
 # ifdef FFTFINE
     integer(8),parameter :: ng=nf ! fine grid fft, for IC, dsp, convert_zip_to_xv
+# elif FFTHALF
+    integer(8),parameter :: ng=nf/2
 # else
     integer(8),parameter :: ng=nc ! coarse grid fft, for N-body main code
 # endif
@@ -55,14 +57,14 @@ module parameters
   real,parameter :: rsoft=0.1 ! PP softening length
   logical,parameter :: np_2n3=.false. ! if there are 2*N**3 particles
   real,parameter :: image_buffer=1.5
-  real,parameter :: tile_buffer=1.5
+  real,parameter :: tile_buffer=2.5
 
   ! cosmological parameters
   real,parameter :: z_i=49.0   ! initial redshift
   real,parameter :: z_i_nu=z_i ! initial redshift for neutrinos
   real,parameter :: a_i=1/(1+z_i) ! initial scale factor
 
-  real,parameter :: box=600.0*nn  ! simulation scale /dim, in unit of Mpc/h
+  real,parameter :: box=75.0*nn  ! simulation scale /dim, in unit of Mpc/h
   real,parameter :: h0=67.74    ! Hubble constant
   real,parameter :: s8=0.8276   ! \sigma_8
   real,parameter :: ratio_nudm_dim=2 ! ratio of number of particles for neutrino/CDM, /dim

### 2017-12-14 CUBE/main/run_mac.sh

@@ -4,24 +4,25 @@ source ../utilities/module_load_mac.sh
 
 cd ../utilities/
 make clean
-make
+make ic.x
+make cicpower.x
+./ic.x
 
 # Main N-body code
 cd ../main/
 make clean
 make
-
-cd ../utilities/
-./ic.x
-cd ../main/
 ./cafcube.x
-./acc.x
 
 cd ../utilities/
-./qspace.x
-./corr.x
-./potential.x # get -grad(phi_q)
-./Jmatrix.x
-./detJ.x # get inv(J), and apply -inv(J)grad(phi_q)
+./cicpower.x
+
+
+#cd ../utilities/
+#./qspace.x
+#./corr.x
+#./potential.x # get -grad(phi_q)
+#./Jmatrix.x
+#./detJ.x # get inv(J), and apply -inv(J)grad(phi_q)
 
 cd ../main/

### 2017-12-14 CUBE/utilities/cicpower.f90

@@ -156,7 +156,7 @@ program cicpower
     write(str_i,'(i6)') image
     write(str_z,'(f7.3)') z_checkpoint(cur_checkpoint)
     !open(15,file=output_name('delta_N'),access='stream')
-    open(15,file='../output/universe1/image'//&
+    open(15,file='../output/universe4/image'//&
                   trim(adjustl(str_i))//'/'//trim(adjustl(str_z))//'delta_N'//&
                   output_suffix(),access='stream')
     read(15) rho0

### 2017-12-14 CUBE/utilities/displacement.f90

@@ -129,7 +129,7 @@ program displacement
     read(15) dsp0(3,1:ng,1:ng,1:ng)
     close(15)
     print*, 'max dsp-dsp0', maxval(abs(dsp-dsp0))
-    
+
   enddo
   print*,'displacement done'
 end

### 2017-12-14 CUBE/utilities/initial_conditions.f90

@@ -1,8 +1,10 @@
 #define mkdir
 #define READ_SEED
 #define WRITE_NOISE
+!#define WRITE_KSPACE
 !#define READ_NOISE
 !#define DO_2LPT
+#define READ_DELTA_L
 
 program initial_conditions
   use pencil_fft
@@ -155,9 +157,9 @@ program initial_conditions
   norm=1
   if (head) print*, 'Normalization factor: norm =', norm
   !Delta^2
-  tf(2,:)=tf(2,:)**2.0 * tf(1,:)**(3*0+n_s) * norm! / (2.0*pi**2)
-  tf(3,:)=tf(3,:)**2.0 * tf(1,:)**(3*0+n_s) * norm! / (2.0*pi**2)
-  tf(6,:)=tf(6,:)**2.0 * tf(1,:)**(3*0+n_s) * norm! / (2.0*pi**2)
+  tf(2,:)=tf(2,:)**2.0 * tf(1,:)**(3+n_s) * norm / (2.0*pi**2)
+  tf(3,:)=tf(3,:)**2.0 * tf(1,:)**(3+n_s) * norm / (2.0*pi**2)
+  tf(6,:)=tf(6,:)**2.0 * tf(1,:)**(3+n_s) * norm / (2.0*pi**2)
   !dk
   tf(4,1)=tf(1,2)/2
   do k=2,nk-1
@@ -176,18 +178,18 @@ program initial_conditions
 !  tf(2:3,:)= scalar_amp*tf(2:3,:)*Dgrow(a)**2 ! for Xin
   sync all
 
-print*, Dgrow(a)
-  open(11,file='tf.bin',access='stream')
-  write(11) tf(1,:)
-  write(11) tf(2,:)/Dgrow(a)**2/Dgrow(1./6)**2
-  close(11)
-  stop
+  !print*, Dgrow(a)
+  !open(11,file='tf.bin',access='stream')
+  !write(11) tf(1,:)
+  !write(11) tf(2,:)/Dgrow(a)**2/Dgrow(1./6)**2
+  !close(11)
+  !stop
 
   ! noisemap -------------------------------------
   if (head) print*,'Generating random noise'
   call random_seed(size=seedsize)
   if (head) print*,'min seedsize =', seedsize
-  seedsize=max(seedsize,12)
+  seedsize=max(seedsize,36)
   allocate(iseed(seedsize))
   allocate(rseed_all(seedsize,nn**3))
[…]

### 2017-12-14 CUBE/utilities/lower_resolution_delta_L.f90

@@ -1,0 +1,28 @@
+! compile with
+! gfortran lower_resolution_delta_L.f90
+implicit none
+
+integer,parameter :: ng=256 ! original resolution
+integer,parameter :: n=128 ! lower resolution
+integer i,j,k
+
+real delta_L(ng,ng,ng)
+real delta_new(n,n,n)
+
+open(11,file='../output/universe3/image1/delta_L_1.bin',status='old',access='stream')
+read(11) delta_L
+close(11)
+
+do k=1,n
+do j=1,n
+do i=1,n
+  delta_new(i,j,k) = 0.125 * sum( delta_L(i*2-1:i*2, j*2-1:j*2, k*2-1:k*2) )
+enddo
+enddo
+enddo
+
+open(11,file='../output/universe4/image1/delta_L_1.bin',status='replace',access='stream')
+write(11) delta_new
+close(11)
+
+end

### 2017-12-14 CUBE/utilities/lower_resolution_delta_k.f90

@@ -1,0 +1,19 @@
+implicit none
+
+integer,parameter :: ng=256 ! original resolution
+integer,parameter :: n=128 ! lower resolution
+
+complex cxyz(ng/2+1,ng,ng)
+complex cnew(n/2+1,n,n)
+
+open(11,file='../output/universe1/image1/delta_k_1.bin',status='old',access='stream')
+read(11) cxyz
+close(11)
+
+
+
+open(11,file='../output/universe2/image1/delta_k_1.bin',status='old',access='stream')
+read(11) cnew
+close(11)
+
+end

### 2017-12-16 CUBE/main/parameters.f90
tried to run CDM IC in CUBEnu

@@ -2,7 +2,7 @@ module parameters
   implicit none
   save
 
-  include 'universe3.fh'
+  include 'universe4.fh'
   !! output directory for both IC and snapshots
   !character(*),parameter :: opath='../output/universe1/'
   !
@@ -20,7 +20,7 @@ module parameters
   integer(8),parameter :: nn=1 ! number of imgages (nodes) /dim
   integer(8),parameter :: ncell=4 ! number of nf in each nc, /dim
   integer(8),parameter :: nnt=2 ! number of tiles /image/dim
-  integer(8),parameter :: nc=64 ! nc/image/dim, in physical volume, >=24
+  integer(8),parameter :: nc=32 ! nc/image/dim, in physical volume, >=24
   integer(8),parameter :: nt=nc/nnt ! nc/tile/dim, in physical volume, >=12
 
   integer(8),parameter :: nf=nc*ncell ! >=96
@@ -64,7 +64,7 @@ module parameters
   real,parameter :: z_i_nu=z_i ! initial redshift for neutrinos
   real,parameter :: a_i=1/(1+z_i) ! initial scale factor
 
-  real,parameter :: box=75.0*nn  ! simulation scale /dim, in unit of Mpc/h
+  real,parameter :: box=100.0*nn  ! simulation scale /dim, in unit of Mpc/h
   real,parameter :: h0=67.74    ! Hubble constant
   real,parameter :: s8=0.8276   ! \sigma_8
   real,parameter :: ratio_nudm_dim=2 ! ratio of number of particles for neutrino/CDM, /dim

### 2017-12-16 CUBEnu/main/checkpoint.f90

@@ -24,7 +24,7 @@ subroutine checkpoint
   sim%cur_halo=cur_checkpoint
 
   sim%mass_p=mass_p
-  sim%vsim2phys=(1.5/a)*box*h0*sqrt(omega_m)/nf_global
+  sim%vsim2phys=(1.5/a)*box*h0*100.*sqrt(omega_m)/nf_global
 
   num_io=(nplocal-1)/blocksize+1
 

### 2017-12-16 CUBEnu/main/parameters.f90

@@ -2,13 +2,13 @@ module parameters
   implicit none
   save
 
-  include 'universe1.fh'
-  !! output directory for both IC and snapshots
-  !character(*),parameter :: opath='../output/universe1/'
-  !
-  !! simulation parameters
-  !integer(8),parameter :: izipx=2
-  !integer(8),parameter :: izipv=2
+  ! output directory for both IC and snapshots
+  character(*),parameter :: opath='../output/universe1/'
+
+  ! simulation parameters
+  integer(8),parameter :: izipx=2 ! size to store xp as
+  integer(8),parameter :: izipv=2 ! size to store vp as
+  integer(8), parameter :: izipi = 2 ! if neutrino ids are on, size to store as
 
   integer(8),parameter :: nvbin=int(2,8)**(8*izipv)
   integer(8),parameter :: ishift=-(int(2,8)**(izipx*8-1))
@@ -53,14 +53,14 @@ module parameters
   integer(8),parameter :: np_nc=ncell ! number of particles / coarse cell / dim
 
   real,parameter :: rsoft=0.1 ! PP softening length
-  logical,parameter :: np_2n3=.false. ! if there are 2*N**3 particles
+  logical,parameter :: np_2n3=.false. ! if there are 2*N**3 particles, body-centered cubic
   real,parameter :: image_buffer=2.0
   real,parameter :: tile_buffer=3.0
 
   real,parameter :: pi=4*atan(1.)
 
   ! cosmological parameters
-  real,parameter :: box=300.0*nn  ! simulation scale /dim, in unit of Mpc/h
+  real,parameter :: box=100.0*nn  ! simulation scale /dim, in unit of Mpc/h
   real,parameter :: s8=0 !not used
 
   real,parameter :: z_i=49.0   ! initial redshift
@@ -72,7 +72,7 @@ module parameters
   ! neutrino parameters
   real, parameter :: Tcmb = 2.7255
   real, parameter :: Tcnb = (4./11.)**(1./3.)*Tcmb ! temperature for active neutrinos
-  
+
   integer, parameter :: Nmnu = 3 ! number of massive neutrinos
   real, dimension(Nmnu), parameter :: Mmnu = (/ 0.05,0.05,0.05  /)
   real, dimension(Nmnu), parameter :: Tmnu = (/ Tcnb,Tcnb,Tcnb /)
@@ -83,12 +83,11 @@ module parameters
[…]

### 2017-12-16 CUBEnu/main/pm.f90

@@ -226,7 +226,7 @@ subroutine particle_mesh
   enddo
   enddo
   enddo
-  sim%vsim2phys=(1.5/a)*box*h0*sqrt(omega_m)/nf_global
+  sim%vsim2phys=(1.5/a)*box*h0*100.*sqrt(omega_m)/nf_global
   sync all
 
   if (head) print*, '  constrain dt'

### 2017-12-16 CUBEnu/main/run_fnl.sh

@@ -1,16 +1,0 @@
-source ../utilities/module_load_mac.sh
-
-cp parameters.f90_fnl1 parameters.f90
-source run_mac.sh > log1
-
-cp parameters.f90_fnl2 parameters.f90
-source run_mac.sh > log2
-
-cp parameters.f90_fnl3 parameters.f90
-source run_mac.sh > log3
-
-cp parameters.f90_fnl4 parameters.f90
-source run_mac.sh > log4
-
-cp parameters.f90_fnl5 parameters.f90
-source run_mac.sh > log5

### 2017-12-16 CUBEnu/main/run_timestep.sh

@@ -1,13 +1,0 @@
-source ../utilities/module_load_mac.sh
-
-cp parameters.f90_1 parameters.f90
-source run_mac.sh > log1
-
-cp parameters.f90_2 parameters.f90
-source run_mac.sh > log2
-
-cp parameters.f90_3 parameters.f90
-source run_mac.sh > log3
-
-cp parameters.f90_4 parameters.f90
-source run_mac.sh > log4

### 2017-12-16 CUBEnu/utilities/cicpower.f90

@@ -44,7 +44,7 @@ program cicpower
   sync all
 
   call create_penfft_plan
-  
+
   do cur_checkpoint= 1,n_checkpoint
     if (head) print*, 'Start analyzing redshift ',z2str(z_checkpoint(cur_checkpoint))
     open(12,file=output_name('zip2'),status='old',action='read',access='stream')
@@ -68,6 +68,7 @@ program cicpower
     close(10)
 
     rho0=0
+    rho_grid=0
     nlast=0
     do itz=1,nnt
     do ity=1,nnt
@@ -120,7 +121,7 @@ program cicpower
     rho1=rho_grid(1:ng,1:ng,1:ng)
     !print*, 'check: min,max,sum of rho_grid = '
     !print*, minval(rho1),maxval(rho1),sum(rho1*1d0)
-    
+
     rho8=sum(rho1*1d0)
     sync all
 
@@ -150,12 +151,12 @@ program cicpower
     sync all
 
 
-    
+
     ! cross correlate correct density field
     write(str_i,'(i6)') image
     write(str_z,'(f7.3)') z_checkpoint(cur_checkpoint)
     !open(15,file=output_name('delta_N'),access='stream')
-    open(15,file='../output/universe5/image'//&
+    open(15,file='../output/universe1/image'//&
                   trim(adjustl(str_i))//'/'//trim(adjustl(str_z))//'delta_N'//&
                   output_suffix(),access='stream')
     read(15) rho0

### 2017-12-16 CUBEnu/utilities/compile_dsp.sh

@@ -1,8 +1,0 @@
-$FC $OFLAG ../main/parameters.f90 -DFFTFINE
-$FC $OFLAG ../main/pencil_fft.f90 $FFTFLAG
-$FC $OFLAG powerspectrum.f90 $FFTFLAG
-$FC $OFLAG displacement.f90 $FFTFLAG
-
-$FC $XFLAG parameters.o pencil_fft.o powerspectrum.o displacement.o $FFTFLAG
-
-export FOR_COARRAY_NUM_IMAGES=1

### 2017-12-16 CUBEnu/utilities/compile_init.sh

@@ -1,7 +1,0 @@
-$FC $OFLAG ../main/parameters.f90 -DFFTFINE
-$FC $OFLAG ../main/pencil_fft.f90 $FFTFLAG
-$FC $OFLAG initial_conditions.f90 $FFTFLAG
-
-$FC $XFLAG parameters.o pencil_fft.o initial_conditions.o $FFTFLAG
-
-export FOR_COARRAY_NUM_IMAGES=1

### 2017-12-16 CUBEnu/utilities/convert_zip_to_xv.f90

@@ -1,243 +1,0 @@
-! convert zip format into xv format (for DTFE & Voronoi)
-! compute CIC density fields: delta_nbody.dat
-! compute Voronoi density fields: delta_voronoi.dat
-program convert_zip_to_xv
-use parameters
-implicit none
-
-integer(8) i,j,k,l,nplocal,nlast
-integer(8),parameter :: npnode=nf**3
-real,parameter :: density_buffer=1.1
-integer(8),parameter :: npmax=npnode*density_buffer
-integer(8) ind,dx,dxy,kg,mg,jg,ig,i0,j0,k0,itx,ity,itz,idx,imove,nf_shake,ibin
-
-integer(4) rhoc(nt,nt,nt,nnt,nnt,nnt)
-real rho_f(0:ng+1,0:ng+1,0:ng+1)
-integer(8) idx1(3),idx2(3),ip,np
-real mass_p,dx1(3),dx2(3),xv(6,npmax),xpos(3)
-integer(izipx) x(3,npmax)
-integer(izipv) v(3,npmax)
-
-integer(8) pp,i_s,j_s,k_s,ii_s,jj_s,kk_s,r
-integer(8) wp(npmax),pp_s
-integer(8) hoc_g(ng,ng,ng),ll_p(npmax) ! linked list
-integer(8) hoc_p(npmax),ll_g(ng**3) ! inverse linked list
-real r2,r2min,gpos(3),hpos(3),dpos(3)
-real(8) rho_tot
-logical search
-
-call geometry
-
-if (head) then
-  print*, 'convert zip to xv'
-  print*, 'checkpoint at:'
-  open(16,file='../main/redshifts.txt',status='old')
-  do i=1,nmax_redshift
-    read(16,end=71,fmt='(f8.4)') z_checkpoint(i)
-    print*, z_checkpoint(i)
-  enddo
-  71 n_checkpoint=i-1
-  close(16)
-endif
-
-
-do cur_checkpoint=1,n_checkpoint
-  open(12,file=output_name('zip2'),status='old',action='read',access='stream')
-  read(12) sim
-  ! check zip format and read rhoc
-  if (sim%izipx/=izipx .or. sim%izipv/=izipv) then
-    print*, 'zip format incompatable'
-    close(12)
-    stop
[…]

### 2017-12-16 CUBEnu/utilities/module_load_mac.sh

@@ -1,8 +1,8 @@
-export FC='mpif90'
+export FC='gfortran'
 #export XFLAG='-O3 -cpp -fcoarray=single -mcmodel=medium'
 export XFLAG='-O3 -cpp -fcoarray=single'
 #export XFLAG='-O3 -cpp -fcoarray=single -fopenmp'
 export OFLAG=${XFLAG}' -c'
-export FFTFLAG='-lfftw3f -lm -ldl'
+export FFTFLAG='-I/usr/local/include/ -L/usr/local/lib/ -lfftw3f -lm -ldl'
 # -fopenmp cause (maybe memory) probelm: Segmentation fault: 11
 # in cumsum6

### 2017-12-17 ms_caf/ms_caf.tex
ms_caf submitted version

@@
 \begin{document}
-\title{CUBE: An \tcb{Information-optimized parallel} Cosmological $N$-body Algorithm}
+\title{CUBE: An Information-optimized parallel Cosmological $N$-body Algorithm}
 \author{
@@
 Cosmological large scale structure $N$-body simulations are computation-light, memory-heavy problems in supercomputing.
-\tcb{The considerable amount of memory is usually dominated by an inefficient way of storing more than sufficient phase space information of particles.} We present a new \tcb{parallel, information-optimized}, particle-mesh based $N$-body code CUBE, \tcb{in which information- and memory-efficiency is increased by nearly an order of magnitude.} This is accomplished by storing particle's relative phase space coordinates \tcb{instead of global values, and in the format of fixed point as light as 1-byte.
-The remaining information is given by complimentary density and velocity fields (negligible in memory space) and proper ordering of particles (no extra memory).} Our numerical experiments show that this \tcb{information-optimized $N$-body algorithm provides accurate results within the error of the particle-mesh algorithm.} This significant lowering of memory-to-computation ratio breaks the bottleneck of scaling up and speeding up large cosmological $N$-body simulations on multi-core and heterogenous computing systems.
+The considerable amount of memory is usually dominated by an inefficient way of storing more than sufficient phase space information of particles.
+We present a new parallel, information-optimized, particle-mesh based $N$-body code CUBE, in which information- and memory-efficiency is increased by nearly an order of magnitude.
+This is accomplished by storing particle's relative phase space coordinates instead of global values, and in the format of fixed point as light as 1-byte.
+The remaining information is given by complimentary density and velocity fields (negligible in memory space) and proper ordering of particles (no extra memory).
+Our numerical experiments show that this information-optimized $N$-body algorithm provides accurate results within the error of the particle-mesh algorithm.
+This significant lowering of memory-to-computation ratio breaks the bottleneck of scaling up and speeding up large cosmological $N$-body simulations on multi-core and heterogenous computing systems.
 
@@
 
-Relative to the optimized computation optimizations, $N$-body simulations use considerable amount of memory to store the \tcb{information of particles}.
+Relative to the optimized computation optimizations, $N$-body simulations use considerable amount of memory to store the information of particles.
 Their phase space coordinates $(x,y,z,v_x,v_y,v_z)$ are stored as at least 6 single-precision floating point numbers (total of 24 bytes).
@@
 For example, although native and offload modes of {\tt CUBEP3M} are able to run on the Intel Xeon-PHI MIC architectures, with the requirement of enough memory, TianNu simulation were done on TianHe-2 with only its CPUs -- 73\% of the total memory but only 13\% of the total computing power.
-\tcb{We investigate how most amount of information of particles can be deducted whereas it still preserves the accuracy of $N$-body simulations, in order that the total memory usage for given $N$ is optimized.}
+We investigate how most amount of information of particles can be deducted whereas it still preserves the accuracy of $N$-body simulations, in order that the total memory usage for given $N$ is optimized.
 
-In the following sections we present a new, \tcb{information-optimized} parallel cosmological $N$-body simulation code {\tt CUBE}, using as low as 6 byte per particle.
-It gives accurate results in cosmological LSS simulations -- \tcb{the error induced by information optimization is below the error from the PM algorithm}.
+In the following sections we present a new, information-optimized parallel cosmological $N$-body simulation code {\tt CUBE}, using as low as 6 byte per particle.
+It gives accurate results in cosmological LSS simulations -- the error induced by information optimization is below the error from the PM algorithm.
 In \S\ref{s.method} we show how the memory can be saved by using a ``integer-based storage'' and how the PM $N$-body algorithm is adapted with this storage format.
@@
 \section{Method}\label{s.method}
-The most memory consuming part of a $N$-body simulation is usually the phase space coordinates of $N$-body particles -- 24 byte per particle \tcb{(here after we use ``bpp'' referring to ``byte per particle'')}, which contains 4 single precision floating numbers, must be used to store each particles' 3-dimensional position and velocity vectors.
+The most memory consuming part of a $N$-body simulation is usually the phase space coordinates of $N$-body particles -- 24 byte per particle (here after we use ``bpp'' referring to ``byte per particle''), which contains 4 single precision floating numbers, must be used to store each particles' 3-dimensional position and velocity vectors.
 {\tt CUBEP3M}, an example of a memory-efficient parallel $N$-body code, can use as low as 40 bpp in sacrificing computing speed \citep{2013MNRAS.436..540H}.
@@
 \centering
-  \includegraphics[width=0.88\linewidth]{vdisp.pdf}
+  \includegraphics[width=1\linewidth]{f1}
  \caption{Variance of the velocity distribution given by Equation \ref{eq.vdisp}.
@@
 \centering
-  \includegraphics[width=0.90\linewidth]{fig_tile.pdf}
+  \includegraphics[width=0.9\linewidth]{f2}
  \caption{Spacial decomposition in {\tt CUBE} in a 2D analogy.
@@
 {\tt CUBE} uses a 2-level PM force calculation, same as {\tt CUBEP3M}, However, in order to apply the integer based format to the $N$-body simulation, substantial structural changes need to be done.
-{\tt CUBE} is written in Coarray Fortran, where Coarray features \tcb{replace} MPI (Message Passing Interface) communications between computation nodes/images\footnote{Images are the concept of computing nodes or MPI tasks in Coarray Fortran.
+{\tt CUBE} is written in Coarray Fortran, where Coarray features replace MPI (Message Passing Interface) communications between computation nodes/images\footnote{Images are the concept of computing nodes or MPI tasks in Coarray Fortran.
 We use this terminology in this paper.}.
@@
 We operate a Runge-Kutta 2 method for time integration.
-i.e., updating positions ($D$=drift) and velocities ($K$=kick) are at every half time step by \tcb{operator splitting}.
+i.e., updating positions ($D$=drift) and velocities ($K$=kick) are at every half time step by operator splitting.
 For $n$ time steps, the operation would be $(DKKD)^n$ which is 2nd order accurate.
@@
 \centering
-\caption{Memory layout for a \tcb{certain} configuration}
+\caption{Memory layout for a certain configuration}
 \label{t.memory}
@@
 \centering
-  \includegraphics[width=0.75\linewidth]{overplotparticles.pdf}
+  \includegraphics[width=0.75\linewidth]{f7}
  \caption{Offset in the particle distribution induced by integer-1 based algorithm (x1v1).
@@
 \label{t.sim}
-\begin{tabular}{lrrrrrr}
+\begin{tabular}{lrrrrr}
 \hline
-& \multicolumn{6}{c}{Configurations}\\
-\cline{2-7}
-Name  & $N_{\rm node}$ & $L/({\rm Mpc}/h)$ & $z_i$ & $z_{\rm ckpt}$ & $N_p$ & $m_p/\Msun$ \\
+& \multicolumn{5}{c}{Configurations}\\
+\cline{2-6}
+Name  & $N_{\rm node}$ & $L/({\rm Mpc\,}h^{-1})$ & $z_i$ & $N_p$ & $m_p/\Msun$ \\
 \hline
-S512   & 8     & 200   & 49 & 0    & $512^3$  & $7.5\times 10^9$    \\
-S256   & 1     & 80    & 49 & 0    & $256^3$  & $3.8\times 10^9$    \\
-S2048S & 64    & 400   & 49 & 1,0  & $2048^3$ & $9.4\times 10^8$    \\
-S2048L & 64    & 1200  & 49 & 1,0  & $2048^3$ & $2.5\times 10^{10}$ \\
+S512   & 8     & 200   & 49   & $512^3$  & $7.5\times 10^9$    \\
+S256   & 1     & 80    & 49   & $256^3$  & $3.8\times 10^9$    \\
+S2048S & 64    & 400   & 49   & $2048^3$ & $9.4\times 10^8$    \\
+S2048L & 64    & 1200  & 49   & $2048^3$ & $2.5\times 10^{10}$ \\
 \hline
-TianNu & 13824 & 1200  & 100& 0.01 & $6912^3$ & $6.9\times 10^8$\\
-       &       &       & 5  & 0.01 & $13824^3$& $3.2\times 10^5$\\
-TianZero & 13824 & 1200 & 100& 0.01 & $6912^3$ & $7.0\times 10^8$\\
+TianNu & 13824 & 1200  & 100  & $6912^3$ & $6.9\times 10^8$\\
+       &       &       & 5    & $13824^3$& $3.2\times 10^5$\\
+TianZero & 13824 & 1200 & 100 & $6912^3$ & $7.0\times 10^8$\\
 \hline
@@
 \end{table}
+
+\begin{figure}[]
+\centering
+  \includegraphics[width=0.99\linewidth]{f8}
+ \caption{Statistics of the error in particle displacement $\Delta{\bs\Psi}$ induced by integer based algorithms in {\tt CUBE}.
+The PDF and CDF of $|\Delta{\bs\Psi}|$ from S256 are shown in the upper and lower axes, as functions of fine grids.
+Black, red, blue and green correspond to x2v2, x1v2, x2v1 and x1v1 respectively.
+The grey lines, marked with ``initial'', shows the distribution of actual displacement of particles in {\tt CUBEP3M}, $|{\bs\Psi}_0|$, which is orders of magnitudes larger than $|\Delta{\bs\Psi}|$.
+}
+\label{fig.dsp}
+\end{figure}
 
@@
 Then the main $N$-body codes run their own initial conditions to redshift $z=0$.
-\tcb{We use the same force kernels as {\tt CUBEP3M} without PP force.
+We use the same force kernels as {\tt CUBEP3M} without PP force.
 Note that near find grid scales there are possibilities to enhance the force kernel to better match the nonlinear power spectrum predictions, however we use the conservative mode of {\tt CUBEP3M} in this paper.
 An extended PP force and an unbiased force matching algorithm will be added to {\tt CUBE}.
-The power spectrum studies are presented in \citet{2013MNRAS.436..540H}, and here we focus on the cross correlations between different integer based methods.}
+The power spectrum studies are presented in \citet{2013MNRAS.436..540H}, and here we focus on the cross correlations between different integer based methods.
 
@@
 
-We list the names and configurations of simulations used in Table \ref{t.sim}, where $N_{\rm node}$, $L$, $z_i$, $z_{\rm ckpt}$, $N_p$ and $m_p$ are respectively the number of computing nodes used, length of the side of the box, initial redshift, redshift where we do checkpoint, total number of particles and particle mass.
+We list the names and configurations of simulations used in Table \ref{t.sim}, where $N_{\rm node}$, $L$, $z_i$, $N_p$ and $m_p$ are respectively the number of computing nodes used, length of the side of the box, initial redshift, total number of particles and particle mass.
 These configurations are run by {\tt CUBEP3M}, x2v2, x1v2, x2v1 and x1v1 versions of {\tt CUBE} with same initial seeds.
@@
 We also list the configurations for TianNu and TianZero \citep{2017NatAs...1E.143Y,2017RAA....17...85E} simulations as a reference.
+
+\begin{figure*}
+\centering
+  \includegraphics[width=1.0\linewidth]{f9}
+ \caption{Dimensionless power spectra $\Delta^2(k)$ and decorrelations $\xi(k)$ (lower axes) induced by the integer-based algorithms in simulations S2048S and S2048L.
+The four solid lines show the results from \{x2v2,x1v2,x2v1,x1v1\}.
+In lower panels, the dashed curves show the decorrelations given by the PM algorithm, measured by cross correlating different resolutions of {\tt CUBE}-x2v2.
+The vertical orange dashed lines show the scale $k=0.2\, k_{\rm Nyquist}$.}
+\label{fig.ccc}
+\end{figure*}
+\subsection{Power spectrum}
 
@@
 
-\begin{figure}[]
-\centering
-  \includegraphics[width=0.99\linewidth]{ddsp_pcdf_256_box80.pdf}
- \caption{Statistics of the error in particle displacement $\Delta{\bs\Psi}$ induced by integer based algorithms in {\tt CUBE}.
-The PDF and CDF of $|\Delta{\bs\Psi}|$ from S256 are shown in the upper and lower axes, as functions of fine grids.
-Black, red, blue and green correspond to x2v2, x1v2, x2v1 and x1v1 respectively.
-The grey lines, marked with ``initial'', shows the distribution of actual displacement of particles in {\tt CUBEP3M}, $|{\bs\Psi}_0|$, which is orders of magnitudes larger than $|\Delta{\bs\Psi}|$.
-}
-\label{fig.dsp}
-\end{figure}
-
-\begin{figure*}
-\centering
-  \includegraphics[width=1.0\linewidth]{ccc.pdf}
- \caption{\tcb{Fractional deviations $\Delta^2(k)/\Delta^2_0(k)-1$ in the power spectra relative to nonlinear predictions $\Delta^2_0(k)$ (upper axes), and {\tt CUBE}'s decorrelations $\xi(k)$ relative to {\tt CUBEP3M} (lower axes) induced by the integer-based algorithms.
-The four panels show the results of two redshifts in S2048S and S2048L: $\{ z=1, z=0 \} \times \{ L=400{\rm Mpc}/h, L=1200{\rm Mpc}/h \}$.
-The four solid lines show the results from \{x2v2,x1v2,x2v1,x1v1\}.
-In lower panels, the dashed grey curves show the decorrelations given by the PM algorithm.
-The orange dashed lines show the scale $k=0.2\ k_{\rm Nyquist}$.}}
-\label{fig.ccc}
-\end{figure*}
-\subsection{Power spectrum}
-
 S2048S and S2048L are two simulations with $2048^3$ particles in small ($L=400\ {\rm Mpc}/h$) and large ($L=1200\ {\rm Mpc}/h$) box sizes.
-We compare their accuracy by their power spectra and their cross correlations with {\tt CUBEP3M} at two checkpoints $z=1$ and $z=0$.
-The physical scale of S2048S is designed such that the particle mass resolution $m_p$ is comparable to TianNu and TianZero simulations (their parameters are also listed in Table \ref{t.sim}) to study halo properties.
+We compare their accuracy by their power spectra and their cross correlations with {\tt CUBEP3M} at $z=0$.
+The physical scale of S2048S is designed such that the particle mass resolution $m_p$ is comparable to TianNu and TianZero simulations (their parameters are also listed in Table \ref{t.sim}).
 On the other hand, S2048L focuses on larger structures, on which scale one can study weak gravitational lensing, baryonic acoustic oscillation (BAO) \citep{2005ApJ...633..560E} and its reconstruction \citep{2007ApJ...664..675E,2017ApJ...841L..29W} etc.
 
-For each simulation, at redshifts $z=1$ and $z=0$, the particles are firstly cloud-in-cell (CIC) interpolated onto the fine mesh grid of simulation, and from the density field $\rho$ we define the density contrast $\delta\equiv \rho/\left\langle \rho \right\rangle-1$.
-More generally we define the cross power spectrum $P_{\alpha\beta}(k)$ between two fields $\delta_\alpha$ and $\delta_\beta$ ($\delta_\alpha=\delta_\beta$ for auto power spectrum) in Fourier space as
+For each simulation the particles are firstly cloud-in-cell (CIC) interpolated onto the fine mesh grid, and from the density field $\rho$ we define the density contrast $\delta\equiv \rho/\left\langle \rho \right\rangle-1$.
+We define the cross power spectrum $P_{\alpha\beta}(k)$ between two fields $\delta_\alpha$ and $\delta_\beta$ ($\delta_\alpha=\delta_\beta$ for auto power spectrum) in Fourier space as
 \begin{equation}\label{eq.cps}
@@
 
-\tcb{We compute the nonlinear prediction of the power spectrum by CLASS \citep{2011JCAP...07..034B}, denoted $\Delta^2_0(k)$, at $z=0,1$.
-Then we study the fractional power spectrum deviations $\Delta^2(k)/\Delta^2_0(k)-1$.
-In the four panels of figure \ref{fig.ccc} we denote the box sizes of the simulation and redshifts at the upper left corners.
-The upper parts of these four panels show $\Delta^2(k)/\Delta^2_0(k)-1$, and the four curves of different colors show the results of x2v2, x1v2, x2v1 and x1v1.
-They almost overlap with each other, and in general agree with the nonlinear predictions.}
+In the upper two panels of figure \ref{fig.ccc} we show the power spectra of {\tt CUBE}.
+In both plots of S2048S and S2048L the four solid curves of different colors show the results of x2v2, x1v2, x2v1 and x1v1, and they almost overlapped with each other.
+The dashed curves are the nonlinear prediction of matter power spectrum by CLASS \citep{2011JCAP...07..034B}.
 
-\tcb{We label $k=0.2\,k_{\rm Nyquist}$ as vertical dashed lines, where $k_{\rm Nyquist}$ is the scale of fine mesh grids, and the scale of average particle separations.
+We label $k=0.2\,k_{\rm Nyquist}$ as vertical dashed lines, where $k_{\rm Nyquist}$ is the scale of fine mesh grids, and the scale of average particle separations.
 On this scale, the power spectra are offset from nonlinear predictions by at least 20\%.
@@
 We do not plot {\tt CUBEP3M} because we examined that {\tt CUBEP3M} and x2v2 of {\tt CUBE} produce same results.
-Thus, the differences between {\tt CUBEP3M} and different integer formats of {\tt CUBE} are negligible compared to the error of the PM algorithm.}
+Thus, the differences between {\tt CUBEP3M} and different integer formats of {\tt CUBE} are negligible compared to the error of the PM algorithm.
 
-\tcb{In the lower parts of these four panels study the cross correlations.
-This time we in turn compare everything with {\tt CUBEP3M}, to emphasize the decorrelation by different integer formats.
-Firstly, x2v2 is perfectly correlated and cross correlation of x1v2 is nearly 100\% in all cases.
-In order to quantify the PM error in terms of decorrelation, we run x2v2 simulations with same initial conditions, but 8 times more particles and cells, and measure the decorrelation caused by coarser resolutions.
-These results are dashed curves (labeled ``PM'') in the lower parts of these four panels.
-We conclude that cross correlation of x1v1, in all cases, is not worse than the PM errors.}
+In the lower parts of these four panels study the cross correlations.
+We compare everything with {\tt CUBE}-x2v2, to emphasize the decorrelation by different integer formats.
+Note that x2v2 is perfectly correlated with {\tt CUBEP3M}.
+In the lower two panels of figure \ref{fig.ccc} the solid curves show the decorrelation $\xi$ of x1v2, x2v1 and x1v1.
+Higher resolution (S2048S) in general comes with more decorrelations, and x1v2 cross correlate with x2v2 better than the other two cases.
+In order to quantify the PM error in terms of decorrelations, we run x2v2 simulations with same initial conditions, but 8 times more particles and cells, and measure the decorrelation caused by coarser resolutions.
+These results are shown in dashed curves (labeled ``PM'').
+We conclude that cross correlation of x1v1, in all cases, is not worse than the PM errors.
 
-\tcb{To summarize from Figure \ref{fig.ccc}, in terms of either power spectrum deviation or cross correlation, the error induced by information optimization (even for x1v1) is lower than the error from the PM algorithm, and we can safely use x1v1 for most of the LSS studies.
-If one still wants to compare the results when PM already contributes significant error ($k\gtrsim 0.2\,k_{\rm Nyquist}$), errors are from nonlinear (lower $m_p$ and/or lower redshift) and are primarily from compressing the velocity information (x2v1 and x1v1).}
+To summarize from Figure \ref{fig.ccc}, in terms of either power spectrum deviation or cross correlation, the error induced by information optimization (even for x1v1) is lower than the error from the PM algorithm, and we can safely use x1v1 for most of the LSS studies.
 
 \section{Discussion and conclusion}\label{s.discussion}
-We present an parallel, \tcb{information-optimized} $N$-body algorithm.
+We present an parallel, information-optimized $N$-body algorithm.
 This open-source code {\tt CUBE}, has recently been used in many studies of LSS, e.g.
@@
 In the case of using 2-byte integers for both positions and velocities (``x2v2'', and memory can be 12 bpp), the algorithm gives exact results given by traditional $N$-body algorithms.
-\tcb{Note that the results are exactly same in that they not only produce same physical statistics of LSS, but also the same error (not physical) from the PM algorithm near Nyquist frequencies.
+Note that the results are exactly same in that they not only produce same physical statistics of LSS, but also the same error (not physical) from the PM algorithm near Nyquist frequencies.
 In another word, the positions and velocities of each particle are exact.
@@
 In Figure \ref{fig.ccc} we see that it is the case even for the most memory efficient configuration x1v1.
-This shows that in most LSS studies, when our scales of interest are smaller than $k\simeq 0.2\,k_{\rm Nyquist}$, six 1-byte fixed point numbers contains sufficient information of every $N$-body particle.}
+This shows that in most LSS studies, when our scales of interest are smaller than $k\simeq 0.2\,k_{\rm Nyquist}$, six 1-byte fixed point numbers contains sufficient information of every $N$-body particle.
 
@@
 Traditional $N$-body codes consumes considerable memory while doing a relatively light computations.
-\tcb{{\tt CUBE} is designed to optimize the efficiency of information and the memory usage in $N$-body simulations.
-{\tt CUBE} is written in Coarray Fortran -- the concise Coarray features are used instead of complicated MPI, and the code itself is much more concise than {\tt CUBEP3M} for future maintenance and development.} The next steps are optimization of the code and adapting it for various kinds of heterogeneous computing systems, e.g.
+{\tt CUBE} is designed to optimize the efficiency of information and the memory usage in $N$-body simulations.
+{\tt CUBE} is written in Coarray Fortran -- the concise Coarray features are used instead of complicated MPI, and the code itself is much more concise than {\tt CUBEP3M} for future maintenance and development.
+The next steps are optimization of the code and adapting it for various kinds of heterogeneous computing systems, e.g.
 MIC and GPUs.
@@
 
-%\begin{figure}
-%\centering
-%  \includegraphics[width=0.95\linewidth]{power_universe1-8_z0and1.pdf}
-% \caption{\tcb{We don't need this figure in submission.
-However we need to compare the results with nonlinear power spectrum by {\tt HALOFIT}.} The solid lines show the power spectra of S2048L and S2048XS ($L=300\ {\rm Mpc}/h$) at $z=0$, and dotted lines show the $z=1$ counterparts.
-For $z=0$, the two data tips are at $k=0.2\ k_{\rm Nyquist}$, where the power spectrum is suppressed by more than 20\% compared to higher resolution simulations.}
-%\label{fig.power}
-%\end{figure}
-
-%\clearpage
-
-%\appendix
-
-%\section{A.}
-
 \acknowledgements
+We acknowledge funding from NSERC.
+H.R.Y. thanks Derek Inman for many helpful discussions.
+The simulations were performed on the GPC supercomputer at the SciNet HPC Consortium.
+The code is publicly available on github.com under {\tt yuhaoran/CUBE}.
 

### 2017-12-17 CUBE/utilities/module_load_mac.sh

@@ -1,0 +1,8 @@
+export FC='gfortran'
+#export XFLAG='-O3 -cpp -fcoarray=single -mcmodel=medium'
+export XFLAG='-O3 -cpp -fcoarray=single'
+#export XFLAG='-O3 -cpp -fcoarray=single -fopenmp'
+export OFLAG=${XFLAG}' -c'
+export FFTFLAG='-I/usr/local/include/ -L/usr/local/lib/ -lfftw3f -lm -ldl'
+# -fopenmp cause (maybe memory) probelm: Segmentation fault: 11
+# in cumsum6

### 2018-03-16 ms_caf/ms_caf.tex
reply referee of CUBE paper

@@
 
-When $N$ is large, the brute force pairwise particle-particle (PP) force brings unaffordable $o(N^2)$ computations.
-However the workload can be alleviated to $o(N\log N)$ by many algorithms.
-For example, the ``tree'' algorithms like {\tt GADGET} \citep{2001NewA....6...79S,2005MNRAS.364.1105S} and its simulation ``Millennium'' \citep{2005Natur.435..629S,2012MNRAS.426.2046A}, {\tt TPM} \citep{1995ApJS...98..355X} and {\tt GOTPM} \citep{2004NewA....9..111D}, adaptive grid algorithm including {\tt HYDRA} \citep{1995ApJ...452..797C} and {\tt RAMSES} \citep{2010ascl.soft11007T}, as well as mesh-refined codes \citep{1991ApJ...368L..23C} and moving adaptive particle-mesh (PM) codes \citep{1995ApJS..100..269P}.
-Among these, the standard PM algorithm \citep{1988csup.book.....H} is most memory and computational efficient if we focus on large cosmological scales.
+When $N$ is large, the brute force pairwise particle-particle (PP) force brings unaffordable $o(N^2)$ computations, so many algorithms were designed to alleviate it.
+Various fast multipole methods \citep{1985JCoPh..60..187R} improve the complexity to $o(N\log N)$ even $o(N)$, among which the most popular one is ``tree'', like {\tt GADGET} \citep{2001NewA....6...79S,2005MNRAS.364.1105S} and its simulation ``Millennium'' \citep{2005Natur.435..629S,2012MNRAS.426.2046A}, {\tt TPM} \citep{1995ApJS...98..355X} and {\tt GOTPM} \citep{2004NewA....9..111D}.
+Other methods include adaptive grid algorithms like {\tt HYDRA} \citep{1995ApJ...452..797C} and {\tt RAMSES} \citep{2010ascl.soft11007T}, as well as mesh-refined codes \citep{1991ApJ...368L..23C} and moving adaptive particle-mesh (PM) codes \citep{1995ApJS..100..269P}.
+The standard PM algorithm \citep{1988csup.book.....H} is most memory and computational efficient if we focus on large cosmological scales.
 The load balancing problem is minimized as the matter distribution is rather homogeneous, and the speed benefits from the fast Fourier transform (FFT) libraries, such as {\tt FFTW3} \citep{Frigo05thedesign}.
@@
 
-In the following sections we present a new, information-optimized parallel cosmological $N$-body simulation code {\tt CUBE}, using as low as 6 byte per particle.
+In the following sections we present a new, information-optimized parallel cosmological $N$-body simulation code {\tt CUBE} \citep{haoran_yu_2018_1198909}, using as low as 6 byte per particle.
 It gives accurate results in cosmological LSS simulations -- the error induced by information optimization is below the error from the PM algorithm.
@@
 Instead of storing global coordinates of each particle, we store its offset relative to its parent cell which contains the particle.
+This is similar with storing the quantities of nodes/clumps (structures of a tree in a tree code) relative to their parents \citep{1985SJSSC...6...85A}.
 We divide the cell, in each dimension $d$, evenly into $2^8=256$ bins, and use a 1-byte (8 bits) integer $\chi_d \in \{-128,-127,...,127\}$ to indicate which bin it locates in this dimension.
@@
 
-In a simulation with volume $L^3$ and $N_c^3$ coarse cells, particle positions are stored with a precision of $L/(256N_c)$.
+In a simulation with volume $L^3$ and $N_c^3$ coarse cells, particle positions are stored with a resolution of $L/(256N_c)$.
 The force calculation (e.g.
@@
 On the other hand, particle position can also be stored as 2-byte (16 bits) integers to increase the resolution.
-In this case, each coarse cell is divided into $2^{16}=65536$ bins and the position precision is $L/(65536N_c)$, precise enough compared to using4-byte global coordinates, see later results.
+In this case, each coarse cell is divided into $2^{16}=65536$ bins and the position resolution is $L/(65536N_c)$, precise enough compared to using4-byte global coordinates, see later results.
 We denote this case ``x2'' and denote using 1-byte integers for positions ``x1''.
@@
 where $[\ ]$ is the operator to take the integer part.
-$n_\chi\in\{1,2\}$ is the number of bytes used for each integer, $x_d$ and $\chi_d$ are floating and integer version of the coordinate.
+$n_\chi\in\{1,2\}$ is the number of bytes used for each integer, $x_d$ and $\chi_d$ are floating and integer versions of the coordinate.
 The velocity counterpart of them are $n_\nu=1,2$, $v_d$ and $\nu_d$.
-$n_c=1,2,...$ is the coarse grid index, given by the ordering of particles and a integer based particle density field (see \ref{sss.spatial_decomposition}).
+%$n_c=1,2,...$ is the coarse grid index, given by the ordering of particles and a integer based particle density field (see \ref{sss.spatial_decomposition}).
 The position resolution for $n_\chi$-byte integer, ``x$n_\chi$'', is $2^{-8n_\chi}L/N_c$.
@@
 \end{equation}
-$v_c$ is always recorded and kept updated, and occupies negligible memory.
+$v_c$ is always recorded and kept updated, and should not occupy considerable memory.
 We then divide velocity space $\Delta v$ into uneven bins, and use a $n_\nu$-byte integer to indicate which $\Delta v$ bin the particle is located.
@@
  \caption{Spacial decomposition in {\tt CUBE} in a 2D analogy.
-In this example, there are 2 images per dimension (${\tt nn}=2$), and 2 tiles per image per dimension (${\tt nnt}=2$).
-The orange boxes show the overlapped physical+buffer=extended tile regions, inside of which one physical regions is indicated with green.}
+In this example, there are 2 images per dimension ($M_g=2$), and 2 tiles per image per dimension ($M_t=2$).
+The orange boxes show the overlapped extended($V_e$)=physical($V_p$)+buffer($V_b$) tile regions.
+One physical region is indicated with green.}
 \label{fig.tile}
@@
 where $\lfloor\ \rfloor$ is the operator to take the nearest integer.
-Tangent functions are convenient and compute very fast.
+Tangent functions are convenient and computing-efficient.
 Compared to error functions used in Gaussian case, they take the same variance at $\nu_d=0$ but resolve high velocities relatively better.
-Note again that proper choice of conversion formulae and $\sigma_\Delta$ only optimizes the velocity space sampling, but does not affect the physics.
+All possible choice of conversion formulae and $\sigma_\Delta$ are unbiased in the conversion Equations (\ref{eq.nu},\ref{eq.v}), however a proper choices optimize the velocity space sampling and can result in more precise results.
 
@@
 \subsection{Code overview}\label{ss.overview}
-{\tt CUBE} uses a 2-level PM force calculation, same as {\tt CUBEP3M}, However, in order to apply the integer based format to the $N$-body simulation, substantial structural changes need to be done.
+{\tt CUBE} uses a 2-level PM force calculation.
+In order to apply the integer based format to the $N$-body simulation, substantial structural changes need to be done.
 {\tt CUBE} is written in Coarray Fortran, where Coarray features replace MPI (Message Passing Interface) communications between computation nodes/images\footnote{Images are the concept of computing nodes or MPI tasks in Coarray Fortran.
@@
 \subsubsection{Spacial decomposition}\label{sss.spatial_decomposition}
-{\tt CUBE} uses cubic decomposition structures.
-The global simulation volume is decomposed into ${\tt nn}^3$ cubic sub-volumes with {\tt nc} coarse grids per side, and each of these are assigned to a coarray {\it image}.
-Inside of an image, the sub-volume is further decomposed into ${\tt nnt}^3$ cubic {\it tiles}, with {\tt nt}={\tt nc}/{\tt nnt} coarse grids per side, which is a essential unit in the calculations.
-Each tile is surrounded by a {\it buffer} region which is ${\tt ncb}$ coarse cells thick.
-The buffer is designed for two reasons: (1) computing the fine mesh force, whose cut-off length ${\tt nforce\_cutoff}\leq{\tt ncb}$, and (2) collecting all possible particles travelling from a tile's buffer region to its center, {\it physical} region.
-Figure \ref{fig.tile} shows the spacial decomposition in a 2-dimensional analogy, with ${\tt nn=2}$ and ${\tt nnt=2}$.
+{\tt CUBE} decomposes the global simulation volume into $M_g^3$ cubic sub-volumes with $N_c$ coarse grids or $N_f=RN_c$ fine grids per side.
+The fine-mesh is usually $R=4$ times finer than the coarse mesh.
+Each of these sub-volumes is assigned to a coarray {\it image}.
+Inside of an image, the sub-volume is further decomposed into $M_t^3$ cubic {\it tiles} (defined as $V_p$, or {\it physical} region) with $N_c/(M_g M_t)$ coarse grids per side.
+Each $V_p$ is surrounded by a {\it buffer} region $V_b$ which is $N_b$ coarse cells thick.
+We define extended tile region $V_e\equiv V_p+V_b$.
+$V_e$ is designed for two reasons: (1) if the short fine-mesh force ${\bs F}_f$ has a cut-off $N_b$, and is computed by $\rho_c$ whose domain $D(\rho_f)=V_e$, then ${\bs F}_f$ in $V_p$ is guaranteed to be correct, i.e.,
+\begin{eqnarray}\label{eq.cond1}
+\nonumber	{\bs F}_f(r>N_b)&=&0 \wedge D(\rho_c)=V_e\\
+	&\Rightarrow& \{{\bs F}_f(\rho_c) |{\bs F}_f\in V_p \} {\rm \ is\ correct},
+\end{eqnarray}
+and (2) to collect all particles being able to travel to $V_p$, i.e.,
+\begin{equation}\label{eq.cond2}
+	{\bs v}\,{\rm d}t \leq N_b \Rightarrow \{ {\bs x} | {\bs x}+{\bs v}\,{\rm d}t\in V_p\} \in V_e.
+\end{equation}
+Figure \ref{fig.tile} shows the spacial decomposition in a 2-dimensional analogy, with $M_g=2$ and $M_t=2$.
 
-According to this spatial decomposition, and as discussed in the last two subsections, we declare $\{\rho_c,\chi_d,v_c,\nu_d\}$, or in variable expression $\{{\tt rho\_c},{\tt xp},{\tt vfield},{\tt vp}\}$, as follows:
+According to this spatial decomposition, and as discussed in the last two subsections, we declare $\{\rho_c,v_c,\chi_d,\nu_d\}$ by using Fortran language:
 \begin{eqnarray*}
-	&&{\tt integer(1)\ rho\_c(nex,nex,nex,nnt,nnt,nnt)[nn,nn,*]}\\
-	&&{\tt integer(}n_\chi{\tt )\ xp(npmax)[nn,nn,*]}\\
-	&&{\tt real(4)\ vfield(nex,nex,nex,nnt,nnt,nnt)[nn,nn,*]}\\
-	&&{\tt integer(}n_\nu{\tt )\ vp(npmax)[nn,nn,*]}
+	&&{\tt integer(1)}\ \ \, \rho_c(N_e,N_e,N_e,M_t,M_t,M_t)[M_g,M_g,*]\\
+	&&{\tt real(4)}\ \ \ \ \ \ \ \, v_c(N_e,N_e,N_e,M_t,M_t,M_t)[M_g,M_g,*]\\
+	&&{\tt integer(}n_\chi{\tt )}\, \chi_d(3,P_{\rm max})[M_g,M_g,*]\\
+	&&{\tt integer(}n_\nu{\tt )}\ \nu_d(3,P_{\rm max})[M_g,M_g,*]
 \end{eqnarray*}
-where ${\tt nex}={\tt nt}+2\times{\tt ncb}$ covers the buffer region on both sides, {\tt nnt} is the tile dimemsions, and {\tt nn} is the image {\it codimensions}\footnote{Coarray Fortran concept.
+where $N_e=N_t+2N_b$ covers the buffer region on both sides, $M_t$ is the tile dimemsions, and $M_g$ is the image {\it codimensions}\footnote{Coarray Fortran concept.
 Codimensions can do communications between images.}.
-We denote actual number of particles in a given image {\tt nplocal}, and ${\tt npmax}>{\tt nplocal}$ is a value (discussed in Section \ref{ss.memory}) large enough to store particles.
-The particle position and velocity arrays {\tt xp} and {\tt vp} (equivalent to $\chi_d$ and $\nu_d$ in \ref{ss.position} and \ref{ss.velocity} respectively) are required to be sorted according to the same memory layout of {\tt rho\_c}, such that $n_c$ and thus global positions of particles $x_d$ can be obtained by equation (\ref{eq.x}).
-$\{{\tt rho\_c},{\tt xp},{\tt vfield},{\tt vp}\}$ provides a complete information of positions and velocities of particles, and we call it a snapshot, or checkpoint.
+We denote actual number of particles in a given image $P_{\rm local}$, and $P_{\rm max}>P_{\rm local}$ is a value (discussed in Section \ref{ss.memory}) large enough to store particles in $V_e$.
+$\chi_d$ and $\nu_d$
+%(equivalent to $\chi_d$ and $\nu_d$ in \ref{ss.position} and \ref{ss.velocity} respectively)
+are required to be sorted according to the same memory layout of $\rho_c$, such that $n_c$ and $x_d$ can be obtained by equation (\ref{eq.x}). 
+$\{\rho_c,v_c,\chi_d,\nu_d\}$ provides a complete information of positions and velocities of particles, and we call it a checkpoint.
 
-An additional particle-ID (PID) array, {\tt PID}, can also be declared to differentiate particle types (e.g.
-CDM and neutrino particles) or differentiate every particle, by using 1-byte, 4-byte or 8-byte integer per particle.
-If PID is turned on, the array {\tt PID} is also included in the checkpoints, and the ordering of PID is same as {\tt xp} and {\tt vp}.
+An additional particle-ID (PID) array, $I_P$, can also be declared to differentiate particle types (e.g.
+CDM and neutrino particles) or differentiate every particle, by using $n_I$-integer per particle.
+If PID is turned on, the array $I_P$ is also included in the checkpoints, and the ordering of $I_P$ is same as $\chi_d$ and $\nu_d$.
 
 \begin{figure}[t]
-\begin{verbatim}
-
-program Initial_Condition_Generator_for_CUBE
-  Get random seed
-  Generate Gaussian random field
-  Multiply by power spectrum in Fourier space
-  Get linear density field
-  Solve Poisson equation
-  Get displacement potential  
-  do (each tile) ! create particles per tile
-    do (each fine grid in the extended tile region)
-      calculate particle's coarse grid position
-      update rho_c_new
-    enddo
-    do (each fine grid in the extended tile region)
-      calculate particle's new accurate position
-      calculate praticle's initial velocity
-      calculate particle's index
-      create xp_new(index) & vp_new(index)
-      if (PID_flag) create PID(index)
-    enddo
-    do (each coarse grid)
-      discard buffer information
-    enddo
-    write to disk
-    sum up nplocal
-  enddo
-  sync all
-  sum up npglobal
-  write to headers of the checkpoint
-end
-\end{verbatim}
+{\tt program Initial\_Condition\_Generator\_for\_CUBE\\
+\indent \ \ calculate $\Phi$ in Fourier space\\
+\indent \ \ do (each tile)\\
+\indent \ \ \ \ do (each particle at ${\bs q}\in V_e$)\\
+\indent \ \ \ \ \ \ ${\bs \Psi}({\bs q})=\nabla\Phi({\bs q})$\\
+\indent \ \ \ \ \ \ $\{{\bs q},{\bs\Psi({\bs q})}\}\rightarrow\{{\bs x},{\bs v}\}$\\
+\indent \ \ \ \ \ \ $\{{\bs x},{\bs v}\}\rightarrow \{\rho_c,v_c\}$\\
+\indent \ \ \ \     enddo\\
+\indent \ \ \ \ do (each particle at ${\bs q}\in V_e$)\\
+\indent \ \ \ \ \ \ ${\bs \Psi}({\bs q})=\nabla\Phi({\bs q})$\\
+\indent \ \ \ \ \ \ $\{{\bs q},{\bs\Psi({\bs q})}\}\rightarrow\{{\bs x},{\bs v}\}$\\
+\indent \ \ \ \ \ \ calculate particle's index $i$ according to $\rho_c$\\
+\indent \ \ \ \ \ \ ${\bs x}\rightarrow\chi_d(:,i);\,\{{\bs v},v_c\}\rightarrow\nu_d(:,i)$\\
+\indent \ \ \ \ \ \ if (PID\_flag) create $I_P(i)$\\
+\indent \ \ \ \ enddo\\
+\indent \ \ \ \ do (each coarse grid $\in V_e$)\\
+\indent \ \ \ \ \ \ delete particles $\in V_b$\\
+\indent \ \ \ \ enddo\\
+\indent \ \ \ \ write $\{\rho_c,v_c,\chi_d,\nu_d\}$ to disk\\
+\indent \ \ \ \ sum up $P_{\rm local}$\\
+\indent \ \ enddo\\
+\indent \ \ sync all\\
+\indent \ \ sum up $P_{\rm global}$\\
+end\\}
 \caption{Pseudocode for the initial condition generator.}
@@
 
-The first step is the calculation of the potential field.
-At a sufficient high initial redshift $z_i$, we first generate a linear density fluctuation $\delta_L({\bs q})$ on Lagrangian grid ${\bs q}$ by multiplying a Gaussian random field with the transfer function $T(k)$ in Fourier space, where $T(k)$ is given by the assumed cosmological model.
-Then we solve for the potential ${\Phi({\bs q}, z_i)}$ of a curl-less displacement field ${\bs\Psi({\bs q})}$ by Poisson equation $-\nabla^2\Phi=\delta_L$.
-At this step, global Pencil FFTs are applied on the fine mesh to transform the initial random Gaussian field to the displacement potential $\Phi$.
-The initial condition generator is applied only once, and in principle, at some costs of speed, one can use only one scalar field on the fine mesh, and FFTs can be done in-place.
-If the number of particles equals the number of fine grids, then the memory consumption is dominated by the single precision fine mesh field, or 4 bpp.
+The first step is the calculation of the displacement potential $\Phi$.
+At an initial redshift $z_i$, we generate a linear density fluctuation $\delta_L({\bs q})$ on Lagrangian grid ${\bs q}$ by multiplying a Gaussian random field with the transfer function $T(k)$ (given by the assumed cosmological model) in Fourier space.
+Then we solve for the potential ${\Phi({\bs q}, z_i)}$ of a curl-less displacement field ${\bs\Psi({\bs q})}$ by Poisson equation $-\nabla^2\Phi=\delta_L$ in Fourier space.
 
-The second step is to generate particles on Lagrangian coordinates ${\bs q}$ and displace them with Zel'dovich approximation (ZA) \citep{1970A&A.....5...84Z}, by locally differentiating $\Phi$ in real space.
-This step is done tile by tile.
-For each tile, we iterate twice over particles' Lagrangian coordinates ${\bs q}$ (usually at the center of every fine grid).
-The first iteration calculates particles' new, coarse grid positions by ZA and get a coarse density field on tile.
-The second iteration, ZA is applied again to calculate the accurate positions of particles and their velocities, and convert them into integer based ones by Equations (\ref{eq.chi},\ref{eq.nu}) and place them in a certain order according to the coarse density field generated by the first iteration.
-Lastly, we discard those particles out of the physical region of the tile, and write only particles in physical region on to the disk.
-As we will see in Section \ref{sss.algorithm}, {\tt update\_xp} paragraph, it is very similar to the ``drift'' step in the main $N$-body code, where an ``buffered'' state of particles is firstly created by two iterations of particles, and the particles in the buffer region are discarded to achieve a ``disjoint'' state.
-If PIDs are needed, they are also generated at the second iteration and are manipulated similarly as integer based velocities.
-Because particles are generated and saved to disk tile by tile, locally, after working on certain tile, the arrays for positions, velocities and PIDs (if any) are freed up and can be reused for the next tile.
-During this step, the only major memory usage is still dominated by the fine grid potential $\Phi$.
-
-After working on all tiles, we sum up the total particle number on this image, {\tt nplocal}, and write it in the header of the checkpoint.
-Finally, $\{{\tt rho\_c},{\tt xp},{\tt vfield},{\tt vp},{\tt PID}({\rm optional})\}$ is rewritten to the disk as the checkpoint of $z=z_i$.
-We summarize the above steps into a pseudocode in Figure \ref{fig.ic}, where the ``{\tt do}'' loop to create particles per tile is very similar to the following {\tt update\_xp} subroutine in the main $N$-body code.
-The corresponding arrays are also discussed below.
+The second step to generate particles and displace them by Zel'dovich approximation (ZA) \citep{1970A&A.....5...84Z}, where the displacement field is obtained by differentiating $\Phi$, ${\bs \Psi}({\bs q})=\nabla\Phi({\bs q})$ in real space.
+This step is done on $V_e$ of each tile.
+Similar to Equation (\ref{eq.cond2}), we have
+\begin{equation}\label{eq.cond3}
+	\max(\Psi_d) \leq N_b \Rightarrow \{ {\bs q} | {\bs q}+{\bs \Psi}({\bs q})\in V_p\} \in V_e.
+\end{equation}
+We iterate twice over particles' ${\bs q}$ in $V_e$.
+The first iteration calculates particles' ${\bs x}$ and ${\bs v}$ by ZA and obtains $\rho_c$ and $v_c$ on the coarse grid.
+The second iteration, ${\bs x}$ and ${\bs v}$ is calculated again and are converted to $\chi_d$ and $\nu_d$ by Equations (\ref{eq.chi},\ref{eq.nu}) and place them in a certain order according to $\rho_c$.
+Lastly, we delete particles in $V_b$, and resort the ones in $V_p$ and write $\{\rho_c,v_c,\chi_d,\nu_d,I_P({\rm optional})\}$ of this tile to disk.
+The above is similar to {\tt update\_xp} of Section \ref{sss.algorithm}.
+If PIDs are needed, they are also generated here.
+After working on all tiles, we sum up $P_{\rm local}$ and $P_{\rm global}$.
+We summarize the above steps into a pseudocode in Figure \ref{fig.ic}.
+During this step, the only major memory usage is $\Phi$ on the fine mesh.
+If the number of particles per fine grid $P_f=1$, the memory consumption of this in-place FFT is 4 bpp.
 
@@
 \subsubsection{Algorithm}\label{sss.algorithm}
-Figure \ref{fig.code} shows the overall structure of the code and these subroutines are described in following paragraphs.
+Figure \ref{fig.code} shows the overall structure of the main code.
 
-{\tt initialize} and {\tt read\_particles} -- The subroutine {\tt initialize} creates necessary FFT plans and read in configuration files telling the program at which redshifts we need to do checkpoints, halofinds, or stop the simulation.
-Force kernels {\tt kern\_c}, {\tt kern\_f} are also created or read in here.
-In {\tt read\_particles}, for each image, we read in all particles in {\it physical} regions of every tile (indicated in Figure \ref{fig.tile}), i.e., $\{{\tt rho\_c},{\tt xp},{\tt vfield},{\tt vp},{\tt PID}({\rm if\ any})\}$.
-These are obtained by the initial condition generator (Section \ref{sss.ic}).
-Because physical regions of tiles are {\it complete} and {\it disjoint} in space, particles at this stage are also complete and disjoint.
-We call it ``disjoint state''.
-At this stage, {\tt rho\_c}'s values in buffer regions are 0's, and the elements of {\tt xp} and {\tt vp} beyond physical number -- {\tt xp(nplocal+1:)} and {\tt vp(nplocal+1:)} can be arbitrary and are not used.
+{\tt initialize} -- creates fine-mesh and coarse mesh FFT plans, and read in configuration files telling the program at which redshifts we need to do checkpoints, halofinds, or stop the simulation.
+Force kernels $K_c$, $K_f$ are also computed or loaded.
 
-{\tt buffer\_density}, {\tt buffer\_x} and {\tt buffer\_v} -- In order to use the integer based format, {\tt xp} and {\tt vp} must always be ordered, and their number density field {\tt rho\_c} must always be present.
-In updating arrays of {\tt xp} and {\tt vp} (not simultaneously) of a local tile, the buffer region of the tile is also needed.
-First, buffer regions of {\tt rho\_c} is synchronized between tiles and images by subroutine {\tt buffer\_density}.
-Then, by subroutines {\tt buffer\_x} and {\tt buffer\_v} respectively, {\tt xp} and {\tt vp} are updated to contain common, buffered particles, in an order according to the new, buffered {\tt rho\_c}.
-If there are PIDs used, then {\tt PID} is also updated in subroutine {\tt update\_v}.
-We call this stage ``buffered state''.
-After {\tt particle\_initialize} is done by all images (synchronized by a ``{\tt sync all}'', which is equivalent to a {\tt mpi\_barrier}), these three subroutines are called and particles are converted from disjoint state to buffered state.
+{\tt read\_particles} -- from disk, reads in checkpoint $\{\rho_c,v_c,\chi_d,\nu_d,I_P({\rm optional})\}$ for each image.
+Because they exist only in $V_p$ of every tile, they are {\it disjoint}; and they provide {\it complete} information of the whole simulation volume -- we call it ``{\it disjoint state}''.
+In this state, $\rho_c$, $v_c$'s values in buffer regions, and $\chi_d(:,P_{\rm local}+1:)$ and $\nu_d(:,P_{\rm local}+1:)$ are 0's.
+Because $I_P$ is generated and manipulated together with $\nu_d$, so we do not explicitly mention $I_P$ in the followings.
+
+{\tt buffer\_density}, {\tt buffer\_x} and {\tt buffer\_v} -- converts ``disjoint'' state to ``{\it buffered state}''.
+In {\tt buffer\_density}, $V_b$ regions of $\rho_c$ is synchronized between tiles and images.
+By {\tt buffer\_x}, $\chi_d$ is updated to contain common, buffered particles, and they are sorted according to the buffered $\rho_c$.
+{\tt buffer\_v} deals with $\nu_d$ in a same manner.
 
 {\tt timestep} --
-We operate a Runge-Kutta 2 method for time integration.
-i.e., updating positions ($D$=drift) and velocities ($K$=kick) are at every half time step by operator splitting.
-For $n$ time steps, the operation would be $(DKKD)^n$ which is 2nd order accurate.
-The actual simulation applies varied time steps.
-In each iteration of the main loop, we firstly call {\tt timestep}, where a increment of time {\tt dt} is controlled by particles' maximum velocities, accelerations, cosmic expansion and any other desired conditions.
+Runge-Kutta 2 method is used in the time integration.
+i.e., for $n$ time steps, we update positions ($D$=drift) and velocities ($K$=kick) at interlaced half time-steps by operator splitting; the operation $(DKKD)^n$ is 2nd order accurate.
+The actual simulation applies varied time steps by {\tt timestep}, where a time increment ${\rm d}t$ is constrained by particles' maximum velocities, accelerations, cosmic expansion and any other desired conditions.
 
-\begin{figure}[]
-\begin{verbatim}
-
-subroutine update_xp
-  do (each tile)
-    do (each coarse grid)
-      do (each particle)
-        calculate particle's new coarse grid position
-        update rho_c_new
-        update vfield_new
-      enddo
-    enddo
-    do (each coarse grid)
-      do (each particle)
-        calculate particle's new accurate position
-        calculate particle's index
-        update xp_new(index) & vp_new(index)
-      enddo
-    enddo
-    do (each coarse grid)
-      discard buffer information
-      replace xp and vp for this tile
-    enddo
-  enddo
-  sync all
-  update velocity dispersion
-  check nplocal & npglobal
-end
-\end{verbatim}
+\begin{figure}[t]
+{\tt subroutine update\_xp\\
+\indent \ \ do (each physical tile)\\
+\indent \ \ \ \ do (each particle)\\
+\indent \ \ \ \ \ \ $\{\rho_c,\chi_d\}\rightarrow {\bs x};\,\{v_c,\nu_d\}\rightarrow {\bs v}$\\
+\indent \ \ \ \ \ \ ${\bs x}={\bs x}+{\bs v}\,{\rm d}t$\\
+\indent \ \ \ \ \ \ update $\rho_c^*,\,v_c^*$ according to ${\bs x}$\\
+\indent \ \ \ \ enddo\\
+\indent \ \ \ \ do (each particle)\\
+\indent \ \ \ \ \ \ $\{\rho_c,\chi_d\}\rightarrow {\bs x};\,\{v_c,\nu_d\}\rightarrow {\bs v}$\\
+\indent \ \ \ \ \ \ ${\bs x}={\bs x}+{\bs v}\,{\rm d}t$\\
+\indent \ \ \ \ \ \ calculate particle's index $i$ according to $\rho_c^*$\\
+\indent \ \ \ \ \ \ ${\bs x}\rightarrow\chi_d^*(:,i);\,\{{\bs v},v_c\}\rightarrow\nu_d^*(:,i)$\\
+\indent \ \ \ \ enddo\\
+\indent \ \ \ \ do (each coarse grid)\\
+\indent \ \ \ \ \ \ discard buffer information\\
+\indent \ \ \ \ enddo\\
+\indent \ \ \ \ replace $\{\rho_c,v_c,\chi_d,\nu_d\}$ with $\{\rho_c^*,v_c^*,\chi_d^*,\nu_d^*\}$\\
+\indent \ \ enddo\\
+\indent \ \ sync all\\
+\indent \ \ update velocity dispersion $\sigma^2_{\Delta}$\\
+\indent \ \ sum up $P_{\rm local},\,P_{\rm global}$\\
+end\\}
 \caption{Pseudocode for subroutine {\tt update\_xp}.}
@@
 {\tt update\_xp} --
-According to {\tt dt}, subroutine {\tt update\_xp} updates the positions of particles in a ``{\it gather}'' algorithm (in contrast, {\tt CUBEP3M} uses ``scatter'' algorithm) tile by tile.
-For each particle, {\tt xp} and {\tt vp} are converted to $x_d$ and $v_d$ by Equations(\ref{eq.x},\ref{eq.v}).
-Because each tile is in the buffered state with buffer depth {\tt ncb}, we are able to collect all possible particles whose $v_d\times{\tt dt}<{\tt ncb}$ from physical+buffer region to its physical region.
+According to ${\rm d}t$, {\tt update\_xp} updates particle positions (drift $D$) in a ``{\it gather}'' algorithm
+%(in contrast, {\tt CUBEP3M} uses ``scatter'' algorithm)
+tile by tile.
+For each particle, $\chi_d$ and $\nu_d$ are converted to $x_d$ and $v_d$ by Equations(\ref{eq.x},\ref{eq.v}).
+Due to Condition (\ref{eq.cond2}), we are able to collect all particles to travel to $V_p$.
 
-In order to keep particles ordered, for each tile, we first do $x_d=x_d+v_d\times{\tt dt}$ on all particles to obtain an updated density and velocity field on the tile, {\tt rho\_c\_new} and {\tt vfield\_new}.
-Then, this calculation in done on the same tile again\footnote{This repetition scales as $o(N)$ and is computational inexpensive.} to generate a new, local particle list {\tt xp\_new} and {\tt vp\_new} (also {\tt PID\_new} if needed) on the tile by Equations(\ref{eq.chi},\ref{eq.nu}).
-Here, the ordering of particles depends on {\tt rho\_c\_new}, and {\tt vfield\_new} is used as $v_c$ in Equation(\ref{eq.nu}).
-Then, another iteration is done on this tile to discard buffer regions of $\{{\tt rho\_c\_new},{\tt xp\_new},{\tt vfield\_new},{\tt vp\_new},{\tt PID\_new}({\rm if\ any})\}$, converting buffer state to disjoint state, and it replaces corresponding part of $\{{\tt rho\_c},{\tt xp},{\tt vfield},{\tt vp},{\tt PID}({\rm if\ any})\}$.
-When all tiles are iterated, the entire image is in disjoint state, and {\tt nplocal} is updated.
-Then all images is synchronized by {\tt sync all} and we sum up {\tt nplocal} over images to check if total number of particles {\tt npglobal} is conserved.
+In order to keep particles ordered, for each tile, we first do $x_d=x_d+v_d\,{\rm d}t$ on all particles to obtain an updated density and velocity field on the tile, $\rho_c^*$ and $v_c^*$.
+Then, this calculation in done on the same tile again\footnote{This repetition scales as $o(N)$ and is computational inexpensive.} to generate a new, local particle list $\chi_d^*$ and $\nu_d^*$ by Equations(\ref{eq.chi},\ref{eq.nu}).
+Here, the ordering of $\chi_d^*$ and $\nu_d^*$ relies on $\rho_c^*$.
+Then, the third iteration is done on this tile to delete buffer regions of $\{\rho_c^*,v_c^*,\chi_d^*,\nu_d^*\}$.
+Then the disjoint state of $\{\rho_c^*,v_c^*,\chi_d^*,\nu_d^*\}$ replaces the old $\{\rho_c,v_c,\chi_d,\nu_d\}$.
+Finally $P_{\rm local}$ and $P_{\rm global}=\sum P_{\rm local}$ are updated.
 These steps are summarized in Figure \ref{fig.update_xp}.
@@
 {\tt update\_vp} -- 
-The particle-mesh (PM) or particle-particle particle-mesh (P$^3$M) algorithm is applied in this subroutine to update particles's velocities.
-After {\tt update\_xp}, or drift $D$, we call {\tt buffer\_density} and {\tt buffer\_xp} in order that particle positions are in buffered state, based on which, we update velocities (kick $K$) of particles in physical region of each tile.
-{\tt CUBE} uses a 2-level particle mesh scheme \citep{2005NewA...10..393M,2013MNRAS.436..540H}.
-Local fine forces have a force cutoff ${\tt nforce\_cutoff}\leq{\tt ncb}$, i.e., $F_{\rm fine}(r>{\tt ncb})=0$.
-So, if we apply a fine-grid particle-mesh on an extended tile with buffered-depth {\tt ncb} (see Figure \ref{fig.tile} and regard it periodic), the force on physical regions does not depend on the false periodic boundary assumption, and the resulting fine force $F_{\rm fine}$ and velocity update on physical region is correct.
+the PM or particle-particle particle-mesh (P$^3$M) algorithm is applied in this subroutine to update particles's velocities (kick $K$).
+We first call {\tt buffer\_density} and {\tt buffer\_xp} in order that particle positions are in buffered state.
+Then according to the particle distributions $\rho_f$ on $V_e$ we calculate the fine-mesh force ${\bs F}_f$ and update the particle velocities in the $V_p$.
+Condition (\ref{eq.cond1}) guarantees the correctness of this step.
+An optional particle-particle (PP) force ${\bs F}_{\rm pp}$ can be called to increase the force resolution.
 
-\begin{figure}[]
-\begin{verbatim}
-
-subroutine update_vp
-  do (each tile)
-    calculate fine mesh density
-    calculate fine mesh force: local FFT
-    update fine mesh velocity
-    update maximum acceleration
-    if (PP force) then
-      PP calculation & velocity update
-      update maximum acceleration
-    endif
-  enddo
-  sync all
-  calculate coarse mesh density
-  calculate coarse mesh force: global pencil FFT
-  update coarse mesh velocity
-  update maximum acceleration
-  update maximum velocity
-end
-\end{verbatim}
+\begin{figure}[t]
+{\tt subroutine update\_vp\\
+\indent \ \ do (each extended tile)\\
+\indent \ \ \ \ calculate $\rho_f$ and solve ${\bs F}_f$ in Fourier space\\
+\indent \ \ \ \ do (each particle in physical region)\\
+\indent \ \ \ \ \ \ $\nu_d\rightarrow{\bs v},\,{\bs v}={\bs v}+{\bs F}_f{\rm d}t,\, {\bs v}\rightarrow\nu_d$\\
+\indent \ \ \ \ enddo\\
+\indent \ \ \ \ update $\max(|{\dot{\bs v}}|)$\\
+\indent \ \ \ \ (PP force)\\
+\indent \ \ enddo\\
+\indent \ \ calculate $\rho_c$ and solve ${\bs F}_c$ in Fourier space\\
+\indent \ \ do (each particle)\\
+\indent \ \ \ \ $\nu_d\rightarrow{\bs v},\,{\bs v}={\bs v}+{\bs F}_c{\rm d}t,\, {\bs v}\rightarrow\nu_d$\\
+\indent \ \ enddo\\
+\indent \ \ update $\max(v_d),\max(|{\dot{\bs v}}|)$\\
+end\\}
 \caption{Pseudocode for subroutine {\tt update\_vp}.}
@@
 
-The compensating coarse grid force $F_{\rm coarse}$ is globally computed by using a coarser (usually by factor of 4) mesh by dimensional splitting -- a distributed-memory cubic decomposed 3D coarse density field is interpolated by particles, and we Fourier transform data in consecutive three dimensions with global transposition in between (known as the pencil decomposition).
-After the multiplication of force kernels, the inverse transform takes place to get the cubic distributed coarse force field $F_{\rm coarse}$, upon which velocities are updated again.
+The compensating coarse grid force ${\bs F}_c$ is globally computed by using a coarser- (usually by factor of $R=4$) mesh by dimensional splitting -- the cubic distributed coarse density field $\rho_c$ is transposed (inter-image) and Fourier transformed (inner-image) in consecutive three dimensions.
+After the multiplication of memory-distributed force kernel $K_c$, the inverse transform takes place to get the cubic distributed coarse force field ${\bs F}_c$, upon which velocities are updated again.
 
-An optional particle-particle (PP) force $F_{\rm pp}$ can be called to increase the force resolution and the velocities are updated again.
-The collective operations of velocity update is Equation(\ref{eq.v}), $v_d=v_d+F_{\rm total}$ and Equation(\ref{eq.nu}).
-
-The maximum accelerations $\max(\dot{v}_{\rm fine})$, $\max(\dot{v}_{\rm coarse})$, $\max(\dot{v}_{\rm pp})$ and maximum of velocities $\max(v_d)$ are collected controlling {\tt dt} for the {\tt timestep} in the next iteration.
+For each type of velocity updates, the collective operations are Equation(\ref{eq.v}), ${\bs v}={\bs v}+{\bs F}_{\rm total}$ and Equation(\ref{eq.nu}).
 We also update $\sigma^2_{\Delta}$ according to the new $v_d-v_c$.
@@
 
-By far, {\tt vp} in physical regions are correctly updated.
-Remember that particle locations in the buffer regions has updated before {\tt PM} and remained unchanged.
-So, in the main code we simply call {\tt buffer\_v} again to bring {\tt vp} and {\tt PID} (if any) into buffered state, such that the {\tt update\_x} in the next iteration will be done correctly.
+By far, $\nu_d$ in physical regions are correctly updated.
+We simply call {\tt buffer\_v} again to bring $\nu_d$ into buffered state, such that the {\tt update\_x} in the next iteration will be done correctly.
 
 {\tt checkpoint} -- 
-If a desired redshift is reached, we execute the last drift step in the $(DKKD)^n$ operation by {\tt update\_xp}, and call {\tt checkpoint} to save the disjoint state of \{{\tt xp}, {\tt vp}, {\tt rho\_c}, {\tt vfield}\} on disk.
+If a desired redshift is reached, we execute the last drift step in the $(DKKD)^n$ operation by {\tt update\_xp}, and call {\tt checkpoint} to save the disjoint state of $\{\rho_c,v_c,\chi_d,\nu_d\}$ on disk.
 Related operations, like run-time halo finder, projections are also done at this point.
-If final desired redshift is reached, {\tt final\_step} let us exit the main loop.
-These corresponding logical variables are controlled in {\tt timestep}.
+%If final desired redshift is reached, {\tt final\_step} let us exit the main loop. These corresponding logical variables are controlled in {\tt timestep}.
 
@@
 \subsubsection{Arrays of particles}
-The arrays of particles contains {\tt xvp(6,npmax)} and a temporary {\tt xvp\_new(6,nptile)}\footnote{We use ``{\tt xvp}'' to represent $\{{\tt xp},{\tt vp}\}$, and use ``{\tt xvp\_new}'' to represent $\{{\tt xp\_new},{\tt vp\_new}\}$.}.
-{\tt npmax} is the number of particles {\tt xvp} can store:
+These are the majority memory usage.
+contains checkpoint arrays $\chi_d$, $\nu_d$, $I_P$, and temporary arrays $\chi_d^*$, $\nu_d^*$, $I_P^*$.
+The former uses memory $\mathcal{M}=(3n_\chi+3n_\nu+n_I)P_{\rm max}\,{\rm byte}$, where
 \begin{equation}\label{eq.npmax}
-	{\tt npmax}=\left\langle {\tt nplocal} \right\rangle \left( 1+\frac{2\times {\tt ncb}}{\tt nt} \right)^3(1+\epsilon_{\rm image}),
+	P_{\rm max}=\left\langle P_{\rm local} \right\rangle \left( 1+\frac{2N_b}{N_t} \right)^3(1+\epsilon_{\rm image}),
 \end{equation}
-where $\left\langle {\tt nplocal} \right\rangle$ is the average number of particles per image.
-The second term above let us store particles in buffer regions, and the third term $1+\epsilon_{\rm image}$ takes into account the inhomogeneity of {\tt nplocal} on different images.
+and $\left\langle P_{\rm local} \right\rangle$ is the average number of particles per image.
+The second term, proportional to $V_e/V_p$,  let us store additional particles in $V_b$, and the third term $1+\epsilon_{\rm image}$ takes into account the inhomogeneity of $P_{\rm local}$ on different images.
 When each image models smaller physical scales, $\epsilon_{\rm image}$ should be set larger.
 
-{\tt xvp\_new} stores temporary particles only on tiles:
+Temporary $\chi_d^*$, $\nu_d^*$, $I_P^*$ store particles only on tiles, and the particle number is set to be
 \begin{equation}\label{eq.nptile}
-	{\tt nptile}=\left\langle {\tt nplocal} \right\rangle \left( \frac{1}{\tt nnt} \right)^3(1+\epsilon_{\rm tile}),
+	P_{\rm max}^*=\left\langle P_{\rm local} \right\rangle \left( \frac{1}{M_t} \right)^3(1+\epsilon_{\rm tile}),
 \end{equation}
-where, recall that {\tt nnt} is the number of tiles per image per dimension, and similarly $\epsilon_{\rm tile}$ controls the inhomogeneity on scale of tiles.
-Larger {\tt nnt} causes more inhomogeneity on smaller tiles, and $\epsilon_{\rm tile}$ can be much larger than $\epsilon_{\rm image}$, however the term ${\tt nnt}^{-3}$ decreases much faster.
-So, the majority memory is occupied by {\tt xvp}.
+where $\epsilon_{\rm tile}$ controls the inhomogeneity on scale of tiles.
+Larger $M_t$ causes more inhomogeneity on smaller tiles, and $\epsilon_{\rm tile}$ can be much larger than $\epsilon_{\rm image}$, however the term $M_t^{-3}$ decreases much faster.
+Practically, the majority memory is occupied by $\chi_d$, $\nu_d$, $I_P$.
+
+Summarizing above we obtain the memory usage per particle (byte per particle) ${\mathcal M}_P \equiv {\mathcal M} / \left\langle P_{\rm local} \right\rangle/{\rm byte}$, given $n_\chi=n_\nu=1$ and $n_I=0$, is
+\begin{eqnarray}\label{eq.Mp_particle}
+\nonumber\mathcal{M}_P^{\rm particle}=6\Big [ (1+2N_b N_t^{-1})^3(1+\epsilon_{\rm image} )\\ 
+	+M_t^{-3}(1+\epsilon_{\rm tile})\Big ].
+\end{eqnarray}
+Because $N_t=N_c/(M_g M_t)$, we can minimize Equation (\ref{eq.Mp_particle}) by tuning $M_t$.
 
 \subsubsection{Coarse mesh arrays}
-On coarse mesh, {\tt rho\_c}, {\tt vfield} and force kernel {\tt kern\_c} should always be kept.
-They are usually configured to be 4 times coarser than fine grids and particle number density.
-In this case, each coarse grid, 3D {\tt vfield} takes only 12 bytes, and 64 particles will at least take 384 bytes ($n_\chi=n_\nu=1$ case).
-Similarly, pencil-FFT arrays, coarse force arrays etc.
-are memory-light and/or temporary.
+On coarse mesh, $\rho_c$ (4-byte integers), $v_c$ and force kernel $K_c$ should always be kept.
+They are usually configured to be $R=4$ times coarser than fine grids and particle number density.
+They have use memory $(1+3)\times 4\times(N_e M_t)^3+3\times 4\times(N_c/M_g)^3/2$ bytes per image, or
+\begin{equation}\label{eq.Mp_coarse}
+	\mathcal{M}_P^{\rm coarse}=P_c^{-1} \left [ 16\left( 1+\frac{2N_b}{N_t} \right)^3+6 \right ],
+\end{equation}
+where $P_c$ is the average number of particles per coarse cell.
+Other coarse grid based arrays are $\rho_c^*$, $F_c$ pencil-FFT arrays.
+$\rho_c^*$ exists only on tiles; $F_c$ and pencil-FFT arrays can be equivalenced with other temporary arrays.
+Thus the major memory usage from coarse FFT arrays are from Equation (\ref{eq.Mp_coarse}).
 
 \subsubsection{Fine mesh arrays}
-On the local fine mesh, only a force kernel array needs to be kept.
+On the local fine mesh, only a force kernel array $K_f$ needs to be kept. It has memory of $3\times 4\times(RN_e M_t)^3/2$ per image, or
+\begin{eqnarray}\label{eq.Mp_fine}
+\nonumber	\mathcal{M}_P^{\rm fine}&=&6 P_f^{-1} N_t^{-3} \left ( 1+\frac{2N_b}{N_t} \right )^3 \\
+					 &=& 6 R^3 P_c^{-1} N_t^{-3} \left ( 1+\frac{2N_b}{N_t} \right )^3,
+\end{eqnarray}
+where $P_f$ is the average number of particles per fine cell.
 The fine mesh density arrays, force field arrays are temporary.
-Since {\tt xvp\_new}, coarse force arrays, pencil-FFT arrays are also temporary and are not used in any calculation simultaneously, they can be overlapped in memory by using equivalent statements.
+Since $\chi_d^*$, $\nu_d^*$, coarse force arrays, pencil-FFT arrays are also temporary and are not used in any calculation simultaneously, they can be overlapped in memory by using equivalent statements.
 
@@
 \hline
-Particles  & \tt xvp          & 29.9   & 8.24   & 83.8\%     \\
-           & \tt xvp\_new     & \underline{3.16}   & \underline{0.872}  & \underline{8.87\%}     \\
-           & \tt PID          & 0      & 0      &    0\%     \\
+Particles  & $\chi_d$, $\nu_d$      & 29.9   & 8.24   & 83.8\%     \\
+           & $\chi_d^*$, $\nu_d^*$  & \underline{3.16}   & \underline{0.872}  & \underline{8.87\%}     \\
+           & $I_P$, $I_P^*$         & 0      & 0      &    0\%     \\
            & Subtotal               & 33.0   & 9.12   & 92.7\%     \\
 \hline
-Coarse mesh& \tt rho\_c       & 0.296  & 0.0818 & 0.831\%    \\
-           & \tt vfield       & 0.889  & 0.245  & 2.49\%     \\
-           & \tt kern\_c      & 0.340  & 0.0939 & 0.954\%    \\
-           & \tt force\_c     & (0.690)  & (0.190)  & (1.94\%)    \\
-           & \tt rho\_c\_new  & 0.0140 & 0.00388  & 0.0394\% \\
+Coarse mesh& $\rho_c$               & 0.296  & 0.0818 & 0.831\%    \\
+           & $v_c$                  & 0.889  & 0.245  & 2.49\%     \\
+           & $K_c$                  & 0.340  & 0.0939 & 0.954\%    \\
+           & $F_c$                  & (0.690)  & (0.190)  & (1.94\%)    \\
+           & $\rho_c^*$             & 0.0140 & 0.00388  & 0.0394\% \\
            & Pencil-FFT             & (0.454)  & (0.125) & (1.27\%)      \\
@@
 \hline
-Fine mesh  & \tt kern\_f      & 1.06   & 0.292 & 2.97\%      \\
-           & \tt force\_f     & (1.63) & (0.450) & (4.57\%)     \\
+Fine mesh  & $K_f$                  & 1.06   & 0.292 & 2.97\%      \\
+           & $F_f$                  & (1.63) & (0.450) & (4.57\%)     \\
            & Fine-FFT               &   (1.41) & (0.389) & (3.96\%)     \\
@@
 \subsubsection{Compare with traditional algorithms}
-To illustrate the improvement of memory usage, we refer to the TianNu simulation \citep{2017NatAs...1E.143Y} on the Tianhe-2 supercomputer, which used a traditional $N$-body code {\tt CUBEP3M}.
-TianNu's particles number is limited by memory per computing node -- for each computing node, an average of $576^3$ neutrino particles and $288^3$ CDM particles are used, and consumes about 40 GB of memory\footnote{Additional memory is used for OpenMP parallelization and for particle IDs to differentiate different particle spices.}, or about 186 bpp.
-A memory efficient case of {\tt CUBEP3M} by using large physical scales and at costs of speed, still uses about 40 bpp.
+To illustrate the improvement of memory usage, we refer to the TianNu simulation \citep{2017NatAs...1E.143Y} run on the Tianhe-2 supercomputer, which used a traditional $N$-body code {\tt CUBEP3M}.
+TianNu's particles number is limited by memory per computing node -- for each computing node, an average of $576^3$ neutrino particles and $288^3$ CDM particles are used, and consumes $\mathcal{M}=40\,{\rm GB}$\footnote{Additional memory is used for OpenMP parallelization and for particle IDs to differentiate different particle spices.}, or about $\mathcal{M}_P=186$.
+A memory efficient configuration of {\tt CUBEP3M} by using large physical scales and at costs of speed, still uses about $\mathcal{M}_P=40$.
 
-If the same amount of memory is allocated to {\tt CUBE}, we can set parameters as $n_\chi=n_\nu=1$, ${\tt nc}=384$, ${\tt nnt}=3$, ${\tt ncb}=6$, $\epsilon_{\rm image}=5\%$, $\epsilon_{\rm tile}=200\%$.
-This gives $\left\langle {\tt nplocal} \right\rangle=1536^3$, and uses about 35.6 GB of memory, corresponding to 9.84 bpp.
+If the same amount of memory is allocated to {\tt CUBE}, we can set parameters as $n_\chi=n_\nu=1$, $N_c/M_g=384$, $N_b=6$, $\epsilon_{\rm image}=5\%$, $\epsilon_{\rm tile}=200\%$ and thus $\left\langle P_{\rm local}\right\rangle=1536^3$.
+Setting $M_t=3$ (27 tiles per image) minimizes $\mathcal{M}$ and uses about $\mathcal{M}=35.6\,{\rm GB}$, corresponding to $\mathcal{M}_P=9.84$.
 This can be done on most of the supercomputers, even modern laptops.
@@
 Table \ref{t.memory} shows the memory consumption for this test simulation.
-The memory-consuming arrays are listed and classified into the three types above mentioned, and their memory usages are in unit of GB ($10^9$ bytes), byte per particle (bpp), and their percentage to the total memory usage.
+The memory-consuming arrays are listed and classified into the three types above mentioned, and their memory usages are in unit of GB ($10^9$ byte), byte per particle (bpp), and their percentage to the total memory usage.
 The parenthesized numbers show overlapped memory, which is saved by equivalencing them with the underscored numbers.
 There are other, unlisted variables which are memory-light, and can also be equivalenced with listed variables.
-For this $1536^3$ per image simulation, using ${\tt nnt}^3=27$ tiles per image optimizes the memory usage, because it well controls the balance between the second terms in Equations (\ref{eq.npmax},\ref{eq.nptile}).
-A great amount of temporary memory in coarse and fine mesh force calculations is equivalenced to {\tt xvp\_new}.
 
-In the bottom of Table \ref{t.memory} we stress that the optimal memory usage is 6 bpp, or 21.7 GB, 61\% of the actual memory usage in this test simulation.
-The most prominent departure from this limit is that {\tt xvp} already uses 8.24 bpp, and all other variables occupy only additional 8\%.
-Of coarse, on modern super computers, the memory per computing node is usually much larger, by scaling up the number of particles per node, the buffer ratio ${\tt ncb}/{\tt nt}$ will be lowered and we can approach closer to the 6 bpp limit.
+In the bottom of Table \ref{t.memory} we stress that the optimal memory usage is 6 bpp, or 21.7 GB, 61\% of the actual $\mathcal{M}$.
+The dominating departure from this limit is that $\chi_d$ and $\nu_d$ already occupy 8.24 bpp, which come from the $(1+2N_b/N_t)^3$ term of Equation (\ref{eq.Mp_particle}).
+All other variables occupy additional 8\%.
+On modern super computers, the memory per computing node is usually much larger, and by scaling up the number of particles per node, the buffer ratio $N_b/N_t$ will be lowered and we can approach closer to the 6 bpp limit.
 
@@
 First, by using different configurations -- different number of computing nodes, box sizes, particle resolutions, different number of tiles per node/image etc., we find that by using 2-byte integers for both positions and velocities ($n_\chi=n_\nu=2$, or x2v2) {\tt CUBE} gives exact results compared to {\tt CUBEP3M}.
-So, if memory usage is not a problem, one can always double the memory usage to get exact results as {\tt CUBEP3M}, and the optimal memory limit of this case is 12 bpp, still much lower than traditional methods.
+So, if sufficient memory is provided, one can always use x2v2 to get exact results as {\tt CUBEP3M}, and the optimal memory limit of this case is 12 bpp, still much lower than traditional methods.
 Next, we focus on the accuracy of the other three cases -- x1v2, x2v1 and x1v1.
@@
 
-\bibliographystyle{hapj}
+\bibliographystyle{aasjournal}
 \bibliography{haoran_ref}

### 2018-06-03 CUBEnu/work/main/initialize.f90
parallel on niagara

@@ -50,7 +50,6 @@ subroutine initialize
   halofind_step=.false.
   final_step=.false.
 
-  if (head) then
     open(16,file='z_checkpoint.txt',status='old')
     do i=1,nmax_redshift-1
       read(16,end=71,fmt='(f8.4)') z_checkpoint(i)
@@ -63,7 +62,6 @@ subroutine initialize
     enddo
     81 n_halofind=i-1
     close(16)
-  endif
   if (n_checkpoint==0) stop 'z_checkpoint.txt empty'
   if (n_halofind==0) stop 'z_halofind.txt empty'
   tau=-3/sqrt(1./(1+z_checkpoint(cur_checkpoint)))

### 2018-06-03 CUBEnu/work/utilities/module_load_niagara_intel.sh

@@ -1,7 +1,7 @@
 module load NiaEnv/2018a intel/2018.2 intelmpi/2018.2 fftw/3.3.7
 
 #export XFLAG=' -O3 -cpp -fopenmp -fcoarray=single -mcmodel=medium'
-export XFLAG=' -O1 -g -xHost -fpp -qopenmp -coarray -mcmodel=medium'
+export XFLAG=' -O1 -g -xHost -fpp -qopenmp -coarray=distributed -mcmodel=medium'
 #export XFLAG='-O3 -cpp -fcoarray=single -fopenmp'
 export OFLAG=${XFLAG}' -c'
 export FFTFLAG='-I$(SCINET_FFTW_ROOT)/include/ -L$(SCINET_FFTW_ROOT)/lib/ -lfftw3f -lm -ldl'
@@ -9,8 +9,8 @@ export FFTFLAG='-I$(SCINET_FFTW_ROOT)/include/ -L$(SCINET_FFTW_ROOT)/lib/ -lfftw
 # in cumsum6
 
 export OMP_STACKSIZE=16000M
-export KMP_STACKSIZE=16000M
+#export KMP_STACKSIZE=16000M
 export OMP_NUM_THREADS=80
 #export OMP_THREAD_LIMIT=4
-export FOR_COARRAY_NUM_IMAGES=1
+export FOR_COARRAY_NUM_IMAGES=8
 ulimit

### 2018-06-03 ms_caf/ms_caf.tex
nu vtf

@@
 \usepackage[english]{babel}
-\usepackage{amsmath,color,aas_macros}
+\usepackage{amsmath,color,aas_macros,ulem}
 
@@
 $^2$Canadian Institute for Theoretical Astrophysics, University of Toronto, Toronto, ON M5H 3H8 Canada\\
-$^3$Department of Physics and Astronomy, Shanghai Jiao Tong University, Shanghai, 200240, China\\
+$^3$Department of Astronomy, Shanghai Jiao Tong University, Shanghai, 200240, China\\
 $^4$Dunlap Institute for Astronomy and Astrophysics, University of Toronto, Toronto, ON M5S 3H4, Canada\\
@@
 Our numerical experiments show that this information-optimized $N$-body algorithm provides accurate results within the error of the particle-mesh algorithm.
-This significant lowering of memory-to-computation ratio breaks the bottleneck of scaling up and speeding up large cosmological $N$-body simulations on multi-core and heterogenous computing systems.
+This significant lowering of memory-to-computation ratio breaks the bottleneck of scaling up and speeding up large cosmological $N$-body simulations on multi-core and heterogeneous computing systems.
 
@@
 When $N$ is large, the brute force pairwise particle-particle (PP) force brings unaffordable $o(N^2)$ computations, so many algorithms were designed to alleviate it.
-Various fast multipole methods \citep{1985JCoPh..60..187R} improve the complexity to $o(N\log N)$ even $o(N)$, among which the most popular one is ``tree'', like {\tt GADGET} \citep{2001NewA....6...79S,2005MNRAS.364.1105S} and its simulation ``Millennium'' \citep{2005Natur.435..629S,2012MNRAS.426.2046A}, {\tt TPM} \citep{1995ApJS...98..355X} and {\tt GOTPM} \citep{2004NewA....9..111D}.
+Various fast multipole methods \tcb{\citep{1985JCoPh..60..187R,2014ComAC...1....1D,2016ascl.soft09016P}} improve the complexity to $o(N\log N)$ even $o(N)$, among which the most popular one is ``tree'', like {\tt GADGET} \citep{2001NewA....6...79S,2005MNRAS.364.1105S} and its simulation ``Millennium'' \citep{2005Natur.435..629S,2012MNRAS.426.2046A}, {\tt TPM} \citep{1995ApJS...98..355X} and {\tt GOTPM} \citep{2004NewA....9..111D}.
 Other methods include adaptive grid algorithms like {\tt HYDRA} \citep{1995ApJ...452..797C} and {\tt RAMSES} \citep{2010ascl.soft11007T}, as well as mesh-refined codes \citep{1991ApJ...368L..23C} and moving adaptive particle-mesh (PM) codes \citep{1995ApJS..100..269P}.
@@
 The load balancing problem is minimized as the matter distribution is rather homogeneous, and the speed benefits from the fast Fourier transform (FFT) libraries, such as {\tt FFTW3} \citep{Frigo05thedesign}.
-{\tt PMFAST} \citep{2005NewA...10..393M} is first codes to use a 2-level PM algorithm, aiming to put PM codes toward speed, memory compactness and scalability.
+{\tt PMFAST} \citep{2005NewA...10..393M} introduces a 2-level PM algorithm, aiming to put PM codes toward speed, memory compactness and scalability.
 After subsequent developments on {\tt PMFAST}, {\tt CUBEP3M} \citep{2013MNRAS.436..540H} uses cubic spatial decomposition, and adds PP force and many other features.
@@
 \section{Method}\label{s.method}
-The most memory consuming part of a $N$-body simulation is usually the phase space coordinates of $N$-body particles -- 24 byte per particle (here after we use ``bpp'' referring to ``byte per particle''), which contains 4 single precision floating numbers, must be used to store each particles' 3-dimensional position and velocity vectors.
+The most memory consuming part of a $N$-body simulation is usually the phase space coordinates of $N$-body particles -- 24 byte per particle (here after we use ``bpp'' referring to ``byte per particle''), which contains 6 single precision floating numbers, must be used to store each particles' 3-dimensional position and velocity vectors.
 {\tt CUBEP3M}, an example of a memory-efficient parallel $N$-body code, can use as low as 40 bpp in sacrificing computing speed \citep{2013MNRAS.436..540H}.
@@
 Sometimes 4-byte real numbers are not necessarily adequate in representing the {\it global} coordinates in simulations.
-If the box size is many orders of magnitude larger than interactive distance between particles, especially in the field of re-simulation of dense subregions, double precision (8-byte) coordinates are needed to avoid truncation errors.
+If the box size is many orders of magnitude larger than interactive distance between particles, especially in the field of re-simulation of dense subregions, double precision (8-byte) coordinates are needed to avoid \tcb{round-off} errors.
 Another solution is to record {\it relative} coordinates for both position and velocity.
@@
 On the other hand, particle position can also be stored as 2-byte (16 bits) integers to increase the resolution.
-In this case, each coarse cell is divided into $2^{16}=65536$ bins and the position resolution is $L/(65536N_c)$, precise enough compared to using4-byte global coordinates, see later results.
+In this case, each coarse cell is divided into $2^{16}=65536$ bins and the position resolution is $L/(65536N_c)$, precise enough compared to using 4-byte global coordinates, see later results.
 We denote this case ``x2'' and denote using 1-byte integers for positions ``x1''.
@@
 The velocity counterpart of them are $n_\nu=1,2$, $v_d$ and $\nu_d$.
-%$n_c=1,2,...$ is the coarse grid index, given by the ordering of particles and a integer based particle density field (see \ref{sss.spatial_decomposition}).
 The position resolution for $n_\chi$-byte integer, ``x$n_\chi$'', is $2^{-8n_\chi}L/N_c$.
@@
 Similarly, actual velocity in $d$th dimension $v_d$ is decomposed into an averaged velocity field on the same coarse grid $v_c$ and a residual $\Delta v$ relative to this field:
-\begin{equation}
+\begin{equation}\label{eq.deltav}
 	v_d=v_c+\Delta v.
@@
 \begin{equation}\label{eq.vdisp}
-	\sigma_v(a,k) = \frac{1}{3}  (a H f D)^2 \int_0^kd^3q\frac{P_{L}(q)}{q^2},
+	\tcb{\sigma^2_v(a,r) = (a H f D)^2 \int_0^\infty d^3k\frac{P(k)}{k^2}W^2(k,r)},
 \end{equation}
-where $a(z)$ is the scale factor, $H(z)$ the Hubble parameter, $D$ is the linear growth factor, $f=d \ln D/d\ln a$, and $P_L(k)$ is the linear power spectrum of density contrast at redshift zero.
-$\sigma_v(a,k)$ is a function of cosmic evolution $a$ and a smoothing scale $k$, or $r$ (see Figure \ref{fig.vdisp}).
-$\Delta v$ is the velocity dispersion relative to the coarse grid, so we approximate its variance as
+where $a(z)$ is the scale factor, $H(z)$ is the Hubble parameter, $D$ is the linear growth factor, $f=d \ln D/d\ln a$, and $P(k)$ is the linear power spectrum of density contrast at redshift zero.
+\tcb{$W(k,R)$ is the Fourier transform of real space top-hat window function with smoothing scale $r$.
+In Figure \ref{fig.vdisp} we plot $\sigma_v(a,r)$ as a function of $a$ for a few smoothing scale $r$.
+$\Delta v$ in equation (\ref{eq.deltav})} is the velocity dispersion relative to the coarse grid, so we approximate its variance as
 \begin{equation}\label{eq.vdelta}
@@
 Initially, particles are generated by initial condition generator, at a higher redshift.
-The coarse grid density field $v_c$ is also generated at this step by averaging all particles in the coarse cell.
+The coarse grid \tcb{velocity} field $v_c$ is also generated at this step by averaging all particles in the coarse cell.
 A global $\sigma_\Delta$ is calculated by equation (\ref{eq.vdelta}), where linear approximation is hold.
@@
 Each $V_p$ is surrounded by a {\it buffer} region $V_b$ which is $N_b$ coarse cells thick.
-We define extended tile region $V_e\equiv V_p+V_b$.
-$V_e$ is designed for two reasons: (1) if the short fine-mesh force ${\bs F}_f$ has a cut-off $N_b$, and is computed by $\rho_c$ whose domain $D(\rho_f)=V_e$, then ${\bs F}_f$ in $V_p$ is guaranteed to be correct, i.e.,
-\begin{eqnarray}\label{eq.cond1}
-\nonumber	{\bs F}_f(r>N_b)&=&0 \wedge D(\rho_c)=V_e\\
-	&\Rightarrow& \{{\bs F}_f(\rho_c) |{\bs F}_f\in V_p \} {\rm \ is\ correct},
-\end{eqnarray}
-and (2) to collect all particles being able to travel to $V_p$, i.e.,
-\begin{equation}\label{eq.cond2}
-	{\bs v}\,{\rm d}t \leq N_b \Rightarrow \{ {\bs x} | {\bs x}+{\bs v}\,{\rm d}t\in V_p\} \in V_e.
-\end{equation}
+We define \tcb{the} extended tile region \tcb{as} \tcb{$V_e\equiv V_p\cup V_b$}.
+$V_e$ is designed for two reasons:
+
+(1) if the short fine-mesh force ${\bs F}_f$ has a cut-off, \tcb{$N_b$ ${\bs F}_f(r>N_b)=0$}, and is computed \tcr{\sout{ by $\rho_c$ whose domain $D(\rho_f)=V_e$}} \tcb{on $V_e$}, then ${\bs F}_f$ in $V_p$ is guaranteed to be correct.
+
+\tcb{(2) $V_e$ is able to collect all particles that are able to travel to $V_p$.} \tcr{\sout{i.e., ... }[deleted previous two equations (8,9)]}
+%\begin{eqnarray}\label{eq.cond1}
+%\nonumber	{\bs F}_f(r>N_b)&=&0 \wedge D(\rho_c)=V_e\\
+%	&\Rightarrow& \{{\bs F}_f(\rho_c) |{\bs F}_f\in V_p \} {\rm \ is\ correct},
+%\end{eqnarray}
+
+%and (2) to collect all particles being able to travel to $V_p$, i.e.,
+%\begin{equation}\label{eq.cond2}
+%	{\bs v}\,{\rm d}t \leq N_b \Rightarrow \{ {\bs x} | {\bs x}+{\bs v}\,{\rm d}t\in V_p\} \in V_e.
+%\end{equation}
 Figure \ref{fig.tile} shows the spacial decomposition in a 2-dimensional analogy, with $M_g=2$ and $M_t=2$.
@@
 $\chi_d$ and $\nu_d$
-%(equivalent to $\chi_d$ and $\nu_d$ in \ref{ss.position} and \ref{ss.velocity} respectively)
 are required to be sorted according to the same memory layout of $\rho_c$, such that $n_c$ and $x_d$ can be obtained by equation (\ref{eq.x}).
@@
 This step is done on $V_e$ of each tile.
-Similar to Equation (\ref{eq.cond2}), we have
-\begin{equation}\label{eq.cond3}
-	\max(\Psi_d) \leq N_b \Rightarrow \{ {\bs q} | {\bs q}+{\bs \Psi}({\bs q})\in V_p\} \in V_e.
-\end{equation}
+\tcr{\sout{Similar to condition (9), we have ...} [deleted previous equation (10).]}
+%\begin{equation}\label{eq.cond3}
+%	\max(\Psi_d) \leq N_b \Rightarrow \{ {\bs q} | {\bs q}+{\bs \Psi}({\bs q})\in V_p\} \in V_e.
+%\end{equation}
+
 We iterate twice over particles' ${\bs q}$ in $V_e$.
@@
 {\tt timestep} --
-Runge-Kutta 2 method is used in the time integration.
+\tcb{The second order Runge-Kutta method} is used in the time integration.
 i.e., for $n$ time steps, we update positions ($D$=drift) and velocities ($K$=kick) at interlaced half time-steps by operator splitting; the operation $(DKKD)^n$ is 2nd order accurate.
@@
 {\tt update\_xp} --
-According to ${\rm d}t$, {\tt update\_xp} updates particle positions (drift $D$) in a ``{\it gather}'' algorithm
-%(in contrast, {\tt CUBEP3M} uses ``scatter'' algorithm)
-tile by tile.
+According to ${\rm d}t$, {\tt update\_xp} updates particle positions (drift $D$) in a ``{\it gather}'' algorithm tile by tile.
 For each particle, $\chi_d$ and $\nu_d$ are converted to $x_d$ and $v_d$ by Equations(\ref{eq.x},\ref{eq.v}).
-Due to Condition (\ref{eq.cond2}), we are able to collect all particles to travel to $V_p$.
+\tcr{\sout{Due to Condition (9), we are able to collect all particles to travel to $V_p$.}}
 
@@
 Then according to the particle distributions $\rho_f$ on $V_e$ we calculate the fine-mesh force ${\bs F}_f$ and update the particle velocities in the $V_p$.
-Condition (\ref{eq.cond1}) guarantees the correctness of this step.
-An optional particle-particle (PP) force ${\bs F}_{\rm pp}$ can be called to increase the force resolution.
+\tcr{\sout{Condition (8) guarantees the correctness of this step.}}  An optional particle-particle (PP) force ${\bs F}_{\rm pp}$ can be called to increase the force resolution.
 
@@
 
-By far, $\nu_d$ in physical regions are correctly updated.
-We simply call {\tt buffer\_v} again to bring $\nu_d$ into buffered state, such that the {\tt update\_x} in the next iteration will be done correctly.
+\tcb{After the updating of $\nu_d$ in $V_p$, we} simply call {\tt buffer\_v} again to bring $\nu_d$ into buffered state, such that the {\tt update\_x} in the next iteration will be done correctly.
 
@@
 Related operations, like run-time halo finder, projections are also done at this point.
-%If final desired redshift is reached, {\tt final\_step} let us exit the main loop. These corresponding logical variables are controlled in {\tt timestep}.
 
@@
 To illustrate the improvement of memory usage, we refer to the TianNu simulation \citep{2017NatAs...1E.143Y} run on the Tianhe-2 supercomputer, which used a traditional $N$-body code {\tt CUBEP3M}.
-TianNu's particles number is limited by memory per computing node -- for each computing node, an average of $576^3$ neutrino particles and $288^3$ CDM particles are used, and consumes $\mathcal{M}=40\,{\rm GB}$\footnote{Additional memory is used for OpenMP parallelization and for particle IDs to differentiate different particle spices.}, or about $\mathcal{M}_P=186$.
+TianNu's particle number \tcb{(shown in Table \ref{t.sim})} is limited by memory per computing node -- for each computing node, an average of $576^3$ neutrino particles and $288^3$ CDM particles are used, and consumes $\mathcal{M}=40\,{\rm GB}$\footnote{Additional memory is used for OpenMP parallelization and for particle IDs to differentiate different particle spices.}, or about $\mathcal{M}_P=186$.
 A memory efficient configuration of {\tt CUBEP3M} by using large physical scales and at costs of speed, still uses about $\mathcal{M}_P=40$.
@@
   \includegraphics[width=1.0\linewidth]{f9}
- \caption{Dimensionless power spectra $\Delta^2(k)$ and decorrelations $\xi(k)$ (lower axes) induced by the integer-based algorithms in simulations S2048S and S2048L.
+ \caption{Dimensionless power spectra $\Delta^2(k)$ and \tcb{the cross-correlations $\xi(k)$ (lower axes) with respect to x2v2} in simulations S2048S and S2048L.
 The four solid lines show the results from \{x2v2,x1v2,x2v1,x1v1\}.
@@
 For x2v2, almost all particles are accurate up to 1/100 of a fine grid, and the worst particle is $\sim 1/10$ of a fine grid away from its counterpart in {\tt CUBEP3M}.
-The difference is caused by truncation errors and is negligible in physical and cosmological applications.
+The difference is caused by \tcb{round-off} errors and is negligible in physical and cosmological applications.
 The accuracy of x1v2 is between x2v2 and x1v1, and x2v1 gives only minor improvement from x1v1.
@@
 
-In the lower parts of these four panels study the cross correlations.
-We compare everything with {\tt CUBE}-x2v2, to emphasize the decorrelation by different integer formats.
+In the lower parts of these four panels study the cross-correlations.
+We compare everything with {\tt CUBE}-x2v2, to emphasize the decorrelation \tcb{(i.e.
+$1-\xi(k)$)} by different integer formats.
 Note that x2v2 is perfectly correlated with {\tt CUBEP3M}.
@@
 \section{Discussion and conclusion}\label{s.discussion}
-We present an parallel, information-optimized $N$-body algorithm.
+We present a parallel, information-optimized $N$-body algorithm.
 This open-source code {\tt CUBE}, has recently been used in many studies of LSS, e.g.
@@
 We acknowledge funding from NSERC.
-H.R.Y. thanks Derek Inman for many helpful discussions.
+H.R.Y. thanks Derek Inman for many helpful discussions. \tcb{We thank the anonymous referee for many helpful suggestions which improve the paper.}
 The simulations were performed on the GPC supercomputer at the SciNet HPC Consortium.

### 2018-06-03 CUBEnu/work/main/update_particle.f90

@@ -218,7 +218,7 @@ subroutine update_xp()
     print*,'  std_vsim_res',std_vsim_res*sim%vsim2phys,'km/s'
     print*,'  sigma_vi    ',sigma_vi,'(simulation unit)'
     print*,'  sigma_vi_new',sigma_vi_new,'(simulation unit)'
-    write(77) a-da,real((/std_vsim,std_vsim_c,std_vsim_res/))
+    write(77) a-da,real([std_vsim,std_vsim_c,std_vsim_res])*sim%vsim2phys
   endif
   sync all
 

### 2018-06-03 CUBEnu/work/utilities/initial_conditions.f90

@@ -50,7 +50,7 @@ program initial_conditions
   integer(izipx) xp(3,npmax)
   integer(izipv) vp(3,npmax)
 #ifdef PID
-    integer(8) pid(npmax)
+    integer(4) pid(npmax)
 #endif
   real grad_max(3)[*],vmax(3),vf
   real(4) svz(500,2),svr(100,2)
@@ -123,7 +123,7 @@ program initial_conditions
   sim%omega_m=omega_m
   sim%omega_l=omega_l
   sim%s8=s8
-  sim%vsim2phys=(150./sim%a)*box*h0*sqrt(omega_m)/nf_global
+  sim%vsim2phys=(150./sim%a)*box*sqrt(omega_m)/nf_global
   sim%z_i=z_checkpoint(cur_checkpoint)
   sim%z_i_nu=z_i_nu
   sync all
@@ -203,7 +203,7 @@ program initial_conditions
   if (head) print*,'Generating random noise'
   call random_seed(size=seedsize)
   if (head) print*,'  min seedsize =', seedsize
-  seedsize=max(seedsize,36)
+  seedsize=max(seedsize,12)
   allocate(iseed(seedsize))
   allocate(rseed_all(seedsize,nn**3))
 #ifdef READ_SEED
@@ -455,10 +455,12 @@ program initial_conditions
   do i=1,nn**3 ! co_max
     grad_max=max(grad_max,grad_max(:)[i])
   enddo
+  sync all
   vmax=grad_max/2/(4*pi)*vf
 
   sim%dt_vmax=vbuf*20./maxval(abs(vmax))
   sim%vz_max=vmax(3)
+  nlayer=2*ceiling(grad_max(3)/8/pi/ncell)+1
   if (head) then
     print*, '  grad_max',grad_max
     print*, '  max dsp',grad_max/2/(4*pi)
@@ -469,8 +471,8 @@ program initial_conditions
       print*, maxval(grad_max)/2/(4*pi),nfb
       stop
     endif
-    nlayer=2*ceiling(grad_max(3)/8/pi/ncell)+1
     print*,'  Thread save nlayer =',nlayer
+    print*,''
   endif
   sync all
 
@@ -634,6 +636,8 @@ program initial_conditions
     close(15)
 # endif
   call system_clock(t2,t_rate)
+  sync all
+
   if (head) print*, '  elapsed time =',real(t2-t1)/t_rate,'secs';
 
   if (head) then

### 2018-06-03 CUBEnu/work/utilities/initial_conditions_neu_derek.f90

@@ -1,0 +1,928 @@
+!#define READ_NOISE
+
+program initial_conditions
+  use pencil_fft
+  !use powerspectrum
+  use iso_fortran_env, only : int64
+  implicit none
+  save
+
+  ! nc: coarse grid per image per dim
+  ! nf: fine grid per image per dim, ng=nf
+  ! nyquest: Nyquest frequency
+  logical,parameter :: correct_kernel=.true.
+  logical,parameter :: write_potential=.true.
+
+  real,parameter :: vsim2phys_zi_nu=(150./a_i_nu)*box*h0*sqrt(omega_m)/nf_global
+  real, parameter :: fdf = 25.8341 !kT/m for T=1K, m=1eV
+  real, parameter :: fd = (1./vsim2phys_zi_nu)*fdf*maxval(Tnu/Mnu)/a_i_nu !kBcTnu/mass with temp in K and mass in eV
+  real, parameter :: sigma_vi_nu = 3.59714*fd/sqrt(3.) !fd velocity dispersion (45/3 * Zeta(5)/Zeta(3))**0.5
+
+  !Fermi-Dirac CDF
+  integer, parameter :: ncdf = 10000
+  real, dimension(2,ncdf) :: cdf
+
+  integer(8),parameter :: nk=132 ! ???
+  integer(8) i,j,k,n
+  integer(4) seedsize
+  real kmax,temp_r,temp_theta,pow,phi8,temp8[*]
+  real(8) v8, norm, xq(3),gradphi(3),gradphiv(3),vreal(3), dvar[*], dvarg
+  integer(int64) :: time64
+
+  integer(8) nplocal[*],npglobal,ip,l,idx_v
+  ! power spectrum arrays
+  real, dimension(20,nk) :: tf, tfv, tfd !CLASS
+  real, dimension(2,nc) :: pkm,pkn
+
+  integer(4),allocatable :: iseed(:)
+  real,allocatable :: rseed_all(:,:)
+
+  integer(8) ind,dx,dxy,kg,mg,jg,ig,ii,jj,kk,itx,ity,itz,idx,imove,nf_shake
+  integer(8) ileft,iright,nlen,nlast,g(3)
+  real kr,kx,ky,kz
+  !real xi(10,nbin)
+
+  complex delta_k(nyquest+1,nf,npen)
+  real phi(-nfb:nf+nfb+1,-nfb:nf+nfb+1,-nfb:nf+nfb+1)[*]
[…]

### 2018-06-03 CUBEnu/work/utilities/initial_conditions_nu.f90

@@ -1,8 +1,12 @@
-!#define READ_NOISE
+#define sigma_8
+#define READ_SEED
+#define READ_NOISE
+#define MTF
 
-program initial_conditions
+program initial_conditions_nu
+  use omp_lib
+  use variables, only: spine_tile
   use pencil_fft
-  !use powerspectrum
   use iso_fortran_env, only : int64
   implicit none
   save
@@ -13,157 +17,191 @@ program initial_conditions
   logical,parameter :: correct_kernel=.true.
   logical,parameter :: write_potential=.true.
 
-  real,parameter :: vsim2phys_zi_nu=(150./a_i_nu)*box*h0*sqrt(omega_m)/nf_global
-  real, parameter :: fdf = 25.8341 !kT/m for T=1K, m=1eV
-  real, parameter :: fd = (1./vsim2phys_zi_nu)*fdf*maxval(Tnu/Mnu)/a_i_nu !kBcTnu/mass with temp in K and mass in eV
-  real, parameter :: sigma_vi_nu = 3.59714*fd/sqrt(3.) !fd velocity dispersion (45/3 * Zeta(5)/Zeta(3))**0.5
+#ifdef sigma_8
+  integer(8),parameter :: nk=1000
+  real tf(7,nk), vtf(7,nk)
+#else
+  integer(8),parameter :: nk=132
+  real tf(14,nk)
+#endif
 
-  !Fermi-Dirac CDF
-  integer, parameter :: ncdf = 10000
-  real, dimension(2,ncdf) :: cdf
-
-  integer(8),parameter :: nk=132 ! ???
-  integer(8) i,j,k,n
-  integer(4) seedsize
-  real kmax,temp_r,temp_theta,pow,phi8,temp8[*]
-  real(8) v8, norm, xq(3),gradphi(3),gradphiv(3),vreal(3), dvar[*], dvarg
-  integer(int64) :: time64
-
-  integer(8) nplocal[*],npglobal,ip,l,idx_v
-  ! power spectrum arrays
-  real, dimension(20,nk) :: tf, tfv, tfd !CLASS
-  real, dimension(2,nc) :: pkm,pkn
+  integer(8) i,j,k,ip,l,nzero
+  integer(8) ind,dx,dxy,kg,mg,jg,ig,ii,jj,kk,itx,ity,itz,idx,imove,g(3),iq(3)
[…]

### 2018-06-03 CUBEnu/work/utilities/module_load_mac.sh

@@ -10,3 +10,5 @@ export FFTFLAG='-I/usr/local/include/ -L/usr/local/lib/ -lfftw3f -lm -ldl'
 export OMP_STACKSIZE=1000M
 export OMP_NUM_THREADS=4
 #export OMP_THREAD_LIMIT=4
+ulimit -s 61000
+ulimit

### 2018-07-29 ms_caf/ms_caf.tex
ApJS publication version of CUBE paper

@@
 \usepackage[english]{babel}
-\usepackage{amsmath,color,aas_macros,ulem}
+\usepackage{amsmath,xcolor,aas_macros,ulem}
 
@@
 \newcommand{\tcb}{\textcolor{blue}}
+\newcommand{\tcv}{\textcolor{violet}}
 \newcommand{\tcr}{\textcolor{red}}
+\newcommand{\tcx}{\textcolor{teal}}
 
@@
 \author{
-Hao-Ran Yu$^{1,2,3}$,
-Ue-Li Pen$^{2,4,5,6}$,
+Hao-Ran Yu$^{1,2,3\star}$,
+Ue-Li Pen$^{2,1,4,5,6}$,
 Xin Wang$^{2}$
@@
 
-\email{haoran@cita.utoronto.ca}
+\email{$\star$ haoran@cita.utoronto.ca}
 
@@
 The considerable amount of memory is usually dominated by an inefficient way of storing more than sufficient phase space information of particles.
-We present a new parallel, information-optimized, particle-mesh based $N$-body code CUBE, in which information- and memory-efficiency is increased by nearly an order of magnitude.
-This is accomplished by storing particle's relative phase space coordinates instead of global values, and in the format of fixed point as light as 1-byte.
-The remaining information is given by complimentary density and velocity fields (negligible in memory space) and proper ordering of particles (no extra memory).
+We present a new parallel, information-optimized, particle-mesh-based $N$-body code CUBE, in which information-efficiency and memory-efficiency are increased by nearly an order of magnitude.
+This is accomplished by storing particle's relative phase space coordinates instead of global values, and in the format of fixed point as light as 1 byte.
+The remaining information is given by complementary density and velocity fields (negligible in memory space) and proper ordering of particles (no extra memory).
 Our numerical experiments show that this information-optimized $N$-body algorithm provides accurate results within the error of the particle-mesh algorithm.
-This significant lowering of memory-to-computation ratio breaks the bottleneck of scaling up and speeding up large cosmological $N$-body simulations on multi-core and heterogeneous computing systems.
+This significant lowering of the memory-to-computation ratio breaks the bottleneck of scaling up and speeding up large cosmological $N$-body simulations on multi-core and heterogeneous computing systems.
 
@@
 The $N$-body simulation, a dynamical simulation of a group of particles, is a powerful tool in physics and astronomy \citep{1988csup.book.....H}.
-It is widely used in cosmology to model the large scale structure (LSS) of the Universe \citep{1985ApJ...292..371D}.
-Current percent and sub-percent level LSS measurements of cosmological parameters, via the matter power spectrum \citep{2005MNRAS.360L..82R,2011ApJ...726....7T}, baryonic acoustic oscillations (BAO) \citep{2005ApJ...633..560E,2009ApJ...700..479T}, weak gravitational lensing \citep{2003ApJ...592..699V,2009A&A...499...31H,2009ApJ...701..945S} etc., require understandings of nonlinear dynamics of the cosmic structure, and rely on high-resolution and high dynamic range $N$-body simulations.
+It is widely used in cosmology to model the large scale structure (LSS) of the universe \citep{1985ApJ...292..371D}.
+Current percent and sub-percent level LSS measurements of cosmological parameters, via the matter power spectrum \citep{2005MNRAS.360L..82R,2011ApJ...726....7T}, baryonic acoustic oscillations (BAO) \citep{2005ApJ...633..560E,2009ApJ...700..479T}, weak gravitational lensing \citep{2003ApJ...592..699V,2009A&A...499...31H,2009ApJ...701..945S} etc., require understandings of the nonlinear dynamics of the cosmic structure, and rely on high-resolution and high dynamic range $N$-body simulations.
 
 When $N$ is large, the brute force pairwise particle-particle (PP) force brings unaffordable $o(N^2)$ computations, so many algorithms were designed to alleviate it.
-Various fast multipole methods \tcb{\citep{1985JCoPh..60..187R,2014ComAC...1....1D,2016ascl.soft09016P}} improve the complexity to $o(N\log N)$ even $o(N)$, among which the most popular one is ``tree'', like {\tt GADGET} \citep{2001NewA....6...79S,2005MNRAS.364.1105S} and its simulation ``Millennium'' \citep{2005Natur.435..629S,2012MNRAS.426.2046A}, {\tt TPM} \citep{1995ApJS...98..355X} and {\tt GOTPM} \citep{2004NewA....9..111D}.
+Various fast-multipole methods \citep{1985JCoPh..60..187R,2014ComAC...1....1D,2016ascl.soft09016P} improve the complexity to $o(N\log N)$ even $o(N)$, among which the most popular one is ``tree'', like {\tt GADGET} \citep{2001NewA....6...79S,2005MNRAS.364.1105S} and its simulation ``Millennium'' \citep{2005Natur.435..629S,2012MNRAS.426.2046A}, {\tt TPM} \citep{1995ApJS...98..355X} and {\tt GOTPM} \citep{2004NewA....9..111D}.
 Other methods include adaptive grid algorithms like {\tt HYDRA} \citep{1995ApJ...452..797C} and {\tt RAMSES} \citep{2010ascl.soft11007T}, as well as mesh-refined codes \citep{1991ApJ...368L..23C} and moving adaptive particle-mesh (PM) codes \citep{1995ApJS..100..269P}.
 The standard PM algorithm \citep{1988csup.book.....H} is most memory and computational efficient if we focus on large cosmological scales.
-The load balancing problem is minimized as the matter distribution is rather homogeneous, and the speed benefits from the fast Fourier transform (FFT) libraries, such as {\tt FFTW3} \citep{Frigo05thedesign}.
-{\tt PMFAST} \citep{2005NewA...10..393M} introduces a 2-level PM algorithm, aiming to put PM codes toward speed, memory compactness and scalability.
+The load-balancing problem is minimized because the matter distribution is rather homogeneous, and the speed benefits from the fast Fourier transform (FFT) libraries, such as {\tt FFTW3} \citep{Frigo05thedesign}.
+{\tt PMFAST} \citep{2005NewA...10..393M} introduces a 2-level PM algorithm, aiming to push PM codes toward speed, memory compactness, and scalability.
 After subsequent developments on {\tt PMFAST}, {\tt CUBEP3M} \citep{2013MNRAS.436..540H} uses cubic spatial decomposition, and adds PP force and many other features.
 
-Beside the methodology, the fast development of parallel supercomputers enable us to simulate a system of more than a trillion ($10^{12}$) $N$-body particles.
-To date the largest $N$-body simulation in application is the ``TianNu'' \citep{2017NatAs...1E.143Y,2017RAA....17...85E} run on the TianHe-2 supercomputer.
+In addition to the new methodology, the fast development of parallel supercomputers enables us to simulate a system of more than a trillion ($10^{12}$) $N$-body particles.
+To date, the largest $N$-body simulation in application is the ``TianNu'' \citep{2017NatAs...1E.143Y,2017RAA....17...85E} run on the TianHe-2 supercomputer.
 With the code {\tt CUBEP3M} adding neutrino modules, it uses $3\times 10^{12}$ particles to simulate the cold dark matter (CDM) and cosmic neutrino evolution through the cosmic age.
 
-Relative to the optimized computation optimizations, $N$-body simulations use considerable amount of memory to store the information of particles.
-Their phase space coordinates $(x,y,z,v_x,v_y,v_z)$ are stored as at least 6 single-precision floating point numbers (total of 24 bytes).
-On the other hand, modern supercomputer systems use multi cores, many integrated cores (MIC) and even densely parallelized GPUs, bringing orders of magnitude higher computing power, whereas these architectures usually have limited memory allocation.
-Thus, these computation-light but memory-heavy applications, compared to matrix multiplication and decomposition calculations, are currently less suitable for fully usage of the computing power of modern supercomputers.
-For example, although native and offload modes of {\tt CUBEP3M} are able to run on the Intel Xeon-PHI MIC architectures, with the requirement of enough memory, TianNu simulation were done on TianHe-2 with only its CPUs -- 73\% of the total memory but only 13\% of the total computing power.
-We investigate how most amount of information of particles can be deducted whereas it still preserves the accuracy of $N$-body simulations, in order that the total memory usage for given $N$ is optimized.
+Relative to the optimized computation optimizations, $N$-body simulations use a considerable amount of memory to store the information of particles.
+Their phase space coordinates $(x,y,z,v_x,v_y,v_z)$ are stored as at least six single-precision floating point numbers (a total of 24 bytes).
+On the other hand, modern supercomputer systems use multi-cores, many integrated cores (MIC) and even densely parallelized GPUs, bringing orders of magnitude higher computing power, whereas these architectures usually have limited memory allocation.
+Thus, these computation-light but memory-heavy applications, compared to matrix multiplication and decomposition calculations, are currently less suitable for full usage of the computing power of modern supercomputers.
+For example, although native and offload modes of {\tt CUBEP3M} are able to run on the Intel Xeon-PHI MIC architectures, with the requirement of enough memory, TianNu simulations were done on TianHe-2 using only its CPUs -- 73\% of the total memory but only 13\% of the total computing power.
+We investigate how the greatest amount of information on particles can be deduced while still preserving the accuracy of $N$-body simulations, with the aim of optimizing the total memory usage for a given $N$.
 
-In the following sections we present a new, information-optimized parallel cosmological $N$-body simulation code {\tt CUBE} \citep{haoran_yu_2018_1198909}, using as low as 6 byte per particle.
+In the following sections we present a new, information-optimized parallel cosmological $N$-body simulation code {\tt CUBE} \citep{2018ascl.soft05018Y}, using as little as 6 bytes per particle (bpp).
 It gives accurate results in cosmological LSS simulations -- the error induced by information optimization is below the error from the PM algorithm.
-In \S\ref{s.method} we show how the memory can be saved by using a ``integer-based storage'' and how the PM $N$-body algorithm is adapted with this storage format.
-In \S\ref{s.results} we quantify the accuracy of this algorithm by groups of simulations by {\tt CUBEP3M} and {\tt CUBE}.
-Discussions and conclusions are in \S\ref{s.discussion}.
+In section \ref{s.method} we show how the memory can be saved by using an ``integer-based storage'' and how the PM $N$-body algorithm is adapted with this storage format.
+In section \ref{s.results} we quantify the accuracy of this algorithm using groups of simulations from {\tt CUBEP3M} and {\tt CUBE}.
+Discussions and conclusions are provided in section \ref{s.discussion}.
 
 \section{Method}\label{s.method}
-The most memory consuming part of a $N$-body simulation is usually the phase space coordinates of $N$-body particles -- 24 byte per particle (here after we use ``bpp'' referring to ``byte per particle''), which contains 6 single precision floating numbers, must be used to store each particles' 3-dimensional position and velocity vectors.
-{\tt CUBEP3M}, an example of a memory-efficient parallel $N$-body code, can use as low as 40 bpp in sacrificing computing speed \citep{2013MNRAS.436..540H}.
-This includes the phase coordinates (24 bpp) for particles in physical domain and buffered region, a linked list (4 bpp), and a global coarse mesh and local fine mesh.
+The most memory-consuming part of an $N$-body simulation is usually the phase space coordinates of $N$-body particles -- 24 bpp -- which contains 6 single-precision floating numbers that must be used to store each particle's 3-dimensional position and velocity vectors.
+{\tt CUBEP3M}, an example of a memory-efficient parallel $N$-body code, can use as little as 40 bpp when sacrificing computing speed \citep{2013MNRAS.436..540H}.
+This includes the phase coordinates (24 bpp) for particles in the physical domain and buffered region, a linked list (4 bpp), and a global coarse mesh and local fine mesh.
 Sometimes 4-byte real numbers are not necessarily adequate in representing the {\it global} coordinates in simulations.
-If the box size is many orders of magnitude larger than interactive distance between particles, especially in the field of re-simulation of dense subregions, double precision (8-byte) coordinates are needed to avoid \tcb{round-off} errors.
+If the box size is many orders of magnitude larger than the interactive distance between particles, especially in the field of resimulation of dense subregions, double-precision (8 byte) coordinates are needed to avoid round-off errors.
 Another solution is to record {\it relative} coordinates for both position and velocity.
-{\tt CUBE} replaces the coordinates and linked list 24+4=28 bpp memory usage with an integer based storage, reduces the basic memory usage from 28 bpp down to 6 bpp, described as following \ref{ss.position} and \ref{ss.velocity}.
+{\tt CUBE} replaces the coordinates and linked list 24+4=28 bpp memory usage with an integer-based storage, thus reducing the basic memory usage from 28 bpp down to 6 bpp, as described in \ref{ss.position} and \ref{ss.velocity}.
 The algorithm is described in \ref{ss.overview}.
@@
 We construct a uniform mesh throughout the space and each particle belongs to its parent cell of the mesh.
-Instead of storing global coordinates of each particle, we store its offset relative to its parent cell which contains the particle.
-This is similar with storing the quantities of nodes/clumps (structures of a tree in a tree code) relative to their parents \citep{1985SJSSC...6...85A}.
-We divide the cell, in each dimension $d$, evenly into $2^8=256$ bins, and use a 1-byte (8 bits) integer $\chi_d \in \{-128,-127,...,127\}$ to indicate which bin it locates in this dimension.
-The global locations of particles are given by cell-ordered format in memory space, and a complimentary number count of particle number in this mesh (density field) will give complete information of particle distribution in the mesh.
-Then the global coordinate in $d$th dimension $x_d$ is given by $x_d=(n_c-1)+(\chi_d+128+1/2)/256$, where $n_c=1,2,...,N_c$ is the index of the coarse grid.
+Instead of storing the global coordinates of each particle, we store its offset relative to its parent cell that contains the particle.
+This is similar to storing the quantities of nodes/clumps (structures of a tree in a tree code) relative to their parents \citep{1985SJSSC...6...85A}.
+We divide the cell, in each dimension $d$, evenly into $2^8=256$ bins, and use a 1 byte (8 bits) integer $\chi_d \in \{-128,-127,...,127\}$ to indicate which bin it locates in this dimension.
+The global locations of particles are given by a cell-ordered format in memory space, and a complementary number count of particle numbers in this mesh (density field) will give complete information on the particle distribution in the mesh.
+Then, the global coordinate in the $d$th dimension $x_d$ is given by $x_d=(n_c-1)+(\chi_d+128+1/2)/256$, where $n_c=1,2,...,N_c$ is the index of the coarse grid.
 The mesh is chosen to be coarse enough such that the density field takes negligible memory.
-This coarse density field can be further compressed into 1-byte integer format, such that a 1-byte integer show the particle number in this coarse cell in range 0 to 255.
-In the densest cells (rarely happened) where there are $\ge 255$ particles, we can just write 255, and write the actual number as a 4-byte integer in another file.
+This coarse density field can be further compressed into a 1-byte integer format, such that a 1-byte integer shows the particle number in this coarse cell in a range from 0 to 255.
+In the densest cells (this rarely happens) where there are $\ge 255$ particles, we can just write 255, and write the actual number as a 4-byte integer in another file.
 
@@
 The force calculation (e.g.
-softening length) should be configured much finer than this resolution, discussed in later sections.
+softening length) should be configured to be much finer than this resolution, as discussed in later sections.
 On the other hand, particle position can also be stored as 2-byte (16 bits) integers to increase the resolution.
-In this case, each coarse cell is divided into $2^{16}=65536$ bins and the position resolution is $L/(65536N_c)$, precise enough compared to using 4-byte global coordinates, see later results.
-We denote this case ``x2'' and denote using 1-byte integers for positions ``x1''.
+In this case, each coarse cell is divided into $2^{16}=65536$ bins and the position resolution is $L/(65536N_c)$, which is precise compared to using 4-byte global coordinates (see later results).
+We denote this case as ``x2'' and denote the case in which we use 1-byte integers for positions as ``x1''.
 
-We collectively write the general position conversion formulae
+We collectively write the general position conversion formulae as
 \begin{equation}\label{eq.chi}
@@
 where $[\ ]$ is the operator to take the integer part.
-$n_\chi\in\{1,2\}$ is the number of bytes used for each integer, $x_d$ and $\chi_d$ are floating and integer versions of the coordinate.
-The velocity counterpart of them are $n_\nu=1,2$, $v_d$ and $\nu_d$.
-The position resolution for $n_\chi$-byte integer, ``x$n_\chi$'', is $2^{-8n_\chi}L/N_c$.
+$n_\chi\in\{1,2\}$ is the number of bytes used for each integer, and $x_d$ and $\chi_d$ are floating and integer versions of the coordinate.
+The velocity counterparts of them are $n_\nu=1,2$, $v_d$ and $\nu_d$.
+The position resolution for an $n_\chi$-byte integer, ``x$n_\chi$'', is $2^{-8n_\chi}L/N_c$.
 
-As a $n_\chi=1$, 1D ($d=1$), 4-coarse-cell ($N_c=4$) example, if $$\chi_1=(-128,127,0,60),$$ particle number density $${\rho_c^{\rm 1D}}=(1,0,2,1),$$ then in unit of coarse cells, the accurate positions of these four particles are $$x_1=(0.001953125, 2.998046875, 2.501953125, 3.736328125).$$
+As a $n_\chi=1$, 1D ($d=1$), 4-coarse-cell ($N_c=4$) example, if $$\chi_1=(-128,127,0,60),$$ and particle number density $${\rho_c^{\rm 1D}}=(1,0,2,1),$$ then in units of coarse cells, the accurate positions of these four particles are $$x_1=(0.001953125, 2.998046875, 2.501953125, 3.736328125).$$
 
@@
  \caption{Variance of the velocity distribution given by Equation \ref{eq.vdisp}.
-The $x$-axis is the scale factor characterizing the cosmic evolution and the $y$-axis shows the $\sigma_v$ in unit of km/s.
-The 5 curves, from top to bottom, shows the $\sigma_v(a)$ with increasing smoothing scale $r$ from 0 to 20 Mpc$/h$.}
+The $x$-axis is the scale factor characterizing the cosmic evolution and the $y$-axis shows the $\sigma_v$ in units of km/s.
+The 5 curves, from top to bottom, show the $\sigma_v(a)$ with increasing smoothing scale $r$ from 0 to 20 Mpc$/h$.}
 \label{fig.vdisp}
@@
 \subsection{Particle velocity storage}\label{ss.velocity}
-Similarly, actual velocity in $d$th dimension $v_d$ is decomposed into an averaged velocity field on the same coarse grid $v_c$ and a residual $\Delta v$ relative to this field:
+Similarly, the actual velocity in the $d$th dimension $v_d$ is decomposed into an averaged velocity field on the same coarse grid $v_c$ and a residual $\Delta v$ relative to this field:
 \begin{equation}\label{eq.deltav}
@@
 
-The reason why we use uneven bins is that, slower particles are more abundant compared to faster ones, and one should better resolve slower particles tracing at least linear evolution.
-On the other hand, there could be extreme scattering particles (in case of particle-particle force), and we can safely ignore or less resolve those nonphysical particles.
-One of the solution is that, if we know the probability distribution function (PDF) $f(\Delta v)$ we divide its cumulative distribution function (CDF) $F(\Delta v)\in(0,1)$ into $2^{8n_\nu}$ bins to determine the boundary of $\Delta v$ bins, and particles should evenly distribute in the corresponding uneven $\Delta v$ bins.
-Practically we find that either $f(v_d)$ or $f(\Delta v)$ is close to Gaussian, so we can use Gaussian CDF, or any convenient analytic functions which are close to Gaussian, to convert velocity between real numbers and integers.
+The reason why we use uneven bins is that slower particles are more abundant compared to faster ones, and one should better resolve slower particles by tracing at least linear evolution.
+On the other hand, there could be extreme scattering particles (in case of PP force), and we can safely ignore or less resolve those nonphysical particles.
+One of the solutions is that, if we know the probability distribution function (PDF) $f(\Delta v)$ we divide its cumulative distribution function (CDF) $F(\Delta v)\in(0,1)$ into $2^{8n_\nu}$ bins to determine the boundary of $\Delta v$ bins, and particles should evenly distribute in the corresponding uneven $\Delta v$ bins.
+Practically we find that either $f(v_d)$ or $f(\Delta v)$ is close to Gaussian, so we can use Gaussian CDF, or any convenient analytic functions that are close to Gaussian, to convert velocity between real numbers and integers.
 
 The essential parameter of the velocity distribution is its variance.
-On non-linear scale, the velocity distribution function is non-Gaussian.
-However, to the first order approximation, we simply assume it as Gaussian and characterized by the variance
+On a nonlinear scale, the velocity distribution function is non-Gaussian.
+However, to the first-order approximation, we simply assume it as Gaussian and characterized it by the variance
 \begin{equation}\label{eq.vdisp}
-	\tcb{\sigma^2_v(a,r) = (a H f D)^2 \int_0^\infty d^3k\frac{P(k)}{k^2}W^2(k,r)},
+	\sigma^2_v(a,r) = (a H f D)^2 \int_0^\infty d^3k\frac{P(k)}{k^2}W^2(k,r),
 \end{equation}
 where $a(z)$ is the scale factor, $H(z)$ is the Hubble parameter, $D$ is the linear growth factor, $f=d \ln D/d\ln a$, and $P(k)$ is the linear power spectrum of density contrast at redshift zero.
-\tcb{$W(k,R)$ is the Fourier transform of real space top-hat window function with smoothing scale $r$.
+$W(k,R)$ is the Fourier transform of the real space top-hat window function with a smoothing scale $r$.
 In Figure \ref{fig.vdisp} we plot $\sigma_v(a,r)$ as a function of $a$ for a few smoothing scale $r$.
-$\Delta v$ in equation (\ref{eq.deltav})} is the velocity dispersion relative to the coarse grid, so we approximate its variance as
+$\Delta v$ in equation (\ref{eq.deltav}) is the velocity dispersion relative to the coarse grid, so we approximate its variance as
 \begin{equation}\label{eq.vdelta}
@@
 \end{equation}
-where $r_c$ is the scale of coarse grid, and $r_p$ is the scale of average particle separation.
-In each dimension of 3D velocity field, we use $\sigma^2_{\Delta}(a)/3$ according to the equipartition therorem.
+where $r_c$ is the scale of the coarse grid, and $r_p$ is the scale of average particle separation.
+In each dimension of the 3D velocity field, we use $\sigma^2_{\Delta}(a)/3$ according to the equipartition theorem.
 On different scales, we measure the statistics of $v_d$, $v_c$ and $\Delta v$ and find good agreement with the above model.
@@
   \includegraphics[width=0.9\linewidth]{f2}
- \caption{Spacial decomposition in {\tt CUBE} in a 2D analogy.
-In this example, there are 2 images per dimension ($M_g=2$), and 2 tiles per image per dimension ($M_t=2$).
-The orange boxes show the overlapped extended($V_e$)=physical($V_p$)+buffer($V_b$) tile regions.
-One physical region is indicated with green.}
+ \caption{Spatial decomposition in {\tt CUBE} in a 2D analogy.
+In this example, there are two images per dimension ($M_g=2$), and two tiles per image per dimension ($M_t=2$).
+The orange boxes show the overlapped extended($V_e$)$\,=\,$physical($V_p$)$\,\cup\,$buffer($V_b$) tile regions.
+One physical region is indicated in green.}
 \label{fig.tile}
@@
 The simulation results are very insensitive if we manually tune the variance of the model $\sigma_\Delta$ within an order of magnitude.
-However, in $n_\nu=1$ case, the method of using uneven bins gets much better results than simply using equal bins between minimum and maximum values $[\min(\Delta v),\max(\Delta v)]$.
+However, in the $n_\nu=1$ case, the method of using uneven bins gets much better results than simply using equal bins between minimum and maximum values $[\min(\Delta v),\max(\Delta v)]$.
 So, one can safely use a standard $\Lambda$CDM (cold dark matter with a cosmological constant as dark energy) for slightly different cosmological models, in equation (\ref{eq.vdisp}).
@@
 Tangent functions are convenient and computing-efficient.
-Compared to error functions used in Gaussian case, they take the same variance at $\nu_d=0$ but resolve high velocities relatively better.
-All possible choice of conversion formulae and $\sigma_\Delta$ are unbiased in the conversion Equations (\ref{eq.nu},\ref{eq.v}), however a proper choices optimize the velocity space sampling and can result in more precise results.
+Compared to the error functions used in the Gaussian case, they take the same variance at $\nu_d=0$ but resolve high velocities relatively better.
+All possible choices of conversion formulae and $\sigma_\Delta$ are unbiased in the conversion Equations (\ref{eq.nu},\ref{eq.v}), however, a proper choices optimize the velocity space sampling and can result in more precise results.
 
-Initially, particles are generated by initial condition generator, at a higher redshift.
-The coarse grid \tcb{velocity} field $v_c$ is also generated at this step by averaging all particles in the coarse cell.
-A global $\sigma_\Delta$ is calculated by equation (\ref{eq.vdelta}), where linear approximation is hold.
-Then velocities are stored by equation (\ref{eq.nu}).
-During the simulation, $v_c$ is updated every time step, and a nonlinear $\sigma_\Delta$ is measured directly from the simulation, and can be simply used in the next time step, after scaled by the ratio of growth factors between two adjacent time steps.
-More details see section \ref{sss.algorithm}.
+Initially, particles are generated by an initial condition generator, at a higher redshift.
+The coarse grid velocity field $v_c$ is also generated at this step by averaging all particles in the coarse cell.
+A global $\sigma_\Delta$ is calculated by equation (\ref{eq.vdelta}), where linear approximation holds.
+Then, velocities are stored by equation (\ref{eq.nu}).
+During the simulation, $v_c$ is updated every time-step, and a nonlinear $\sigma_\Delta$ is measured directly from the simulation, and can be simply used in the next time step, after scaled by the ratio of growth factors between two adjacent time-steps.
+For more details see section \ref{sss.algorithm}.
 
@@
 {\tt CUBE} uses a 2-level PM force calculation.
-In order to apply the integer based format to the $N$-body simulation, substantial structural changes need to be done.
+In order to apply the integer-based format to the $N$-body simulation, substantial structural changes need to be done.
 {\tt CUBE} is written in Coarray Fortran, where Coarray features replace MPI (Message Passing Interface) communications between computation nodes/images\footnote{Images are the concept of computing nodes or MPI tasks in Coarray Fortran.
@@
 
-\subsubsection{Spacial decomposition}\label{sss.spatial_decomposition}
+\subsubsection{Spatial decomposition}\label{sss.spatial_decomposition}
 {\tt CUBE} decomposes the global simulation volume into $M_g^3$ cubic sub-volumes with $N_c$ coarse grids or $N_f=RN_c$ fine grids per side.
-The fine-mesh is usually $R=4$ times finer than the coarse mesh.
+The fine mesh is usually $R=4$ times finer than the coarse mesh.
 Each of these sub-volumes is assigned to a coarray {\it image}.
-Inside of an image, the sub-volume is further decomposed into $M_t^3$ cubic {\it tiles} (defined as $V_p$, or {\it physical} region) with $N_c/(M_g M_t)$ coarse grids per side.
+Inside of an image, the sub-volume is further decomposed into $M_t^3$ cubic {\it tiles} (defined as $V_p$, or a {\it physical} region) with $N_c/(M_g M_t)$ coarse grids per side.
 Each $V_p$ is surrounded by a {\it buffer} region $V_b$ which is $N_b$ coarse cells thick.
-We define \tcb{the} extended tile region \tcb{as} \tcb{$V_e\equiv V_p\cup V_b$}.
-$V_e$ is designed for two reasons:
+We define the extended tile region as $V_e\equiv V_p\cup V_b$.
+$V_e$ is designed for two purposes:
 
-(1) if the short fine-mesh force ${\bs F}_f$ has a cut-off, \tcb{$N_b$ ${\bs F}_f(r>N_b)=0$}, and is computed \tcr{\sout{ by $\rho_c$ whose domain $D(\rho_f)=V_e$}} \tcb{on $V_e$}, then ${\bs F}_f$ in $V_p$ is guaranteed to be correct.
+(1) If the short fine mesh force ${\bs F}_f$ has a cut-off, $N_b$ ${\bs F}_f(r>N_b)=0$, and is computed on $V_e$, then ${\bs F}_f$ in $V_p$ is guaranteed to be correct.
 
-\tcb{(2) $V_e$ is able to collect all particles that are able to travel to $V_p$.} \tcr{\sout{i.e., ... }[deleted previous two equations (8,9)]}
+(2) $V_e$ is able to collect all particles that are able to travel to $V_p$.
 %\begin{eqnarray}\label{eq.cond1}
@@
 %\end{equation}
-Figure \ref{fig.tile} shows the spacial decomposition in a 2-dimensional analogy, with $M_g=2$ and $M_t=2$.
+Figure \ref{fig.tile} shows the spatial decomposition in a 2-dimensional analogy, with $M_g=2$ and $M_t=2$.
 
@@
 \begin{eqnarray*}
-	&&{\tt integer(1)}\ \ \, \rho_c(N_e,N_e,N_e,M_t,M_t,M_t)[M_g,M_g,*]\\
-	&&{\tt real(4)}\ \ \ \ \ \ \ \, v_c(N_e,N_e,N_e,M_t,M_t,M_t)[M_g,M_g,*]\\
-	&&{\tt integer(}n_\chi{\tt )}\, \chi_d(3,P_{\rm max})[M_g,M_g,*]\\
-	&&{\tt integer(}n_\nu{\tt )}\ \nu_d(3,P_{\rm max})[M_g,M_g,*]
+	&&\tcv{{\tt integer(1)}}\ \ \, \rho_c(N_e,N_e,N_e,M_t,M_t,M_t)[M_g,M_g,*]\\
+	&&\tcv{{\tt real(4)}}\ \ \ \ \ \ \ \, v_c(N_e,N_e,N_e,M_t,M_t,M_t)[M_g,M_g,*]\\
+	&&\tcv{{\tt integer(}n_\chi{\tt )}}\, \chi_d(3,P_{\rm max})[M_g,M_g,*]\\
+	&&\tcv{{\tt integer(}n_\nu{\tt )}}\ \nu_d(3,P_{\rm max})[M_g,M_g,*]
 \end{eqnarray*}
-where $N_e=N_t+2N_b$ covers the buffer region on both sides, $M_t$ is the tile dimemsions, and $M_g$ is the image {\it codimensions}\footnote{Coarray Fortran concept.
-Codimensions can do communications between images.}.
-We denote actual number of particles in a given image $P_{\rm local}$, and $P_{\rm max}>P_{\rm local}$ is a value (discussed in Section \ref{ss.memory}) large enough to store particles in $V_e$.
+where $N_e=N_t+2N_b$ covers the buffer region on both sides, $M_t$ is the tile dimensions, and $M_g$ is the image {\it co-dimensions}.\footnote{Coarray Fortran concept.
+Co-dimensions can enable communications between images.} We denote the actual number of particles in a given image as $P_{\rm local}$, and $P_{\rm max}>P_{\rm local}$ is a value (discussed in Section \ref{ss.memory}) large enough to store particles in $V_e$.
 $\chi_d$ and $\nu_d$
-are required to be sorted according to the same memory layout of $\rho_c$, such that $n_c$ and $x_d$ can be obtained by equation (\ref{eq.x}).
-$\{\rho_c,v_c,\chi_d,\nu_d\}$ provides a complete information of positions and velocities of particles, and we call it a checkpoint.
+must be sorted according to the same memory layout as $\rho_c$, such that $n_c$ and $x_d$ can be obtained from equation (\ref{eq.x}).
+$\{\rho_c,v_c,\chi_d,\nu_d\}$ provides a complete information on the positions and velocities of particles, and we call it a checkpoint.
 
 An additional particle-ID (PID) array, $I_P$, can also be declared to differentiate particle types (e.g.
-CDM and neutrino particles) or differentiate every particle, by using $n_I$-integer per particle.
-If PID is turned on, the array $I_P$ is also included in the checkpoints, and the ordering of $I_P$ is same as $\chi_d$ and $\nu_d$.
+CDM and neutrino particles) or to differentiate every particle, by using an $n_I$-byte integer per particle.
+If PID is turned on, the array $I_P$ is also included in the checkpoints, and the ordering of $I_P$ is the same as that for $\chi_d$ and $\nu_d$.
 
 \begin{figure}[t]
-{\tt program Initial\_Condition\_Generator\_for\_CUBE\\
+{\tt \tcv{program} Initial\_Condition\_Generator\_for\_CUBE\\
 \indent \ \ calculate $\Phi$ in Fourier space\\
-\indent \ \ do (each tile)\\
-\indent \ \ \ \ do (each particle at ${\bs q}\in V_e$)\\
+\indent \ \ \tcv{do} (each tile)\\
+\indent \ \ \ \ \tcv{do} (each particle at ${\bs q}\in V_e$)\\
 \indent \ \ \ \ \ \ ${\bs \Psi}({\bs q})=\nabla\Phi({\bs q})$\\
@@
 \indent \ \ \ \ \ \ $\{{\bs x},{\bs v}\}\rightarrow \{\rho_c,v_c\}$\\
-\indent \ \ \ \     enddo\\
-\indent \ \ \ \ do (each particle at ${\bs q}\in V_e$)\\
+\indent \ \ \ \     \tcv{enddo}\\
+\indent \ \ \ \ \tcv{do} (each particle at ${\bs q}\in V_e$)\\
 \indent \ \ \ \ \ \ ${\bs \Psi}({\bs q})=\nabla\Phi({\bs q})$\\
@@
 \indent \ \ \ \ \ \ ${\bs x}\rightarrow\chi_d(:,i);\,\{{\bs v},v_c\}\rightarrow\nu_d(:,i)$\\
-\indent \ \ \ \ \ \ if (PID\_flag) create $I_P(i)$\\
-\indent \ \ \ \ enddo\\
-\indent \ \ \ \ do (each coarse grid $\in V_e$)\\
+\indent \ \ \ \ \ \ \tcv{if} (PID\_flag) create $I_P(i)$\\
+\indent \ \ \ \ \tcv{enddo}\\
+\indent \ \ \ \ \tcv{do} (each coarse grid $\in V_e$)\\
 \indent \ \ \ \ \ \ delete particles $\in V_b$\\
-\indent \ \ \ \ enddo\\
+\indent \ \ \ \ \tcv{enddo}\\
 \indent \ \ \ \ write $\{\rho_c,v_c,\chi_d,\nu_d\}$ to disk\\
 \indent \ \ \ \ sum up $P_{\rm local}$\\
-\indent \ \ enddo\\
-\indent \ \ sync all\\
+\indent \ \ \tcv{enddo}\\
+\indent \ \ \tcv{sync all}\\
 \indent \ \ sum up $P_{\rm global}$\\
-end\\}
+\tcv{end}\\}
 \caption{Pseudocode for the initial condition generator.}
@@
 \subsubsection{Initial conditions}\label{sss.ic}
-The cosmological initial condition generator is compiled and run separately from the main $N$-body code. Here we briefly describe it for completeness.
+The cosmological initial condition generator is compiled and run separately from the main $N$-body code. Here, we briefly describe it for completeness.
 
@@
 At an initial redshift $z_i$, we generate a linear density fluctuation $\delta_L({\bs q})$ on Lagrangian grid ${\bs q}$ by multiplying a Gaussian random field with the transfer function $T(k)$ (given by the assumed cosmological model) in Fourier space.
-Then we solve for the potential ${\Phi({\bs q}, z_i)}$ of a curl-less displacement field ${\bs\Psi({\bs q})}$ by Poisson equation $-\nabla^2\Phi=\delta_L$ in Fourier space.
+Then, we solve for the potential ${\Phi({\bs q}, z_i)}$ of a curlless displacement field ${\bs\Psi({\bs q})}$ by Poisson equation $-\nabla^2\Phi=\delta_L$ in Fourier space.
 
-The second step to generate particles and displace them by Zel'dovich approximation (ZA) \citep{1970A&A.....5...84Z}, where the displacement field is obtained by differentiating $\Phi$, ${\bs \Psi}({\bs q})=\nabla\Phi({\bs q})$ in real space.
+The second step to generate particles and displace them by Zel'dovich approximation (ZA) \citep{1970A&A.....5...84Z}, where the displacement field is obtained by differentiating $\Phi$, is ${\bs \Psi}({\bs q})=\nabla\Phi({\bs q})$ in real space.
 This step is done on $V_e$ of each tile.
-\tcr{\sout{Similar to condition (9), we have ...} [deleted previous equation (10).]}
 %\begin{equation}\label{eq.cond3}
@@
 The first iteration calculates particles' ${\bs x}$ and ${\bs v}$ by ZA and obtains $\rho_c$ and $v_c$ on the coarse grid.
-The second iteration, ${\bs x}$ and ${\bs v}$ is calculated again and are converted to $\chi_d$ and $\nu_d$ by Equations (\ref{eq.chi},\ref{eq.nu}) and place them in a certain order according to $\rho_c$.
-Lastly, we delete particles in $V_b$, and resort the ones in $V_p$ and write $\{\rho_c,v_c,\chi_d,\nu_d,I_P({\rm optional})\}$ of this tile to disk.
-The above is similar to {\tt update\_xp} of Section \ref{sss.algorithm}.
+The second iteration's ${\bs x}$ and ${\bs v}$ are calculated again and are converted to $\chi_d$ and $\nu_d$ by Equations (\ref{eq.chi},\ref{eq.nu}) and placed in a certain order according to $\rho_c$.
+Lastly, we delete particles in $V_b$, and re-sort the ones in $V_p$ and write $\{\rho_c,v_c,\chi_d,\nu_d,I_P({\rm optional})\}$ of this tile to disk.
+The above is similar to {\tt \tcx{update\_xp}} of Section \ref{sss.algorithm}.
 If PIDs are needed, they are also generated here.
@@
 \begin{figure}[]
-\begin{verbatim}
-
-    program CUBE
-      call initialize
-      call read_particles
-      call buffer_density
-      call buffer_xp
-      call buffer_vp
-      do
-        call timestep
-        call update_xp
-        call buffer_density
-        call buffer_xp
-        call update_vp
-        call buffer_vp
-        if(checkpoint_step) then
-           call update_xp
-           call checkpoint
-           if (final_step) exit
-           call buffer_density
-           call buffer_xp
-           call buffer_vp
-        endif
-      enddo
-      call finalize
-    end
-\end{verbatim}
+{\tt \tcv{program} CUBE\\
+\indent \ \ \tcv{call} \tcx{initialize}\\
+\indent \ \ \tcv{call} \tcx{read\_particles}\\
+\indent \ \ \tcv{call} \tcx{buffer\_density}\\
+\indent \ \ \tcv{call} \tcx{buffer\_xp}\\
+\indent \ \ \tcv{call} \tcx{buffer\_vp}\\
+\indent \ \ \tcv{do}\\
+\indent \ \ \ \ \tcv{call} \tcx{timestep}\\
+\indent \ \ \ \ \tcv{call} \tcx{update\_xp}\\
+\indent \ \ \ \ \tcv{call} \tcx{buffer\_density}\\
+\indent \ \ \ \ \tcv{call} \tcx{buffer\_xp}\\
+\indent \ \ \ \ \tcv{call} \tcx{update\_vp}\\
+\indent \ \ \ \ \tcv{call} \tcx{buffer\_vp}\\
+\indent \ \ \ \ \tcv{if}(checkpoint\_step) \tcv{then}\\
+\indent \ \ \ \ \ \ \tcv{call} \tcx{update\_xp}\\
+\indent \ \ \ \ \ \ \tcv{call} \tcx{checkpoint}\\
+\indent \ \ \ \ \ \ \tcv{if} (final\_step) \tcv{exit}\\
+\indent \ \ \ \ \ \ \tcv{call} \tcx{buffer\_density}\\
+\indent \ \ \ \ \ \ \tcv{call} \tcx{buffer\_xp}\\
+\indent \ \ \ \ \ \ \tcv{call} \tcx{buffer\_vp}\\
+\indent \ \ \ \ \tcv{endif}\\
+\indent \ \ \tcv{enddo}\\
+\indent \ \ \tcv{call} \tcx{finalize}\\
+\tcv{end}\\}
 \caption{Overall structure of {\tt CUBE}. Sections of the code are grouped into Fortran subroutines, which are described in paragraphs of Section \ref{sss.algorithm}.}
@@
 
-{\tt initialize} -- creates fine-mesh and coarse mesh FFT plans, and read in configuration files telling the program at which redshifts we need to do checkpoints, halofinds, or stop the simulation.
+{\tt \tcx{initialize}} creates fine mesh and coarse mesh FFT plans, and reads in configuration files telling the program at which redshifts we need to do checkpoints, halofinds, or stop the simulation.
 Force kernels $K_c$, $K_f$ are also computed or loaded.
 
-{\tt read\_particles} -- from disk, reads in checkpoint $\{\rho_c,v_c,\chi_d,\nu_d,I_P({\rm optional})\}$ for each image.
-Because they exist only in $V_p$ of every tile, they are {\it disjoint}; and they provide {\it complete} information of the whole simulation volume -- we call it ``{\it disjoint state}''.
+{\tt \tcx{read\_particles}}, from the disk, reads in a checkpoint $\{\rho_c,v_c,\chi_d,\nu_d,I_P({\rm optional})\}$ for each image.
+Because they exist only in the $V_p$ of every tile, they are {\it disjoint}; and they provide {\it complete} information on the whole simulation volume -- we call it ``{\it disjoint state}''.
 In this state, $\rho_c$, $v_c$'s values in buffer regions, and $\chi_d(:,P_{\rm local}+1:)$ and $\nu_d(:,P_{\rm local}+1:)$ are 0's.
@@
 
-{\tt buffer\_density}, {\tt buffer\_x} and {\tt buffer\_v} -- converts ``disjoint'' state to ``{\it buffered state}''.
-In {\tt buffer\_density}, $V_b$ regions of $\rho_c$ is synchronized between tiles and images.
-By {\tt buffer\_x}, $\chi_d$ is updated to contain common, buffered particles, and they are sorted according to the buffered $\rho_c$.
-{\tt buffer\_v} deals with $\nu_d$ in a same manner.
+{\tt \tcx{buffer\_density}}, {\tt \tcx{buffer\_x}} and {\tt \tcx{buffer\_v}} convert the ``disjoint'' state to the ``{\it buffered state}''.
+In {\tt \tcx{buffer\_density}}, $V_b$ regions of $\rho_c$ are synchronized between tiles and images.
+By {\tt \tcx{buffer\_x}}, $\chi_d$ is updated to contain common, buffered particles, and they are sorted according to the buffered $\rho_c$.
+{\tt \tcx{buffer\_v}} deals with $\nu_d$ in a similar manner.
 
-{\tt timestep} --
-\tcb{The second order Runge-Kutta method} is used in the time integration.
-i.e., for $n$ time steps, we update positions ($D$=drift) and velocities ($K$=kick) at interlaced half time-steps by operator splitting; the operation $(DKKD)^n$ is 2nd order accurate.
-The actual simulation applies varied time steps by {\tt timestep}, where a time increment ${\rm d}t$ is constrained by particles' maximum velocities, accelerations, cosmic expansion and any other desired conditions.
+{\tt \tcx{timestep}} is
+a second order Runge-Kutta method is used in the time integration, i.e., for $n$ time-steps, we update positions ($D$=drift) and velocities ($K$=kick) at interlaced half time-steps by operator splitting; the operation $({\rm DKKD})^n$ is second-order accurate.
+The actual simulation applies varied time-steps by {\tt \tcx{timestep}}, where a time increment ${\rm d}t$ is constrained by particles' maximum velocities, accelerations, cosmic expansion, and any other desired conditions.
 
 \begin{figure}[t]
-{\tt subroutine update\_xp\\
-\indent \ \ do (each physical tile)\\
-\indent \ \ \ \ do (each particle)\\
+{\tt \tcv{subroutine} \tcx{update\_xp}\\
+\indent \ \ \tcv{do} (each physical tile)\\
+\indent \ \ \ \ \tcv{do} (each particle)\\
 \indent \ \ \ \ \ \ $\{\rho_c,\chi_d\}\rightarrow {\bs x};\,\{v_c,\nu_d\}\rightarrow {\bs v}$\\
@@
 \indent \ \ \ \ \ \ update $\rho_c^*,\,v_c^*$ according to ${\bs x}$\\
-\indent \ \ \ \ enddo\\
-\indent \ \ \ \ do (each particle)\\
+\indent \ \ \ \ \tcv{enddo}\\
+\indent \ \ \ \ \tcv{do} (each particle)\\
 \indent \ \ \ \ \ \ $\{\rho_c,\chi_d\}\rightarrow {\bs x};\,\{v_c,\nu_d\}\rightarrow {\bs v}$\\
@@
 \indent \ \ \ \ \ \ ${\bs x}\rightarrow\chi_d^*(:,i);\,\{{\bs v},v_c\}\rightarrow\nu_d^*(:,i)$\\
-\indent \ \ \ \ enddo\\
-\indent \ \ \ \ do (each coarse grid)\\
+\indent \ \ \ \ \tcv{enddo}\\
+\indent \ \ \ \ \tcv{do} (each coarse grid)\\
 \indent \ \ \ \ \ \ discard buffer information\\
-\indent \ \ \ \ enddo\\
+\indent \ \ \ \ \tcv{enddo}\\
 \indent \ \ \ \ replace $\{\rho_c,v_c,\chi_d,\nu_d\}$ with $\{\rho_c^*,v_c^*,\chi_d^*,\nu_d^*\}$\\
-\indent \ \ enddo\\
-\indent \ \ sync all\\
+\indent \ \ \tcv{enddo}\\
+\indent \ \ \tcv{sync all}\\
 \indent \ \ update velocity dispersion $\sigma^2_{\Delta}$\\
 \indent \ \ sum up $P_{\rm local},\,P_{\rm global}$\\
-end\\}
-\caption{Pseudocode for subroutine {\tt update\_xp}.}
+\tcv{end}\\}
+\caption{Pseudocode for subroutine {\tt \tcx{update\_xp}}.}
 \label{fig.update_xp}
@@
 
-{\tt update\_xp} --
-According to ${\rm d}t$, {\tt update\_xp} updates particle positions (drift $D$) in a ``{\it gather}'' algorithm tile by tile.
+{\tt \tcx{update\_xp}}
+is used, according to ${\rm d}t$, to update particle positions (drift $D$) in a ``{\it gather}'' algorithm tile by tile.
 For each particle, $\chi_d$ and $\nu_d$ are converted to $x_d$ and $v_d$ by Equations(\ref{eq.x},\ref{eq.v}).
-\tcr{\sout{Due to Condition (9), we are able to collect all particles to travel to $V_p$.}}
 
-In order to keep particles ordered, for each tile, we first do $x_d=x_d+v_d\,{\rm d}t$ on all particles to obtain an updated density and velocity field on the tile,  $\rho_c^*$ and $v_c^*$.
-Then, this calculation in done on the same tile again\footnote{This repetition scales as $o(N)$ and is computational inexpensive.} to generate a new, local particle list $\chi_d^*$ and $\nu_d^*$ by Equations(\ref{eq.chi},\ref{eq.nu}).
+In order to keep particles ordered, for each tile, we first perform $x_d=x_d+v_d\,{\rm d}t$ on all particles to obtain an updated density and velocity field on the tile,  $\rho_c^*$ and $v_c^*$.
+Then, this calculation is done on the same tile again to generate a new, local particle list $\chi_d^*$ and $\nu_d^*$ by Equations(\ref{eq.chi},\ref{eq.nu}).
 Here, the ordering of  $\chi_d^*$ and $\nu_d^*$ relies on  $\rho_c^*$.
@@
 Then the disjoint state of $\{\rho_c^*,v_c^*,\chi_d^*,\nu_d^*\}$ replaces the old $\{\rho_c,v_c,\chi_d,\nu_d\}$.
-Finally $P_{\rm local}$ and $P_{\rm global}=\sum P_{\rm local}$ are updated.
+Finally, $P_{\rm local}$ and $P_{\rm global}=\sum P_{\rm local}$ are updated.
 These steps are summarized in Figure \ref{fig.update_xp}.
 
-{\tt update\_vp} --
-the PM or particle-particle particle-mesh (P$^3$M) algorithm is applied in this subroutine to update particles's velocities (kick $K$).
-We first call {\tt buffer\_density} and {\tt buffer\_xp} in order that particle positions are in buffered state.
-Then according to the particle distributions $\rho_f$ on $V_e$ we calculate the fine-mesh force ${\bs F}_f$ and update the particle velocities in the $V_p$.
-\tcr{\sout{Condition (8) guarantees the correctness of this step.}}  An optional particle-particle (PP) force ${\bs F}_{\rm pp}$ can be called to increase the force resolution.
+In {\tt \tcx{update\_vp}}
+the PM or PP particle-mesh (P$^3$M) algorithm is applied in this subroutine to update particles' velocities (kick $K$).
+We first call {\tt \tcx{buffer\_density}} and {\tt \tcx{buffer\_xp}} to place the particle positions in the buffered state.
+Then, according to the particle distributions $\rho_f$ on $V_e$, we calculate the fine mesh force ${\bs F}_f$ and update the particle velocities in the $V_p$.
+An optional PP force ${\bs F}_{\rm pp}$ (subroutine {\tt \tcx{PP\_force}}) can be called to increase the force resolution.
 
 \begin{figure}[t]
-{\tt subroutine update\_vp\\
-\indent \ \ do (each extended tile)\\
+{\tt \tcv{subroutine} \tcx{update\_vp}\\
+\indent \ \ \tcv{do} (each extended tile)\\
+\indent \ \ \ \ \tcv{call} \tcx{PP\_force}\\
 \indent \ \ \ \ calculate $\rho_f$ and solve ${\bs F}_f$ in Fourier space\\
-\indent \ \ \ \ do (each particle in physical region)\\
+\indent \ \ \ \ \tcv{do} (each particle in physical region)\\
 \indent \ \ \ \ \ \ $\nu_d\rightarrow{\bs v},\,{\bs v}={\bs v}+{\bs F}_f{\rm d}t,\, {\bs v}\rightarrow\nu_d$\\
-\indent \ \ \ \ enddo\\
+\indent \ \ \ \ \tcv{enddo}\\
 \indent \ \ \ \ update $\max(|{\dot{\bs v}}|)$\\
-\indent \ \ \ \ (PP force)\\
-\indent \ \ enddo\\
+\indent \ \ \tcv{enddo}\\
 \indent \ \ calculate $\rho_c$ and solve ${\bs F}_c$ in Fourier space\\
-\indent \ \ do (each particle)\\
+\indent \ \ \tcv{do} (each particle)\\
 \indent \ \ \ \ $\nu_d\rightarrow{\bs v},\,{\bs v}={\bs v}+{\bs F}_c{\rm d}t,\, {\bs v}\rightarrow\nu_d$\\
-\indent \ \ enddo\\
+\indent \ \ \tcv{enddo}\\
 \indent \ \ update $\max(v_d),\max(|{\dot{\bs v}}|)$\\
-end\\}
-\caption{Pseudocode for subroutine {\tt update\_vp}.}
+\tcv{end}\\}
+\caption{Pseudocode for subroutine {\tt \tcx{update\_vp}}.}
 \label{fig.update_vp}
@@
 
-The compensating coarse grid force ${\bs F}_c$ is globally computed by using a coarser- (usually by factor of $R=4$) mesh by dimensional splitting -- the cubic distributed coarse density field $\rho_c$ is transposed (inter-image) and Fourier transformed (inner-image) in consecutive three dimensions.
+The compensating coarse grid force ${\bs F}_c$ is globally computed by using a coarser- (usually by a factor of $R=4$) mesh by dimensional splitting -- the cubic distributed coarse density field $\rho_c$ is transposed (inter-image) and Fourier transformed (inner-image) in three consecutive dimensions.
 After the multiplication of memory-distributed force kernel $K_c$, the inverse transform takes place to get the cubic distributed coarse force field ${\bs F}_c$, upon which velocities are updated again.
 
-For each type of velocity updates, the collective operations are Equation(\ref{eq.v}), ${\bs v}={\bs v}+{\bs F}_{\rm total}$ and Equation(\ref{eq.nu}).
+For each type of velocity update, the collective operations are Equation(\ref{eq.v}), ${\bs v}={\bs v}+{\bs F}_{\rm total}$, and Equation(\ref{eq.nu}).
 We also update $\sigma^2_{\Delta}$ according to the new $v_d-v_c$.
@@
 
-\tcb{After the updating of $\nu_d$ in $V_p$, we} simply call {\tt buffer\_v} again to bring $\nu_d$ into buffered state, such that the {\tt update\_x} in the next iteration will be done correctly.
+After the updating of $\nu_d$ in $V_p$, we simply call {\tt \tcx{buffer\_v}} again to bring $\nu_d$ into the buffered state, such that the {\tt \tcx{update\_x}} in the next iteration will be done correctly.
 
-{\tt checkpoint} --
-If a desired redshift is reached, we execute the last drift step in the $(DKKD)^n$ operation by {\tt update\_xp}, and call {\tt checkpoint} to save the disjoint state of $\{\rho_c,v_c,\chi_d,\nu_d\}$ on disk.
-Related operations, like run-time halo finder, projections are also done at this point.
+For {\tt \tcx{checkpoint}},
+if a desired redshift is reached, we execute the last drift step in the $({\rm DKKD})^n$ operation by {\tt \tcx{update\_xp}}, and call {\tt \tcx{checkpoint}} to save the disjoint state of $\{\rho_c,v_c,\chi_d,\nu_d\}$ on the disk.
+Other operations like run-time halo-finder or density projections are also done at this point.
 
-{\tt finalize} --
-Finally, in {\tt finalize} subroutine we destroy all the FFT plans and finish up any timing or statistics taken in the simulation.
+Finally, in the {\tt \tcx{finalize}} subroutine
+we destroy all the FFT plans and finish up any timing or statistics taken in the simulation.
 
 \subsection{Memory layout}\label{ss.memory}
-Here we list the memory-consuming arrays and how they scale with different configurations of the simulation.
-We classify them into (1) arrays of particles, (2) coarse mesh arrays and (3) fine mesh arrays.
+Here, we list the memory-consuming arrays and how they scale with different configurations of the simulation.
+We classify them into (1) arrays of particles, (2) coarse mesh arrays, and (3) fine mesh arrays.
 
 \subsubsection{Arrays of particles}
-These are the majority memory usage.
-contains checkpoint arrays $\chi_d$, $\nu_d$, $I_P$, and temporary arrays $\chi_d^*$, $\nu_d^*$, $I_P^*$.
+These arrays comprise the majority of the memory usage, and contain checkpoint arrays $\chi_d$, $\nu_d$, $I_P$, and temporary arrays $\chi_d^*$, $\nu_d^*$, $I_P^*$.
 The former uses memory $\mathcal{M}=(3n_\chi+3n_\nu+n_I)P_{\rm max}\,{\rm byte}$, where
@@
 and $\left\langle P_{\rm local} \right\rangle$ is the average number of particles per image.
-The second term, proportional to $V_e/V_p$,  let us store additional particles in $V_b$, and the third term $1+\epsilon_{\rm image}$ takes into account the inhomogeneity of $P_{\rm local}$ on different images.
+The second term, proportional to $V_e/V_p$,  lets us store additional particles in $V_b$, and the third term $1+\epsilon_{\rm image}$ takes into account the inhomogeneity of $P_{\rm local}$ on different images.
 When each image models smaller physical scales, $\epsilon_{\rm image}$ should be set larger.
@@
 \end{equation}
-where $\epsilon_{\rm tile}$ controls the inhomogeneity on scale of tiles.
-Larger $M_t$ causes more inhomogeneity on smaller tiles, and $\epsilon_{\rm tile}$ can be much larger than $\epsilon_{\rm image}$, however the term $M_t^{-3}$ decreases much faster.
+where $\epsilon_{\rm tile}$ controls the inhomogeneity on scales of tiles.
+Larger $M_t$ causes more inhomogeneity on smaller tiles, and $\epsilon_{\rm tile}$ can be much larger than $\epsilon_{\rm image}$; however, the term $M_t^{-3}$ decreases much faster.
 Practically, the majority memory is occupied by $\chi_d$, $\nu_d$, $I_P$.
 
-Summarizing above we obtain the memory usage per particle (byte per particle) ${\mathcal M}_P \equiv {\mathcal M} / \left\langle P_{\rm local} \right\rangle/{\rm byte}$, given $n_\chi=n_\nu=1$ and $n_I=0$, is
+Summarizing the above, we find that the memory usage per particle (bpp) ${\mathcal M}_P \equiv {\mathcal M} / \left\langle P_{\rm local} \right\rangle/{\rm byte}$, given $n_\chi=n_\nu=1$ and $n_I=0$, is
 \begin{eqnarray}\label{eq.Mp_particle}
@@
 \subsubsection{Coarse mesh arrays}
-On coarse mesh, $\rho_c$ (4-byte integers), $v_c$ and force kernel $K_c$ should always be kept.
+On coarse mesh, $\rho_c$ (4-byte integers), $v_c$, and force kernel $K_c$ should always be kept.
 They are usually configured to be $R=4$ times coarser than fine grids and particle number density.
-They have use memory $(1+3)\times 4\times(N_e M_t)^3+3\times 4\times(N_c/M_g)^3/2$ bytes per image, or
+They have use memories of $(1+3)\times 4\times(N_e M_t)^3+3\times 4\times(N_c/M_g)^3/2$ bytes per image, or
 \begin{equation}\label{eq.Mp_coarse}
@@
 where $P_c$ is the average number of particles per coarse cell.
-Other coarse grid based arrays are $\rho_c^*$, $F_c$ pencil-FFT arrays.
+Other coarse-grid-based arrays are $\rho_c^*$, $F_c$, and pencil-FFT arrays.
 $\rho_c^*$ exists only on tiles; $F_c$ and pencil-FFT arrays can be equivalenced with other temporary arrays.
-Thus the major memory usage from coarse FFT arrays are from Equation (\ref{eq.Mp_coarse}).
+Thus, the majority of the memory usage from coarse FFT arrays comes from Equation (\ref{eq.Mp_coarse}).
 
@@
 where $P_f$ is the average number of particles per fine cell.
-The fine mesh density arrays, force field arrays are temporary.
-Since $\chi_d^*$, $\nu_d^*$, coarse force arrays, pencil-FFT arrays are also temporary and are not used in any calculation simultaneously, they can be overlapped in memory by using equivalent statements.
+For the fine mesh density arrays, force field arrays are temporary.
+Since $\chi_d^*$, $\nu_d^*$, coarse force arrays, and pencil-FFT arrays are also temporary and are not used in any calculation simultaneously, they can be overlapped in memory by using equivalent statements.
 
@@
   \includegraphics[width=0.75\linewidth]{f7}
- \caption{Offset in the particle distribution induced by integer-1 based algorithm (x1v1).
-In S512's subregion of volume $15.625^2\times 3.90625\ ({\rm Mpc}/h)^3$, particles from {\tt CUBEP3M} (bigger grey dots in the background) and {\tt CUBE}-x1v1 (smaller red dots) are projected onto the plane of $(15.625\ {\rm Mpc}/h)^2$.
-The comparing rules show 1 ${\rm Mpc}/h$, fine grid and coarse grid respectively, and position resolution of x1v1 is 1/64 of a fine grid.}
+ \caption{Offset in the particle distribution induced by an integer-1 based algorithm (x1v1).
+In S512's subregion of volume $15.625^2\times 3.90625\ ({\rm Mpc}/h)^3$, particles from {\tt CUBEP3M} (the larger gray dots in the background) and {\tt CUBE}-x1v1 (the smaller red dots) are projected onto the plane of $(15.625\ {\rm Mpc}/h)^2$.
+The comparing rules show 1 ${\rm Mpc}/h$, for fine and coarse grids respectively, and position resolution of x1v1 is 1/64 of a fine grid.}
 \label{fig.particles}
@@
 To illustrate the improvement of memory usage, we refer to the TianNu simulation \citep{2017NatAs...1E.143Y} run on the Tianhe-2 supercomputer, which used a traditional $N$-body code {\tt CUBEP3M}.
-TianNu's particle number \tcb{(shown in Table \ref{t.sim})} is limited by memory per computing node -- for each computing node, an average of $576^3$ neutrino particles and $288^3$ CDM particles are used, and consumes $\mathcal{M}=40\,{\rm GB}$\footnote{Additional memory is used for OpenMP parallelization and for particle IDs to differentiate different particle spices.}, or about $\mathcal{M}_P=186$.
-A memory efficient configuration of {\tt CUBEP3M} by using large physical scales and at costs of speed, still uses about $\mathcal{M}_P=40$.
+TianNu's particle number (shown in Table \ref{t.sim}) is limited by memory per computing node -- for each computing node, an average of $576^3$ neutrino particles and $288^3$ CDM particles are used, and consumes $\mathcal{M}=40\,{\rm GB}$\footnote{Additional memory is used for OpenMP parallelization and for particle IDs to differentiate different particle spices.}, or about $\mathcal{M}_P=186$.
+A memory-efficient configuration of {\tt CUBEP3M} by using large physical scales and at costs of speed, still uses about $\mathcal{M}_P=40$.
 
@@
 Table \ref{t.memory} shows the memory consumption for this test simulation.
-The memory-consuming arrays are listed and classified into the three types above mentioned, and their memory usages are in unit of GB ($10^9$ byte), byte per particle (bpp), and their percentage to the total memory usage.
+The memory-consuming arrays are listed and classified into the three types above mentioned, and their memory usages are in units of GB ($10^9$ byte), bpp, and their percentage of the total memory usage.
 The parenthesized numbers show overlapped memory, which is saved by equivalencing them with the underscored numbers.
-There are other, unlisted variables which are memory-light, and can also be equivalenced with listed variables.
+There are other unlisted variables that are memory-light, and can also be equivalenced with the listed variables.
 
@@
 The dominating departure from this limit is that $\chi_d$ and $\nu_d$ already occupy 8.24 bpp, which come from the $(1+2N_b/N_t)^3$ term of Equation (\ref{eq.Mp_particle}).
-All other variables occupy additional 8\%.
-On modern super computers, the memory per computing node is usually much larger, and by scaling up the number of particles per node, the buffer ratio $N_b/N_t$ will be lowered and we can approach closer to the 6 bpp limit.
+All other variables occupy an additional 8\%.
+On modern supercomputers, the memory per computing node is usually much larger, and by scaling up the number of particles per node, the buffer ratio $N_b/N_t$ will be lowered and we can approach closer to the 6 bpp limit.
 
@@
   \includegraphics[width=0.99\linewidth]{f8}
- \caption{Statistics of the error in particle displacement $\Delta{\bs\Psi}$ induced by integer based algorithms in {\tt CUBE}.
+ \caption{Statistics of the error in particle displacement $\Delta{\bs\Psi}$ induced by integer-based algorithms in {\tt CUBE}.
 The PDF and CDF of $|\Delta{\bs\Psi}|$ from S256 are shown in the upper and lower axes, as functions of fine grids.
-Black, red, blue and green correspond to x2v2, x1v2, x2v1 and x1v1 respectively.
-The grey lines, marked with ``initial'', shows the distribution of actual displacement of particles in {\tt CUBEP3M}, $|{\bs\Psi}_0|$, which is orders of magnitudes larger than $|\Delta{\bs\Psi}|$.
+Black, red, blue, and green correspond to x2v2, x1v2, x2v1, and x1v1, respectively.
+The gray lines, marked with ``initial'', show the distribution of the actual displacement of particles in {\tt CUBEP3M}, $|{\bs\Psi}_0|$, which is orders of magnitudes larger than $|\Delta{\bs\Psi}|$.
 }
@@
 We run a group of simulations to test the accuracy of {\tt CUBE}.
-We use same seeds to generate same Gaussian random fields in the initial condition generators of {\tt CUBEP3M} and {\tt CUBE}, and then they produce initial conditions of their own format respectively.
-Then the main $N$-body codes run their own initial conditions to redshift $z=0$.
+We use the same seeds to generate the same Gaussian random fields in the initial condition generators of {\tt CUBEP3M} and {\tt CUBE}, and then they produce initial conditions of their own formats.
+Then, the main $N$-body codes run their own initial conditions to redshift $z=0$.
 We use the same force kernels as {\tt CUBEP3M} without PP force.
-Note that near find grid scales there are possibilities to enhance the force kernel to better match the nonlinear power spectrum predictions, however we use the conservative mode of {\tt CUBEP3M} in this paper.
+Note that near the find grid scales, it is possible to enhance the force kernel to better match the nonlinear power spectrum predictions; however, we use the conservative mode of {\tt CUBEP3M} in this paper.
 An extended PP force and an unbiased force matching algorithm will be added to {\tt CUBE}.
-The power spectrum studies are presented in \citet{2013MNRAS.436..540H}, and here we focus on the cross correlations between different integer based methods.
+The power spectrum studies are presented in \citet{2013MNRAS.436..540H}, while here we focus on the cross-correlations between different integer-based methods.
 
-First, by using different configurations -- different number of computing nodes, box sizes, particle resolutions, different number of tiles per node/image etc., we find that by using 2-byte integers for both positions and velocities ($n_\chi=n_\nu=2$, or x2v2) {\tt CUBE} gives exact results compared to {\tt CUBEP3M}.
-So, if sufficient memory is provided, one can always use x2v2 to get exact results as {\tt CUBEP3M}, and the optimal memory limit of this case is 12 bpp, still much lower than traditional methods.
-Next, we focus on the accuracy of the other three cases -- x1v2, x2v1 and x1v1.
+First, by using different configurations -- different numbers of computing nodes, box sizes, particle resolutions, different number of tiles per node/image, etc., we find that by using 2-byte integers for both positions and velocities ($n_\chi=n_\nu=2$, or x2v2) allows {\tt CUBE} to give exact results compared to {\tt CUBEP3M}.
+So if sufficient memory is provided, one can always use x2v2 to get exact results as {\tt CUBEP3M}, and the optimal memory limit of this case is 12 bpp, which is still much lower than traditional methods.
+Next, we focus on the accuracy of the other three cases -- x1v2, x2v1, and x1v1.
 
-We list the names and configurations of simulations used in Table \ref{t.sim}, where $N_{\rm node}$, $L$, $z_i$, $N_p$ and $m_p$ are respectively the number of computing nodes used, length of the side of the box, initial redshift, total number of particles and particle mass.
-These configurations are run by {\tt CUBEP3M}, x2v2, x1v2, x2v1 and x1v1 versions of {\tt CUBE} with same initial seeds.
-Using different number of tiles per image gives exact same results.
+We list the names and configurations of the simulations used in Table \ref{t.sim}, where $N_{\rm node}$, $L$, $z_i$, $N_p$, and $m_p$ are respectively the number of computing nodes used, the length of the side of the box, initial redshift, total number of particles, and particle mass.
+These configurations are run by {\tt CUBEP3M}, x2v2, x1v2, x2v1, and x1v1 versions of {\tt CUBE} with the same initial seeds.
+Using different numbers of tiles per image gives the exact same results.
 We also list the configurations for TianNu and TianZero \citep{2017NatAs...1E.143Y,2017RAA....17...85E} simulations as a reference.
@@
   \includegraphics[width=1.0\linewidth]{f9}
- \caption{Dimensionless power spectra $\Delta^2(k)$ and \tcb{the cross-correlations $\xi(k)$ (lower axes) with respect to x2v2} in simulations S2048S and S2048L.
+ \caption{Dimensionless power spectra $\Delta^2(k)$ and the cross-correlations $\xi(k)$ (lower axes) with respect to x2v2 in simulations S2048S and S2048L.
 The four solid lines show the results from \{x2v2,x1v2,x2v1,x1v1\}.
-In lower panels, the dashed curves show the decorrelations given by the PM algorithm, measured by cross correlating different resolutions of {\tt CUBE}-x2v2.
+In the lower panels, the dashed curves show the decorrelations given by the PM algorithm, measured by cross-correlating different resolutions of {\tt CUBE}-x2v2.
 The vertical orange dashed lines show the scale $k=0.2\, k_{\rm Nyquist}$.}
@@
 \subsection{Displacement of particles}
-In S512, we zoom into a small region of $15.625^2\times 3.90625\ ({\rm Mpc}/h)^3$ and compare the particle distribution between {\tt CUBEP3M} and {\tt CUBE}-x1v1 in Figure \ref{fig.particles}.
-For clarity, {\tt CUBEP3M} particles are marked with bigger, grey dots, whereas smaller red dots are {\tt CUBE}-x1v1 particles over-plotted onto them.
-One can see the difference between them.
-The 1 ${\rm Mpc}/h$, fine grid and coarse grid scales are shown in the figure.
+In S512, we zoom-in on a small region of $15.625^2\times 3.90625\ ({\rm Mpc}/h)^3$ and compare the particle distribution between {\tt CUBEP3M} and {\tt CUBE}-x1v1 in Figure \ref{fig.particles}.
+For clarity, {\tt CUBEP3M} particles are marked with the larger gray dots, whereas the smaller red dots are {\tt CUBE}-x1v1 particles overplotted onto them.
+The 1 ${\rm Mpc}/h$, fine grid, and coarse grid scales are shown in the figure.
 The position resolution of particles in {\tt CUBE}-x1v1 is 1/256 of a coarse grid, or 1/64 of a fine grid.
@@
 To quantify the offset in the final particle distributions, we use PIDs to track the displacement ${\bs \Psi}({\bs q})\equiv{\bs x}-{\bs q}$ of every particle \citep{2017PhRvD..95d3501Y}, where ${\bs x}$ and ${\bs q}$ are Eulerian and Lagrangian coordinates of the particle.
-Then we calculate the absolute value of the offset vector
+Then, we calculate the absolute value of the offset vector
 \begin{equation}\label{eq.offset}
@@
 \end{equation}
-Here, ${\bs\Psi}_0$ stands for {\tt CUBEP3M} and subscript $_i$ can stand for x2v2, x1v2, x2v1 or x1v1.
-The probability distribution functions (PDFs) and cumulative distribution functions (CDFs) of $\Delta\bs\Psi$ in S256 are shown in Figure \ref{fig.dsp}.
-Results from x2v2, x1v2, x2v1 or x1v1 are in black, red, blue and green respectively.
-The results from absolute displacement of particles (by replacing ${\bs\Psi}_i$ with ${\bs q}$ in Equation (\ref{eq.offset})) are shown in grey for comparison.
+Here, ${\bs\Psi}_0$ stands for {\tt CUBEP3M} and subscript $_i$ can stand for x2v2, x1v2, x2v1, or x1v1.
+The PDFs and CDFs of $\Delta\bs\Psi$ in S256 are shown in Figure \ref{fig.dsp}.
+Results from x2v2, x1v2, x2v1, or x1v1 are in black, red, blue, and green respectively.
+The results from absolute displacement of particles (by replacing ${\bs\Psi}_i$ with ${\bs q}$ in Equation (\ref{eq.offset})) are shown in gray for comparison.
 
 For x2v2, almost all particles are accurate up to 1/100 of a fine grid, and the worst particle is $\sim 1/10$ of a fine grid away from its counterpart in {\tt CUBEP3M}.
-The difference is caused by \tcb{round-off} errors and is negligible in physical and cosmological applications.
-The accuracy of x1v2 is between x2v2 and x1v1, and x2v1 gives only minor improvement from x1v1.
-We also run a simulation with same number of particles but with $L=600\ {\rm Mpc}/h$ ($m_p=1.2\times 10^{12}\Msun$), and find that the accuracy of x1v2 is in turn between x2v1 and x1v1.
-We interpret that, in this latter case, particles have lower mass resolution, so they move slower and need higher position resolution but need lower velocity resolution, then x2v1 outperforms x1v2.
+The difference is caused by round-off errors and is negligible in physical and cosmological applications.
+The accuracy of x1v2 is between x2v2 and x1v1, and x2v1 gives only a minor improvement from x1v1.
+We also run a simulation with the same number of particles but with $L=600\ {\rm Mpc}/h$ ($m_p=1.2\times 10^{12}\Msun$), and find that the accuracy of x1v2 is in turns between x2v1 and x1v1.
+We interpret that, in this latter case, particles have lower mass resolution, so they move slower and need higher position resolution but need lower velocity resolution, thus x2v1 outperforms x1v2.
 
 S2048S and S2048L are two simulations with $2048^3$ particles in small ($L=400\ {\rm Mpc}/h$) and large ($L=1200\ {\rm Mpc}/h$) box sizes.
-We compare their accuracy by their power spectra and their cross correlations with {\tt CUBEP3M} at $z=0$.
+We compare their accuracy by their power spectra and their cross-correlations with {\tt CUBEP3M} at $z=0$.
 The physical scale of S2048S is designed such that the particle mass resolution $m_p$ is comparable to TianNu and TianZero simulations (their parameters are also listed in Table \ref{t.sim}).
-On the other hand, S2048L focuses on larger structures, on which scale one can study weak gravitational lensing, baryonic acoustic oscillation (BAO) \citep{2005ApJ...633..560E} and its reconstruction \citep{2007ApJ...664..675E,2017ApJ...841L..29W} etc.
+On the other hand, S2048L focuses on larger structures, on which scale one can study weak gravitational lensing, BAO \citep{2005ApJ...633..560E}, and its reconstruction \citep{2007ApJ...664..675E,2017ApJ...841L..29W}, etc.
 
 For each simulation the particles are firstly cloud-in-cell (CIC) interpolated onto the fine mesh grid, and from the density field $\rho$ we define the density contrast $\delta\equiv \rho/\left\langle \rho \right\rangle-1$.
-We define the cross power spectrum $P_{\alpha\beta}(k)$ between two fields $\delta_\alpha$ and $\delta_\beta$ ($\delta_\alpha=\delta_\beta$ for auto power spectrum) in Fourier space as
+We define the cross-power spectrum $P_{\alpha\beta}(k)$ between two fields $\delta_\alpha$ and $\delta_\beta$ ($\delta_\alpha=\delta_\beta$ for auto-power spectrum) in Fourier space as
 \begin{equation}\label{eq.cps}
@@
 \end{equation}
-where ${\bs \delta}_{\rm 3D}$ is the three-dimensional Dirac delta function.
+where ${\bs \delta}_{\rm 3D}$ is the 3D Dirac delta function.
 In cosmology we usually consider the dimensionless power spectrum $\Delta^2_{\alpha\beta}(k)\equiv k^3 P_{\alpha\beta}(k)/(2\pi^2)$.
-The cross correlation coefficient is defined as
+The cross-correlation coefficient is defined as
 \begin{equation}\label{eq.ccc}
@@
 In the upper two panels of figure \ref{fig.ccc} we show the power spectra of {\tt CUBE}.
-In both plots of S2048S and S2048L the four solid curves of different colors show the results of x2v2, x1v2, x2v1 and x1v1, and they almost overlapped with each other.
-The dashed curves are the nonlinear prediction of matter power spectrum by CLASS \citep{2011JCAP...07..034B}.
+In both plots of S2048S and S2048L the four solid curves of different colors show the results of x2v2, x1v2, x2v1, and x1v1, and they almost overlapped with each other.
+The dashed curves are the nonlinear prediction of the matter power spectrum by CLASS \citep{2011JCAP...07..034B}.
 
@@
 This error is from the PM algorithm and one has to increase the resolution of the simulation to correct these offsets.
-We do not plot {\tt CUBEP3M} because we examined that {\tt CUBEP3M} and x2v2 of {\tt CUBE} produce same results.
+We do not plot {\tt CUBEP3M} because we found that {\tt CUBEP3M} and x2v2 of {\tt CUBE} produce same results.
 Thus, the differences between {\tt CUBEP3M} and different integer formats of {\tt CUBE} are negligible compared to the error of the PM algorithm.
 
-In the lower parts of these four panels study the cross-correlations.
-We compare everything with {\tt CUBE}-x2v2, to emphasize the decorrelation \tcb{(i.e.
-$1-\xi(k)$)} by different integer formats.
+In the lower parts of these four panels we study the cross-correlations.
+We compare everything with {\tt CUBE}-x2v2, to emphasize the decorrelation (i.e.
+$1-\xi(k)$) by different integer formats.
 Note that x2v2 is perfectly correlated with {\tt CUBEP3M}.
-In the lower two panels of figure \ref{fig.ccc} the solid curves show the decorrelation $\xi$ of x1v2, x2v1 and x1v1.
-Higher resolution (S2048S) in general comes with more decorrelations, and x1v2 cross correlate with x2v2 better than the other two cases.
-In order to quantify the PM error in terms of decorrelations, we run x2v2 simulations with same initial conditions, but 8 times more particles and cells, and measure the decorrelation caused by coarser resolutions.
-These results are shown in dashed curves (labeled ``PM'').
-We conclude that cross correlation of x1v1, in all cases, is not worse than the PM errors.
+In the lower two panels of figure \ref{fig.ccc} the solid curves show the decorrelation $\xi$ of x1v2, x2v1, and x1v1.
+The higher resolution (S2048S) in general comes with more decorrelations, and x1v2 cross-correlates with x2v2 better than the other two cases.
+In order to quantify the PM error in terms of decorrelations, we run x2v2 simulations with the same initial conditions, but 8 times more particles and cells, and measure the decorrelation caused by coarser resolutions.
+These results are shown as dashed curves (labeled ``PM'').
+We conclude that the cross-correlation of x1v1, in all cases, is not worse than the PM errors.
 
-To summarize from Figure \ref{fig.ccc}, in terms of either power spectrum deviation or cross correlation, the error induced by information optimization (even for x1v1) is lower than the error from the PM algorithm, and we can safely use x1v1 for most of the LSS studies.
+To summarize from Figure \ref{fig.ccc}, in terms of either power spectrum deviation or cross-correlation, the error induced by information optimization (even for x1v1) is lower than the error from the PM algorithm, and we can safely use x1v1 for most of the LSS studies.
 
@@
 We present a parallel, information-optimized $N$-body algorithm.
-This open-source code {\tt CUBE}, has recently been used in many studies of LSS, e.g.
+This open-source code, {\tt CUBE}, has recently been used in many studies of LSS, e.g.
 \cite{2017PhRvD..95d3501Y,2017ApJ...841L..29W,2017MNRAS.469.1968P}.
-It needs very low memory usage, approaching 6 byte per particle (6 bpp).
+It requires very low memory usage, approaching 6 bpp.
 
 The accuracy of this code is adjustable in that we can choose 1-byte/2-byte integers separately for positions and velocities of particles.
-In the case of using 2-byte integers for both positions and velocities (``x2v2'', and memory can be 12 bpp), the algorithm gives exact results given by traditional $N$-body algorithms.
-Note that the results are exactly same in that they not only produce same physical statistics of LSS, but also the same error (not physical) from the PM algorithm near Nyquist frequencies.
-In another word, the positions and velocities of each particle are exact.
-In practice, we only require that the errors from information optimization is much lower than the errors from PM algorithm.
-In Figure \ref{fig.ccc} we see that it is the case even for the most memory efficient configuration x1v1.
-This shows that in most LSS studies, when our scales of interest are smaller than $k\simeq 0.2\,k_{\rm Nyquist}$, six 1-byte fixed point numbers contains sufficient information of every $N$-body particle.
+In the case of using 2 byte integers for both positions and velocities (``x2v2'', and memory can be 12 bpp), the algorithm gives the exact results given by traditional $N$-body algorithms.
+Note that the results are exactly the same in that they not only produce the same physical statistics of LSS, but also the same error (not physical) from the PM algorithm near Nyquist frequencies.
+In other words, the positions and velocities of each particle are exact.
+In practice, we only require that the errors from information optimization is much lower than the errors from the PM algorithm.
+In Figure \ref{fig.ccc} we see that this is the case even for the most memory-efficient configuration, x1v1.
+This shows that in most LSS studies, when our scales of interest are smaller than $k\simeq 0.2\,k_{\rm Nyquist}$, six 1-byte fixed point numbers contain sufficient information of every $N$-body particle.
 
-Another benefit for this algorithm is LSS simulations with neutrinos, although the neutrino modules of {\tt CUBE} is in development.
-Neutrinos have a high velocity dispersion and move much faster than CDM, and their small scale errors are dominated by their Poisson noise.
-We expect that, compared to CDM, x1v1 gives less power spectrum deviation to neutrinos, as they behave more Gaussian and less clustered.
-LSS-neutrino simulations, like TianNu, can contain much more (for TianNu, 8 times more) neutrino $N$-body particles than CDM, which dominates the memory.
+Another benefit of this algorithm is LSS simulations with neutrinos, although the neutrino modules of {\tt CUBE} are in development.
+Neutrinos have a high velocity dispersion and move much faster than CDM, and their small-scale errors are dominated by their Poisson noise.
+We expect that, compared to CDM, x1v1 gives less power spectrum deviation to neutrinos, as they behave more Gaussian and are less clustered.
+LSS-neutrino simulations, like TianNu, can contain much more (for TianNu, 8 times more) neutrino $N$-body particles than CDM, which dominate the memory.
 For a TianNu-like simulation, one can safely use x1v2 or x2v2 for CDM and x1v1 for neutrinos, and the memory usage can approach 6 bpp.
-This allows much more particles in the simulation and can lower the Poisson noise of neutrinos prominently.
-A 1-byte PID per particle increases minor memory usage and can differentiate 8 kinds of particles, or we can store different particles in different arrays without using PID.
+This allows much more particles to be included in the simulation and can lower the Poisson noise of neutrinos prominently.
+A 1-byte PID per particle increases minor memory usage and can differentiate eight kinds of particles, or we can store different particles in different arrays without using PID.
 
 We did not include PP force in {\tt CUBE} and {\tt CUBEP3M} in this paper.
-If one wants to focus on smaller scales, like halo masses and profiles, an extended PP forces, which act up to adjacent fine cells, should take into account.
+If one wants to focus on smaller scales, like halo masses and profiles, an extended PP forces, which act up to adjacent fine cells, should be taken into account.
 The memory consumption for PP force is only local and is negligible compared to particles.
-In these cases, even x2v2 is used, a 12 bpp memory usage is still much lower than traditional $N$-body algorithms.
+In these cases where even x2v2 is used, a 12 bpp memory usage is still much lower than that of traditional $N$-body algorithms.
 
-Traditional $N$-body codes consumes considerable memory while doing a relatively light computations.
+Traditional $N$-body codes consume considerable memory while performing relatively light computations.
 {\tt CUBE} is designed to optimize the efficiency of information and the memory usage in $N$-body simulations.
-{\tt CUBE} is written in Coarray Fortran -- the concise Coarray features are used instead of complicated MPI, and the code itself is much more concise than {\tt CUBEP3M} for future maintenance and development.
+{\tt CUBE} is written in Coarray Fortran -- concise Coarray features are used instead of complicated MPI -- and the code itself is much more concise than {\tt CUBEP3M} for future maintenance and development.
 The next steps are optimization of the code and adapting it for various kinds of heterogeneous computing systems, e.g.
 MIC and GPUs.
-Optimizing the velocity storage may further improve the accuracy of x1v1, and whose effects on neutrino-LSS simulation is going to be discovered.
+Optimizing the velocity storage may further improve the accuracy of x1v1, and whose effects on neutrino-LSS simulation are yet to be discovered.
 
@@
 We acknowledge funding from NSERC.
-H.R.Y. thanks Derek Inman for many helpful discussions. \tcb{We thank the anonymous referee for many helpful suggestions which improve the paper.}
+H.R.Y. thanks Derek Inman for many helpful discussions. We thank the anonymous referee for many helpful suggestions that improve the paper.
 The simulations were performed on the GPC supercomputer at the SciNet HPC Consortium.

### 2019-02-03 CUBEnu/cube_fft/compile.sh
spin-correlation, force field power, RSD density field etc.

@@ -1,0 +1,4 @@
+rm -f *.o *.out
+
+gfortran -O3 -cpp -fcoarray=single -mcmodel=medium test.f90 -I/usr/local/include/ -L/usr/local/lib/ -lfftw3f -lm -ldl
+#gfortran -fopenmp -cpp -fcoarray=single test_parallel.f90 -I/usr/local/include/ -L/usr/local/lib/ -lfftw3f -lm -ldl

### 2019-02-03 CUBEnu/cube_fft/compile_intel.sh

@@ -1,7 +1,0 @@
-module purge
-module load intel/16.0.3 intelmpi/5.0.3.048 fftw/3.3.4-intel-impi
-module list
-
-rm -f *.o *.out
-
-ifort -i8 -O3 -xHost -fpp -mcmodel=medium test.f90 -lfftw3f -I -L -lm -ldl

### 2019-02-03 CUBEnu/cube_fft/compile_mac.sh

@@ -1,4 +1,9 @@
+#module purge
+#module load intel/16.0.3 intelmpi/5.0.3.048 fftw/3.3.4-intel-impi
+#module list
+
 rm -f *.o *.out
 
-gfortran -O3 -cpp -fcoarray=single -mcmodel=medium test.f90 -I/usr/local/include/ -L/usr/local/lib/ -lfftw3f -lm -ldl
-#gfortran -fopenmp -cpp -fcoarray=single test_parallel.f90 -I/usr/local/include/ -L/usr/local/lib/ -lfftw3f -lm -ldl
+gfortran -O3 -cpp -fopenmp -mcmodel=medium test.f90 -lfftw3f -I/usr/local/include/ -L/usr/local/lib/ -lm -ldl -o fft1.x
+gfortran -O3 -cpp -fopenmp -mcmodel=medium test_parallel.f90 -lfftw3_omp -lfftw3f_omp  -I/usr/local/include/ -L/usr/local/lib/ -lm -ldl -o fft32.x
+export OMP_NUM_THREADS=32

### 2019-02-03 CUBEnu/cube_fft/test.f90

@@ -3,7 +3,7 @@ program test
   use iso_fortran_env, only : int64
   implicit none
   include 'fftw3.f'
-  integer,parameter :: ng=512 ! number of grid per dimension
+  integer,parameter :: ng=304 ! number of grid per dimension
   real,parameter :: pi=3.14159
   integer(8) plan_fft,plan_ifft ! fft plans
   real den(ng+2,ng,ng) ! fft array, additional "+2" for storing Nyquist frequency

### 2019-02-03 CUBEnu/cube_fft/test_parallel.f90

@@ -1,0 +1,117 @@
+program test
+  !use iso_fortran_env, only : int64
+  use,intrinsic :: ISO_C_BINDING
+  use omp_lib
+  implicit none
+  include 'fftw3.f03'
+  integer,parameter :: ng=304 ! number of grid per dimension
+  real,parameter :: pi=3.14159
+  integer(8) plan_fft,plan_ifft ! fft plans
+  real den(ng+2,ng,ng) ! fft array, additional "+2" for storing Nyquist frequency
+  real temp_r,temp_theta
+  real total_start_cpu, total_finish_cpu, totaltime_cpu
+  double precision :: start, finish, total_start, total_finish
+  double precision :: createtime, initialtime, setuptime, executime, totaltime
+  integer i,j,k,l,seedsize,stat
+  integer(c_int64_t) time64
+  integer(4) i4,t1,t2,t_rate,ncore
+
+  integer,allocatable :: iseed(:)
+
+
+
+
+  !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
+  call cpu_time(total_start_cpu)
+  total_start=omp_get_wtime() 
+  start = omp_get_wtime()
+  print*, 'ilp64 test: default integer'
+  i=2147483647
+  i=i+1
+  print*,i
+
+  call sfftw_init_threads(stat)
+  write(*,*) "Status:", stat
+  write(*,*) "ng=",ng
+  ncore=omp_get_max_threads()
+  print*, 'max threads =',ncore
+  !ncore=4
+  call sfftw_plan_with_nthreads(ncore)
+  write(*,*) "Using ", ncore, "threads."
+
+  ! fft plan
+  call system_clock(t1,t_rate) 
+  print*, 'call forward plan'
+  call sfftw_plan_dft_r2c_3d(plan_fft,ng,ng,ng,den,den,FFTW_MEASURE)
+  print*, 'call backward plan'
+  call sfftw_plan_dft_c2r_3d(plan_ifft,ng,ng,ng,den,den,FFTW_MEASURE)
+  call system_clock(t2,t_rate)
+  print*, '  elapsed time =',real(t2-t1)/t_rate,'secs';
+
+  ! initialize matrix
+  call system_clock(t1,t_rate)
+  print*, 'initialize den'
+  call random_seed(size=seedsize)
+  !print*,'min seedsize =', seedsize
+  seedsize=max(seedsize,12)
[…]

### 2019-02-03 CUBEnu/work/main/checkpoint.f90

@@ -25,8 +25,6 @@ subroutine checkpoint
   sim%t=t
   sim%tau=tau
 
-  sim%istep=istep
-
   sim%cur_checkpoint=cur_checkpoint
 
   sim%vsim2phys=(1.5/a)*box*100.*sqrt(omega_m)/nf_global
@@ -48,6 +46,11 @@ subroutine checkpoint
     !$omp section
     open(15,file=fn15,status='replace',access='stream'); write(15) pid(:sim%nplocal); close(15)
     print*, 'check PID range: ',minval(pid(:sim%nplocal)),maxval(pid(:sim%nplocal))
+    print*, 'check PID range:',minval(pid(:sim%nplocal)),maxval(pid(:sim%nplocal))
+    if (minval(pid(:sim%nplocal))<1) then
+      print*, 'pid are not all positive'
+      
+    endif
 # endif
   !$omp endparallelsections
   print*,'  image',this_image(),'wrote',sim%nplocal,'CDM particles'

### 2019-02-03 CUBEnu/work/main/halofind.f90

@@ -1,3 +1,4 @@
+#define HID
 module halo_output
   implicit none
 
@@ -9,7 +10,8 @@ module halo_output
   type type_halo_info
     real hpos(3)
     real mass_odc,radius_odc,v_disp
-    real x_mean(3),v_mean(3),ang_mom(3),var_x(3)
+    real x_mean(3),v_mean(3),ang_mom(3),var_x(3),inertia(3,3)
+    real q_mean(3),inertia_q(3,3)
   endtype
 endmodule
 
@@ -41,9 +43,9 @@ subroutine halofind
   integer(4) ilist_odc(max_halo_np),i_odc,GroupOffset,temp_halo(2,200000)
   integer idist(3,nlist),isortdist(nlist),idist_tmp(3,nlist),isortpeak(n_peak_max)
   integer isortpos_odc(max_halo_np)
-  integer iloc,np_odc,np_search
+  integer iloc,np_odc,np_search,nptemp,qtemp
   integer crbox(3,2),csbox(3,2),frbox(3,2),itile(3),newbox(3),imax(3)
-  real den_odc,d1,d2,r1,r2,w1,w2,den1,den2,dx(3),dv(3)
+  real den_odc,d1,d2,r1,r2,w1,w2,den1,den2,dx(3),dv(3),qpos(3),dx_mean(3),qpos_mean(3),inert(3,3)
   real finegrid(0:newbox_max+1,0:newbox_max+1,0:newbox_max+1)
   real xv_odc(6,max_halo_np),r_odc(max_halo_np)
   integer(4) oct(max_halo_np)
@@ -108,8 +110,8 @@ subroutine halofind
   ! open halo file
   open(11,file=output_name_halo('halo'),status='replace',access='stream')
   open(12,file=output_name_halo('halo_pid'),status='replace',access='stream')
-  open(101,file=output_name_halo('group_tab'),status='replace',access='stream')
-  open(102,file=output_name_halo('group_ids'),status='replace',access='stream')
+  !open(101,file=output_name_halo('group_tab'),status='replace',access='stream')
+  !open(102,file=output_name_halo('group_ids'),status='replace',access='stream')
   ! Determine which candidates are to be considered halos and write their properties to file.
 # ifdef HID
     open(21,file=output_name_halo('hid'),status='replace',access='stream')
@@ -125,8 +127,8 @@ subroutine halofind
   header_halo_tab%TotNgroups=0
   header_halo_tab%Nids=0
[…]

### 2019-02-03 CUBEnu/work/main/main.f90

@@ -3,6 +3,7 @@
 !   haoran@cita.utoronto.ca   !
 !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
 
+!#define record_Fc
 program main
   use omp_lib
   use variables
@@ -18,9 +19,16 @@ program main
 
 # ifdef halo_spin_correlation
     integer(2) halo_id,hid(nf**3),ih
-    integer(4) nhalo,nhalo_tot,nwrite,i0
+    integer(4) nwrite,i0,j0,ipos(3)
     integer(4),allocatable :: isort_mass(:)
-    real,allocatable :: hpos(:,:),spin_q(:,:,:),x_mean(:,:),v_mean(:,:),hp(:),hcorr(:,:,:),scale_fac(:),spin_x(:,:),halonp(:)
+    real,allocatable :: hpos(:,:),spin_q(:,:,:),x_mean(:,:),v_mean(:,:),hp(:),scale_fac(:),spin_x(:,:),halonp(:)
+    real,allocatable :: dspinx(:,:),corrdj(:),ratiodj(:),jtot(:,:,:)
+    real,allocatable :: inertia(:,:,:),tide(:,:,:),tidu(:,:,:),torque(:,:,:),eIT(:,:,:)
+    real,allocatable :: haloloc(:,:),haloforce(:,:)
+    real,allocatable :: force_cdm(:,:),force_neu(:,:)
+    real,allocatable :: phi(:,:,:)[:],phu(:,:,:)[:]
+    real,allocatable :: hinfo(:,:,:)
+    real ztfnu(10),atfnu(10)
     real den_odc,dx(3),dv(3)
     type(type_halo_info) halo_info
 # endif
@@ -29,13 +37,26 @@ program main
   call initialize
   call particle_initialization
 # ifdef halo_spin_correlation
+    open(21,file='tf_nu.txt',status='old')
+    read(21,fmt='(f8.4)') ztfnu
+    close(21)
+    atfnu=1.0/(1+ztfnu)
+
     open(19,file=output_dir()//'0.000_hid'//output_suffix(),access='stream',status='old')
     read(19) hid
     close(19)
+
     open(19,file=output_dir()//'0.000_halo'//output_suffix(),access='stream',status='old')
     read(19) nhalo_tot,nhalo,den_odc
[…]

### 2019-02-03 CUBEnu/work/main/parameters.f90

@@ -22,7 +22,7 @@ module parameters
   ! (hereafter 'number of coarse cells' = 'nc')
   ! (hereafter 'per dimension' = '/dim')
   integer(8),parameter :: nn=1 ! number of imgages (nodes) /dim
-  integer(8),parameter :: ncore=4
+  integer(8),parameter :: ncore=8
   integer(8),parameter :: n_nest=1 ! number of nested threads
   integer(8),parameter :: ncell=4 ! number of nf in each nc, /dim
   integer(8),parameter :: nnt=2 ! number of tiles /image/dim
@@ -53,24 +53,24 @@ module parameters
 
   integer(8),parameter :: nfe=nft+2*nfb ! 96
 
-  logical,parameter :: body_centered_cubic=.false.
-  integer(8),parameter :: np_nc=ncell ! number of particles / coarse cell / dim
+  logical,parameter :: body_centered_cubic=.true.
+  integer(8),parameter :: np_nc=ncell/2 ! number of particles / coarse cell / dim
   integer, parameter :: np_nc_nu = ncell/2 ! number of neutrinos per dim per coarse cell
 
   logical,parameter :: Extended_pp_force=.false.
-  real,parameter :: rsoft=0.6 ! PP softening length
+  real,parameter :: rsoft=0.3 ! PP softening length
   integer,parameter :: pp_range=1 ! set <=4
-  real,parameter :: image_buffer=1.2
-  real,parameter :: tile_buffer=1.5
+  real,parameter :: image_buffer=1.5
+  real,parameter :: tile_buffer=2.5
   real,parameter :: vbuf=0.9
 
   real,parameter :: pi=4*atan(1.)
 
   ! cosmological parameters
-  real,parameter :: box=600.0*nn  ! simulation scale /dim, in unit of Mpc/h
-  real,parameter :: s8=0.8276 ! use -Dsigma_8 in initial_conditions
+  real,parameter :: box=100*nn  ! simulation scale /dim, in unit of Mpc/h
+  real,parameter :: s8=0.8 ! use -Dsigma_8 in initial_conditions
 
-  real,parameter :: z_i_nu=100.0 ! initial redshift for neutrinos
+  real,parameter :: z_i_nu=5.0 ! initial redshift for neutrinos
   real,parameter :: a_i_nu=1./(1.+z_i_nu) ! initial scale factor for neutrinos
[…]

### 2019-02-03 CUBEnu/work/main/particle_initialization.f90

@@ -56,7 +56,11 @@ subroutine particle_initialization
 # ifdef PID
   !$omp section
     open(15,file=fn15,status='old',access='stream'); read(15) pid(:sim%nplocal); close(15)
-    print*, 'check PID range: ',minval(pid(:sim%nplocal)),maxval(pid(:sim%nplocal))
+    print*, '  check PID range: ',minval(pid(:sim%nplocal)),maxval(pid(:sim%nplocal))
+    if (minval(pid(:sim%nplocal))<1) then
+      print*, 'warning: pid are not all positive'
+      !stop
+    endif
 # endif
   print*,'  from image',this_image(),'read',sim%nplocal,' CDM particles'
 

### 2019-02-03 CUBEnu/work/main/pm.f90

@@ -1,3 +1,4 @@
+!#define record_Fc
 subroutine particle_mesh
   use omp_lib
   use variables
@@ -6,6 +7,10 @@ subroutine particle_mesh
   use neutrinos
   implicit none
   save
+
+#ifdef record_Fc
+  character(10) :: str_a
+#endif
 
   ! force settings
   !logical,parameter :: fine_force=.true.
@@ -22,6 +27,8 @@ subroutine particle_mesh
     print*, ''
     print*, 'particle mesh'
     print*, '  mass_p cdm/nu =',sim%mass_p_cdm,sim%mass_p_nu
+    print*, '  a_mid =',a_mid
+    print*, '  dt =',dt
     call system_clock(tt1,t_rate)
   endif
 
@@ -272,6 +279,17 @@ subroutine particle_mesh
     call pencil_fft_backward
     force_c(i_dim,1:nc,1:nc,1:nc)=r3
   enddo
+
+#ifdef record_Fc
+  write(str_a,'(f7.3)') a_mid
+  write(65,*) str_a
+  open(66,file=opath//'image1/'//trim(adjustl(str_a))//'_Fc'//output_suffix(),status='replace',access='stream')
+  write(66) force_c(1,1:nc,1:nc,1:nc)
+  write(66) force_c(2,1:nc,1:nc,1:nc)
+  write(66) force_c(3,1:nc,1:nc,1:nc)
+  close(66)
+#endif
+
   call system_clock(t2,t_rate)
   print*, '      elapsed time =',real(t2-t1)/t_rate,'secs';
   sync all

### 2019-02-03 CUBEnu/work/main/run_halo_corr_mac.sh

@@ -6,7 +6,7 @@ make
 
 cd ../main/
 make clean
-make EXTRA=-DHID
+make EXTRA=-DHALOFIND
 
 cd ../utilities/
 ./ic.x
@@ -17,14 +17,15 @@ cd ../main/
 
 cd ../utilities/
 ./cicpower.x
+#./dsp.x
 ./ang_mom_corr.x # generate halo_init_spin
 
 cd ../main/
-make clean
-make EXTRA=-Dhalo_spin_correlation
-./main.x
+#make clean
+#make EXTRA=-Dhalo_spin_correlation
+#./main.x
 
 cd ../utilities/
-./ang_mom_corr.x
+#./ang_mom_corr.x
 
 cd ../main/

### 2019-02-03 CUBEnu/work/main/run_mac.sh

@@ -19,9 +19,10 @@ cd ../main/
 ./main.x
 
 cd ../utilities/
+./cicrsd.x
 ./cicpower.x
 ./halofinder.x
-./ang_mom_corr.x
+#./ang_mom_corr.x
 #source findhalos.sh
 
 cd ../main/

### 2019-02-03 CUBEnu/work/main/update_particle.f90

@@ -30,6 +30,7 @@ subroutine update_xp()
   if (head) print*,'update_xp'
   call system_clock(t1,t_rate)
   dt_mid=(dt_old+dt)/2
+  if (head) print*,'  dt_mid =',dt_mid
   overhead_tile=0
   np_prev=0
 

### 2019-02-03 CUBEnu/work/utilities/ang_mom_corr.f90

@@ -1,5 +1,3 @@
-!! add -DNEUTRINOS to compute cross_power(delta_c,delta_nu)
-!! otherwise compute cross_power(delta_L,delta_c)
 program ang_mom_corr
   use parameters
   use halo_output, only: type_halo_catalog, type_halo_info
@@ -19,7 +17,7 @@ program ang_mom_corr
   integer pid(max_halo_np),ipos(3)
   real qpos(3,max_halo_np),qpos_mean(3),vq(3),dx(3),vf,scale_factor,dx_mean(3)
   real spin_q(3),spin_x(3),spin_t(3),spin_u(3)
-  real inertia(3,3),tide(3,3),torque(3,3)
+  real inertia(3,3),tide(3,3),tidu(3,3),torque(3,3)
   real force_cdm(3),force_neu(3),torque_c,torque_u
   real theta_qx(3000),theta_qt(3000),theta_tx(3000),theta_cu(3000),theta_ff(3000)
   real phi(0:nf+1,0:nf+1,0:nf+1)[*]
@@ -55,10 +53,16 @@ program ang_mom_corr
   sync all
 
   cur_checkpoint=1
+  print*,output_name('phi1')
   open(21,file=output_name('phi1'),status='old',access='stream')
-  read(21) phi(1:nf,1:nf,1:nf)
+  !read(21) phi(1:nf,1:nf,1:nf)
+  !open(21,file='../../output/universe2/image1/5.000_phi_z5_1.bin',status='old',access='stream')
+  !open(21,file='../../output/universe2/image1/0.000_phiE_1.bin',status='old',access='stream')
+  read(21) phi(1:nf,1:nf,1:nf)  !! attention 1 layer buffer also loaded
   close(21)
+  print*, output_name('phi1_nu')
   open(21,file=output_name('phi1_nu'),status='old',access='stream')
+  !open(21,file='../../output/universe2/image1/5.000_tf5.000_phi1_nu_1.bin',status='old',access='stream')
   read(21) phu(1:nf,1:nf,1:nf)
   close(21)
   vf=vfactor(1/(1+z_checkpoint(cur_checkpoint)))
@@ -89,15 +93,8 @@ program ang_mom_corr
     print*,output_name('halo')
     open(11,file=output_name('halo'),status='old',access='stream')
[…]

### 2019-02-03 CUBEnu/work/utilities/cicpower.f90

@@ -1,5 +1,8 @@
 !! add -DNEUTRINOS to compute cross_power(delta_c,delta_nu)
 !! otherwise compute cross_power(delta_L,delta_c)
+
+!#define write_xreal
+!#define analysis
 program cicpower
   use parameters
   use pencil_fft
@@ -52,7 +55,7 @@ program cicpower
 
   call create_penfft_plan
 
-  do cur_checkpoint= 1,n_checkpoint
+  do cur_checkpoint= n_checkpoint,n_checkpoint
     if (head) print*, 'Start analyzing redshift ',z2str(z_checkpoint(cur_checkpoint))
 
     !call particle_initialization
@@ -84,6 +87,10 @@ program cicpower
     read(11) rhoc
     close(11)
 
+#ifdef write_xreal
+    open(12,file=output_name('xreal'),status='replace',access='stream')
+#endif
+
     rho_grid=0
     nlast=0
     do itz=1,nnt
@@ -96,6 +103,11 @@ program cicpower
         do l=1,np
           ip=nlast+l
           pos1=nt*((/itx,ity,itz/)-1)+ ((/i,j,k/)-1) + (int(xp(:,ip)+ishift,izipx)+rshift)*x_resolution
+
+#ifdef write_xreal
+    write(12) pos1*real(ng)/real(nc)
+#endif
+
           pos1=pos1*real(ng)/real(nc) - 0.5
 
           idx1=floor(pos1)+1
@@ -130,9 +142,20 @@ program cicpower
     rho_grid(:,ng,:)=rho_grid(:,ng,:)+rho_grid(:,0,:)[image1d(icx,ipy,icz)]; sync all
     rho_grid(:,:,1)=rho_grid(:,:,1)+rho_grid(:,:,ng+1)[image1d(icx,icy,inz)]
     rho_grid(:,:,ng)=rho_grid(:,:,ng)+rho_grid(:,:,0)[image1d(icx,icy,ipz)]; sync all
-    rho_c=rho_grid(1:ng,1:ng,1:ng)
-    !print*, 'check: min,max,sum of rho_grid = '
-    !print*, minval(rho_c),maxval(rho_c),sum(rho_c*1d0)
+    !rho_c=rho_grid(1:ng,1:ng,1:ng)
+    do i=1,ng
+    do j=1,ng
+    do k=1,ng
+      rho_c(k,j,i)=rho_grid(k,j,i)
+    enddo
+    enddo
+    enddo
+print*,rho_grid(1:2,1:2,1)
[…]

### 2019-02-03 CUBEnu/work/utilities/cicrsd.f90

@@ -1,0 +1,192 @@
+!! Include RSD effect of particles, and CIC interpolate into grid to get the redshift-space density field
+program cicrsd
+  use parameters
+  implicit none
+  save
+  ! nc: coarse grid per node per dim
+  ! nf: fine grid per node per dim
+  real,parameter :: density_buffer=1.2
+  integer,parameter :: zdim=2 ! the dimension being the redhisft direction
+  integer(8) i,j,k,l,i_dim,iq(3),nplocal,nplocal_nu,itx,ity,itz
+  integer(8) nlast,ip,np,idx1(3),idx2(3)
+
+  real(4) rho_grid(0:ng+1,0:ng+1,0:ng+1)[*]
+  real(4) rho_c(ng,ng,ng),sigma_vi,zshift
+  real(4) mass_p,pos1(3),dx1(3),dx2(3)
+  real(8) rho8[*]
+  integer(izipx),allocatable :: xp(:,:)
+  integer(izipv),allocatable :: vp(:,:)
+  integer(4) rhoc(nt,nt,nt,nnt,nnt,nnt)
+  real(4) vc(3,nt,nt,nt,nnt,nnt,nnt)
+
+  character(20) str_z,str_i
+  call geometry
+  if (head) then
+    print*, 'cicrsd on resolution:'
+    print*, 'ng=',ng
+    print*, 'ng*nn=',ng*nn
+  endif
+  sync all
+
+  if (head) then
+    print*, 'checkpoint at:'
+    open(16,file='../main/z_checkpoint.txt',status='old')
+    do i=1,nmax_redshift
+      read(16,end=71,fmt='(f8.4)') z_checkpoint(i)
+      print*, z_checkpoint(i)
+    enddo
+    71 n_checkpoint=i-1
+    close(16)
+    print*,''
+  endif
+
+  sync all
+  n_checkpoint=n_checkpoint[1]
+  z_checkpoint(:)=z_checkpoint(:)[1]
+  sync all
+
+  do cur_checkpoint= n_checkpoint,n_checkpoint
+    if (head) print*, 'Start analyzing redshift ',z2str(z_checkpoint(cur_checkpoint))
+
+    !call particle_initialization
+
+    open(11,file=output_name('info'),status='old',action='read',access='stream')
+    read(11) sim
+    close(11)
[…]

### 2019-02-03 CUBEnu/work/utilities/displacement.f90

@@ -1,4 +1,6 @@
 !! add -DRECONSTRUCTION to compute delta_E from dsp
+#define Emode
+#define potential
 program displacement
   use parameters
 #ifdef Emode
@@ -8,16 +10,13 @@ program displacement
   save
   ! nc: coarse grid per node per dim
   ! nf: fine grid per node per dim
-  integer(8),parameter :: npnode=nf**3 ! only true for this project
-  real,parameter :: density_buffer=1.2
-  integer(8),parameter :: npmax=npnode*density_buffer
   integer(8) i,j,k,l,i_dim,iq(3),pid8,itx,ity,itz,nlast,ip,np
   integer(4) rhoc(nt,nt,nt,nnt,nnt,nnt),rhoc0(nt,nt,nt,nnt,nnt,nnt)
-  integer(4) rho0(ng,ng,ng) !!!! for checking there is one and only one particle per fine grid
+  integer(1) rho0(ng,ng,ng) !!!! for checking there is one and only one particle per fine grid
   real dsp(3,ng,ng,ng),pos0(3),pos1(3),dpos(3)
 
   integer(izipx),allocatable :: xp(:,:)
-  integer(8),allocatable :: pid(:)
+  integer(4),allocatable :: pid(:)
 
 #ifdef Emode
   integer dim_1,dim_2,dim_3,ig,jg,kg
@@ -30,6 +29,12 @@ program displacement
   real(4) dx1(3),dx2(3)
   complex kpsi(ng*nn/2+1,ng,npen)
   real(4) rho_grid(0:ng+1,0:ng+1,0:ng+1)
+#endif
+
+#ifdef potential
+  real temp8[*],phi8
+  real(8) dvar[*]
+  complex delta_k(nyquest+1,nf,npen)
 #endif
 
   call geometry
@@ -118,8 +123,83 @@ program displacement
[…]

### 2019-02-03 CUBEnu/work/utilities/halofinder.f90

@@ -7,7 +7,7 @@ program halofinder
 #ifdef write_snapshot
   use variables
 #else
-  use variables, only : xp,vp,rhoc,vfield,i
+  use variables, only : xp,vp,rhoc,pid,vfield,i
 #endif
   implicit none
 

### 2019-02-03 CUBEnu/work/utilities/initial_conditions.f90

@@ -2,12 +2,21 @@
 #define sigma_8
 !#define READ_SEED
 !#define READ_NOISE
+!#define filter_phi
+!#define force_power
+
+!#define ELUCID
+!#define IniC_R0
+!#define IniC_E
 
 program initial_conditions
   use omp_lib
   use variables, only: spine_tile
   use pencil_fft
   use iso_fortran_env, only : int64
+# ifdef force_power
+    use powerspectrum
+# endif
   implicit none
   save
 
@@ -28,6 +37,29 @@ program initial_conditions
 #ifdef onehalo
   integer pid_halo(100000),np_halo,ipos(3),nw,iw
   real xv(6,100000),xv_mean(6),ang_mom(3)
+#endif
+
+  complex, allocatable :: cxyz_r(:,:,:)
+#ifdef ELUCID
+# define delta_big
+  integer,parameter :: nr=500
+  character(*),parameter :: file_deltak='../../S500_5001/cxyz_251_500_500.bin'
+#endif
+
+#ifdef IniC_R0
+# define delta_big
+  integer,parameter :: nr=400
+  character(*),parameter :: file_deltak='../../IniC/cxyz_201_400_400.bin'
+#endif
+#ifdef IniC_R1
+# define delta_big
+  integer,parameter :: nr=200
+  character(*),parameter :: file_deltak='../../IniC/cxyz_101_200_200.bin'
+#endif
+#ifdef IniC_R2
+# define delta_big
+  integer,parameter :: nr=300
+  character(*),parameter :: file_deltak='../../IniC/cxyz_151_300_300.bin'
 #endif
 
   integer(8) i,j,k,ip,l,nzero
@@ -58,6 +90,9 @@ program initial_conditions
 #ifdef PID
     integer(4) pid(npmax)
 #endif
+#ifdef force_power
+  real xi(10,nbin)[*]
+#endif
   real grad_max(3)[*],vmax(3),vf
   real(4) svz(500,2),svr(100,2)
   real(8) sigma_vc,sigma_vf
@@ -75,7 +110,10 @@ program initial_conditions
 
[…]

### 2019-02-03 CUBEnu/work/utilities/initial_conditions_nu.f90

@@ -3,6 +3,8 @@
 #define READ_NOISE
 #define MTF
 
+!#define only_phi
+
 program initial_conditions_nu
   use omp_lib
   use variables, only: spine_tile
@@ -17,12 +19,19 @@ program initial_conditions_nu
   logical,parameter :: correct_kernel=.true.
   logical,parameter :: write_potential=.true.
 
+#ifdef only_phi
+  integer(8),parameter :: nk=117
+  real tf(7,nk), vtf(7,nk)
+#else
+
 #ifdef sigma_8
   integer(8),parameter :: nk=1000
   real tf(7,nk), vtf(7,nk)
 #else
   integer(8),parameter :: nk=132
   real tf(14,nk)
+#endif
+
 #endif
 
   integer(8) i,j,k,ip,l,nzero,n_min
@@ -118,12 +127,19 @@ program initial_conditions_nu
   if (head) print*,'Transfer function'
   call system_clock(t1,t_rate)
 #ifdef sigma_8
+#ifdef only_phi
+  open(11,file='../../../../projects/caf/camb_transfer_out_z100.dat.txt',form='formatted')
+  read(11,*) tf
+  close(11)
+  vtf=tf
+#else
   open(11,file='../../tf/ith2_mnu0p05_z5_tk.dat',form='formatted')
   read(11,*) tf
   close(11)
   open(11,file='../../tf/ith2_mnu0p05_z5_v_tk.dat',form='formatted')
   read(11,*) vtf
   close(11)
+#endif
   ! normalization
   ! norm=2.*pi**2.*(h0/100.)**4*(h0/100./0.05)**(n_s-1) ! for Xin
   norm=1
@@ -150,6 +166,10 @@ program initial_conditions_nu
   if (head) print*, 's8**2/v8:', v8, s8**2/v8,nyquest
   tf(2:3,:)=tf(2:3,:)*(s8**2/v8)*Dgrow(a_i_nu)**2 ! CDM
   tf(6,:)=tf(6,:)*(s8**2/v8)*Dgrow(a_i_nu)**2 !
+  !print*,tf(1,:)
+  !print*,''
+  !print*,tf(6,:)
+  !stop
   !tf(2,:)=tf(6,:)*(s8**2/v8)*DgrowRatio(z_i,z_tf)**2 ! T_cb rather than T_c
   !tf(2:3,:)= scalar_amp*tf(2:3,:)*Dgrow(a)**2 ! for Xin
   sync all
[…]

### 2019-02-03 CUBEnu/work/utilities/module_load_mac.sh

@@ -1,13 +1,13 @@
 export FC='gfortran'
 #export XFLAG=' -O3 -cpp -fopenmp -fcoarray=single -mcmodel=medium'
-export XFLAG=' -O3 -cpp -fopenmp -fcoarray=single'
-#export XFLAG='-O3 -cpp -fcoarray=single -fopenmp'
+#export XFLAG=' -cpp -fopenmp -fcoarray=single -fcheck=all'
+export XFLAG='-O3 -cpp -fcoarray=single -fopenmp'
 export OFLAG=${XFLAG}' -c'
 export FFTFLAG='-I/usr/local/include/ -L/usr/local/lib/ -lfftw3f -lm -ldl'
 # -fopenmp cause (maybe memory) probelm: Segmentation fault: 11
 # in cumsum6
 
-export OMP_STACKSIZE=1000M
+export OMP_STACKSIZE=6000M
 export OMP_NUM_THREADS=4
 #export OMP_THREAD_LIMIT=4
 ulimit -s 61000

### 2019-02-03 CUBEnu/work/main/buffer_particle.f90
CITA version RSD

@@ -45,6 +45,7 @@ subroutine buffer_xvp(dox,dov)
   ! sync x- buffer with node on the left
   myzl=pprl(:,:,nnt,:,:)[image1d(inx,icy,icz)]
   myzr=ppr0(:,:,nnt,:,:)[image1d(inx,icy,icz)]
+  sync all
   do itz=1,nnt
   do ity=1,nnt
   do itx=1,1 ! do only tile_x=1
@@ -74,7 +75,7 @@ subroutine buffer_xvp(dox,dov)
   do itz=1,nnt
   do ity=1,nnt
   do itx=2,nnt ! skip tile_x=1
-    !$omp paralleldo default(shared) private(iz,iy)
+    !$omp paralleldo default(shared) schedule(static,2) private(iz,iy)
     do iz=1,nt
     do iy=1,nt
       if (dox) &
@@ -99,6 +100,7 @@ subroutine buffer_xvp(dox,dov)
   ! sync x+ buffer with node on the right
   myzl=ppl0(:,:,1,:,:)[image1d(ipx,icy,icz)]
   myzr=pplr(:,:,1,:,:)[image1d(ipx,icy,icz)]
+  sync all
   do itz=1,nnt
   do ity=1,nnt
   do itx=nnt,nnt ! do only tile_x=nnt
@@ -128,7 +130,7 @@ subroutine buffer_xvp(dox,dov)
   do itz=1,nnt
   do ity=1,nnt
   do itx=1,nnt-1 ! skip tile_x=nnt
-    !$omp paralleldo default(shared) private(iz,iy)
+    !$omp paralleldo default(shared) schedule(static,2) private(iz,iy)
     do iz=1,nt
     do iy=1,nt
       if (dox) &
@@ -154,6 +156,7 @@ subroutine buffer_xvp(dox,dov)
   ! sync y-
   mzl=idx_b_l(nt-ncb+1,:,:,nnt,:)[image1d(icx,iny,icz)]
   mzr=idx_b_r(nt,:,:,nnt,:)[image1d(icx,iny,icz)]
+  sync all
   do itz=1,nnt
   do ity=1,1 ! do only ity=1
   do itx=1,nnt
@@ -181,7 +184,7 @@ subroutine buffer_xvp(dox,dov)
   do itz=1,nnt
   do ity=2,nnt ! skip ity=1
   do itx=1,nnt
-    !$omp paralleldo default(shared) private(iz)
[…]

### 2019-02-03 CUBEnu/work/main/cubefft.f90

@@ -1,21 +1,38 @@
 module cubefft
   contains
   subroutine create_cubefft_plan
+    use,intrinsic :: ISO_C_BINDING
     use variables
+    !use intrinsic :: ISO_C_BINDING
     implicit none
     save
-    include 'fftw3.f'
+    !include 'fftw3.f'
+    include 'fftw3.f03'
+    integer istat,icore
+
+    !call sfftw_init_threads(istat)
+    !if(head) print*, 'sfftw_init_threads status',istat
+    !icore=omp_get_max_threads()
+    !if(head) print*, 'omp_get_max_threads() =',icore
+    !call sfftw_plan_with_nthreads(icore)
+
+    !call sfftw_plan_with_nthreads(32)
+
+
     call sfftw_plan_dft_r2c_3d(plan_fft_fine,nfe,nfe,nfe,rho_f,rho_f,FFTW_MEASURE)
     call sfftw_plan_dft_c2r_3d(plan_ifft_fine,nfe,nfe,nfe,rho_f,rho_f,FFTW_MEASURE)
   endsubroutine create_cubefft_plan
 
   subroutine destroy_cubefft_plan
+    use,intrinsic :: ISO_C_BINDING
     use variables
     implicit none
     save
-    include 'fftw3.f'
+    !include 'fftw3.f'
+    include 'fftw3.f03'
     call sfftw_destroy_plan(plan_fft_fine)
     call sfftw_destroy_plan(plan_ifft_fine)
+    !call fftw_cleanup_threads()
   endsubroutine destroy_cubefft_plan
 
 endmodule

### 2019-02-03 CUBEnu/work/main/ext_pp_force.f90

@@ -17,6 +17,7 @@ subroutine ext_pp_force
   integer(8) ntest,ip_offset,ipll1,ipll2
   integer hoc(1-ncell:nft+ncell,1-ncell:nft+ncell,1-ncell:nft+ncell)
   integer ll(np_pp_max)
+  real ftemp
 
   if (head) then
     print*, ''
@@ -35,7 +36,7 @@ subroutine ext_pp_force
   do itz=1,nnt
   do ity=1,nnt
   do itx=1,nnt
-    !! if (head) print*, '  tile:',int(itx,1),int(ity,1),int(itz,1)
+    if (head) print*, '  tile:',int(itx,1),int(ity,1),int(itz,1)
     call system_clock(t1,t_rate)
     if (np_pp_max<sum(rhoc(0:nt+1,0:nt+1,0:nt+1,itx,ity,itz))) then
       print*, 'np_pp_max too small'
@@ -48,6 +49,7 @@ subroutine ext_pp_force
     !print*,'           ',idx_b_l(0,0,itx,ity,itz)+sum(rhoc(:-1,0,0,itx,ity,itz))
     ip_offset=idx_b_l(0,0,itx,ity,itz)+sum(rhoc(:-1,0,0,itx,ity,itz))
     hoc=0; ll=0
+    print*,'    linked list'
     do igz=0,nt+1
     do igy=0,nt+1
     do igx=0,nt+1
@@ -72,12 +74,13 @@ subroutine ext_pp_force
     enddo
     enddo
 
+    print*,'    force'
     !$omp paralleldo &
-    !$omp& default(shared) &
+    !$omp& default(shared) schedule(static,2)&
     !$omp& private(igz,igy,igx,ipll1,ip1,xvec1,vreal,f_tot,ntest,kk,jj,ii) &
     !$omp& private(ipll2,ip2,xvec2,xvec21,rmag,rcut,pcut,force_pp) &
     !$omp& reduction(+:itest1,npairs) &
-    !$omp& reduction(max:f2_max_pp)
+    !$omp& reduction(max:ftemp)
     do igz=1,nft
     do igy=1,nft
     do igx=1,nft
@@ -116,15 +119,17 @@ subroutine ext_pp_force
           print*,'vfield',vfield(:,i,j,k,itx,ity,itz)
 #       endif
[…]

### 2019-02-03 CUBEnu/work/main/initialize.f90

@@ -20,14 +20,14 @@ subroutine initialize
 
   call geometry
 
-  if (head) print*, '  call create_cubefft_plan ng = ',ng
+  if (head) print*, '  call create_cubefft_plan ng = ',nfe
   call system_clock(t1,t_rate)
   call create_cubefft_plan
   call system_clock(t2,t_rate)
   print*, '  elapsed time =',real(t2-t1)/t_rate,'secs'
   sync all
 
-  if (head) print*, '  call create_penfft_plan nfe = ',nfe
+  if (head) print*, '  call create_penfft_plan nfe = ',ng
   call system_clock(t1,t_rate)
   call create_penfft_plan
   call system_clock(t2,t_rate)
@@ -37,7 +37,6 @@ subroutine initialize
   call kernel_f
   call kernel_c
 
-  istep=0
   t=0
   dt=0
   dt_old=0

### 2019-02-03 CUBEnu/work/main/pm.f90

@@ -17,6 +17,9 @@ subroutine particle_mesh
   !logical,parameter :: coarse_force=.true.
   !logical,parameter :: pp_force=.false.
   !logical,parameter :: ext_pp_force=.false.
+
+integer(4) t01,t02,t0rate
+
   integer(4),parameter :: nlayer=3 ! thread save for CIC integpolation
   integer(4) ilayer
   integer(8) idx1(3),idx2(3)
@@ -47,7 +50,7 @@ subroutine particle_mesh
     rho_f=0
     crho_f=0
     do ilayer=0,nlayer-1
-      !$omp paralleldo default(shared) &
+      !$omp paralleldo default(shared) schedule(static,2)&
       !$omp& private(k,j,i,np,nzero,l,ip,tempx,idx1,idx2,dx1,dx2)
       do k=2-ncb+ilayer,nt+ncb-1,nlayer
       !do k=2-ncb,nt+ncb-1
@@ -105,7 +108,10 @@ subroutine particle_mesh
     testrhof=testrhof+sum(rho_f(nfb+1:nft+nfb,nfb+1:nft+nfb,nfb+1:nft+nfb)*1d0)
     ! fine force ---------------------------------------------------------------
     !if (head) print*,'      fine_fft'
+    !call system_clock(t01,t0rate)
     call sfftw_execute(plan_fft_fine)
+    !call system_clock(t02,t0rate)
+    !print*, '  elapsed time =',real(t02-t01)/t0rate,'secs'; stop
     crho_f(:,:,:)=rho_f(:,:,:) ! back up
     do i_dim=1,3
       !if (head) print*,'      fine_ifft dim',int(i_dim,1)
@@ -120,7 +126,7 @@ subroutine particle_mesh
     f2_max_fine(itx,ity,itz)=maxval(sum(force_f(:,:,:,:)**2,1))
     ! fine velocity ------------------------------------------------------------
     !if (head) print*,'      fine velocity'
-    !$omp paralleldo default(shared) &
+    !$omp paralleldo default(shared) schedule(static,2)&
     !$omp& private(k,j,i,nzero,np,l,ip,tempx,idx1,idx2,dx1,dx2,vreal)
     do k=1,nt
     do j=1,nt
@@ -202,7 +208,7 @@ subroutine particle_mesh
   do itx=1,nnt ! loop over tile
     r3t=0
     do ilayer=0,nlayer-1
[…]

### 2019-02-03 CUBEnu/work/main/timestep.f90

@@ -7,11 +7,12 @@ subroutine timestep
 
   dt_old=dt
   sync all
+  sim%timestep=sim%timestep+1
 
   if (head) then
     print*, ''
     print*, '-------------------------------------------------------'
-    print*, 'timestep    :',istep
+    print*, 'timestep    :',sim%timestep
 
     dt_e=dt_max
     ntry=0
@@ -29,7 +30,7 @@ subroutine timestep
     enddo
 
     dt=min(dt_e,sim%dt_fine,sim%dt_coarse,sim%dt_pp,sim%dt_vmax,merge(sim%dt_vmax_nu,1000.,neutrino_flag))
-!dt=dt/4
+!dt=dt/2
     call expansion(a,dt,da_1,da_2)
 
     da=da_1+da_2
@@ -79,7 +80,6 @@ subroutine timestep
   endif
   sync all
 
-  ! broadcast timestep variables
   a=a[1]
   a_mid=a_mid[1]
   dt=dt[1]

### 2019-02-03 CUBEnu/work/main/update_particle.f90

@@ -53,7 +53,7 @@ subroutine update_xp()
 #   endif
     !if (head) print*,'    density loop'
     do ilayer=0,nlayer-1
-      !$omp paralleldo default(shared) &
+      !$omp paralleldo default(shared) schedule(static,2)&
       !$omp& private(k,j,i,np,nzero,l,ip,xq,vreal,deltax,ig)
       do k=1-ncb+ilayer,nt+ncb,nlayer
       !do k=1-ncb,nt+ncb
@@ -95,7 +95,7 @@ subroutine update_xp()
     ! create x_new v_new for this local tile
     !if (head) print*,'    xvnew loop'
     do ilayer=0,nlayer-1
-      !$omp paralleldo default(shared) &
+      !$omp paralleldo default(shared) schedule(static,2)&
       !$omp& private(k,j,i,np,nzero,l,ip,xq,vreal,deltax,ig,idx)
       do k=1-ncb+ilayer,nt+ncb,nlayer
       !do k=1-ncb,nt+ncb
@@ -127,7 +127,7 @@ subroutine update_xp()
     sync all
     ! delete particles
     !if (head) print*,'    delete_particle loop'
-    !$omp paralleldo default(shared) &
+    !$omp paralleldo default(shared) schedule(static,2)&
     !$omp& private(k,j)
     do k=1,nt
     do j=1,nt
@@ -153,6 +153,13 @@ subroutine update_xp()
   vp(:,sim%nplocal+1:)=0
 # ifdef PID
     pid(sim%nplocal+1:)=0
+    print*, '  check PID range:',minval(pid(:sim%nplocal)),maxval(pid(:sim%nplocal))
+    if (minval(pid(:sim%nplocal))<1) then
+      print*, '  pid are not all positive'
+      print*, '  number of 0s =',count(pid(:sim%nplocal)==0)
+      print*, minloc(pid(:sim%nplocal))
+      !stop
+    endif
 # endif
   !sim%nplocal=sum(rhoc(1:nt,1:nt,1:nt,:,:,:))
   !print*, iright,sum(rhoc(1:nt,1:nt,1:nt,:,:,:))
@@ -162,35 +169,37 @@ subroutine update_xp()
   ! calculate std of the velocity field
   !cum=cumsum6(rhoc)
[…]

### 2019-02-03 CUBEnu/work/main/variables.f90

@@ -30,12 +30,14 @@ module variables
   real(8) xq(3),deltax(3),deltav(3),vreal(3)
 
   ! FFT plans
+  integer stat
   integer(8) plan_fft_fine,plan_ifft_fine
   real vmax(3),vmax_nu(3),overhead_tile[*],overhead_image[*]
   real sigma_vi,sigma_vi_new,sigma_vi_nu,sigma_vi_new_nu
   !real vdisp(506,2),sigma_vi_old,sigma_vi
   real(4) svz(500,2),svr(100,2)
   real(8) sigma_vci,sigma_vfi,sigma_vres,sigma_vci_old,sigma_vfi_old,sigma_vres_old
+  real(8) sum_c,sum_r,sum_s
   real(8) std_vsim_c[*],std_vsim_res[*],std_vsim[*]
   real(8) std_vsim_c_nu[*],std_vsim_res_nu[*],std_vsim_nu[*]
   ! n^3

### 2019-02-03 CUBEnu/work/utilities/ang_mom_corr.f90

@@ -1,3 +1,5 @@
+!#define eig
+!#define Voronoi
 program ang_mom_corr
   use parameters
   use halo_output, only: type_halo_catalog, type_halo_info
@@ -7,25 +9,25 @@ program ang_mom_corr
   ! nf: fine grid per node per dim
 
   integer,parameter :: nc_halo_max=128
-  integer,parameter :: newbox_max=300
+  integer,parameter :: newbox_max=30
   integer,parameter :: nlist=5*(nc_halo_max+1)**3
   integer,parameter :: n_peak_max=5*nc_halo_max**3
   integer,parameter :: max_halo_np=5*(nc_halo_max+1)**3/3
-  integer(8) i,j,k,l
-  integer(4) ihalo,nhalo,ip,np,ncheck
+  integer(8) i,j,k,l,ntemp
+  integer(4) i0,ihalo,nhalo,n_compute,ip,np,nptemp,ncheck
   character(20) str_z,str_i
-  integer pid(max_halo_np),ipos(3)
+  integer pid(max_halo_np),ipos(3),ieig
   real qpos(3,max_halo_np),qpos_mean(3),vq(3),dx(3),vf,scale_factor,dx_mean(3)
-  real spin_q(3),spin_x(3),spin_t(3),spin_u(3)
-  real inertia(3,3),tide(3,3),tidu(3,3),torque(3,3)
-  real force_cdm(3),force_neu(3),torque_c,torque_u
-  real theta_qx(3000),theta_qt(3000),theta_tx(3000),theta_cu(3000),theta_ff(3000)
+  real spin_q(3),spin_t(3),spin_x(3),spin_u(3),spin_v(3),spin_r(3),spin_e(3,3),spin_vor(3)
+  real inertia(3,3),tide(3,3),tidu(3,3),torque(3,3),torquu(3,3),torqur(3,3),torque_c,torque_u
+  real force_cdm(3),force_neu(3),torque_v(3,3)
   real phi(0:nf+1,0:nf+1,0:nf+1)[*]
   real phu(0:nf+1,0:nf+1,0:nf+1)[*]
 
   type(type_halo_info) halo_info
   type(type_halo_catalog) halo_catalog
-  real,allocatable :: corr_info(:,:)
+  integer(4),allocatable :: isort_mass(:)
[…]

### 2019-02-03 CUBEnu/work/utilities/displacement.f90

@@ -13,7 +13,7 @@ program displacement
   integer(8) i,j,k,l,i_dim,iq(3),pid8,itx,ity,itz,nlast,ip,np
   integer(4) rhoc(nt,nt,nt,nnt,nnt,nnt),rhoc0(nt,nt,nt,nnt,nnt,nnt)
   integer(1) rho0(ng,ng,ng) !!!! for checking there is one and only one particle per fine grid
-  real dsp(3,ng,ng,ng),pos0(3),pos1(3),dpos(3)
+  real dsp(3,ng,ng,ng),pos0(3),pos1(3),dpos(3),pow
 
   integer(izipx),allocatable :: xp(:,:)
   integer(4),allocatable :: pid(:)
@@ -210,12 +210,8 @@ program displacement
     enddo
     deallocate(xp,pid)
 
-    print*, 'check: min,max,mean of rho0 = '
-    print*, minval(rho0),maxval(rho0),sum(rho0)/ng/ng/ng
-    if (minval(rho0)<1 .or. maxval(rho0)>1 .or. sum(rho0)/=sim%nplocal) then
-      print*, 'Warning'
-      stop
-    endif
+    print*, 'check: min,max of rho0 = '
+    print*, minval(rho0),maxval(rho0)
 
     do i_dim=1,3
       print*, 'dsp: dimension',int(i_dim,1),'min,max values ='
@@ -515,6 +511,29 @@ program displacement
       cxyz=real(cxyz)*delta_k
       delta_k=cxyz  ! backup phi(k)
       call pencil_fft_backward
+
+      ! add a 1.8 filter on phiE
+  cxyz=0
+  do k=1,npen
+  do j=1,nf
+  do i=1,nyquest+1
+    ! global grid in Fourier space for i,j,k
+    kg=(nn*(icz-1)+icy-1)*npen+k
+    jg=(icx-1)*nf+j
+    ig=i
+    kx(3)=mod(kg+nyquest-1,nf_global)-nyquest
+    kx(2)=mod(jg+nyquest-1,nf_global)-nyquest
+    kx(1)=ig-1
+    kr=sqrt(sum(kx**2))
+    kr=max(kr,1.0)
+    kr=2*pi*kr/box
+    pow=exp(-kr**2*1.8**2/2)**0.25 ! apply E-mode window function
+    cxyz(i,j,k)=delta_k(i,j,k)*pow
+  enddo
+  enddo
+  enddo
[…]

### 2019-02-03 CUBEnu/work/utilities/initial_conditions_nu.f90

@@ -3,7 +3,7 @@
 #define READ_NOISE
 #define MTF
 
-!#define only_phi
+#define only_phi
 
 program initial_conditions_nu
   use omp_lib
@@ -20,6 +20,7 @@ program initial_conditions_nu
   logical,parameter :: write_potential=.true.
 
 #ifdef only_phi
+  character(*),parameter :: string_ztf="5"
   integer(8),parameter :: nk=117
   real tf(7,nk), vtf(7,nk)
 #else
@@ -128,7 +129,7 @@ program initial_conditions_nu
   call system_clock(t1,t_rate)
 #ifdef sigma_8
 #ifdef only_phi
-  open(11,file='../../../../projects/caf/camb_transfer_out_z100.dat.txt',form='formatted')
+  open(11,file='../camb_transfer_out_z'//string_ztf//'.dat.txt',form='formatted')
   read(11,*) tf
   close(11)
   vtf=tf
@@ -198,12 +199,12 @@ program initial_conditions_nu
   if (head) print*,'Generating random noise'
   call random_seed(size=seedsize)
   if (head) print*,'  min seedsize =', seedsize
-  seedsize=max(seedsize,36)
+  seedsize=max(seedsize,12)
   allocate(iseed(seedsize))
   allocate(rseed_all(seedsize,nn**3))
 #ifdef READ_SEED
-    if (head) print*, '  Copy and read seeds from ../confings/'
-    call system('cp ../../configs/seed_'//image2str(image)//'.bin '//opath//'image'//image2str(image))
+    !if (head) print*, '  Copy and read seeds from ../confings/'
+    !call system('cp ../../configs/seed_'//image2str(image)//'.bin '//opath//'image'//image2str(image))
     open(11,file=output_dir()//'seed'//output_suffix(),status='old',access='stream')
     read(11) iseed
     close(11)
@@ -396,18 +397,6 @@ program initial_conditions_nu
   delta_k=cxyz  ! backup phi(k)
   call pencil_fft_backward
 
-
-  ! Primordial Non-Gaussianity
-  dvar=sum((r3*1d0)**2)
-  dvarg=0
-  do i=1,nn**3 ! co_sum
-    dvarg=dvarg+dvar[i]
-  enddo
-  dvarg=dvarg/nf_global/nf_global/nf_global
-
[…]

### 2019-02-03 CUBEnu/work/utilities/powerspectrum.f90

@@ -27,7 +27,7 @@ subroutine cross_power(xi,cube1,cube2)
   real cube1(ng,ng,ng),cube2(ng,ng,ng)
   real xi(10,nbin)[*]
   real amp11,amp12,amp22
-  !complex cx1(ng*nn/2+1,ng,npen),cx2(ng*nn/2+1,ng,npen)
+  complex cx1(ng*nn/2+1,ng,npen),cx2(ng*nn/2+1,ng,npen)
 
   real,parameter :: nexp=4.0 ! CIC kernel
 
@@ -35,11 +35,11 @@ subroutine cross_power(xi,cube1,cube2)
 
   r3=cube1
   call pencil_fft_forward
-  cx1=cxyz
+  cx1=cxyz/ng_global/ng_global/ng_global
 
   r3=cube2
   call pencil_fft_forward
-  cx2=cxyz
+  cx2=cxyz/ng_global/ng_global/ng_global
 
   xi=0
   sync all
@@ -67,9 +67,9 @@ subroutine cross_power(xi,cube1,cube2)
 #   endif
     xi(1,ibin)=xi(1,ibin)+1 ! number count
     xi(2,ibin)=xi(2,ibin)+kr ! k count
-    amp11=real(cx1(i,j,k)*conjg(cx1(i,j,k)))/(ng_global**3)/(ng_global**3)/(sinc**4.0)*4*pi*kr**3
-    amp22=real(cx2(i,j,k)*conjg(cx2(i,j,k)))/(ng_global**3)/(ng_global**3)/(sinc**4.0)*4*pi*kr**3
-    amp12=real(cx1(i,j,k)*conjg(cx2(i,j,k)))/(ng_global**3)/(ng_global**3)/(sinc**4.0)*4*pi*kr**3
+    amp11=real(cx1(i,j,k)*conjg(cx1(i,j,k)))/(sinc**4.0)*4*pi*kr**3
+    amp22=real(cx2(i,j,k)*conjg(cx2(i,j,k)))/(sinc**4.0)*4*pi*kr**3
+    amp12=real(cx1(i,j,k)*conjg(cx2(i,j,k)))/(sinc**4.0)*4*pi*kr**3
 
     xi(3,ibin)=xi(3,ibin)+amp11 ! auto power 1
     xi(4,ibin)=xi(4,ibin)+amp22 ! auto power 2
